### 2014-07-03 LiamWright_cos-cal.tex
Adding my file to the repo

\documentclass[a4paper]{article}

\title{Modifications to Calibration files}
\author{Liam Wright}

\begin{document}
\maketitle

\section*{Initial}

Modifications to the inspiraltest.m file:
\begin{enumerate}
   \item Move definition of $l2p = -\frac{1}{2}log(2\pi)$ from draw\_mcmc\_sample.m to mcmc\_sampler.m
   \item Changing scale frequencies from a range of $F_{min}$ to $F_{max}$ to 125Hz.
   \item In freqdomaininspiral.m had to change the scaleint to work for only one scale frequency.
   \item Running the code with every variable as a set value, and then continuely adding more variables into the prior; each variable being:
   \begin{enumerate}
      \item Masses of each spiraling star having a value between 1 and 2.5 $M_{\odot}$. Uniform prior.
      \item psi having a uniform prior with a value from $-\frac{\pi}{4}$ to $\frac{\pi}{4}$
      \item phic having a uniform prior with a value from 0 to $2\pi$
      \item Luminosity distance having a gaussian prior with a width of about 10\% of its injection distance.
      \item Iota value with a gaussian prior. Width about 30\% of a radian and an injection angle of 0 radians.

   \end{enumerate}
   Remebering to change update in extraparams(...) to 1. \\

   Comparing the standard deviations of post\_samples(:,1), the scale frequency.

\end{enumerate}

Next I used the effective\_sample\_size.m to work out the autocorrected length and effective number of uncorrelated samples in the post\_samples matrix with N parameters (same  from the prior). \\

\end{document}

### 2014-10-07 docs/cosmic_cal.tex
Added a skeleton paper draft.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\usepackage {latexsym}
\usepackage {graphicx}
\usepackage{iopams}
\usepackage {color}

%\documentclass[12pt]{iopart}
%\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
%\usepackage {amsmath}
%\usepackage {amssymb}
%\usepackage {amsfonts}
%\usepackage {amsthm}
%\usepackage {mathrsfs}
%\usepackage {natbib}
%\usepackage {latexsym}
%\usepackage {graphicx}
%\usepackage {dsfont}
%\usepackage {times}
%\usepackage {txfonts}
%\usepackage {rotating}
%\usepackage {wasysym}
%\usepackage {multirow}
%\usepackage {hhline}
\usepackage {hyperref}
%\usepackage {color}
\usepackage {bm}
%\usepackage{appendix}
\usepackage{acronym}
%\usepackage{dcolumn}   % needed for some tables
\usepackage {url}
%\usepackage[normalem]{ulem}   % temporal one in draft.

% variable shortcuts 
\newcommand{\Hub}{H_{0}}
\newcommand{\DL}{D_{\mathrm{L}}}
\newcommand{\lam}{\bm{\lambda}}
%\newcommand{\data}{\bm{d}}
%\newcommand{\sigpar}{\bm{\theta}}
%\newcommand{\nuispar}{\bm{\lambda}}
%\newcommand{\globpar}{\bm{\gamma}}
%\newcommand{\model}{\mathcal{H}}

%% ----- input git-version tag
\input{tag.tex}

\newcommand{\dcc}{LIGO-PXXXXXXXX}
\newcommand{\CM}[1]{\textcolor{red}{CM: #1}}
\newcommand{\MH}[1]{\textcolor{blue}{MH: #1}}
\newcommand{\XF}[1]{\textcolor{green}{XF: #1}}

\newcommand{\aap}{A\&A }
\newcommand{\apj}{ApJ  }
\newcommand{\apjl}{ApJL  }
\newcommand{\apjs}{ApJS  }
\newcommand{\prd}{Phys. Rev. D  }
\newcommand{\nat}{Nature}
\newcommand{\araa}{ARA\&A}
\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society  }
\begin{document}

\title{Requirements and usefulness of galaxy catalogues for gravitational wave astrophysics}

\author{L.~Wright$^1$, M.~Pitkin$^1$ \& C.~Messenger$^1$}
\address{$^1$ SUPA, School of Physics and Astronomy, University of
  Glasgow, Glasgow G12 8QQ, United Kingdom}
\ead{1002990W@student.gla.ac.uk}

\begin{abstract}
  We present an investigation into the potential of calibrating
  gravitational wave detector outputs through the use of standard
  sirens. Such signals, as measured via gravitational wave
  observations provide an estimated luminosity distance which is
  subject to uncertainties in the calibration of the data.  If a host
  galaxy is identified for a given source then its redshift can be
  obtained and combined with current knowledge of the cosmologic
  parameters.  This will yield the true luminosity distance and allow
  the direct comparison with the estimated value.  Discrepancies can
  then be used to correct the originbal calibration.
\end{abstract}

% acronym definitions
\acrodef{GW}[GW]{gravitational wave}
\acrodef{BNS}[BNS]{binary neutron star}
\acrodef{MWEG}[MWEG]{Milky Way equivelent galaxies}
\acrodef{BBH}[BBH]{binary black hole}
\acrodef{NSBH}[NSBH]{neutron star black hole}
\acrodef{SNR}[SNR]{signal-to-noise ratio}
\acrodef{GRB}[GRB]{gamma-ray burst}

%Uncomment for PACS numbers title message
%\pacs{00.00, 20.00, 42.10}
% Keywords required only for MST, PB, PMB, PM, JOA, JOB? 
%\vspace{2pc}
%\noindent{\it Keywords}: Article preparation, IOP journals
% Uncomment for Submitted to journal title message
%\submitto{\JPA}
% Comment out if separate title page not required
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{sec:intro}}

Explain the basic idea.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Gravitational wave detector calibration\label{sec:calibration}}

We need to define our model for the calibration, i.e. are we talking
about amplitude only and are we breaking this up over frequency.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Binary neutron star standard sirens\label{sec:sirens}}

We should explain what this means starting with reference
to~\cite{1986Natur.323..310S}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{GRB counterparts\label{sec:GRB}}

We should discuss the key points with regards to

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A single event with GRB counterpart\label{sec:single}}

This should be the simplest case we can do (and maybe the only case).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multiple events without counterpart\label{sec:multiple}}

An analysis or discussion of the issues regarding multiple events
without EM counterparts.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion\label{sec:discussion}}

What did we learn.

\ack

We would like to acknowledge the useful discussions with a whole bunch
of people.

\section*{References}

\bibliographystyle{unsrt}
\bibliography{masterbib}
\end{document}

### 2015-03-27 liam_src/Liam_report/GRB.tex
Adding my report. Everything needed for the .tex file to run shouwlkd be contained in the directory

\usetikzlibrary{positioning,arrows}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{decorations.markings}

\tikzset{
photon/.style={decorate, draw=black,
    decoration={coil,aspect=0}}
 }

\begin{tikzpicture}

\shade[inner color=black, outer color=white] (0,0) circle (6);

\begin{scope}[scale=0.4]
\fill[top color=yellow!40!white,bottom color=yellow!40!white ,middle color=white,,opacity=.8] (0,-10) circle (2cm and 0.5cm);
\fill[left color=yellow!50!white,right color=yellow!50!white,middle color=white,opacity=.8] (2,-10) -- (0,-2) -- (-2,-10) arc (180:360:2cm and 0.5cm);

\fill[top color=yellow!40!white,bottom color=yellow!40!white ,middle color=white,,opacity=.8] (0,10) circle (2cm and 0.5cm);
\fill[left color=yellow!30!white,right color=yellow!30!white,middle color=white,opacity=.8] (2,10) -- (0,2) -- (-2,10) arc (180:360:2cm and 0.5cm);

\end{scope}

\begin{scope}[rotate=0]
\draw[photon, color=red, ->] (0,3) -- (0,4); 
\draw[photon, color=red, ->] (.1,2) --++(80:0.9); 
\draw[photon, color=red, ->] (-0.1,2) --++ (100:0.9); 
\draw[photon, color=red, ->] (0.3,3) --++ (80:1); 
\draw[photon, color=red, ->] (-0.3,3) --++ (100:1); 
\end{scope}

\begin{scope}[rotate=180]
\draw[photon, color=red, ->] (0,3) -- (0,4); 
\draw[photon, color=red, ->] (.1,2) --++(80:0.9); 
\draw[photon, color=red, ->] (-0.1,2) --++ (100:0.9); 
\draw[photon, color=red, ->] (0.3,3) --++ (80:1); 
\draw[photon, color=red, ->] (-0.3,3) --++ (100:1);
 \end{scope}

\draw[->, thick, color=white] (2,0) arc (0:110:2cm and 0.5cm);

\shade[ball color=white!40!red] (-2,0) circle (0.5);

\shade[ball color=white!40!red] (2,0) circle (0.5);
\draw[->, thick, color=white] (-2,0) arc (180:290:2cm and 0.5cm);

\node[text width=3cm, font = \scriptsize ] at (4.5,-5.5) {\textcopyright Liam Wright};

\end{tikzpicture}

### 2015-03-27 liam_src/Liam_report/project_report.tex
Adding my report. Everything needed for the .tex file to run shouwlkd be contained in the directory

\documentclass{jpconf}

\usepackage{amsmath}
\usepackage{citesort}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{float}
\usepackage{caption}
\usepackage{tikz}
\usepackage{wrapfig}
\usepackage{fancyhdr}
\usepackage{textcomp}

\usepackage{pgfplots}
\pgfplotsset{width=7cm,compat=1.8}

\pagestyle{fancy}

\lhead{\thepage}
\chead{}
\rhead{\bfseries Cosmic Calibration}
\lfoot{}
\cfoot{}
\rfoot{}

\newcommand{\curlH}{\mathcal{H}}
\newcommand{\gws}{\tilde{h}}

\bibliographystyle{unsrt}

\graphicspath{{images/}}

\begin{document}

\title{Cosmic Calibration of Gravitational Wave Detectors}

\author{Liam Wright}

\ead{1002990w@student.gla.ac.uk}
\address{University of Glasgow, Glasgow, G12 8QQ, United Kingdom}

\begin{abstract}
  With the second generation of ground-based gravitational wave detectors on the verge of being switched on, and a predicted detection rate of between $0.4 yr^{-1}$ and $400 yr^{-1}$, the importance in quantifying the calibration errors in the detectors is vital. There may exist a source of gravitational waves with known parameters through counterpart electromagnetic observations (standard siren) which can be used as a way of independently assessing the detectors calibration. In this project, the aim was to recover the uncertainty in the calibration of the detectors using Bayesian inference on a standard siren. Throughout this project, Bayesian analysis was performed on simulated GW signal generated from an inspiralling source detected by either a single detector or a network of detectors. For a single detector and a GW signal analysed at random sky positions, the median uncertainty in the calibration error was roughly 10\%, which is comparable in accuracy compared to the current calibration method. %Something about network detectors

\end{abstract}

\section{Introduction}
\label{sec:intro}

Gravitational waves were first predicted by Albert Einstein as a result of his theory of general relativity~\cite{einstein}. It is well known that an object with mass curves the space-time around it: one would bring to mind the image of a heavy stone placed onto a flat sheet of rubber. If this mass begins to accelerate, the surrounding space-time around it will start to \textit{ripple}. This is the general concept for gravitational waves. Mathematical rigorousness and full derivations can be found in ~\cite{genrel}. Observations of GWs have so far eluded astrophysicists but the search continues. However, there have been an indirect observation in the Hulse-Taylor pulsar for example~\cite{Hulse,hulseweis}. Corrections for the energy lost during the emission of GWs explains the orbital decay in the observed pulsar orbit.

Binary systems of heavy, compact objects such as neutron stars or black holes, are predicted to be the most frequently observable source of GWs~\cite{JVei}. The waves emitted from these types of sources are an extensive study in themselves~\cite{peadvanced}. Understanding the \textit{waveforms} and the algorithms used in analysis of data collected from the detector are vital in obtaining a direct detection, and continue to be import areas of research in the field~\cite{pN,TF2}. The source considered throughout this project will be coalescing binary neutron stars with equal masses, described in section~\ref{sec:cbc}.

Searches for GWs are currently carried out through ground-based interferometers, Virgo, LIGO, and GEO 600~\cite{barish,virgo,geo}. This project concentrated on Virgo and LIGO only. Previous science runs of ground-based GW detectors at initial sensitivity have so far been unsuccessful in directly detecting a GW~\cite{nodetect}. However, an indirect detection has been observed through predicted energy loss of binary pulsars. With the $2^{nd}$ of GW detectors about to begin collecting data, and to be at full sensitivity in 2018~\cite{Harry}, a direct observation is predicted to occur in the not too distant future.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Data mining in GW research is a colossal task due to the signals generated being incredibly weak. Probability statistics are used throughout the field and have proven capable of being able to detect sources and parameterise the defining GW variables through blind testing of a data set~\cite{blind}. Bayesian inference is the method used in data analysis and has two main areas of research: hypothesis testing, which is used in detecting a GW signal in a data set, and parameter estimation of the GW variables. This project concentrated on parameter estimation.

In order for a certain GW detection, all sources of error in the must be accounted for. Proposed in this project is a new method of calibration to be used throughout the analysis itself; an external calibrator. Through an assumed known source of GWs, the error in the calibration is quantified using parameter estimation techniques~\cite{JVei}.

%%Cosmology and effect of a detection

%%% rewrite %%%%%%%%%%
Being able to determine GW parameters of the source gives one that ability to describe the underlying physics of the system. For the binary neutron star system being analysed in this project, a detection would allow for the discovery of the equation of state for such a system and the conditions for the evolution of a binary system~\cite{dynamic}. A detection of gravitational waves would also open up a whole new window into our astrophysical and cosmological understanding of our universe. Allowing for testing genuinely strong-field general relativity and can even be used to detect``dark matter"~\cite{darkmatter}, objects unable to be seen by electromagnetic observations. Gravitational wave detection would also prove vital for determining cosmological parameters~\cite{Schutz,KThorne,chernofffinn}.
%%%%%%%%%%%%%%%%%%%%

In section~\ref{sec:methofcal} the proposed new method of calibration is discussed, using parameter estimation techniques described in section~\ref{sec:bpe}. Section~\ref{sec:nest} gives a brief overview of the algorithm used throughout the project. The results and analysis of the different simulations are given in section~\ref{sec:results}. The conclusion and discussion of the continuation of this project are given in section~\ref{sec:concl}.

\section{Gamma Ray Bursts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A GW detection on its own right would be ground breaking, allowing astrophysicists to observe the sky above in a whole new spectrum. However, to increase the scientific return on a direct GW detection, a counterpart EM observation of the source itself would need to be seen.
There are two approaches to observing a source both in EM and GW spectrum's~\cite{grb}: 
\begin{enumerate}
\item locating and observing a source with an EM follow-up to a GW detection
\item searching for a GW signal generated by a source seen through EM observations
\end{enumerate}
Both of these methods have their pros and cons. Firstly, an EM follow-up would be very difficult due to the duration of GW source system detectable by ground-based interferometers. A neutron star binary system (NS-NS binary) has a duration of $\sim15$ minutes~\cite{bnstime}. Errors in the sky position from a GW detection would also create issues for an EM follow-up. The sky localisation obtained from the ground-based detectors gives a large field of view for the EM observatories to search, roughly an area spanning $100$'s of degrees~\cite{grb}.

The more promising detection approach then is a follow-up GW search from an EM trigger, especially for the source of GW considered in this paper. The merger time for NS-NS binaries are known with an uncertainty within a few seconds for the EM radiation emitted from this system~\cite{grb}.

 Identifying the source of GWs redshift would then lead to calculating the Hubble constant and other cosmological parameters~\cite{Schutz}. The evolution of the system could be better understood, allowing for searches with greater accuracy. Counterpart EM observations of a GW source will also remove uncertainty in the defining GW parameters, a fact which was taken advantage of for this project.

\begin{wrapfigure}{r}{0.5\textwidth}
  \tikzset{every picture/.style={scale=.7}}%
  \centering
  \input{GRB}
  \caption{\textit{A graphical representation of EM radiation emitted from a binary neutron star merger, given as the red spheres. The gamma ray photons (red arrows) are concentrated into the radiation beams (yellow regions). }}
  \label{fig:grb}
\end{wrapfigure}

The source of GW considered in this paper will be a coalescing binary system consisting of neutron stars. As well as being the most probable source of GWs detectable by ground-based interferometers, they are also expected to emit EM radiation. As shown in figure~\ref{fig:grb}, the emission occurs due to the high velocity with which the stars coalesce. The coalescences are thought to be the progenerators of short-duration gamma ray bursts (SGRBs)~\cite{grb}. Observations of this radiation has been see, through the Swift satellite and follow-up ground-based telescopes~\cite{swift}. Unfortunately, no observation of SGRBs have been seen within the range of advanced generation of LIGO and Virgo. It is probable, however, that such a source exists well within the range of advanced LIGO/ Virgo, but has simply not been observed~\cite{grb}. This then validates the assumption made throughout the project.

In this paper, a key assumption has to be made in order to test this new method of calibration. It is assumed that there exists a source of GWs with enough information known through EM observation such that it can be used as a \textit{standard candle}, i.e. an accurately known and well modelled source of GWs. In this project (for the source of GWs being analysed), a GRB observation is assumed to have been seen. This allows for discrepancies in the detector calibration itself to be parameterised. This is the proposed new method of calibration for the detectors, for all types of detectors.

\section{Method of Calibration}
\label{sec:methofcal}

The technique used throughout GW research in calibrating the detectors is carried out through a complicated system of physical manipulations of the Fabret-Perot Michelson interferometer; where an elaborate feedback system is used to sustain a defined measurement in arm length difference between the moving mirrors~\cite{LIGOCal}. In general, the interferometers are set to move at a frequency corresponding to a GW signal. An error signal is then found through the Pound-Driver-Hall technique~\cite{Black} for mirror stabilisation which is proportional to the difference in mirror separation of the Fabret-Perot Michelson interferometer. The Pound-Drever-Hall technique is essentially a control loop feedback system in which the difference output (in GW detectors, this is the distance the mirrors move due to the GW) and the the output set is reduced to the highest degree of certainty. This is just a brief overview of the calibration method currently used, and is an extensive study in itself, see~\cite{Vitale} and references therein. For advanced LIGO and Virgo, at the frequency range used in this project, the error in the calibration using the 'hardware injection method' is roughly $10\%$~\cite{Vitale}. This is a benchmark set for the error estimation using the proposed new method in this project.

The hardware method of calibrating the ground-based detectors is incredibly complex in its application, with many sources of error which have to be accounted for. The new method of calibration proposed and tested in this project is a lot simpler. The method acts as an external calibrator, to be carried out using the data collected by the detector, $\tilde{d}(f)$. We assume the collected data contains some unknown and unaccounted for calibration error. The uncertainty in $\tilde{d}(f)$ can then be evaluated using methods of parameterisation already carried out in GW research. This method of parameterising GW defining variables is then carried out on a known source, a standard candle as described above, which allows one to measure the calibration uncertainty in the data set.

One can estimate the error in the calibration using parameter estimation methods, or more precisely Bayesian inference. In this project the calibration offset used throughout is simply a scale factor, $s$. This effects the data set collected by a simple multiplication:
\begin{equation}
  \label{eq:scaledata}
  \tilde{d}_m(f) = \tilde{d}_e(f) \times s
\end{equation}

where $\tilde{d}_m(f)$ is the data-set measured by the detector, and $\tilde{d}_e(f)$ is the exact data set, with no error. A description of the collected data is given in section~\ref{sec:cbc}. The scale factor being parameterised, $s$, represents all known sources of error in the system. This is not entirely realistic as $s$ is known to be a function of frequency. Testing a more realistic calibration error is work to be done in the future.

Being able to resolve the calibration scale factor with the least amount of uncertainty was the aim of this project; to show that this method is competitive with current techniques. If the calibration error can be resolved with known uncertainty, GW data analysis can then be carried out with more certainty. This is useful as an entirely independent method of assessing the uncertainty in the current calibration method.

\section{Bayesian Inference \& Gravitational Wave Data Analysis}
\label{sec:bpe}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Basic Introduction

Bayesian inference is a branch of probability statistics used throughout scientific research today. In its generality, Bayesian statistics can be applied to many fields from particle physics to predicting the weather. The framework for this method of probability statistic is based on Bayes' Theorem, published in 1763~\cite{Bayes}. Bayes' theorem, given as 
\begin{equation}
  \label{eq:bayesthm}
  p(A|B) \propto p(A) p(B|A),
\end{equation}

states that the \textit{posterior} probability, the probability obtaining statement A given statement B (where probability is represented by $p(...)$ and '$|$' means 'given'), is proportional to the the \textit{prior probability}, $p(A)$, and the \textit{likelihood function}, $p(B|A)$. The prior probability is the known (or unknown) information about statement A prior to any analysis. The likelihood function represents the probability of obtaining statement B assuming statement A is true~\cite{Sivia}.

Bayesian inference is used throughout GW research and has many applications. In this project however, parameter estimation is the primary concentration. In parameter estimation, one wishes to paramaterise the set of unknown parameters, $\vec{\theta}$ which represents all the unknown variables defining the GW (including the scale factor error, $s$), a description of which is given in section~\ref{sec:cbc}.

Bayes' theorem in the context of GW parameter estimation then becomes
\begin{equation}
\label{eq:bayesgw}
  p(\vec{\theta} | \vec{d}, \curlH) = \frac{p(\vec{\theta} | \curlH) p(\vec{d} | \vec{\theta}, \curlH)}{p(\vec{d} | \curlH)}
\end{equation}

where $\vec{d}$, and $\curlH$ are the data set being analysed and the hypothesis respectively. Both of these terms are described in section~\ref{sec:cbc}. The prior, $p(\vec{\theta}| \curlH)$, and likelihood, $p(\vec{d} | \vec{\theta}, \curlH)$ are described in section~\ref{sec:pandl}. A new term has been introduced in equation~\ref{eq:bayesgw}: the denominator $p(\vec{d}|\curlH)$, or more commonly known as the Bayesian \textit{evidence} given as
\begin{equation}
  \label{eq:evidence}
   Z = \int \limits_{\vec{\Theta}} p(\vec{d}|\vec{\theta}, \mathcal{H}) p(\vec{\theta} | \vec{d}, \mathcal{H})  \;d\vec{\theta} 
\end{equation}

where $\vec{\Theta}$ is the vector representing all the parameters (the vector representing all unknown parameter $\vec{\theta}$ is a subset of $\vec{\Theta}$). Since the prior is just a value between $0$ and $1$, the evidence is essentially the expectation value of the likelihood, weighted by the prior. It is then clear that calculating equation~\ref{eq:evidence} can become computationally demanding when the dimensionality of $\vec{\Theta}$ becomes large; as it does in the case of GW research. To overcome this problem the method of solving equation~\ref{eq:evidence} and indeed equation~\ref{eq:bayesgw} used was \textit{nested sampling}. A description of which is given in section~\ref{sec:nest}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hypothesis and Waveform\label{sec:cbc}}

In this project the hypothesis considered, $\curlH$, simply states that the data, $\tilde{d}(f),$ collected by a generic detector contains the GW signal, $\gws(f)$, and some noise, $\tilde{n}(f)$~\cite{JVei},
\begin{equation}
  \label{eq:data}
  \tilde{d}^D(f) = \gws^D(f) + \tilde{n}^D(f)
\end{equation}
where $D$ represents the detector site. In this project there will be three detectors in question: LIGO Livingston (L), LIGO Hanford (H), and Virgo (V); $D = \{L,H,V\}$. Testing different hypotheses is the method used to determine whether a data set contains a GW signal or not~\cite{SathSchutz}. This was not explored in this project so the above hypothesis is consistent throughout this paper.

 Coalescing binary systems of compact objects, such as neutron stars or black holes, are the most probable source of GWs detectable by ground-based interferometers~\cite{peadvanced}. A direct detection of GWs would allow for a fuller understanding of the mechanics of these systems. Compact binary coalescence's (CBCs) are systems of two compact object inspiralling, with reducing separation, until the inevitable merger of the objects. This orbital motion creates ripples in space-time; this is the GW from this motion. For gravitational wave research, the interests lie in the GW signal generated, or the \textit{waveform}, shown in figure~\ref{fig:signal}.

%Need to talk about the noise being a gaussian with <.> = 0 and stuff

In this paper, a brief overview of the mathematical breakdown for the waveform is shown. For a more rigorous derivation, see~\cite{SathSchutz} and references there in. The waveform used throughout this project is a simplified system where the neutron stars are non rotating about their axes, in a perfectly circular inspiralling orbit; known as the TaylorF2 model~\cite{TF2,pN}. A more complex and realistic model could be tested in future for this calibration method. Parameter estimation using a variety of complex waveforms are carried out in~\cite{peadvanced}.

The general equation for a GW signal detected at a generic detector, in the reference frame of the detector site, is given as~\cite{SathSchutz}

\begin{equation}
  \label{eq:gravsig}
   \gws^D(f) = \big[ F_{+}^D(\alpha, \delta, \psi)\gws_{+}(f) + F_{\times}^D(\alpha, \delta, \psi)\gws_{\times}(f)\big]
\end{equation}

where $F_{+,\times}(\alpha, \delta, \psi)$ are the antenna response function of the detector, dependent on the right ascension, $\alpha$, declination, $\delta$, the polarisation angle $\psi$, and time. Testing the antenna patterns is carried out in section~\ref{sec:rndsp} of this report when the sky positions, $(\alpha, \delta)$, are randomised across the entire field of view for the detector.

The polarisation amplitudes of the GW, given by $\gws_{+}(f)$, and $\gws_{\times}(f)$, represent the direction with which the GW compresses and stretches space-time. The form of these two amplitudes depends on the source being analysed. As mentioned above, the model used throughout this project was the TaylorF2 model. The parameters defining the GW polarisation amplitudes using the TaylorF2 model are:

\begin{itemize}
\item the total mass, $M = m_{1}+m_{2}$, where $m_{1,2}$ are the masses of the binary stars
\item the symmetric mass ratio $\eta = \frac{m_{1}m_{2}}{(m_{1}+m_{2})}$
\item chirp mass, $\mathcal{M} = \eta^{3/5}M$
\item luminosity distance, $D_{L}$, measured in Mpc ($1 Mpc \approx   3.1\times10^{22}m  $)
\item the angle of inclination $\iota$. This is the angle from the source, taken as the angle between the normal of the orbital plane and the observer orientation.
\item phase of the wave, $\psi$, which defines the polarisation of the angle of the GW, shown in equation~\ref{eq:psi}.
\item the coalescence time, $t_c$, and phase of coalescence, $\phi_c$.
\end{itemize}

Together with the sky position, these parameters form the vector $\vec{\Theta} = \{\mathcal{M}, \eta, D_{L}, \iota, \alpha, \delta, \psi, t_{c}, \phi_{c}\}$ which defines the entire GW signal. 
The form of the polarisation amplitudes using the TaylorF2 model are given as~\cite{SathSchutz,pN,TF2}

\begin{equation}
  \gws_{+}(f) = \frac{1+\cos^{2}\iota}{2D_{L}} \frac{\mathcal{M}^{5/6}}{\pi^{2/3}} \sqrt{\frac{5}{24}}f^{-7/6}e^{-i\psi(f)}
\end{equation}
\begin{equation}
  \gws_{\times}(f) = i\frac{\cos^{2}\iota}{D_{L}}\frac{\mathcal{M}^{5/6}}{\pi^{2/3}} \sqrt{\frac{5}{24}}f^{-7/6}e^{-i\psi(f)}
\end{equation}

where the phase of the GW, $\psi$, is 
\begin{equation}
  \label{eq:psi}
  \psi(f) = 2\pi f t_c - \phi_c + \psi_N(\eta)\sum\limits_{k=0}^4{\psi_k(\eta) (\pi(m_1 + m_2)f)^{(k-5)/3}}.
\end{equation}

The terms $\psi_N$ and $\psi_k$ are the standard Newtonian and post-Newtonian coefficients. Derivations of these two equations can be found in~\cite{pN} and references there-in. The simulated gravitational wave signal used throughout this project can be seen in figure~\ref{fig:signal}. As one can see, the signal strength is very faint (an order of magnitude around $10^{-24}$). This is the the biggest problem in detecting GWs.

As mentioned above, the source considered throughout this project thought to produce EM radiation; SGRBs. In this project, we assume that SGRBs of the GW source described above have been observed. If a GRB has been seen, then the redshift of the merger can be calculated. This then allows for the luminosity distance to be calculated and in turn other defining parameters can be obtained (or known to a degree of uncertainty)~\cite{SathSchutz}.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{grav_signal}
  \caption{\textit{The gravitational waveform produced using coalescing binary neutron stars. The wave is represented in the frequency domain with masses $m_{1,2} = 1.4M_{\odot}$ at a luminosity distance $D_{L} = 200 \,Mpc$.}}
  \captionsetup{justification=centering}
  \label{fig:signal}
\end{figure}

%More explination
Finally, the last term in equation~\ref{eq:data} must be discussed; the noise of the model, $\tilde{n}(f)$. Throughout this project, we have added no noise to the signal, although as discussed later we do include a noise estimate in the likelihood.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Prior and Likelihood}
\label{sec:pandl}

The prior probability function, $p(\vec{\theta} | \curlH)$ in equation~\ref{eq:bayesgw}, represents our prior knowledge in $\theta$ before the analysis.The range of parameters used throughout this paper were chosen due to their inability to be justifiably known through a GRB assumption. The parameters were drawn from the ranges for any signals generated in this analysis, given in table 1.

\begin{table}[H]
  \centering
  \begin{tabular}{|c|c|}
%  \caption{Prior parameters}
    \hline
    Parameter & Distribution \\
    \hline
    $\psi$   &  U($0,\pi/2$)\\
    $\iota$   & N($0,0.1$) \\
    $\phi_c$  & U($0,\pi/2$)\\
    $t_c$     & U($t_{o}-0.01, t_{o}+0.01$) \\
    $s$       & U(0,2) \\
    \hline
  \end{tabular}
  \label{tab:prior}
  \caption{\textit{The parameters used throughout the analysis with some degree of uncertainty. U(a,b) is a uniform distribution between a and b. N($\mu, \sigma$) is a normal distribution, with a mean $\mu$ and standard deviation $\sigma$. $t_{o} = 9\times10^{8}$}}
\end{table}

Through the assumption made above allows us to constrain the system, maintaining a small number of parameters in the prior. For a realistic parameter estimation of the source, a larger range of values with a bigger distribution. This was not tested due to time constraints but is proposed as work for the future.

The likelihood function used throughout this project is given for a generic detector D as~\cite{JVei}, 
\begin{equation}
  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp(-(d^{D}(f) - h^{D}(f)|d^{D}(f) - h^{D}(f))/2)
\end{equation}
where $(\cdot |\cdot )$ is a representation of the inner product. This derivation is given in the appendix. Using the inner product of two real functions, and approximating for discrete integer values, the likelihood then becomes

\begin{equation}
  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp\Bigg[ 2\Delta f\sum_{k>0}\frac{|\tilde{d}(f_{k}) - \tilde{h}(f_{k})|^{2}}{S(f_{k})} \Bigg]
\end{equation}

\begin{wrapfigure}{l}{0.5\textwidth}
  \centering
  \includegraphics[width=0.5\textwidth]{PSD}
  \caption{\textit{Noise spectral density for advanced LIGO and Virgo.}}
  \label{fig:psd}
\end{wrapfigure}

where $S(f_{k})$ is the \textit{power spectral density} (PSD). This defines the sensitivity of the detector as it a measure of the strain in the detector due to noise~\cite{SathSchutz,sens}. The PSD for advanced LIGO and Virgo are shown in figure~\ref{fig:psd}. The figure shows that the strain effect for advanced LIGO is less than it is for Virgo; LIGO has a better sensitivity. This is a factor in analysing the calibration error, given in section~\ref{sec:results}. The subscript $k$ represents discrete values of relevant frequency bins (As carrying out the integral would become computationally costly). This likelihood will be used throughout the project. The factor of $\Delta f$ comes from the approximation of the inner product of two real functions (as it is integrated over frequency space).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Nested Sampling\label{sec:nest}}

In this paper the nested sampling method was used to solve equation~\ref{eq:bayesgw}~\cite{Sivia}. This algorithm is predominantly used in hypothesis testing, i.e. comparing a noise only data set with one possibly containing a GW to show whether there was a direct detection. As stated above, there lies a problem when solving Bayes' theorem in high dimensional parameter space. It becomes very difficult and computationally demanding when solving the integral in equation~\ref{eq:evidence}. It becomes far to complicated to compute analytically, and cannot be numerically calculated without sever computational cost.

Nested sampling can handle this complex integral by simply changing it from a $|\vec{\Theta}|$ dimension integral into a parameter \textit{volume} integral. Equation~\ref{eq:bayesgw} becomes:

\begin{equation*}
  \Rightarrow Z = \int \limits_{\vec{\Theta}} p(\vec{d}|\vec{\theta}, \mathcal{H}) \pi \;d\pi 
\end{equation*}
\begin{equation*}
  \Rightarrow Z \approx \sum \limits_i^N p(\vec{d}| \vec{\theta_i}, \curlH) \pi_i
\end{equation*}

\begin{equation*}
  \Rightarrow Z \approx \sum \limits_i^N L_i\pi_i
\end{equation*}

where $\pi_i$ represents the \textit{prior volume} of the $i^{th}$ point in N randomly chosen positions on parameter space; N \textit{livepoints}. $L_{i}$ is the likelihood of the $i^{th}$ position. In the presence of a GW signal, the evidence is usually concentrated into one region of parameter space defined by the \textit{information} of the data set~\cite{Sivia}. It is then key to not waste computational time by analysis regions of parameter space which do not contain any information about the signal.

N livepoints are chosen randomly in parameter space from the prior volume and $L_{i}$ and $\pi_{i}$ are calculated for each point. In this description, the prior parameters defining $\pi$ are assumed to be uniformly distributed, hence the $i^{th}$ point can be said to occupy a volume of $V_{tot}/N$ where $V_{tot}$ is the entire parameter volume.   
The process of the algorithm is carried out as~\cite{JVei}:
\begin{enumerate}
\item set $j = 0$, $Z_{j} = 0$, $\pi_{j} = 1/N$ 
\item While $Z_{j} < Z_{thresh}$:
  \begin{itemize}
  \item $j = j + 1$
  \item $L_{min} = min(\{L_{i}\})$
  \item $\pi_{j} = \pi_{j-1} - 1/N $
  \item $Z_{j} = Z_{j-1} + L_{min}\times \pi_{j}$
  \item Discard the position giving $L_{min}$
  \item Chose a new random point where $L_{new}>L_{min}$ and concatenate into $\{L_{i}\}$
  \end{itemize}

\end{enumerate}

As stated in the pseudo-code, the algorithm is dependent on a threshold term, $Z_{thresh}$. This is a value determined by the change in evidence as the algorithm is carried out. When the difference in the evidence integral between iterations becomes significantly small, then the algorithm can stop. Once the algorithm has finished the discarded livepoints can be re-sampled to give points drawn from the posterior distribution of the parameters. The accuracy of the algorithm and the parameterisation is dependent on the number of livepoints being analysed. Balancing accuracy with computational cost is one of the debated topics in GWs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results\label{sec:results}}

The aim of this project was to determine how well the calibration error could be parameterised in a software injected gravitational wave source. The error was estimated through numerous situations but the basis of the simulations remained the same throughout the project: a single source of GWs at varying signal to noise ratios. The GRB assumption was analysed as well by relaxing some of the defining GW paramaters and adding some uncertainty into their value, given in section~\ref{sec:pandl}.

These values where chosen due to the assumed counterpart EM observation. These values could not be justified by that assumption so they were added into the prior with reasonable uncertainty.

The estimation of the error, given in its simplest form as a scale factor, was performed for numerous systems: single detector, a network of detectors, random sky positions, etc. Before the parameter estimation was carried out, the relationship between distance of the source relative to the detector and the signal to noise ratio (SNR) needed to be found.

The SNR is given by
\begin{equation}
  \label{eq:snr}
  SNR = \sqrt{\Bigg( \frac{2 |\gws^D(f)|^2}{S^D(f)} df \Bigg)}
\end{equation}

where the noise spectral density, $S^D(f)$ is not a function of distance, shown in section~\ref{sec:pandl}. For coalescing binary neutron stars of masses $m_{1,2} = 1.4\,M_{\odot}$ and  $D_{L}200\,Mpc$, the expected SNR is roughly 8. The SNR was determined from $D_{L} = [10,200]\,Mpc$  using a randomly chosen sky position which defines the antenna patterns and the GW signal. The range of luminosity distance used in this paper represented a high level of SNR up to the distance at which the first predicted detectable GW source; $D_{L} = 200\,Mpc$. The detector limits for the advanced era is a luminosity distance of $\sim 450\,Mpc$. This was not initially tested as the SNR was low enough to cause problems with computational time. It was however tested in the random sky position simulations, see section~\ref{sec:rndsp}. The initial value for the calibration error used in the analysis was chosen randomly.

\begin{figure}[h]
  \centering
  \includegraphics[width = \textwidth]{DvsSNR-mult-det}
  \captionsetup{justification=centering}
  \caption{\textit{The distance and SNR relationship for all three detectors: LIGO Hanford, H1, LIGO Livingston, L1, and VIRGO.}}
  \label{fig:distsnr}
\end{figure}

\subsection{Single Detector with GRB Observation \label{sec:sdgrb}}

To begin with, parameter estimation was carried out on the simplest case: a single source of GWs using a single detector, LIGO Hanford, at different levels of SNR. The aim was to resolve the calibration scale factor with the highest degree of certainty. 
Before testing the uncertainty resolution using a the full range of prior parameters, we fixed all values to known signal values and only searched over the scale factor. Results shown for with and without a full prior range are given in figures~\ref{fig:sd-empty-D} and~\ref{fig:sd-non-empty-D} respectively.

\begin{figure}[h]
  \centering
  \includegraphics[width = \textwidth]{SD_empty_D10_200}
  \caption{\textit{Posterior distribution on the calibration scale factor for LIGO Hanford. Simulations carried out using an entirely known source of GWs.}}
  \label{fig:sd-empty-D}
\end{figure}

The figure above is a histogram of the calibration error, with each curve representing a different distance (SNR). The black line at $s = 0.8$ is the initial scale factor value used to multiply the unanalysed data set. This representation of the initial calibration scale factor will be continuous throughout the paper. 
One would expect with lower SNR (greater $D_{L}$), the estimation of the parameter would become less certain. This is clearly shown in figure~\ref{fig:sd-empty-D} as the standard deviations of the curves increases with smaller SNR/larger luminosity distances. For this entirely known source of GWs, we would expect to find the uncertainty in $s$ to be approximately $1/$SNR; this is what we see in figure~\ref{fig:sd-empty-D}. The error in estimating $s$ for $D_{L} = 200 \,Mpc$ is roughly $\sim 20\%$. This is larger than the current method of obtaining the uncertainty in calibration. A full test of the new method proposed is carried out in section~\ref{sec:rndsp}.

The next estimation of the calibration scale factor was performed with a full range of prior parameters, giving a more realistic system. As before, the simulations were carried out for differing levels of SNR and a histogram of the calibration error was made, seen in figure~\ref{fig:sd-non-empty-D}.

\begin{figure}[h]
  \centering 
  \includegraphics[width = \textwidth]{SD_non_empty_D10_200}
  \caption{\textit{Posterior distribution on the calibration scale factor for LIGO Hanford. Simulations carried out with some parameters in the prior.}}
  \label{fig:sd-non-empty-D}
\end{figure}

The error in the calibration scale factor, as before, increases with decreasing SNR (increasing $D_{L}$). The magnitude of this error however is larger than in the previous simulation. With a full prior range given in section~\ref{sec:pandl}, the standard deviation increases; at $D_{L} = 200\,Mpc$ the error in $s$ is $\sim 24\%$. When comparing this new method of calibration scale factor estimation with the current, feedback loop system method used today, this method is not as accurate. However, it should again be noted that this method is an independent estimate of the calibration accuracy.

\subsection{Multiple Detectors with GRB Observation}
\label{sec:mdgrb}
A network of detectors are used throughout GW research. Being able to show the effect of the calibration error and recovering its uncertainty in the analysis would help in giving a detection more confidence. In this section, the simulations were carried out using multiple detectors; LIGO Hanford, LIGO Livingston, and Virgo. In reality, multiple detectors are used in order to determine the (rough) sky position of a signal. Due to their locations, if one site detects a signal before the others, then the position of the source can be roughly worked out. The aim was, as before, to estimate the calibration scale factor with the least amount of uncertainty for differing levels of SNR for a single source of GWs.

For a network of detectors, the overall likelihood function becomes a product of each likelihood for all detectors, i.e.

\begin{equation}
  \label{eq:mult-likeli}
  p(\vec{d}| \vec{\theta}, \curlH) = \prod \limits_D p(\vec{d}^D| \vec{\theta}, \curlH)
\end{equation}

where $D = \{H,L,V\}$. The initial calibration scale factors where again chosen at random ($s = [4.6, 2.5, 6.1]$ for LIGO Hanford, LIGO Livingston, and Virgo respectively). The simulations were the same as before, using the same parameter values. The GRB assumption was again tested in these simulations as a handful of parameters were relaxed and given some uncertainty. The results of these simulations are shown in figure~\ref{fig:mult-empty-D} and figure~\ref{fig:mult-non-D}.

\begin{figure}[h]
  \centering
  \includegraphics[width = \textwidth]{MD_empty_D10_200}
  \captionsetup{justification=centering}
  \caption{\textit{Posterior distribution on the calibration scale factor for all three detectors. Simulations carried out using an entirely known source of GWs.}}
    \label{fig:mult-empty-D}
\end{figure}

Figure~\ref{fig:mult-empty-D} shows the histograms of the calibration scale factor for each detector. These simulations used an entirely known source of GWs, at a randomly chosen sky position (the same sky position as before). As the figure shows both LIGO sites, Hanford and Livingston, are able to resolve the error with a measurable degree of uncertainty. The standard deviation for $s$ at $D_{L}=200\,Mpc$ is $24\%$ and $41\%$ of the true scale factor for Hanford and Livingston respectively. The estimation for the calibration error using VIRGO however, was not as accurate. It can be seen that, for increasing luminosity distances, the resolution of $s$ becomes increasingly uncertain until the point at which the distribution becomes almost uniform; $D_{L} \approx 120\,Mpc$. There are a few reasons why this may be. Firstly, the sky position used may not be 'ideal' for generating the antenna patterns. The most probable cause of the estimation being very poor will lie in the detector itself. The interferometer itself is not as sensitive the two LIGO sites, hence the estimation of parameters using this site will not be as precise.

\begin{figure}[h]
  \centering
  \includegraphics[width = \textwidth]{MD_non_empty_D10_200}
  \captionsetup{justification=centering}
  \caption{\textit{Posterior distribution on the calibration scale factor for all three detectors. Simulations carried out with some prior parameters. } }
  \label{fig:mult-non-D}
\end{figure}

The histograms shown in figure~\ref{fig:mult-non-D} were made using the full range of prior parameters. The uncertainty in the calibration error is larger with a less known source of GWs. VIRGO, again, has the least certain calibration error resolution with the distribution becoming even more uniform. The standard deviation for both LIGO curves become larger with increasing luminosity distance as before. For $D_{L} = 200Mpc$,  the standard deviation is $\tilde 30\%$ of the true value. When comparing this to the current method, this again is not as accurate.

\subsection{Random Sky Position}
\label{sec:rndsp}
Throughout the duration of this project the antenna patterns were defined by a randomly chosen sky position, as seen in equation~\ref{eq:gravsig}. This did not give a true representation of the estimation capable by algorithm as this position could be a bad one; making parameterisation more difficult and less precise. To then test this new method of calibration fully, the entire field of view of the detector must be analysed. The sky positions $\alpha$, $\delta$, were distributed as:
\begin{table}[H]
\centering
  \begin{tabular}{|c|c|}
    \hline
    Parameter & Distribution \\
    \hline
    $\alpha$       & U($0$,$2\pi$) \\
    $\sin(\delta)$ & U($-1$,$1$)    \\
    \hline
  \end{tabular}
  \caption{\textit{The distribution of sky position. U(a,b) indicates a uniform distribution between a and b.}}
  \label{tab:sp}
\end{table}

Figures~\ref{fig:snr-rndsp} and~\ref{fig:std-rndsp} were obtained using a luminosity distance $D_{L} = 200 Mpc$ as this is the expected luminosity distance predicted to host the first observable GW~\cite{peadvanced}. $100$ different sky positions were tested using the LIGO Hanford site. The number of locations was limited by the length of time these simulations took to carry out. Increasing the number of locations would give a more accurate representation of the estimation ability; this can be carried out in future. Finally, a full range of prior parameters were also used in the simulation set-up.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{rand_sp_snr_D200}
  \caption{\textit{Signal to noise ratio for every sky position analysed at $D_{L} = 200 \,Mpc$. The vertical black line represents the SNR for the sky position used throughout previous simulations}}
  \label{fig:snr-rndsp}
\end{figure}

Figure~\ref{fig:snr-rndsp} shows the histogram of SNRs for the entire field of view for the detector. Predictions state that at a luminosity distance of $D_{L}=200\,Mpc$, the SNR will be roughly 8. This is consistent with the results obtained. The vertical black line shown in figure~\ref{fig:snr-rndsp} represents the SNR used throughout the project using the randomly chosen $\alpha$ and $\delta$. One can see that this is a less-than-average SNR and may explain the resolution of the calibration error not being too precise. This then is a promising result and the full testing of this new calibration method can be carried out.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{rand_sp_std_D200}
  \caption{\textit{Standard Deviation (as a $\%$ of the true scale factor value) for every sky position analysed at $D_{L}=\,200Mpc$. The vertical black line indicates the median of the distribution.}}
  \label{fig:std-rndsp}
\end{figure}

Figure~\ref{fig:std-rndsp} is the distribution of fractional standard deviations on s for each sky position analysed, given as a percentage. The vertical black line in this figure shows the \textit{median} of the distribution. The results show that the algorithm can resolve the calibration error in the data within the range of $\sim 7\%$ to $\sim 14\%$ of the true value of $s$, for $D_{L}=200\,Mpc$. Being able to recover the scale factor error to this accuracy range confirms that the new method of calibration method is very competative with current techniques. These results also confirm that the sky position used throughout the project was indeed a bad choice with an error estimation of $\sim 24\%$.

Other simulations relating to the randomisation of $\alpha$ and $\delta$ were carried out in this project. Different luminosity distances were simulated to test different levels of SNR. The expected trend was that the standard deviation scales with SNR; with high SNR yielding a better resolution of the scale factor error. The results show this trend, given in table

\begin{table}[H]
  \centering
  \begin{tabular}{|c|c|c|}
    \hline
    Distance ($Mpc$) &$\langle$ SNR $\rangle$ & median(Standard Deviation)($\%$) \\
    \hline 
    $20$ & 74 & 1.5 \\
    $120$ & 12  & 7 \\
    $450$ & 3 & 27\\
    \hline
  \end{tabular}
  \caption{\textit{Random sky positions for different luminosity distances; $\langle$SNR$\rangle$ is the mean SNR for the distribution. }}
  \label{tab:dist-randsp}
\end{table}

The last, and lengthiest, set of simulations carried out used a network of detectors. Analysis on a luminosity distance of $200 \,Mpc$ was the only one carried out in this paper. This was due to time taken to process these simulations. With a full prior range and $100$ random sky positions as before, the standard deviations and SNRs were obtained. Results are shown in~\ref{tab:mult-rndsp}.

\begin{table}[H]
\centering
\begin{tabular}[H]{|c|c|c|}
\hline
  Detector   & $\langle$SNR$\rangle$ & median(Standard Deviation)($\%$)\\
\hline
LIGO Hanford & 20& 3 \\
LIGO Livingston & 20 & 10 \\
Virgo & 5 & 30\\
\hline
\end{tabular}
\caption{\textit{Results from randomising the sky position using a network of detectors with a full range of prior parameters at $D_{L} = 200\, Mpc$.}}
\label{tab:mult-rndsp}
\end{table}

Table~\ref{tab:mult-rndsp} gives promising results for both LIGO sites. One can see that the error estimation for LIGO Hanford is very good, with half of the positions analysed having an uncertainty $<3\%$. One would expect the gap in estimation accuracy between the two LIGO sites to be smaller as they are built to the same sensitivity. This leads to the conclusion that not enough sky positions were analysed. Virgo again has a less accurate recovery of the calibration scale factor. As mentioned above, the larger uncertainty is expected to originate from the sensitivity of the detector itself.

\section{Conclusion \& Discussion}
\label{sec:concl}

This paper has shown that for differing levels of SNR, the error in the calibration is resolved within a certain accuracy range. Initial simulations used a bad sky position, resulting a low SNR which created larger uncertainty in resolving $s$. It was shown that for this sky position, the calibration scale factor estimation was recovered with an accuracy of roughly a $27\%$ for $D_{L} = 200\,Mpc$ for LIGO Hanford. The resolution of $s$ for a network of detectors was also shown.

The main results obtained in this project were from analysing randomly chosen sky positions from the entire field of view for the detector. Through these simulations, it was shown that this new method of calibration can reach and surpass the current method in terms of accuracy. The results in figure~\ref{fig:std-rndsp} show that for half of the randomly chosen sky positions relate to an estimation of $s$ within a range comparable with the current methods.

These results are reassuring for parameter estimation of the defining GW signal and would help in reducing the uncertainty in a direct detection. This project was; however, time constrained. There are many, many things that could be done with this project. A true noise realisation with each frequency bin having a random Gaussian noise would be the next step in calibration error estimation.

This project is only the beginning in testing this new method of calibration. More rigorous testing has to be carried out for a number of different systems. As mentioned above, a larger number of random sky positions have to be analysed to give better understanding of the resolution capabilities of the method. The current technique of calibration defines the error in the detector to be a complex function of frequency and phase of the GW. Instead of the simple scale factor used so far, a more realistic error can be tested to ensure that the estimation in its value is still comparable with current methods.

For applications of this calibration method, hypothesis testing could be carried out. Testing different hypothesis is the technique used in determining a GW signal in a data-set when compared to a noise-only data-set. Being able to show with confidence that the data contains a signal is the primary focus of GW research at the moment. Therefore if uncertainty can be removed from the analysis this will ensure a more confident detection. Using this new method is resolving and removing the uncertainty is an area which has to be explored.

\section{Acknowledgements}
\label{sec:ackn}

I would like to thank my supervisors, Dr. Matthew Pitkin and Dr Christopher Messenger, for their help and guidance throughout the project. Without their help, bug fixing in code would have been an arduous task.

\section*{Appendix}
The inner production of two real functions, $a$ and $b$, is given as
\begin{equation}
  (a|b) = 2 \int_0^{\infty} \frac{a(f)b^*(f) + a^*(f)b(f)}{S(f)} \, df
\end{equation}
\begin{equation}
  \label{eq:blah}
  \approx 2 \Delta f \sum_{k>0} \frac{a(f_{k})b^*(f_{k}) + a^*(f_{k})b(f_{k})}{S(f_{k})}
\end{equation}
 where the integral has been approximated to a summation over discrete frequency bins, $f_{k}$. When the inner product of one real function $a(f)$, 15 becomes
 \begin{equation}
   (a|a) = 2 \Delta f \sum_{k>0} \frac{|a(f_k)|^{2}}{S(f_{k})}
 \end{equation}

For the likelihood considered in this project,
 \begin{equation}
    p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp\Bigg[ 2\Delta f\sum_{k>0}\frac{|\tilde{d}(f_{k}) - \tilde{h}(f_{k})|^{2}}{S(f_{k})} \Bigg]
 \end{equation}

\section{References}
\label{sec:ref}

\bibliography{references}

\end{document}

### 2015-04-07 docs/cosmic_cal.tex
Started filling out the paper draft from Liam's report.

@@
 
+\newcommand{\curlH}{\mathcal{H}}
+\newcommand{\gws}{\tilde{h}}
+
 %% ----- input git-version tag
@@
 \newcommand{\dcc}{LIGO-PXXXXXXXX}
-\newcommand{\CM}[1]{\textcolor{red}{CM: #1}}
-\newcommand{\MH}[1]{\textcolor{blue}{MH: #1}}
-\newcommand{\XF}[1]{\textcolor{green}{XF: #1}}
+\newcommand{\cm}[1]{\textcolor{red}{CM: #1}}
+\newcommand{\MP}[1]{\textcolor{blue}{MP: #1}}
+\newcommand{\lw}[1]{\textcolor{green}{LW: #1}}
 
@@
 
-\title{Requirements and usefulness of galaxy catalogues for gravitational wave astrophysics}
+\title{Astrophysical calibration of gravitational-wave detectors}
 
@@
   the direct comparison with the estimated value.  Discrepancies can
-  then be used to correct the originbal calibration.
+  then be used to correct the original calibration.
 \end{abstract}
@@
 
-Explain the basic idea.
+\cm{Explain the basic idea.}
 
@@
 
-We need to define our model for the calibration, i.e. are we talking
-about amplitude only and are we breaking this up over frequency.
+\cm{We need to define our model for the calibration, i.e. are we talking
+about amplitude only and are we breaking this up over frequency.}
+
+The technique used throughout GW research in calibrating the detectors is
+carried out through a complicated system of physical manipulations of the
+Fabret-Perot Michelson interferometer; where an elaborate feedback system is
+used to sustain a defined measurement in arm length difference between the
+moving mirrors~\cite{LIGOCal}. In general, the interferometers are set to move
+at a frequency corresponding to a GW signal. An error signal is then found
+through the Pound-Driver-Hall technique~\cite{Black} for mirror stabilisation
+which is proportional to the difference in mirror separation of the
+Fabret-Perot Michelson interferometer. The Pound-Drever-Hall technique is
+essentially a control loop feedback system in which the difference output (in
+GW detectors, this is the distance the mirrors move due to the GW) and the the
+output set is reduced to the highest degree of certainty. This is just a brief
+overview of the calibration method currently used, and is an extensive study in
+itself, see~\cite{Vitale} and references therein. For advanced LIGO and Virgo,
+at the frequency range used in this project, the error in the calibration using
+the 'hardware injection method' is roughly $10\%$~\cite{Vitale}. This is a
+benchmark set for the error estimation using the proposed new method in this
+project.
+
+The hardware method of calibrating the ground-based detectors is incredibly
+complex in its application, with many sources of error which have to be
+accounted for. The new method of calibration proposed and tested in this
+project is a lot simpler. The method acts as an external calibrator, to be
+carried out using the data collected by the detector, $\tilde{d}(f)$. We assume
+the collected data contains some unknown and unaccounted for calibration error.
+The uncertainty in $\tilde{d}(f)$ can then be evaluated using methods of
+parameterisation already carried out in GW research. This method of
+parameterising GW defining variables is then carried out on a known source, a
+standard candle as described above, which allows one to measure the calibration
+uncertainty in the data set.
+
+One can estimate the error in the calibration using parameter estimation
+methods, or more precisely Bayesian inference. In this project the calibration
+offset used throughout is simply a scale factor, $s$. This effects the data set
+collected by a simple multiplication:
+
+\begin{equation}
+  \label{eq:scaledata}
+  \tilde{d}_m(f) = \tilde{d}_e(f) \times s
+\end{equation}
+
+where $\tilde{d}_m(f)$ is the data-set measured by the detector, and
+$\tilde{d}_e(f)$ is the exact data set, with no error. A description of the
+collected data is given in section~\ref{sec:cbc}. The scale factor being
+parameterised, $s$, represents all known sources of error in the system. This
+is not entirely realistic as $s$ is known to be a function of frequency.
+Testing a more realistic calibration error is work to be done in the future.
+
+Being able to resolve the calibration scale factor with the least amount of
+uncertainty was the aim of this project; to show that this method is
+competitive with current techniques. If the calibration error can be resolved
+with known uncertainty, GW data analysis can then be carried out with more
+certainty. This is useful as an entirely independent method of assessing the
+uncertainty in the current calibration method.
 
@@
 
-We should discuss the key points with regards to
+A GW detection on its own right would be ground breaking, allowing
+astrophysicists to observe the sky above in a whole new spectrum. However, to
+increase the scientific return on a direct GW detection, a counterpart EM
+observation of the source itself would need to be seen.  There are two
+approaches to observing a source both in EM and GW spectrum's~\cite{grb}:
+
+\begin{enumerate}
+  \item locating and observing a source with an EM follow-up to
+a GW detection
+  \item searching for a GW signal generated by a source seen
+through EM observations
+\end{enumerate}
+
+Both of these methods have their pros and cons. Firstly, an EM follow-up would
+be very difficult due to the duration of GW source system detectable by
+ground-based interferometers. A neutron star binary system (NS-NS binary) has a
+duration of $\sim15$ minutes~\cite{bnstime}.  Errors in the sky position from a
+GW detection would also create issues for an EM follow-up. The sky localisation
+obtained from the ground-based detectors gives a large field of view for the EM
+observatories to search, roughly an area spanning $100$'s of
+degrees~\cite{grb}.
+
+The more promising detection approach then is a follow-up GW search from an EM
+trigger, especially for the source of GW considered in this paper. The merger
+time for NS-NS binaries are known with an uncertainty within a few seconds for
+the EM radiation emitted from this system~\cite{grb}.
+
+Identifying the source of GWs redshift would then lead to calculating the
+Hubble constant and other cosmological parameters~\cite{Schutz}. The evolution
+of the system could be better understood, allowing for searches with greater
+accuracy. Counterpart EM observations of a GW source will also remove
+uncertainty in the defining GW parameters, a fact which was taken advantage of
+for this project.
+
+The source of GW considered in this paper will be a coalescing binary system
+consisting of neutron stars. As well as being the most probable source of GWs
+detectable by ground-based interferometers, they are also expected to emit EM
+radiation. As shown in figure~\ref{fig:grb}, the emission occurs due to the
+high velocity with which the stars coalesce. The coalescences are thought to be
+the progenerators of short-duration gamma ray bursts (SGRBs)~\cite{grb}.
+Observations of this radiation has been see, through the Swift satellite and
+follow-up ground-based telescopes~\cite{swift}. Unfortunately, no observation
+of SGRBs have been seen within the range of advanced generation of LIGO and
+Virgo. It is probable, however, that such a source exists well within the range
+of advanced LIGO/ Virgo, but has simply not been observed~\cite{grb}. This then
+validates the assumption made throughout the project.
+
+In this paper, a key assumption has to be made in order to test this new method
+of calibration. It is assumed that there exists a source of GWs with enough
+information known through EM observation such that it can be used as a
+\textit{standard candle}, i.e. an accurately known and well modelled source of
+GWs. In this project (for the source of GWs being analysed), a GRB observation
+is assumed to have been seen. This allows for discrepancies in the detector
+calibration itself to be parameterised. This is the proposed new method of
+calibration for the detectors, for all types of detectors.
 
@@
 
-This should be the simplest case we can do (and maybe the only case).
+\cm{This should be the simplest case we can do.}
+
+In this project the hypothesis considered, $\curlH$, simply states that the
+data, $\tilde{d}(f),$ collected by a generic detector contains the GW signal,
+$\gws(f)$, and some noise, $\tilde{n}(f)$~\cite{JVei},
+
+\begin{equation}
+  \label{eq:data}
+  \tilde{d}^D(f) = \gws^D(f) + \tilde{n}^D(f)
+\end{equation}
+
+where $D$ represents the detector site. In this project there will be three
+detectors in question: LIGO Livingston (L), LIGO Hanford (H), and Virgo (V); $D
+= \{L,H,V\}$. Testing different hypotheses is the method used to determine
+whether a data set contains a GW signal or not~\cite{SathSchutz}. This was not
+explored in this project so the above hypothesis is consistent throughout this
+paper.
+
+Coalescing binary systems of compact objects, such as neutron stars or black
+holes, are the most probable source of GWs detectable by ground-based
+interferometers~\cite{peadvanced}. A direct detection of GWs would allow for a
+fuller understanding of the mechanics of these systems. Compact binary
+coalescence's (CBCs) are systems of two compact object inspiralling, with
+reducing separation, until the inevitable merger of the objects. This orbital
+motion creates ripples in space-time; this is the GW from this motion. For
+gravitational wave research, the interests lie in the GW signal generated, or
+the \textit{waveform}, shown in figure~\ref{fig:signal}.
+
+In this paper, a brief overview of the mathematical breakdown for the waveform
+is shown. For a more rigorous derivation, see~\cite{SathSchutz} and references
+there in. The waveform used throughout this project is a simplified system
+where the neutron stars are non rotating about their axes, in a perfectly
+circular inspiralling orbit; known as the TaylorF2 model~\cite{TF2,pN}. A more
+complex and realistic model could be tested in future for this calibration
+method. Parameter estimation using a variety of complex waveforms are carried
+out in~\cite{peadvanced}.
+
+The general equation for a GW signal detected at a generic detector, in the
+reference frame of the detector site, is given as~\cite{SathSchutz}
+
+\begin{equation}
+  \label{eq:gravsig}
+   \gws^D(f) = \big[ F_{+}^D(\alpha, \delta, \psi)\gws_{+}(f) +
+F_{\times}^D(\alpha, \delta, \psi)\gws_{\times}(f)\big]
+\end{equation}
+
+where $F_{+,\times}(\alpha, \delta, \psi)$ are the antenna response function of
+the detector, dependent on the right ascension, $\alpha$, declination,
+$\delta$, the polarisation angle $\psi$, and time. Testing the antenna patterns
+is carried out in section~\ref{sec:rndsp} of this report when the sky
+positions, $(\alpha, \delta)$, are randomised across the entire field of view
+for the detector.
+
+The polarisation amplitudes of the GW, given by $\gws_{+}(f)$, and
+$\gws_{\times}(f)$, represent the direction with which the GW compresses and
+stretches space-time. The form of these two amplitudes depends on the source
+being analysed. As mentioned above, the model used throughout this project was
+the TaylorF2 model. The parameters defining the GW polarisation amplitudes
+using the TaylorF2 model are:
+
+\begin{itemize}
+\item the total mass, $M = m_{1}+m_{2}$, where $m_{1,2}$ are the masses of the
+binary stars
+\item the symmetric mass ratio $\eta = \frac{m_{1}m_{2}}{(m_{1}+m_{2})}$
+\item chirp mass, $\mathcal{M} = \eta^{3/5}M$
+\item luminosity distance, $D_{L}$, measured in Mpc ($1 Mpc \approx
+3.1\times10^{22}m  $)
+\item the angle of inclination $\iota$. This is the angle from the source,
+taken as the angle between the normal of the orbital plane and the observer
+orientation.
+\item phase of the wave, $\psi$, which defines the polarisation of the angle of
+the GW, shown in equation~\ref{eq:psi}.
+\item the coalescence time, $t_c$, and phase of coalescence, $\phi_c$.
+\end{itemize}
+
+Together with the sky position, these parameters form the vector $\vec{\Theta}
+= \{\mathcal{M}, \eta, D_{L}, \iota, \alpha, \delta, \psi, t_{c}, \phi_{c}\}$
+which defines the entire GW signal.  The form of the polarisation amplitudes
+using the TaylorF2 model are given as~\cite{SathSchutz,pN,TF2}
+
+\begin{equation}
+  \gws_{+}(f) = \frac{1+\cos^{2}\iota}{2D_{L}}
+\frac{\mathcal{M}^{5/6}}{\pi^{2/3}} \sqrt{\frac{5}{24}}f^{-7/6}e^{-i\psi(f)}
+\end{equation}
+\begin{equation}
+  \gws_{\times}(f) =
+i\frac{\cos^{2}\iota}{D_{L}}\frac{\mathcal{M}^{5/6}}{\pi^{2/3}}
+\sqrt{\frac{5}{24}}f^{-7/6}e^{-i\psi(f)}
+\end{equation}
+
+where the phase of the GW, $\psi$, is
+\begin{equation}
+  \label{eq:psi}
+  \psi(f) = 2\pi f t_c - \phi_c + \psi_N(\eta)\sum\limits_{k=0}^4{\psi_k(\eta)
+(\pi(m_1 + m_2)f)^{(k-5)/3}}.
+\end{equation}
+
+The terms $\psi_N$ and $\psi_k$ are the standard Newtonian and post-Newtonian
+coefficients. Derivations of these two equations can be found in~\cite{pN} and
+references there-in. The simulated gravitational wave signal used throughout
+this project can be seen in figure~\ref{fig:signal}. As one can see, the signal
+strength is very faint (an order of magnitude around $10^{-24}$). This is the
+the biggest problem in detecting GWs.
+
+As mentioned above, the source considered throughout this project thought to
+produce EM radiation; SGRBs. In this project, we assume that SGRBs of the GW
+source described above have been observed. If a GRB has been seen, then the
+redshift of the merger can be calculated. This then allows for the luminosity
+distance to be calculated and in turn other defining parameters can be obtained
+(or known to a degree of uncertainty)~\cite{SathSchutz}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{grav_signal}
+  \caption{The gravitational waveform produced using coalescing binary
+neutron stars. The wave is represented in the frequency domain with masses
+$m_{1,2} = 1.4M_{\odot}$ at a luminosity distance $D_{L} = 200 \,Mpc$.}
+  \label{fig:signal}
+\end{figure}
+
+%More explination
+Finally, the last term in equation~\ref{eq:data} must be discussed; the noise
+of the model, $\tilde{n}(f)$. Throughout this project, we have added no noise
+to the signal, although as discussed later we do include a noise estimate in
+the likelihood.
+
+The prior probability function, $p(\vec{\theta} | \curlH)$ in
+equation~\ref{eq:bayesgw}, represents our prior knowledge in $\theta$ before
+the analysis.The range of parameters used throughout this paper were chosen due
+to their inability to be justifiably known through a GRB assumption. The
+parameters were drawn from the ranges for any signals generated in this
+analysis, given in table 1.
+
+\begin{table}
+  \centering
+  \begin{tabular}{|c|c|}
+%  \caption{Prior parameters}
+    \hline
+    Parameter & Distribution \\
+    \hline
+    $\psi$   &  U($0,\pi/2$)\\
+    $\iota$   & N($0,0.1$) \\
+    $\phi_c$  & U($0,\pi/2$)\\
+    $t_c$     & U($t_{o}-0.01, t_{o}+0.01$) \\
+    $s$       & U(0,2) \\
+    \hline
+  \end{tabular}
+  \label{tab:prior}
+  \caption{\textit{The parameters used throughout the analysis with some degree
+of uncertainty. U(a,b) is a uniform distribution between a and b. N($\mu,
+\sigma$) is a normal distribution, with a mean $\mu$ and standard deviation
+$\sigma$. $t_{o} = 9\times10^{8}$}}
+\end{table}
+
+Through the assumption made above allows us to constrain the system,
+maintaining a small number of parameters in the prior. For a realistic
+parameter estimation of the source, a larger range of values with a bigger
+distribution. This was not tested due to time constraints but is proposed as
+work for the future.
+
+The likelihood functiona used throughout this project is given for a generic
+detector D as~\cite{JVei},
+
+\begin{equation}
+  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp(-(d^{D}(f) -
+h^{D}(f)|d^{D}(f) - h^{D}(f))/2)
+\end{equation}
+
+where $(\cdot |\cdot )$ is a representation of the inner product. This
+derivation is given in the appendix. Using the inner product of two real
+functions, and approximating for discrete integer values, the likelihood then
+becomes
+
+\begin{equation}
+  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp\Bigg[ 2\Delta
+f\sum_{k>0}\frac{|\tilde{d}(f_{k}) - \tilde{h}(f_{k})|^{2}}{S(f_{k})} \Bigg]
+\end{equation}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{PSD}
+  \caption{\textit{Noise spectral density for advanced LIGO and Virgo.}}
+  \label{fig:psd}
+\end{figure}
+
+where $S(f_{k})$ is the \textit{power spectral density} (PSD). This defines the
+sensitivity of the detector as it a measure of the strain in the detector due
+to noise~\cite{SathSchutz,sens}. The PSD for advanced LIGO and Virgo are shown
+in figure~\ref{fig:psd}. The figure shows that the strain effect for advanced
+LIGO is less than it is for Virgo; LIGO has a better sensitivity. This is a
+factor in analysing the calibration error, given in section~\ref{sec:results}.
+The subscript $k$ represents discrete values of relevant frequency bins (As
+carrying out the integral would become computationally costly). This likelihood
+will be used throughout the project. The factor of $\Delta f$ comes from the
+approximation of the inner product of two real functions (as it is integrated
+over frequency space).
+
+The aim of this project was to determine how well the calibration error could
+be parameterised in a software injected gravitational wave source. The error
+was estimated through numerous situations but the basis of the simulations
+remained the same throughout the project: a single source of GWs at varying
+signal to noise ratios. The GRB assumption was analysed as well by relaxing
+some of the defining GW paramaters and adding some uncertainty into their
+value, given in section~\ref{sec:pandl}.
+
+These values where chosen due to the assumed counterpart EM observation. These
+values could not be justified by that assumption so they were added into the
+prior with reasonable uncertainty.
+
+The estimation of the error, given in its simplest form as a scale factor, was
+performed for numerous systems: single detector, a network of detectors, random
+sky positions, etc. Before the parameter estimation was carried out, the
+relationship between distance of the source relative to the detector and the
+signal to noise ratio (SNR) needed to be found.
+
+The SNR is given by
+\begin{equation}
+  \label{eq:snr}
+  SNR = \sqrt{\Bigg( \frac{2 |\gws^D(f)|^2}{S^D(f)} df \Bigg)}
+\end{equation}
+
+where the noise spectral density, $S^D(f)$ is not a function of distance, shown
+in section~\ref{sec:pandl}. For coalescing binary neutron stars of masses
+$m_{1,2} = 1.4\,M_{\odot}$ and  $D_{L}200\,Mpc$, the expected SNR is roughly 8.
+The SNR was determined from $D_{L} = [10,200]\,Mpc$  using a randomly chosen
+sky position which defines the antenna patterns and the GW signal. The range of
+luminosity distance used in this paper represented a high level of SNR up to
+the distance at which the first predicted detectable GW source; $D_{L} =
+200\,Mpc$. The detector limits for the advanced era is a luminosity distance of
+$\sim 450\,Mpc$. This was not initially tested as the SNR was low enough to
+cause problems with computational time. It was however tested in the random sky
+position simulations, see section~\ref{sec:rndsp}. The initial value for the
+calibration error used in the analysis was chosen randomly.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{DvsSNR-mult-det}
+  \caption{The distance and SNR relationship for all three detectors:
+LIGO Hanford, H1, LIGO Livingston, L1, and VIRGO.}
+  \label{fig:distsnr}
+\end{figure}
+
+To begin with, parameter estimation was carried out on the simplest case: a
+single source of GWs using a single detector, LIGO Hanford, at different levels
+of SNR. The aim was to resolve the calibration scale factor with the highest
+degree of certainty.  Before testing the uncertainty resolution using a the
+full range of prior parameters, we fixed all values to known signal values and
+only searched over the scale factor. Results shown for with and without a full
+prior range are given in figures~\ref{fig:sd-empty-D}
+and~\ref{fig:sd-non-empty-D} respectively.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{SD_empty_D10_200}
+  \caption{\textit{Posterior distribution on the calibration scale factor for
+LIGO Hanford. Simulations carried out using an entirely known source of GWs.}}
+  \label{fig:sd-empty-D}
+\end{figure}
+
+The figure above is a histogram of the calibration error, with each curve
+representing a different distance (SNR). The black line at $s = 0.8$ is the
+initial scale factor value used to multiply the unanalysed data set. This
+representation of the initial calibration scale factor will be continuous
+throughout the paper.  One would expect with lower SNR (greater $D_{L}$), the
+estimation of the parameter would become less certain. This is clearly shown in
+figure~\ref{fig:sd-empty-D} as the standard deviations of the curves increases
+with smaller SNR/larger luminosity distances. For this entirely known source of
+GWs, we would expect to find the uncertainty in $s$ to be approximately
+$1/$SNR; this is what we see in figure~\ref{fig:sd-empty-D}. The error in
+estimating $s$ for $D_{L} = 200 \,Mpc$ is roughly $\sim 20\%$. This is larger
+than the current method of obtaining the uncertainty in calibration. A full
+test of the new method proposed is carried out in section~\ref{sec:rndsp}.
+
+The next estimation of the calibration scale factor was performed with a full
+range of prior parameters, giving a more realistic system. As before, the
+simulations were carried out for differing levels of SNR and a histogram of the
+calibration error was made, seen in figure~\ref{fig:sd-non-empty-D}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{SD_non_empty_D10_200}
+  \caption{\textit{Posterior distribution on the calibration scale factor for
+LIGO Hanford. Simulations carried out with some parameters in the prior.}}
+  \label{fig:sd-non-empty-D}
+\end{figure}
+
+The error in the calibration scale factor, as before, increases with decreasing
+SNR (increasing $D_{L}$). The magnitude of this error however is larger than in
+the previous simulation. With a full prior range given in
+section~\ref{sec:pandl}, the standard deviation increases; at $D_{L} =
+200\,Mpc$ the error in $s$ is $\sim 24\%$. When comparing this new method of
+calibration scale factor estimation with the current, feedback loop system
+method used today, this method is not as accurate. However, it should again be
+noted that this method is an independent estimate of the calibration accuracy.
+
+A network of detectors are used throughout GW research. Being able to show the
+effect of the calibration error and recovering its uncertainty in the analysis
+would help in giving a detection more confidence. In this section, the
+simulations were carried out using multiple detectors; LIGO Hanford, LIGO
+Livingston, and Virgo. In reality, multiple detectors are used in order to
+determine the (rough) sky position of a signal. Due to their locations, if one
+site detects a signal before the others, then the position of the source can be
+roughly worked out. The aim was, as before, to estimate the calibration scale
+factor with the least amount of uncertainty for differing levels of SNR for a
+single source of GWs.
+
+For a network of detectors, the overall likelihood function becomes a product
+of each likelihood for all detectors, i.e.
+
+\begin{equation}
+  \label{eq:mult-likeli}
+  p(\vec{d}| \vec{\theta}, \curlH) = \prod \limits_D p(\vec{d}^D| \vec{\theta},
+\curlH)
+\end{equation}
+
+where $D = \{H,L,V\}$. The initial calibration scale factors where again chosen
+at random ($s = [4.6, 2.5, 6.1]$ for LIGO Hanford, LIGO Livingston, and Virgo
+respectively). The simulations were the same as before, using the same
+parameter values. The GRB assumption was again tested in these simulations as a
+handful of parameters were relaxed and given some uncertainty. The results of
+these simulations are shown in figure~\ref{fig:mult-empty-D} and
+figure~\ref{fig:mult-non-D}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{MD_empty_D10_200}
+  \caption{Posterior distribution on the calibration scale factor for
+all three detectors. Simulations carried out using an entirely known source of
+GWs.}
+    \label{fig:mult-empty-D}
+\end{figure}
+
+Figure~\ref{fig:mult-empty-D} shows the histograms of the calibration scale
+factor for each detector. These simulations used an entirely known source of
+GWs, at a randomly chosen sky position (the same sky position as before). As
+the figure shows both LIGO sites, Hanford and Livingston, are able to resolve
+the error with a measurable degree of uncertainty. The standard deviation for
+$s$ at $D_{L}=200\,Mpc$ is $24\%$ and $41\%$ of the true scale factor for
+Hanford and Livingston respectively. The estimation for the calibration error
+using VIRGO however, was not as accurate. It can be seen that, for increasing
+luminosity distances, the resolution of $s$ becomes increasingly uncertain
+until the point at which the distribution becomes almost uniform; $D_{L}
+\approx 120\,Mpc$. There are a few reasons why this may be. Firstly, the sky
+position used may not be 'ideal' for generating the antenna patterns. The most
+probable cause of the estimation being very poor will lie in the detector
+itself. The interferometer itself is not as sensitive the two LIGO sites, hence
+the estimation of parameters using this site will not be as precise.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{MD_non_empty_D10_200}
+  \caption{Posterior distribution on the calibration scale factor for
+all three detectors. Simulations carried out with some prior parameters.}
+  \label{fig:mult-non-D}
+\end{figure}
+
+The histograms shown in figure~\ref{fig:mult-non-D} were made using the full
+range of prior parameters. The uncertainty in the calibration error is larger
+with a less known source of GWs. VIRGO, again, has the least certain
+calibration error resolution with the distribution becoming even more uniform.
+The standard deviation for both LIGO curves become larger with increasing
+luminosity distance as before. For $D_{L} = 200Mpc$,  the standard deviation is
+$\tilde 30\%$ of the true value. When comparing this to the current method,
+this again is not as accurate.
+
+Throughout the duration of this project the antenna patterns were defined by a
+randomly chosen sky position, as seen in equation~\ref{eq:gravsig}. This did
+not give a true representation of the estimation capable by algorithm as this
+position could be a bad one; making parameterisation more difficult and less
+precise. To then test this new method of calibration fully, the entire field of
+view of the detector must be analysed. The sky positions $\alpha$, $\delta$,
+were distributed as: \begin{table}[H] \centering \begin{tabular}{|c|c|} \hline
+Parameter & Distribution \\ \hline $\alpha$       & U($0$,$2\pi$) \\
+$\sin(\delta)$ & U($-1$,$1$)    \\ \hline \end{tabular} \caption{\textit{The
+distribution of sky position. U(a,b) indicates a uniform distribution between a
+and b.}} \label{tab:sp} \end{table}
+
+Figures~\ref{fig:snr-rndsp} and~\ref{fig:std-rndsp} were obtained using a
+luminosity distance $D_{L} = 200 Mpc$ as this is the expected luminosity
+distance predicted to host the first observable GW~\cite{peadvanced}. $100$
+different sky positions were tested using the LIGO Hanford site. The number of
+locations was limited by the length of time these simulations took to carry
+out. Increasing the number of locations would give a more accurate
+representation of the estimation ability; this can be carried out in future.
+Finally, a full range of prior parameters were also used in the simulation
+set-up.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{rand_sp_snr_D200}
+  \caption{Signal to noise ratio for every sky position analysed at
+$D_{L} = 200 \,Mpc$. The vertical black line represents the SNR for the sky
+position used throughout previous simulations}
+  \label{fig:snr-rndsp}
+\end{figure}
+
+Figure~\ref{fig:snr-rndsp} shows the histogram of SNRs for the entire field of
+view for the detector. Predictions state that at a luminosity distance of
+$D_{L}=200\,Mpc$, the SNR will be roughly 8. This is consistent with the
+results obtained. The vertical black line shown in figure~\ref{fig:snr-rndsp}
+represents the SNR used throughout the project using the randomly chosen
+$\alpha$ and $\delta$. One can see that this is a less-than-average SNR and may
+explain the resolution of the calibration error not being too precise. This
+then is a promising result and the full testing of this new calibration method
+can be carried out.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{rand_sp_std_D200}
+  \caption{Standard Deviation (as a $\%$ of the true scale factor
+value) for every sky position analysed at $D_{L}=\,200Mpc$. The vertical black
+line indicates the median of the distribution.}
+  \label{fig:std-rndsp}
+\end{figure}
+
+Figure~\ref{fig:std-rndsp} is the distribution of fractional standard
+deviations on s for each sky position analysed, given as a percentage. The
+vertical black line in this figure shows the \textit{median} of the
+distribution. The results show that the algorithm can resolve the calibration
+error in the data within the range of $\sim 7\%$ to $\sim 14\%$ of the true
+value of $s$, for $D_{L}=200\,Mpc$. Being able to recover the scale factor
+error to this accuracy range confirms that the new method of calibration method
+is very competative with current techniques. These results also confirm that
+the sky position used throughout the project was indeed a bad choice with an
+error estimation of $\sim 24\%$.
+
+Other simulations relating to the randomisation of $\alpha$ and $\delta$ were
+carried out in this project. Different luminosity distances were simulated to
+test different levels of SNR. The expected trend was that the standard
+deviation scales with SNR; with high SNR yielding a better resolution of the
+scale factor error. The results show this trend, given in table
+
+\begin{table}
+  \centering
+  \begin{tabular}{|c|c|c|}
+    \hline
+    Distance ($Mpc$) &$\langle$ SNR $\rangle$ & median(Standard
+Deviation)($\%$) \\
+    \hline
+    $20$ & 74 & 1.5 \\
+    $120$ & 12  & 7 \\
+    $450$ & 3 & 27\\
+    \hline
+  \end{tabular}
+  \caption{\textit{Random sky positions for different luminosity distances;
+$\langle$SNR$\rangle$ is the mean SNR for the distribution. }}
+  \label{tab:dist-randsp}
+\end{table}
+
+The last, and lengthiest, set of simulations carried out used a network of
+detectors. Analysis on a luminosity distance of $200 \,Mpc$ was the only one
+carried out in this paper. This was due to time taken to process these
+simulations. With a full prior range and $100$ random sky positions as before,
+the standard deviations and SNRs were obtained. Results are shown
+in~\ref{tab:mult-rndsp}.
+
+\begin{table}
+\centering
+\begin{tabular}[H]{|c|c|c|}
+\hline
+  Detector   & $\langle$SNR$\rangle$ & median(Standard Deviation)($\%$)\\
+\hline
+LIGO Hanford & 20& 3 \\
+LIGO Livingston & 20 & 10 \\
+Virgo & 5 & 30\\
+\hline
+\end{tabular}
+\caption{\textit{Results from randomising the sky position using a network of
+detectors with a full range of prior parameters at $D_{L} = 200\, Mpc$.}}
+\label{tab:mult-rndsp}
+\end{table}
+
+Table~\ref{tab:mult-rndsp} gives promising results for both LIGO sites. One can
+see that the error estimation for LIGO Hanford is very good, with half of the
+positions analysed having an uncertainty $<3\%$. One would expect the gap in
+estimation accuracy between the two LIGO sites to be smaller as they are built
+to the same sensitivity. This leads to the conclusion that not enough sky
+positions were analysed. Virgo again has a less accurate recovery of the
+calibration scale factor. As mentioned above, the larger uncertainty is
+expected to originate from the sensitivity of the detector itself.
 
@@
 
-An analysis or discussion of the issues regarding multiple events
-without EM counterparts.
+\cm{A discussion of the issues regarding multiple events
+without EM counterparts.}
 
@@
 
-What did we learn.
+\cm{What did we learn.}
+
+This paper has shown that for differing levels of SNR, the error in the
+calibration is resolved within a certain accuracy range. Initial simulations
+used a bad sky position, resulting a low SNR which created larger uncertainty
+in resolving $s$. It was shown that for this sky position, the calibration
+scale factor estimation was recovered with an accuracy of roughly a $27\%$ for
+$D_{L} = 200\,Mpc$ for LIGO Hanford. The resolution of $s$ for a network of
+detectors was also shown.
+
+The main results obtained in this project were from analysing randomly chosen
+sky positions from the entire field of view for the detector. Through these
+simulations, it was shown that this new method of calibration can reach and
+surpass the current method in terms of accuracy. The results in
+figure~\ref{fig:std-rndsp} show that for half of the randomly chosen sky
+positions relate to an estimation of $s$ within a range comparable with the
+current methods.
+
+These results are reassuring for parameter estimation of the defining GW signal
+and would help in reducing the uncertainty in a direct detection. This project
+was; however, time constrained. There are many, many things that could be done
+with this project. A true noise realisation with each frequency bin having a
+random Gaussian noise would be the next step in calibration error estimation.
+
+This project is only the beginning in testing this new method of calibration.
+More rigorous testing has to be carried out for a number of different systems.
+As mentioned above, a larger number of random sky positions have to be analysed
+to give better understanding of the resolution capabilities of the method. The
+current technique of calibration defines the error in the detector to be a
+complex function of frequency and phase of the GW. Instead of the simple scale
+factor used so far, a more realistic error can be tested to ensure that the
+estimation in its value is still comparable with current methods.
+
+For applications of this calibration method, hypothesis testing could be
+carried out. Testing different hypothesis is the technique used in determining
+a GW signal in a data-set when compared to a noise-only data-set. Being able to
+show with confidence that the data contains a signal is the primary focus of GW
+research at the moment. Therefore if uncertainty can be removed from the
+analysis this will ensure a more confident detection. Using this new method is
+resolving and removing the uncertainty is an area which has to be explored.
 

### 2015-04-08 docs/cosmic_cal.tex
Added stuff and tidied things a bit.

@@
   the direct comparison with the estimated value.  Discrepancies can
-  then be used to correct the original calibration.
+  then be used to correct the original calibration.~\cm{Update at end}
 \end{abstract}
@@
 \acrodef{GRB}[GRB]{gamma-ray burst}
+\acrodef{sGRB}[sGRB]{short gamma-ray burst}
+\acrodef{CBC}[CBC]{compact binary coalescence}
+\acrodef{EM}[EM]{electro-magnetic}
 
@@
 
-\cm{Explain the basic idea.}
+% introduce GWs and BNS systems
+It is expected that the advanced generation of interferometric \ac{GW}
+detectors will detect such waves emitted from $O(10s)$ of compact binary
+coalescences.  One such class of these cataclysmic events, the inspiral and
+merger of \ac{BNS} systems will be detected out to a maximum range of $\approx
+450$ Mpc. Assuming the current best estimates for the cosmological parameters,
+this is equivelent to a redshift $z\approx 0.1$. As noted by~\cite{} \ac{CBC}
+systems can be used as cosmiological distance markers, otherwise known as
+"standard sirens" (analogous to the \ac{EM} standard candles). Standard sirens
+allow ...
+
+% BNS systems are GRBs
+It is considered likely that the merger of \ac{BNS} systems, in addition to
+emitting detectable \acp{GW}, is also the mechanism for producing \acp{sGRB}.
+In this scenario these \ac{EM} events are produced through the ... and result
+in tightly beamed emission parallel to the orbital angular momentum vector of
+the \ac{BNS} system. Additional evidence for the coincidence of \acp{sGRB} with
+\ac{GW} events are the estimated astrophysical rates of both phenomena.
+Observations of \acp{sGRB} give a rate of XXX Mpc$^{-3}$yr$^{-1}$ which can be
+compared to estimates of \ac{BNS} merger rates of XXX-XXX Mpc$^{-3}$yr$^{-1}$.
+This latter range is obtained from population synthesis models together with
+knowledge of the X known \ac{BNS} systems in our galaxy.
+
+% talk about the calibration idea
+If a single \ac{GW} event is observed in coincidence with a \ac{sGRB} it will
+be possible to identify the host galaxy of the \ac{BNS} source.  With this
+information a spectroscopic redshift can be very accurately obtained. With
+knowledge of the redshift, using current best estimates for the Hubble constant
+and other cosmological parameters, the true luminosity distance can be
+estimated to $\sim 1\%$ accuracy  The \ac{GW} measurement acts as a standard
+siren also giving us a direct measurement of the luminosity distance to the source.
+The accuracy of such a measurement depends on a number of factors including the
+accuracy with which the \ac{GW} detector has been calibrated.  Hence comparison
+with the distance estimate from the \ac{sGRB} we can re-calibrate (or validate)
+the existing experimentally obtained calibration.
+
+% What do we do in this paper
+In this paper we investigate the feasibility of this approach for single
+coincident \ac{GW}--\ac{sGRB} events and establish the validation power of such
+a calibration technique.  In Sec.~\ref{sec:calibration} we briefly summarize
+the existing experimental technique for \ac{GW} detector calibration and its
+expected accuracy.  We then review the concept of \ac{GW} standard sirens and
+their proposed coincident \ac{EM} signatures, the \acp{sGRB} in
+Secs.~\ref{sec:sirens} and~\ref{sec:GRB}. In Sec.~\ref{sec:single} we present
+the results of Monte-Carlo simulations for the purpose of luminosity distance
+estimation using \ac{GW} events. In Sec.~\ref{sec:cosmo} we show the likely
+accuracy of \ac{sGRB} distances obtained from their host galaxy redshift and in
+Sec.~\ref{sec:discussion} we conclude.
 
@@
 benchmark set for the error estimation using the proposed new method in this
-project.
+project.~\cm{Clean this up and expand.}
 
-The hardware method of calibrating the ground-based detectors is incredibly
-complex in its application, with many sources of error which have to be
-accounted for. The new method of calibration proposed and tested in this
-project is a lot simpler. The method acts as an external calibrator, to be
-carried out using the data collected by the detector, $\tilde{d}(f)$. We assume
-the collected data contains some unknown and unaccounted for calibration error.
-The uncertainty in $\tilde{d}(f)$ can then be evaluated using methods of
-parameterisation already carried out in GW research. This method of
-parameterising GW defining variables is then carried out on a known source, a
-standard candle as described above, which allows one to measure the calibration
-uncertainty in the data set.
-
-One can estimate the error in the calibration using parameter estimation
-methods, or more precisely Bayesian inference. In this project the calibration
-offset used throughout is simply a scale factor, $s$. This effects the data set
-collected by a simple multiplication:
-
-\begin{equation}
-  \label{eq:scaledata}
-  \tilde{d}_m(f) = \tilde{d}_e(f) \times s
+% Describe our model for the calibration
+In this work the calibration correction factor used throughout is simply a scale
+factor, $s$. This effects the data set collected by a simple multiplication:
+%
+\begin{equation}\label{eq:scaledata}
+  h(t) = s\times h_{\mathrm{m}}(t)
 \end{equation}
-
-where $\tilde{d}_m(f)$ is the data-set measured by the detector, and
-$\tilde{d}_e(f)$ is the exact data set, with no error. A description of the
+%
+where $h_{\mathrm{m}}(t)$ is the measured detector strain, and
+$h(t)$ is the exact perfectly calibrated strain. A description of the
 collected data is given in section~\ref{sec:cbc}. The scale factor being
@@
 is not entirely realistic as $s$ is known to be a function of frequency.
-Testing a more realistic calibration error is work to be done in the future.
-
-Being able to resolve the calibration scale factor with the least amount of
-uncertainty was the aim of this project; to show that this method is
-competitive with current techniques. If the calibration error can be resolved
-with known uncertainty, GW data analysis can then be carried out with more
-certainty. This is useful as an entirely independent method of assessing the
-uncertainty in the current calibration method.
+Testing a more realistic calibration error is work to be done in the
+future.~\cm{Need to clean this up and expand the discussion}
 
@@
 
-A GW detection on its own right would be ground breaking, allowing
-astrophysicists to observe the sky above in a whole new spectrum. However, to
-increase the scientific return on a direct GW detection, a counterpart EM
-observation of the source itself would need to be seen.  There are two
-approaches to observing a source both in EM and GW spectrum's~\cite{grb}:
+% what are sGRBs 
+\cm{What are sGRBs?}\\
 
-\begin{enumerate}
-  \item locating and observing a source with an EM follow-up to
-a GW detection
-  \item searching for a GW signal generated by a source seen
-through EM observations
-\end{enumerate}
+% discuss joint-event rates
+\cm{Discuss the expected rate of joint detections}\\
 
-Both of these methods have their pros and cons. Firstly, an EM follow-up would
-be very difficult due to the duration of GW source system detectable by
-ground-based interferometers. A neutron star binary system (NS-NS binary) has a
-duration of $\sim15$ minutes~\cite{bnstime}.  Errors in the sky position from a
-GW detection would also create issues for an EM follow-up. The sky localisation
-obtained from the ground-based detectors gives a large field of view for the EM
-observatories to search, roughly an area spanning $100$'s of
-degrees~\cite{grb}.
+% observing a coincident event
+The most likely scenario in which a coincident \ac{GW}--\ac{sGRB} event would
+be identified is through the targeted follow-up of an observed \ac{sGRB} or by
+post-facto matching of \ac{GW} trigger lists with known \ac{sGRB} events.  The
+likelihood of being able to follow-up \ac{GW} events using gamma-ray telescopes
+with low enough latency to catch a \ac{sGRB} is low.  Compounding this issue is
+the relatively large \ac{GW} sky error-box giving a field-of-view for the
+\ac{EM} observatories to search spanning $\sim 100$'s of square
+degrees~\cite{grb}.  For the \ac{GW} follow-up of \ac{sGRB} scenario the merger
+time for \ac{BNS} systems will be estimated from the \ac{sGRB} to within a few
+seconds~\cite{grb}.  This makes the follow-up search less computationally
+expensive since it is performed over a smaller range of data using potentially
+fewer numbers of waveform templates.  This computational saving enables the use
+of a more computationally expensive multi-detector coherent scheme rather than
+the cheaper coincidence methods used in the untargeted searches.
 
-The more promising detection approach then is a follow-up GW search from an EM
-trigger, especially for the source of GW considered in this paper. The merger
-time for NS-NS binaries are known with an uncertainty within a few seconds for
-the EM radiation emitted from this system~\cite{grb}.
-
-Identifying the source of GWs redshift would then lead to calculating the
-Hubble constant and other cosmological parameters~\cite{Schutz}. The evolution
-of the system could be better understood, allowing for searches with greater
-accuracy. Counterpart EM observations of a GW source will also remove
-uncertainty in the defining GW parameters, a fact which was taken advantage of
-for this project.
-
-The source of GW considered in this paper will be a coalescing binary system
-consisting of neutron stars. As well as being the most probable source of GWs
-detectable by ground-based interferometers, they are also expected to emit EM
-radiation. As shown in figure~\ref{fig:grb}, the emission occurs due to the
-high velocity with which the stars coalesce. The coalescences are thought to be
-the progenerators of short-duration gamma ray bursts (SGRBs)~\cite{grb}.
-Observations of this radiation has been see, through the Swift satellite and
-follow-up ground-based telescopes~\cite{swift}. Unfortunately, no observation
-of SGRBs have been seen within the range of advanced generation of LIGO and
-Virgo. It is probable, however, that such a source exists well within the range
-of advanced LIGO/ Virgo, but has simply not been observed~\cite{grb}. This then
-validates the assumption made throughout the project.
-
-In this paper, a key assumption has to be made in order to test this new method
-of calibration. It is assumed that there exists a source of GWs with enough
-information known through EM observation such that it can be used as a
-\textit{standard candle}, i.e. an accurately known and well modelled source of
-GWs. In this project (for the source of GWs being analysed), a GRB observation
-is assumed to have been seen. This allows for discrepancies in the detector
-calibration itself to be parameterised. This is the proposed new method of
-calibration for the detectors, for all types of detectors.
+% discuss the inclination angle issue
+A fortunate consequence of a joint \ac{GW}--\ac{sGRB} observation will be the
+fact that in order for such an event to be observed, the \ac{BNS} system must
+have had its orbital angular momentum vector pointing towards (or away) from
+the detector.  The actual inclination angle of the system, defined as the angle
+between the orbital angular momentum vector and the line of sight, must be $<$
+half of the \ac{sGRB} beaming angle.  This prerequisite property limits us to
+systems that are approximately "face-on" and therefore biases us to higher
+\ac{SNR} signals.  However, as discussed earlier, the property of beaming
+severely impacts the probable rate of such joint observations.
 
@@
 
-\cm{This should be the simplest case we can do.}
+\cm{This should be the simplest case we can do.  We need to cut a lot of this.}
+
+% Describe what we are going to do in this section
+
+% Describe the signal model (Taylor F2 etc...)
+
+% Define the detectors and the noise epoch we will be using
+
+% Define the likelihood function we will be using
+
+% describe what parameters we will be considering
+
+% describe the priors
+
+% mention the no-noise issue
+
+% Describe the ensemble of simulations used
 
@@
 paper.
-
-Coalescing binary systems of compact objects, such as neutron stars or black
-holes, are the most probable source of GWs detectable by ground-based
-interferometers~\cite{peadvanced}. A direct detection of GWs would allow for a
-fuller understanding of the mechanics of these systems. Compact binary
-coalescence's (CBCs) are systems of two compact object inspiralling, with
-reducing separation, until the inevitable merger of the objects. This orbital
-motion creates ripples in space-time; this is the GW from this motion. For
-gravitational wave research, the interests lie in the GW signal generated, or
-the \textit{waveform}, shown in figure~\ref{fig:signal}.
 
@@
 (or known to a degree of uncertainty)~\cite{SathSchutz}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{grav_signal}
-  \caption{The gravitational waveform produced using coalescing binary
-neutron stars. The wave is represented in the frequency domain with masses
-$m_{1,2} = 1.4M_{\odot}$ at a luminosity distance $D_{L} = 200 \,Mpc$.}
-  \label{fig:signal}
-\end{figure}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Distance estimates from \acp{sGRB}\label{sec:cosmo}}
+
+~\cm{Describe the cosmological analysis used to obtain the distance from the
+GRB}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Multiple events without counterpart\label{sec:multiple}}

### 2015-04-09 docs/cosmic_cal.tex
A bit more tidying up.

@@
 % Describe what we are going to do in this section
+Let us consider a single \ac{GW} detection of the inspiral stage of a \ac{BNS}
+merger and assume that this signal has been associated with a potential
+\ac{sGRB} detection. We therefore also assume that the inclination angle
+of the binary is consistent with a given \ac{sGRB} beaming angle.  Finally we
+assume that a host galaxy has been identified for the source and that the sky
+position is therefore known to high precision.
 
 % Describe the signal model (Taylor F2 etc...)
+We use the post-Newtonian Taylor F2 waveform approximation for modelling the
+signal whereby the \ac{GW} phase for 2 point-particles of mass $m_{1}$ and
+$m_{2}$ is given by
+%
+\begin{equation}
+\label{eq:gwphase}
+  \Psi(f) = 2\pi f t_\mathrm{c} - \phi_{\mathrm{c}} - \frac{\pi}{4} +
+\frac{3}{128\eta\nu^{5}}\sum\limits_{k=0}^{7}\alpha_{k}\nu^{k}.
+\end{equation}
+%
+Here $\eta=m_{1}m_{2}/M^{2}$ is the symmetric mass-ratio with $M=m_{1}+m_{2}$
+the total mass.  The post-Newtonian expansion parameter is $\nu=(\pi M
+f)^{1/3}$ and the coefficients are $\alpha$.
+
+The frequency-domain polarisation amplitudes are then defined as
+%
+\begin{eqnarray}
+  \gws_{+}(f) &=& \frac{1+\cos^{2}\iota}{2D_{L}}
+\frac{\mathcal{M}^{5/6}}{\pi^{2/3}}
+\sqrt{\frac{5}{24}}f^{-7/6}e^{-i\Psi(f)}\\
+  \gws_{\times}(f) &=&
+ i\frac{\cos^{2}\iota}{D_{L}}\frac{\mathcal{M}^{5/6}}{\pi^{2/3}}
+\sqrt{\frac{5}{24}}f^{-7/6}e^{-i\Psi(f)}
+\end{eqnarray}
+%
+where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=M\eta^{3/5}$,
+and $\iota$ is the inclination angle.  The \ac{GW} strain measured
+at the $k$'th detector is then given by
+%
+\begin{equation}
+  \label{eq:gravsig}
+   \gws^{k}(f) = \big[ F_{+}^{k}(\alpha, \delta, \psi)\gws_{+}(f) +
+F_{\times}^D(\alpha, \delta, \psi)\gws_{\times}(f)\big]
+\end{equation}
+%
+where $F_{+},F_{\times}$ are the antenna response functions which are dependent
+upon the polarisation angle $\psi$ and the sky position of the source
+$\alpha,\delta$.
 
 % Define the detectors and the noise epoch we will be using
+For this analysis we will consider advanced generation instruments operating
+at their design sensitivity.  The network we consider is the 3 detector
+H1--L1--V1 configuration.
 
 % Define the likelihood function we will be using
+We consider that the measured data in any detector is defined as
+%
+\begin{equation}
+  d^{k}(f)=\tilde{n}^{k}(f)+s\gws_{k}(f)
+\end{equation}
+%
+where $\tilde{n}(f)$ is the Fourier transform of Gaussian noise drawn from a
+distribution consistent with the power-spectral-density of the advanced
+detector noise.  The likelihood function for application to discretely sampled
+data is 
+%
+\begin{equation}
+  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp\Bigg[ 2\Delta
+f\sum_{k>0}\frac{|\tilde{d}(f_{k}) - \tilde{h}(f_{k})|^{2}}{S(f_{k})}\Bigg]
+\end{equation}
+% 
+where $\Delta f$ is the frequency resolution of the frequency domain data.
 
-% describe what parameters we will be considering
-
-% describe the priors
+% describe what parameters we will be considering and their priors
+We assume uniform prior distributions on the unknown nuisance parameters
+$\phi_{\mathrm{c}},\psi$ on their standard ranges. The time of coalescence is
+restricted to a prior range of $\pm 10$ msec around the true value.  Given our
+assumption of an \ac{sGRB} counterpart and host galaxy, we treat the sky
+position as exactly known. We also consider the cosine of the inclination angle to be drawn
+from a zero-mean Gaussian with standard deviation $0.1$ consistent with an
+uncertainty on \ac{sGRB} opening of XXX~\cm{What was exactly done here?}.  The
+main parameter of interest, the calibration scale factor $s$ is assumed to have
+a uniform prior distribution on the range $(0,2)$.
 
 % mention the no-noise issue
+We note that the analysis performed here used zero-noise realisations in which
+no actual noise was added to our signals.  The likelihood functions still
+contain the correct \ac{PSD} and hence the accuracy with which the parameters
+are recovered (indicated by the widths of the posterior distributions) is
+representative of the statistical errors in realistic noise scenarios.
 
 % Describe the ensemble of simulations used
-
-In this project the hypothesis considered, $\curlH$, simply states that the
-data, $\tilde{d}(f),$ collected by a generic detector contains the GW signal,
-$\gws(f)$, and some noise, $\tilde{n}(f)$~\cite{JVei},
-
-\begin{equation}
-  \label{eq:data}
-  \tilde{d}^D(f) = \gws^D(f) + \tilde{n}^D(f)
-\end{equation}
-
-where $D$ represents the detector site. In this project there will be three
-detectors in question: LIGO Livingston (L), LIGO Hanford (H), and Virgo (V); $D
-= \{L,H,V\}$. Testing different hypotheses is the method used to determine
-whether a data set contains a GW signal or not~\cite{SathSchutz}. This was not
-explored in this project so the above hypothesis is consistent throughout this
-paper.
-
-In this paper, a brief overview of the mathematical breakdown for the waveform
-is shown. For a more rigorous derivation, see~\cite{SathSchutz} and references
-there in. The waveform used throughout this project is a simplified system
-where the neutron stars are non rotating about their axes, in a perfectly
-circular inspiralling orbit; known as the TaylorF2 model~\cite{TF2,pN}. A more
-complex and realistic model could be tested in future for this calibration
-method. Parameter estimation using a variety of complex waveforms are carried
-out in~\cite{peadvanced}.
-
-The general equation for a GW signal detected at a generic detector, in the
-reference frame of the detector site, is given as~\cite{SathSchutz}
-
-\begin{equation}
-  \label{eq:gravsig}
-   \gws^D(f) = \big[ F_{+}^D(\alpha, \delta, \psi)\gws_{+}(f) +
-F_{\times}^D(\alpha, \delta, \psi)\gws_{\times}(f)\big]
-\end{equation}
-
-where $F_{+,\times}(\alpha, \delta, \psi)$ are the antenna response function of
-the detector, dependent on the right ascension, $\alpha$, declination,
-$\delta$, the polarisation angle $\psi$, and time. Testing the antenna patterns
-is carried out in section~\ref{sec:rndsp} of this report when the sky
-positions, $(\alpha, \delta)$, are randomised across the entire field of view
-for the detector.
-
-The polarisation amplitudes of the GW, given by $\gws_{+}(f)$, and
-$\gws_{\times}(f)$, represent the direction with which the GW compresses and
-stretches space-time. The form of these two amplitudes depends on the source
-being analysed. As mentioned above, the model used throughout this project was
-the TaylorF2 model. The parameters defining the GW polarisation amplitudes
-using the TaylorF2 model are:
-
-\begin{itemize}
-\item the total mass, $M = m_{1}+m_{2}$, where $m_{1,2}$ are the masses of the
-binary stars
-\item the symmetric mass ratio $\eta = \frac{m_{1}m_{2}}{(m_{1}+m_{2})}$
-\item chirp mass, $\mathcal{M} = \eta^{3/5}M$
-\item luminosity distance, $D_{L}$, measured in Mpc ($1 Mpc \approx
-3.1\times10^{22}m  $)
-\item the angle of inclination $\iota$. This is the angle from the source,
-taken as the angle between the normal of the orbital plane and the observer
-orientation.
-\item phase of the wave, $\psi$, which defines the polarisation of the angle of
-the GW, shown in equation~\ref{eq:psi}.
-\item the coalescence time, $t_c$, and phase of coalescence, $\phi_c$.
-\end{itemize}
-
-Together with the sky position, these parameters form the vector $\vec{\Theta}
-= \{\mathcal{M}, \eta, D_{L}, \iota, \alpha, \delta, \psi, t_{c}, \phi_{c}\}$
-which defines the entire GW signal.  The form of the polarisation amplitudes
-using the TaylorF2 model are given as~\cite{SathSchutz,pN,TF2}
-
-\begin{equation}
-  \gws_{+}(f) = \frac{1+\cos^{2}\iota}{2D_{L}}
-\frac{\mathcal{M}^{5/6}}{\pi^{2/3}} \sqrt{\frac{5}{24}}f^{-7/6}e^{-i\psi(f)}
-\end{equation}
-\begin{equation}
-  \gws_{\times}(f) =
-i\frac{\cos^{2}\iota}{D_{L}}\frac{\mathcal{M}^{5/6}}{\pi^{2/3}}
-\sqrt{\frac{5}{24}}f^{-7/6}e^{-i\psi(f)}
-\end{equation}
-
-where the phase of the GW, $\psi$, is
-\begin{equation}
-  \label{eq:psi}
-  \psi(f) = 2\pi f t_c - \phi_c + \psi_N(\eta)\sum\limits_{k=0}^4{\psi_k(\eta)
-(\pi(m_1 + m_2)f)^{(k-5)/3}}.
-\end{equation}
-
-The terms $\psi_N$ and $\psi_k$ are the standard Newtonian and post-Newtonian
-coefficients. Derivations of these two equations can be found in~\cite{pN} and
-references there-in. The simulated gravitational wave signal used throughout
-this project can be seen in figure~\ref{fig:signal}. As one can see, the signal
-strength is very faint (an order of magnitude around $10^{-24}$). This is the
-the biggest problem in detecting GWs.
-
-As mentioned above, the source considered throughout this project thought to
-produce EM radiation; SGRBs. In this project, we assume that SGRBs of the GW
-source described above have been observed. If a GRB has been seen, then the
-redshift of the merger can be calculated. This then allows for the luminosity
-distance to be calculated and in turn other defining parameters can be obtained
-(or known to a degree of uncertainty)~\cite{SathSchutz}.
-
-%More explination
-Finally, the last term in equation~\ref{eq:data} must be discussed; the noise
-of the model, $\tilde{n}(f)$. Throughout this project, we have added no noise
-to the signal, although as discussed later we do include a noise estimate in
-the likelihood.
-
-The prior probability function, $p(\vec{\theta} | \curlH)$ in
-equation~\ref{eq:bayesgw}, represents our prior knowledge in $\theta$ before
-the analysis.The range of parameters used throughout this paper were chosen due
-to their inability to be justifiably known through a GRB assumption. The
-parameters were drawn from the ranges for any signals generated in this
-analysis, given in table 1.
-
-\begin{table}
-  \centering
-  \begin{tabular}{|c|c|}
-%  \caption{Prior parameters}
-    \hline
-    Parameter & Distribution \\
-    \hline
-    $\psi$   &  U($0,\pi/2$)\\
-    $\iota$   & N($0,0.1$) \\
-    $\phi_c$  & U($0,\pi/2$)\\
-    $t_c$     & U($t_{o}-0.01, t_{o}+0.01$) \\
-    $s$       & U(0,2) \\
-    \hline
-  \end{tabular}
-  \label{tab:prior}
-  \caption{\textit{The parameters used throughout the analysis with some degree
-of uncertainty. U(a,b) is a uniform distribution between a and b. N($\mu,
-\sigma$) is a normal distribution, with a mean $\mu$ and standard deviation
-$\sigma$. $t_{o} = 9\times10^{8}$}}
-\end{table}
-
-Through the assumption made above allows us to constrain the system,
-maintaining a small number of parameters in the prior. For a realistic
-parameter estimation of the source, a larger range of values with a bigger
-distribution. This was not tested due to time constraints but is proposed as
-work for the future.
-
-The likelihood functiona used throughout this project is given for a generic
-detector D as~\cite{JVei},
-
-\begin{equation}
-  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp(-(d^{D}(f) -
-h^{D}(f)|d^{D}(f) - h^{D}(f))/2)
-\end{equation}
-
-where $(\cdot |\cdot )$ is a representation of the inner product. This
-derivation is given in the appendix. Using the inner product of two real
-functions, and approximating for discrete integer values, the likelihood then
-becomes
-
-\begin{equation}
-  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp\Bigg[ 2\Delta
-f\sum_{k>0}\frac{|\tilde{d}(f_{k}) - \tilde{h}(f_{k})|^{2}}{S(f_{k})} \Bigg]
-\end{equation}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.5\textwidth]{PSD}
-  \caption{\textit{Noise spectral density for advanced LIGO and Virgo.}}
-  \label{fig:psd}
-\end{figure}
-
-where $S(f_{k})$ is the \textit{power spectral density} (PSD). This defines the
-sensitivity of the detector as it a measure of the strain in the detector due
-to noise~\cite{SathSchutz,sens}. The PSD for advanced LIGO and Virgo are shown
-in figure~\ref{fig:psd}. The figure shows that the strain effect for advanced
-LIGO is less than it is for Virgo; LIGO has a better sensitivity. This is a
-factor in analysing the calibration error, given in section~\ref{sec:results}.
-The subscript $k$ represents discrete values of relevant frequency bins (As
-carrying out the integral would become computationally costly). This likelihood
-will be used throughout the project. The factor of $\Delta f$ comes from the
-approximation of the inner product of two real functions (as it is integrated
-over frequency space).
-
-The aim of this project was to determine how well the calibration error could
-be parameterised in a software injected gravitational wave source. The error
-was estimated through numerous situations but the basis of the simulations
-remained the same throughout the project: a single source of GWs at varying
-signal to noise ratios. The GRB assumption was analysed as well by relaxing
-some of the defining GW paramaters and adding some uncertainty into their
-value, given in section~\ref{sec:pandl}.
-
-These values where chosen due to the assumed counterpart EM observation. These
-values could not be justified by that assumption so they were added into the
-prior with reasonable uncertainty.
-
-The estimation of the error, given in its simplest form as a scale factor, was
-performed for numerous systems: single detector, a network of detectors, random
-sky positions, etc. Before the parameter estimation was carried out, the
-relationship between distance of the source relative to the detector and the
-signal to noise ratio (SNR) needed to be found.
-
-The SNR is given by
-\begin{equation}
-  \label{eq:snr}
-  SNR = \sqrt{\Bigg( \frac{2 |\gws^D(f)|^2}{S^D(f)} df \Bigg)}
-\end{equation}
 

### 2015-04-23 docs/cosmic_cal.tex
Some very minor changes - mainly adding some references

@@
 \begin{abstract}
-  We present an investigation into the potential of calibrating
+  We present an investigation into the potential of calibrating \MP{change to: ``assessing the validity of the calibration''?}
   gravitational wave detector outputs through the use of standard
@@
 It is expected that the advanced generation of interferometric \ac{GW}
-detectors will detect such waves emitted from $O(10s)$ of compact binary
+detectors will detect waves emitted from $O(10s)$ of compact binary
 coalescences.  One such class of these cataclysmic events, the inspiral and
@@
 systems can be used as cosmiological distance markers, otherwise known as
-"standard sirens" (analogous to the \ac{EM} standard candles). Standard sirens
+``standard sirens'' (analogous to the \ac{EM} standard candles). Standard sirens
 allow ...
@@
 % talk about the calibration idea
-If a single \ac{GW} event is observed in coincidence with a \ac{sGRB} it will
+If a single \ac{GW} event is observed in coincidence with a \ac{sGRB} it will \MP{may?}
 be possible to identify the host galaxy of the \ac{BNS} source.  With this
@@
 and other cosmological parameters, the true luminosity distance can be
-estimated to $\sim 1\%$ accuracy  The \ac{GW} measurement acts as a standard
+estimated to $\sim 1\%$ accuracy.  The \ac{GW} measurement acts as a standard
 siren also giving us a direct measurement of the luminosity distance to the source.
@@
 
+\MP{In this paper I think we stick to just talking about a single overall amplitude calibration scale.}
+
 The technique used throughout GW research in calibrating the detectors is
@@
 overview of the calibration method currently used, and is an extensive study in
-itself, see~\cite{Vitale} and references therein. For advanced LIGO and Virgo,
+itself, see~\cite{Vitale:2012} and references therein. For advanced LIGO and Virgo,
 at the frequency range used in this project, the error in the calibration using
-the 'hardware injection method' is roughly $10\%$~\cite{Vitale}. This is a
+the `hardware injection method' is roughly $10\%$~\cite{Vitale:2012}. This is a
 benchmark set for the error estimation using the proposed new method in this
@@
 $h(t)$ is the exact perfectly calibrated strain. A description of the
-collected data is given in section~\ref{sec:cbc}. The scale factor being
+collected data is given in Section~\ref{sec:cbc}. The scale factor being
 parameterised, $s$, represents all known sources of error in the system. This
@@
 half of the \ac{sGRB} beaming angle.  This prerequisite property limits us to
-systems that are approximately "face-on" and therefore biases us to higher
+systems that are approximately ``face-on'' and therefore biases us to higher
 \ac{SNR} signals.  However, as discussed earlier, the property of beaming
@@
 
+\MP{We emphasise that our work has assumed that the PSD used in the likelihood function
+for estimating the calibration scale factor is an accurate representation of the
+PSD at the time of the observed CBC signal. In reality the PSD may be estimated from
+a period close to, but not overlapping, the signal time and therefore may be slightly
+different (see e.g. \cite{2013PhRvD..88h4044L}). Our result would therefore be completely correlated
+with any uncertainty in the PSD estimate, but would still offer an upper limit
+on the calibration uncertainty.}
+
+\MP{The BayesLine algorithm \cite{2015PhRvD..91h4034L} may provide a natural
+way to perform this analysis in particular when accounting for a frequency
+dependent calibration.}
+
 \ack

### 2015-04-24 docs/cosmic_cal.tex
Some minor changes, but importantly I have correctly the likelihood function definition to be consistent with what is (correctly) actually used

@@
 \begin{equation}
-  d^{k}(f)=\tilde{n}^{k}(f)+s\gws_{k}(f)
+  \tilde{d}_{k}(f)=\tilde{n}_{k}(f)+s\gws_{k}(f,\vec{\theta})
 \end{equation}
@@
 distribution consistent with the power-spectral-density of the advanced
-detector noise.  The likelihood function for application to discretely sampled
-data is 
+detector noise.
+
+\MP{Our aim is to estimate the marginal posterior probability distribution function (pdf)
+for the unknown calibration scale factor. We use Bayes' theorem for which we need to define
+a likelihood function and prior pdfs on the signal parameters being estimated. In this
+analysis we use a nested sampling algorithm \cite{Skilling} to provide samples from the
+marginal posteriors that we require.}
+We use a Gaussian likelihood function given by
 %
+%\begin{equation}
+%  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp\Bigg[ 2\Delta
+%f\sum_{k>0}\frac{|\tilde{d}(f_{k}) - \tilde{h}(f_{k})|^{2}}{S(f_{k})}\Bigg]
+%\end{equation}
 \begin{equation}
-  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp\Bigg[ 2\Delta
-f\sum_{k>0}\frac{|\tilde{d}(f_{k}) - \tilde{h}(f_{k})|^{2}}{S(f_{k})}\Bigg]
+  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm det}}
+  \sum_{i=i_{f_{\rm low}}}^{i_{f_{\rm high}}}\frac{\Re{\left\{ \tilde{d}^{*}_{k,i} \gws_{k,i}(\vec{\theta})
+  - \frac{1}{2}\left(\gws^*_{k,i}(\vec{\theta})\gws_{k,i}(\vec{\theta}) + \tilde{d}^*_{k,i}\tilde{d}_{k,i}\right) \right\}}}{S_{k,i}}\right]}
 \end{equation}
 % 
-where $\Delta f$ is the frequency resolution of the frequency domain data.
+where $\Delta f$ is the frequency resolution of the data, $S_k$ is the
+$k^{\rm th}$ detector's one sided noise power spectral density, and the $i$ indices
+increment over frequency with $i_{f_{\rm low}}$ and $i_{f_{\rm high}}$ corresponding to
+the lower and upper range in frequencies used for our
+analsysis. \MP{Need to say the
+range of frequencies we integrate over, which I think was 50\,Hz to 1600\,Hz from looking
+at Liam's code.} The source parameters, and the unknown scale factor, have been subsumed
+into a vector $\vec{\theta}$.
 
@@
 We assume uniform prior distributions on the unknown nuisance parameters
-$\phi_{\mathrm{c}},\psi$ on their standard ranges. The time of coalescence is
+$\phi_{\mathrm{c}},\psi$ on the ranges $[0, 2\pi]$ and $[0, \pi/2]$ respectively.
+The time of coalescence is
 restricted to a prior range of $\pm 10$ msec around the true value.  Given our

### 2015-04-24 docs/cosmic_cal.tex
Some minor changes in discussion about calibration

@@
 \usepackage {url}
-%\usepackage[normalem]{ulem}   % temporal one in draft.
+\usepackage[normalem]{ulem}   % temporal one in draft.
 
@@
 
+\MP{The gravitational wave strain is measured through the differential arm length changes
+of the interferometer via $h = \Delta L(f) / L$, where $L$ is the total arm length.
+Calibration is required to relate the actual measured interferometer error signal ouptut
+$e(f)$ to this differential arm length $\Delta L$. This relation
+is known as the length response function, $R(f,t)$, defined such that
+\begin{equation}
+\Delta L(f) = R(f,t) e(f),
+\end{equation}
+where $R$ various slowly in time (in comparison to transient signal timescales).
+Calculation of $R$ requires the measurement of various functions within a control feedback loop
+(see \cite{Vitale:2012} for a brief overview of this), which are subject to
+uncertainties. In this study we will assume an estimate of $R$ is available (although
+in theory we could take on the role of estimating $R$ itself), but that it differs from the
+truth through some unknown scale factor, $s$, so that
+\begin{equation}
+h_{\rm true}(f) = s h_{\rm measured}(f) = s \frac{R(f,t) e(f)}{L}.
+\end{equation}
+In this analysis we will simplify the situation by assuming that $s$ is a constant with
+respect to frequency, but in future studies that assumption could be dropped and $s$
+could take some functional form with respect to $f$.
+}
+
 % Describe our model for the calibration
@@
 where $h_{\mathrm{m}}(t)$ is the measured detector strain, and
-$h(t)$ is the exact perfectly calibrated strain. A description of the
-collected data is given in Section~\ref{sec:cbc}. The scale factor being
+$h(t)$ is the exact perfectly calibrated strain. \sout{A description of the
+collected data is given in Section~\ref{sec:cbc}}. The scale factor being
 parameterised, $s$, represents all known sources of error in the system. This

### 2015-04-27 docs/cosmic_cal.tex
Some more minor changes

@@
 \usepackage[normalem]{ulem}   % temporal one in draft.
+
+%\usepackage{harvard}
 
@@
 450$ Mpc. Assuming the current best estimates for the cosmological parameters,
-this is equivelent to a redshift $z\approx 0.1$. As noted by~\cite{} \ac{CBC}
+this is equivelent to a redshift $z\approx 0.1$. As noted by \cite{1986Natur.323..310S} \ac{CBC}
 systems can be used as cosmiological distance markers, otherwise known as
@@
 overview of the calibration method currently used, and is an extensive study in
-itself, see~\cite{Vitale:2012} and references therein. For advanced LIGO and Virgo,
+itself, see \cite{Vitale:2012} and references therein. For advanced LIGO and Virgo,
 at the frequency range used in this project, the error in the calibration using
-the `hardware injection method' is roughly $10\%$~\cite{Vitale:2012}. This is a
+the `hardware injection method' is roughly $10\%$ \cite{Vitale:2012}. This is a
 benchmark set for the error estimation using the proposed new method in this
@@
 
-\MP{The gravitational wave strain is measured through the differential arm length changes
+\MP{The gravitational wave strain is measured through the differential arm length, $\Delta L$, changes
 of the interferometer via $h = \Delta L(f) / L$, where $L$ is the total arm length.
 Calibration is required to relate the actual measured interferometer error signal ouptut
-$e(f)$ to this differential arm length $\Delta L$. This relation
-is known as the length response function, $R(f,t)$, defined such that
+$e(f)$ to $\Delta L$. This relation is known as the length response function, $R(f,t)$,
+defined such that
 \begin{equation}
@@
 \end{equation}
-where $R$ various slowly in time (in comparison to transient signal timescales).
+where $R$ varies slowly in time (in comparison to transient signal timescales).
 Calculation of $R$ requires the measurement of various functions within a control feedback loop
@@
 \begin{equation}
-h_{\rm true}(f) = s h_{\rm measured}(f) = s \frac{R(f,t) e(f)}{L}.
+h(f) = s h_{\rm m}(f) = s \frac{R(f,t) e(f)}{L},
 \end{equation}
+where $h(f)$ is the true strain and $h_{\rm m}(f)$ is the measured strain.
 In this analysis we will simplify the situation by assuming that $s$ is a constant with
 respect to frequency, but in future studies that assumption could be dropped and $s$
-could take some functional form with respect to $f$.
+could take some functional form, or piece-wise fit, with respect to $f$.
 }
@@
 
+\MP{We want to assess how well a calibration scale factor can be estimated
+from a single observed \ac{GW} associated with a particular \ac{sGRB}. As we {\it a priori}
+have no knowledge of the likely location and distance of such an event we have performed
+simulations of multiple events to see the expected distribution the accuracy of the
+calibration scale factor recovery.}
+
 % Describe what we are going to do in this section
@@
 position is therefore known to high precision.
+
+\MP{We perform 100 simulations at each of a range of distances from 10\,Mpc out
+to 200\,Mpc. For the simulations at each distance range we have allowed the
+source's sky position to be randomly drawn from a uniform distribution over
+the sky, the polarisation angle $\psi$ to drawn from a uniform distribution in the
+range $[0, \pi/2]$, the coalescence phase to be drawn from a uniform distribution
+in the range $[0, 2\pi]$, and the inclination angle drawn from a Gaussian with range XX?.
+The binary system masses have been fixed as $1.4\,{\rm M}_\odot$ (based on the
+very good assumption that whatever masses are chosen in these cases they will be
+recovered to high accuracy) and the time of coalescence is fixed.}
 
@@
 For this analysis we will consider advanced generation instruments operating
-at their design sensitivity.  The network we consider is the 3 detector
-H1--L1--V1 configuration.
+at their design sensitivity, \MP{with noise power spectral densities taken from
+\cite{2013arXiv1304.0670L}}. The network we consider is the 3 detector H1--L1--V1 configuration.
 
@@
 
+\subsection{Analysis method}
+
 \MP{Our aim is to estimate the marginal posterior probability distribution function (pdf)
@@
 a likelihood function and prior pdfs on the signal parameters being estimated. In this
-analysis we use a nested sampling algorithm \cite{Skilling} to provide samples from the
-marginal posteriors that we require.}
+analysis we use a nested sampling algorithm \cite{Skilling}\footnote{The code used was based
+on {\tt matlabmultinest} found at \url{https://github.com/mattpitkin/matlabmultinest}.}
+to provide samples from the marginal posteriors that we require.}
 We use a Gaussian likelihood function given by
@@
 analsysis. \MP{Need to say the
-range of frequencies we integrate over, which I think was 50\,Hz to 1600\,Hz from looking
+range of frequencies we integrate over, which I think was 40\,Hz to 1600\,Hz from looking
 at Liam's code.} The source parameters, and the unknown scale factor, have been subsumed
@@
 
+\subsection{Simulations}
+
 % mention the no-noise issue
@@
 
-where the noise spectral density, $S^D(f)$ is not a function of distance, shown
-in section~\ref{sec:pandl}. For coalescing binary neutron stars of masses
+For coalescing binary neutron stars of masses
 $m_{1,2} = 1.4\,M_{\odot}$ and  $D_{L}200\,Mpc$, the expected SNR is roughly 8.
@@
 200\,Mpc$. The detector limits for the advanced era is a luminosity distance of
-$\sim 450\,Mpc$. This was not initially tested as the SNR was low enough to
+$\sim 450\,Mpc$. \sout{This was not initially tested as the SNR was low enough to
 cause problems with computational time. It was however tested in the random sky
 position simulations, see section~\ref{sec:rndsp}. The initial value for the
-calibration error used in the analysis was chosen randomly.
+calibration error used in the analysis was chosen randomly.}
 
@@
 
-A network of detectors are used throughout GW research. Being able to show the
+\sout{A network of detectors are used throughout GW research. Being able to show the
 effect of the calibration error and recovering its uncertainty in the analysis
@@
 factor with the least amount of uncertainty for differing levels of SNR for a
-single source of GWs.
+single source of GWs.}
 
@@
 \MP{We emphasise that our work has assumed that the PSD used in the likelihood function
-for estimating the calibration scale factor is an accurate representation of the
+when estimating the calibration scale factor is an accurate representation of the
 PSD at the time of the observed CBC signal. In reality the PSD may be estimated from
 a period close to, but not overlapping, the signal time and therefore may be slightly
-different (see e.g. \cite{2013PhRvD..88h4044L}). Our result would therefore be completely correlated
+different \cite{2013PhRvD..88h4044L}. Our result would therefore be completely correlated
 with any uncertainty in the PSD estimate, but would still offer an upper limit
@@
 We would like to acknowledge the useful discussions with a whole bunch
-of people.
+of people. LW was part funded for this work through a Royal Astronomical
+Society Undergraduate Research Bursary.
+MP is funded by the STFC under grant number ST/L000946/1.
 
@@
 
-\bibliographystyle{unsrt}
+%\biaxbliographystyle{jphysicsB}
+\bibliographystyle{iopart-num}
 \bibliography{masterbib}

### 2015-04-28 src/python/scale_factor_mcmc.py
Added initial draft python script for rerunning the analysis

@@ -1,0 +1,384 @@
+#!/usr/bin/env python
+
+"""
+A script to run the calibration scale factor check using the emcee MCMC sampler
+"""
+import numpy as np
+import copy
+import sys
+from optparse import OptionParser
+
+# swig lal modules for signal generation
+import lal
+import lalsimulation
+
+# MCMC code
+import emcee
+
+# pyCBC code for generating coloured frequency noise
+from pycbc.types import frequencyseries
+from pycbc import noise
+
+# a function to compute the antenna response
+def antenna_response( gpsTime, ra, dec, psi, det ):
+  gps = lal.LIGOTimeGPS( gpsTime )
+  gmst_rad = lal.GreenwichMeanSiderealTime(gps)
+
+  # create detector-name map
+  detMap = {'H1': lal.LALDetectorIndexLHODIFF, \
+            'H2': lal.LALDetectorIndexLHODIFF, \
+            'L1': lal.LALDetectorIndexLLODIFF, \
+            'G1': lal.LALDetectorIndexGEO600DIFF, \
+            'V1': lal.LALDetectorIndexVIRGODIFF, \
+            'T1': lal.LALDetectorIndexTAMA300DIFF, \
+            'AL1': lal.LALDetectorIndexLLODIFF, \
+            'AH1': lal.LALDetectorIndexLHODIFF, \
+            'AV1': lal.LALDetectorIndexVIRGODIFF}
+
+  try:
+    detector=detMap[det]
+  except KeyError:
+    raise ValueError, "ERROR. Key %s is not a valid detector name." % (det)
+
+  # get detector
+  detval = lal.CachedDetectors[detector]
+
+  response = detval.response
+
+  # actual computation of antenna factors
+  fp, fc = lal.ComputeDetAMResponse(response, ra, dec, psi, gmst_rad)
+
+  return fp, fc
+
+"""
+a function to generate a frequency domain inspiral waveform (see e.g. 
+https://www.lsc-group.phys.uwm.edu/ligovirgo/cbcnote/WaveformsSanityChecks?action=AttachFile&do=view&target=
+lalsimutils.py)
+This function sets spin parameters to zero, so all output are non-spinning waveforms. The reference frequency
+is set to fmin.
+
+Inputs are:
+ - phiref - the reference phase (rads)
+ - deltaF - the frequency bin size (Hz)
+ - m1 - the mass of the first component (solar masses)
+ - m2 - the mass of the second component (solar masses)
[…]

### 2015-04-29 src/python/scale_factor_mcmc.py
Editted python script so data and PSD can be noisy

@@ -6,7 +6,10 @@ A script to run the calibration scale factor check using the emcee MCMC sampler
 import numpy as np
 import copy
 import sys
+import os
 from optparse import OptionParser
+
+from matplotlib import pyplot as pl
 
 # swig lal modules for signal generation
 import lal
@@ -14,10 +17,6 @@ import lalsimulation
 
 # MCMC code
 import emcee
-
-# pyCBC code for generating coloured frequency noise
-from pycbc.types import frequencyseries
-from pycbc import noise
 
 # a function to compute the antenna response
 def antenna_response( gpsTime, ra, dec, psi, det ):
@@ -30,10 +29,7 @@ def antenna_response( gpsTime, ra, dec, psi, det ):
             'L1': lal.LALDetectorIndexLLODIFF, \
             'G1': lal.LALDetectorIndexGEO600DIFF, \
             'V1': lal.LALDetectorIndexVIRGODIFF, \
-            'T1': lal.LALDetectorIndexTAMA300DIFF, \
-            'AL1': lal.LALDetectorIndexLLODIFF, \
-            'AH1': lal.LALDetectorIndexLHODIFF, \
-            'AV1': lal.LALDetectorIndexVIRGODIFF}
+            'T1': lal.LALDetectorIndexTAMA300DIFF}
 
   try:
     detector=detMap[det]
@@ -83,6 +79,48 @@ def fdwaveform(phiref, deltaF, m1, m2, fmin, fmax, dist, incl):
   return hptilde.data.data, hctilde.data.data
 
 
+# A function to generate frequency domain coloured noise
+# This function is adapted from pyCBC in the noise module:
+# https://github.com/ligo-cbc/pycbc/blob/master/pycbc/noise/gaussian.py
+# copyright of Alex Nitz (2012)
+def frequency_noise_from_psd(psd, deltaF, seed = None):
+    """ Create noise with a given psd.
+    
+    Return noise coloured with the given psd. The returned noise 
+    has the same length and frequency step as the given psd. 
+    Note that if unique noise is desired a unique seed should be provided.
+    Parameters
+    ----------
+    psd : np.array
+        The noise weighting to color the noise.
+    deltaF: float
+        The frequency step size
+    seed : {0, int} or None
+        The seed to generate the noise. If None specified,
[…]

### 2015-04-29 src/python/scale_factor_mcmc.py
some additional outputs for script

@@ -8,6 +8,7 @@ import copy
 import sys
 import os
 from optparse import OptionParser
+import json
 
 from matplotlib import pyplot as pl
 
@@ -203,6 +204,24 @@ def lnlike(theta, y, dd, m1, m2, dist, fmin, fmax, deltaF, resps, asds):
   return L
 
 
+# a function to get the credible intervals using a greedy binning method
+def credible_interval(dsamples, ci):
+    n, binedges = np.histogram(dsamples, bins=100)
+    dbins = binedges[1]-binedges[0] # width of a histogram bin
+    bins = binedges[0:-1]+dbins/2. # centres of bins
+
+    histIndices=np.argsort(n)[::-1]  # indices of the points of the histogram in decreasing order
+
+    frac = 0.0
+    j = 0
+    for i in histIndices:
+        frac += float(n[i])/float(len(dsamples))
+        j = j+1
+        if frac >= ci:
+            break
+
+    return (np.min(bins[histIndices[:j]]), np.max(bins[histIndices[:j]]))
+
 
 if __name__=='__main__':
   
@@ -287,6 +306,11 @@ a uniform distribution between 0 and pi/2.")
                      help="If this flag is non-zero the PSD is estimated as the average of \
 the given number of noisy PSD estimates.")
   
+  parser.add_option("-P", "--plot", dest="plot", default=False, action="store_true",
+                    help="If this flag is set the posteriors will be plotted (requires triangle.py).")
+
+  parser.add_option("-T", "--seed", dest="seed", type="int",
+                     help="A numpy random number generator seed value.")
 
   # parse input options
   (opts, args) = parser.parse_args()
@@ -320,8 +344,6 @@ the given number of noisy PSD estimates.")
   # injected source options
   dist = opts.dist
   t0 = opts.t0
-  phi0 = opts.phi0
-  psi = opts.psi
   
   if not opts.__dict__['ra']:
     ra = 2.*np.pi*np.random.rand() # generate RA uniformly between 0 and 2pi
@@ -353,6 +375,9 @@ the given number of noisy PSD estimates.")
     scales = [1.]
   else:
     scales = opts.scales
+
+  if opts.__dict__['seed']:
+    np.random.seed(opts.seed) # set the random seed
 
[…]

### 2015-04-29 src/python/simulation_check.py
Add python script to just double check that the injections are correct

@@ -1,0 +1,289 @@
+#!/usr/bin/env python
+
+"""
+A script to check that the injections are as expected
+"""
+import numpy as np
+import copy
+import sys
+import os
+from optparse import OptionParser
+import json
+
+from matplotlib import pyplot as pl
+
+# swig lal modules for signal generation
+import lal
+import lalsimulation
+
+# a function to compute the antenna response
+def antenna_response( gpsTime, ra, dec, psi, det ):
+  gps = lal.LIGOTimeGPS( gpsTime )
+  gmst_rad = lal.GreenwichMeanSiderealTime(gps)
+
+  # create detector-name map
+  detMap = {'H1': lal.LALDetectorIndexLHODIFF, \
+            'H2': lal.LALDetectorIndexLHODIFF, \
+            'L1': lal.LALDetectorIndexLLODIFF, \
+            'G1': lal.LALDetectorIndexGEO600DIFF, \
+            'V1': lal.LALDetectorIndexVIRGODIFF, \
+            'T1': lal.LALDetectorIndexTAMA300DIFF}
+
+  try:
+    detector=detMap[det]
+  except KeyError:
+    raise ValueError, "ERROR. Key %s is not a valid detector name." % (det)
+
+  # get detector
+  detval = lal.CachedDetectors[detector]
+
+  response = detval.response
+
+  # actual computation of antenna factors
+  fp, fc = lal.ComputeDetAMResponse(response, ra, dec, psi, gmst_rad)
+
+  return fp, fc
+
+"""
+a function to generate a frequency domain inspiral waveform (see e.g. 
+https://www.lsc-group.phys.uwm.edu/ligovirgo/cbcnote/WaveformsSanityChecks?action=AttachFile&do=view&target=
+lalsimutils.py)
+This function sets spin parameters to zero, so all output are non-spinning waveforms. The reference frequency
+is set to fmin.
+
+Inputs are:
+ - phiref - the reference phase (rads)
+ - deltaF - the frequency bin size (Hz)
+ - m1 - the mass of the first component (solar masses)
+ - m2 - the mass of the second component (solar masses)
+ - fmin - the lower bound on frequency (Hz)
+ - fmax - the upper bound on frequency (Hz)
+ - dist - the source distance (in Mpc)
+ - incl - the source inclination angle (rads)
+
+This function does not apply the antenna pattern to the output.
+"""
[…]

### 2015-07-01 src/python/scale_factor_mcmc.py
scale_factor_mcmc.py: changes include:  - use lognormal prior on the scale factors with a mode of 1 and a standard deviation of 1 (which gives a mean of 1)  - set the code so that to run the MCMC for a single detector analysis the SNR must be greater than sqrt(2.*5.5^2), while for a multi-detector analysis there must be an SNR of > 5.5 in at least two detector (this is the CBC candidate selection criterion from the S6 analysis) [This criterion an be overridden if required with an input flag]

@@ -109,7 +109,7 @@ def frequency_noise_from_psd(psd, deltaF, seed = None):
     """
     sigma = 0.5 * np.sqrt(psd / deltaF)
     if seed is not None:
-        np.random.seed(seed)
+      np.random.seed(seed)
     
     not_zero = (sigma != 0) & np.isfinite(sigma)
     
@@ -168,12 +168,25 @@ def lnprior(theta, iotawidth, tccentre, ndets):
     return -np.inf
   
   # work in log scale factors so a Jeffrey's prior of scale factor within range 0.1 to 10 is flat in log space
+  #for s in ss:
+  #  if np.log(0.1) < s < np.log(10.):
+  #    lp = 0.
+  #  else:
+  #    return -np.inf
+  
+  # use a log normal prior on the scale factor with a mode of 1 and a standard deviation of 1
+  # (using a standard deviation of 1 means that the probability at 0.1 and 10 will be a factor
+  # of 14.17 less than at 1) - for this I will continue working in log of scale factors
+  lognormalsigma = 1.
+  lognormalmu = 1.
   for s in ss:
-    if np.log(0.1) < s < np.log(10.):
-      lp = 0.
+    #lp += -(s + 0.5*np.log(2.*np.pi*lognormalmu**2)) - 0.5*((s - lognormalmu)**2/lognormalsigma**2)
+    if s > 0.:
+      lns = np.log(s)
+      lp += -(lns + 0.5*np.log(2.*np.pi*lognormalsigma**2)) - 0.5*((lns - lognormalmu)**2/lognormalsigma**2)
     else:
       return -np.inf
-  
+ 
   # add iota prior
   lp -= 0.5*(iota/iotawidth)**2
 
@@ -207,7 +220,8 @@ def lnlike(theta, y, dd, dist, fmin, fmax, deltaF, resps, asds):
   for i in range(len(resps)):
     Ap, Ac = resps[i]
 
-    H = (hp*(Ap*cpsi + Ac*spsi) + hc*(Ac*cpsi - Ap*spsi))*np.exp(ss[i])
+    #H = (hp*(Ap*cpsi + Ac*spsi) + hc*(Ac*cpsi - Ap*spsi))*np.exp(ss[i])
+    H = (hp*(Ap*cpsi + Ac*spsi) + hc*(Ac*cpsi - Ap*spsi))*ss[i]
 
     Hs = H/asds[i]
     Hs[~np.isfinite(Hs)] = 0.
[…]

### 2015-07-01 src/python/results_script.py
Remove whiskers from boxplots

@@ -182,7 +182,8 @@ for dist in dists:
   positions.append(dist)
   positions.append(dist+10)
  
-bp = pl.boxplot(data, notch=0, sym='+', positions=positions, widths=8, showfliers=False) 
+#bp = pl.boxplot(data, notch=0, sym='+', positions=positions, widths=8)
+bp = pl.boxplot(data, notch=0, sym='', positions=positions, widths=8) 
 
 pl.setp(bp['boxes'], color='black')
 pl.setp(bp['whiskers'], color='black')

### 2015-07-01 src/python/results_script.py
results_script.py: some changes to the results plotting script

@@ -71,7 +71,6 @@ pl.rc('font', family='serif')
 pl.rc('font', size=14)
 
 dists = [50, 100, 150, 200, 250, 300, 350, 400, 450]
-colours = ['r', 'm', 'c', 'g', 'b', 'p', 'k']
 
 prefix = indir
 
@@ -107,10 +106,10 @@ for i, d in enumerate(dirs):
   nprelsf = np.array(relsf)
   print nprelsf.shape
   
-  # divide by two to get the half widths
-  nprelsf[:,0] = nprelsf[:,0]/(2.*info['InjectionParameters']['scales'][0])
-  nprelsf[:,1] = nprelsf[:,1]/(2.*info['InjectionParameters']['scales'][1]) 
-  nprelsf[:,2] = nprelsf[:,2]/(2.*info['InjectionParameters']['scales'][2])
+  # divide by two to get the half widths and convert to percentage
+  nprelsf[:,0] = 100.*nprelsf[:,0]/(2.*info['InjectionParameters']['scales'][0])
+  nprelsf[:,1] = 100.*nprelsf[:,1]/(2.*info['InjectionParameters']['scales'][1]) 
+  nprelsf[:,2] = 100.*nprelsf[:,2]/(2.*info['InjectionParameters']['scales'][2])
 
   data.append(nprelsf[:,0])
   data.append(nprelsf[:,1])
@@ -138,44 +137,6 @@ for i, d in enumerate(dirs):
   #  pl.plot([dists[i], dists[i]], ci2, 'r', lw=2)
   #  pl.plot([dists[i]+7.5, dists[i]+7.5], ci3, 'g', lw=2)
 
-
-# check for directory at distance 10000 (i.e. noise only) and if so plot the
-# *minimum* value from that, to show what to expect from noise
-
-minH1 = None
-minL1 = None
-minV1 = None
-
-ldir = os.path.join(prefix, '10000Mpc')
-if os.path.isdir(ldir):
-  relsf = []
-
-  files = [os.path.join(ldir, f) for f in os.listdir(ldir) if os.path.isfile(os.path.join(ldir, f)) and fnpre in f]
-
-  # go through files and extract the relative standard devaition on the scale factors for each detector
-  for f in files:
-    fo = open(f, 'r')
-    info = json.load(fo)
-    fo.close()
-
-    vals = []
-    for k in range(len(info['InjectionParameters']['scales'])):
[…]

### 2015-07-02 docs/cosmic_cal.tex
Some major changes to the paper

@@
 %
-\documentclass[12pt]{iopart}
-\usepackage {latexsym}
-\usepackage {graphicx}
+\documentclass[10pt]{iopart}
+\usepackage{latexsym}
+\usepackage{graphicx}
 \usepackage{iopams}
-\usepackage {color}
+\usepackage{color}
 
@@
 %\usepackage {hhline}
-\usepackage {hyperref}
+\usepackage{hyperref}
 %\usepackage {color}
-\usepackage {bm}
+\usepackage{bm}
 %\usepackage{appendix}
@@
 %\usepackage{dcolumn}   % needed for some tables
-\usepackage {url}
+\usepackage{url}
 \usepackage[normalem]{ulem}   % temporal one in draft.
+\usepackage{xspace}
 
 %\usepackage{harvard}
+\usepackage{amsopn} % for using DeclareMathOperator
 
@@
 
-\newcommand{\aap}{A\&A }
-\newcommand{\apj}{ApJ  }
-\newcommand{\apjl}{ApJL  }
-\newcommand{\apjs}{ApJS  }
-\newcommand{\prd}{Phys. Rev. D  }
-\newcommand{\nat}{Nature}
-\newcommand{\araa}{ARA\&A}
-\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society  }
+\DeclareMathOperator{\erf}{erf}
+
 \begin{document}
@@
   Glasgow, Glasgow G12 8QQ, United Kingdom}
-\ead{1002990W@student.gla.ac.uk}
+\eads{\mailto{matthew.pitkin@glasgow.ac.uk}, \mailto{christopher.messenger@glasgow.ac.uk}}
 
@@
   galaxy is identified for a given source then its redshift can be
-  obtained and combined with current knowledge of the cosmologic
+  obtained and combined with current knowledge of the cosmological
   parameters.  This will yield the true luminosity distance and allow
@@
 \acrodef{BNS}[BNS]{binary neutron star}
-\acrodef{MWEG}[MWEG]{Milky Way equivelent galaxies}
+\acrodef{MWEG}[MWEG]{Milky Way equivalent galaxies}
 \acrodef{BBH}[BBH]{binary black hole}
@@
 450$ Mpc. Assuming the current best estimates for the cosmological parameters,
-this is equivelent to a redshift $z\approx 0.1$. As noted by \cite{1986Natur.323..310S} \ac{CBC}
-systems can be used as cosmiological distance markers, otherwise known as
+this is equivalent to a redshift $z\approx 0.1$. As noted by \cite{1986Natur.323..310S} \ac{CBC}
+systems can be used as cosmological distance markers, otherwise known as
 ``standard sirens'' (analogous to the \ac{EM} standard candles). Standard sirens
@@
 
+% Describe our model for the calibration
 \MP{The gravitational wave strain is measured through the differential arm length, $\Delta L$, changes
-of the interferometer via $h = \Delta L(f) / L$, where $L$ is the total arm length.
-Calibration is required to relate the actual measured interferometer error signal ouptut
+of the interferometer via $h = \Delta L(f) / L$, where $L$ is the full arm length.
+Calibration is required to relate the actual measured interferometer error signal output
 $e(f)$ to $\Delta L$. This relation is known as the length response function, $R(f,t)$,
@@
 respect to frequency, but in future studies that assumption could be dropped and $s$
-could take some functional form, or piece-wise fit, with respect to $f$.
-}
-
-% Describe our model for the calibration
-In this work the calibration correction factor used throughout is simply a scale
-factor, $s$. This effects the data set collected by a simple multiplication:
-%
-\begin{equation}\label{eq:scaledata}
-  h(t) = s\times h_{\mathrm{m}}(t)
-\end{equation}
-%
-where $h_{\mathrm{m}}(t)$ is the measured detector strain, and
-$h(t)$ is the exact perfectly calibrated strain. \sout{A description of the
-collected data is given in Section~\ref{sec:cbc}}. The scale factor being
-parameterised, $s$, represents all known sources of error in the system. This
-is not entirely realistic as $s$ is known to be a function of frequency.
-Testing a more realistic calibration error is work to be done in the
-future.~\cm{Need to clean this up and expand the discussion}
+could take some functional form, or piece-wise fit, with respect to $f$.}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A single event with GRB counterpart\label{sec:single}}
+\section{Analysis}
 
@@
 %
-\begin{eqnarray}
+\begin{eqnarray}\label{eq:signal}
   \gws_{+}(f) &=& \frac{1+\cos^{2}\iota}{2D_{L}}
 \frac{\mathcal{M}^{5/6}}{\pi^{2/3}}
-\sqrt{\frac{5}{24}}f^{-7/6}e^{-i\Psi(f)}\\
+\sqrt{\frac{5}{24}}f^{-7/6}e^{-i\Psi(f)} \nonumber \\
   \gws_{\times}(f) &=&
@@
 \begin{equation}
-  \tilde{d}_{k}(f)=\tilde{n}_{k}(f)+s\gws_{k}(f,\vec{\theta})
+  \tilde{d}_{k}(f)=\tilde{n}_{k}(f)+s\gws_{k}(f,\btheta)
 \end{equation}
@@
 
-\subsection{Analysis method}
+\subsection{Method}
 
@@
 a likelihood function and prior pdfs on the signal parameters being estimated. In this
-analysis we use a nested sampling algorithm \cite{Skilling}\footnote{The code used was based
-on {\tt matlabmultinest} found at \url{https://github.com/mattpitkin/matlabmultinest}.}
-to provide samples from the marginal posteriors that we require.}
-We use a Gaussian likelihood function given by
+analysis we use the Markov chain Monte Carlo code {\tt emcee} \cite{2013PASP..125..306F} to sample 
+the posterior distribution.} We use a Gaussian likelihood function given by
 %
@@
 \begin{equation}
-  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm det}}
-  \sum_{i=i_{f_{\rm low}}}^{i_{f_{\rm high}}}\frac{\Re{\left\{ \tilde{d}^{*}_{k,i} \gws_{k,i}(\vec{\theta})
-  - \frac{1}{2}\left(\gws^*_{k,i}(\vec{\theta})\gws_{k,i}(\vec{\theta}) + \tilde{d}^*_{k,i}\tilde{d}_{k,i}\right) \right\}}}{S_{k,i}}\right]}
+\fl p(\bi{d} | \btheta, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm det}}
+  \sum_{i=i_{f_{\rm low}}}^{i_{f_{\rm high}}}\frac{\Re{\left\{ \tilde{d}^{*}_{k,i} 
+\gws_{k,i}(\btheta) - \frac{1}{2}\left(\gws^*_{k,i}(\btheta)\gws_{k,i}(\btheta) + 
+\tilde{d}^*_{k,i}\tilde{d}_{k,i}\right) \right\}}}{S_{k,i}}\right]}
 \end{equation}
 % 
-where $\Delta f$ is the frequency resolution of the data, $S_k$ is the
-$k^{\rm th}$ detector's one sided noise power spectral density, and the $i$ indices
-increment over frequency with $i_{f_{\rm low}}$ and $i_{f_{\rm high}}$ corresponding to
-the lower and upper range in frequencies used for our
-analsysis. \MP{Need to say the
-range of frequencies we integrate over, which I think was 40\,Hz to 1600\,Hz from looking
-at Liam's code.} The source parameters, and the unknown scale factor, have been subsumed
-into a vector $\vec{\theta}$.
+where $\Delta f$ is the frequency resolution of the data, $S_k$ is the $k^{\rm th}$ detector's one 
+sided noise power spectral density, and the $i$ indices increment over frequency with $i_{f_{\rm 
+low}}$ and $i_{f_{\rm high}}$ corresponding to the lower and upper range in frequencies used for 
+our analysis (which were 40\,Hz to 1600\,Hz respectively \MP{Should we actually go down to 20\,Hz -- 
+I imagine for BNS systems it doesn't make much difference?}). The source parameters, and the 
+unknown scale factor, have been subsumed into a vector $\vec{\theta}$. For a network of detectors, 
+the overall likelihood function becomes a product of the likelihoods for each detectors, i.e.\
+\begin{equation}
+  \label{eq:mult-likeli}
+  p(\bi{d}| \btheta, \curlH) = \prod \limits_D p(\bi{d}^D| \vec{\theta}, \curlH)
+\end{equation}
+where $D = \{H,L,V\}$.
 
@@
 
-\subsection{Simulations}
+\subsubsection{Prior ranges}\label{sec:priors}
 
-% mention the no-noise issue
-We note that the analysis performed here used zero-noise realisations in which
-no actual noise was added to our signals.  The likelihood functions still
-contain the correct \ac{PSD} and hence the accuracy with which the parameters
-are recovered (indicated by the widths of the posterior distributions) is
-representative of the statistical errors in realistic noise scenarios.
+The primary reason why we can use coincident observations with a GRB as a potential check of 
+detector calibration is that the GRB allows us to constrain our priors on various parameters that 
+would normally be highly correlated with the signal amplitude. The most important of these is that 
+the GRB can provide a very tight constraint on the source distance\footnote{A consequence of our 
+choice of BNS signals is that they will only be observable within the local Universe ($z \sim 
+0.1$), so the relation between the measured GRB redshift and distance is simple and is negligably 
+effected by the choice of cosmology.}. In this analysis we therefore assume that the source distance 
+is known (i.e.\ has a $\delta$-function prior). Another piece of information that we can use from 
+having a coincident GRB is that the system must be close to face-on, which allows us to place prior 
+constraints on the system inclination angle. In our analysis we take two scenarios: the first use a 
+prior on the beam opening angle (half-width of the beam) with a Gaussian distribution with zero mean 
+and $\sigma = 5^{\circ}$, whilst the second uses $\sigma = 20^{\circ}$. Beam opening angles for 
+short GRBs are relatively poorly constrained as they are based on only a few jet break detection, 
+however \cite{2014ApJ...780..118F} give a median opening angle value of $\sim 10^{\circ}$. If we 
+assume that the beam opening angle is a half-normal distribution, for which the median is given by 
+$\sqrt{2}\erf{}^{-1}(1/2)\sigma \sim 0.67\sigma$, then a median of $\sim 10^{\circ}$ corresponds to 
+$\sigma \sim 14.8^{\circ}$.
+
+As we have assumed that the GRB progenitor is a binary neutron star system the prior we use for the 
+system component masses is a Gaussian distribution for both components with means 
+of $1.35\,\textrm{M}_{\odot}$ and standard deviations of $0.13\,\textrm{M}_{\odot}$ 
+\MP{(reference)}. In reality, for all the systems we use in our study, the chirp mass, which 
+we see from equation~\ref{eq:signal} plays a role in the overall signal amplitude, is very well 
+constrained by the systems phase evolution, so our prior could be expanded with minimal effect on 
+the results. In the future neutron star-black hole systems could be included, although this add the 
+complication of having to include the black hole spin (by using BNS systems we have assumed that 
+the spins are small enough that they will introduce only a negligible effect).
+
+The final important piece of information that we can make use of from the GRB observation is the 
+sky position of the source, which we take to be known precisely.
+
+For the time of coalescence we use a uniform prior of spanning $\pm0.01$\,seconds around the 
+recorded time of the signal that would be returned by the detection pipeline. For the reference 
+phase a uniform prior between 0 and $2\pi$ is used, and for the polarisation angle a uniform prior 
+between 0 and $\pi/2$ is used. As the signals we use are generally close to being circularly 
+polarised (face-on) the phase and polarisation angle will be degenerate meaning the exact 
+combination of them will have very little effect of the result.
+
+Finally, we require a prior on the calibration scale factors for each detector. We make the 
+assumption that in general the calibration will be correct, so want to use a prior that is peaked 
+at unity. We also assume that the probability density that the calibration scale factor is either 
+larger or smaller than unity by an equivalent factor, e.g.\ $s=10$ or $s=0.1$, is the same. For this 
+we therefore use a log-normal distribution as our prior
+\begin{equation}
+ p(s|I) = \frac{1}{s\sigma\sqrt{2\pi}}\exp{\left( -\frac{(\ln{s} - \mu)^2}{2\sigma^2} \right)},
+\end{equation}
+where we chose $\sigma = 1$, which in turn means that having the mode at unity gives $\mu = 1$. 
+With these parameters the prior probability density at 0.1 and 10 is $\sim 1/14$ of that at 1. 
+Note that this prior does not provide produce a symmetric amount of probability about unity, but 
+given our simulation criterion described below in Section~\ref{sec:simulations} we find that the 
+likelihood generally overwhelms the prior. As $s$ is a scale factor we also have the further 
+constraint that it must be positive.
+
+\subsection{Simulations}\label{sec:simulations}
+
+To estimate how well we can constrain the calibration scaling each detector in an advanced 
+detector network we have performed simulations with injected BNS signals at a range of distances 
+(50\,Mpc to 450\,Mpc on 50\,Mpc increments). The simulations use the two Advanced LIGO detectors 
+(one at Hanford -- H1 -- and one at Livingston -- L1) and the Advanced Virgo detector (V1). We 
+assume all are expected to be operating at their design sensitivity.
 
 % Describe the ensemble of simulations used
+As described in Section~\ref{sec:priors} we had two different scenarios for the prior distribution 
+of source system inclination angles, so we have produced separate simulations for both cases. In 
+each case 1000 sources were created with distributions on the inclination, component masses,
+reference phase and polarisation angle drawn from their respective prior distributions. For the 
+time of coalescence the times were held fixed within the centre of the prior window. When drawing 
+the sky position points were drawn uniformly over the sky. However, with all these an additional 
+constraint was imposed based on the fact that we require a distributions sources that could 
+actually have been detected. We therefore include the requirement from \cite{2012PhRvD..85h2002A} 
+that their be at least two detectors in the network for which the SNR is $\gtrsim 5.5$ (as these 
+would be signals coincident with a GRB we do not include the often used further constraint that 
+the total coherent network SNR is greater than 12). This generally means that as we go to larger
+distances our population of sources that fulfil this criterion will not be uniformly distributed
+over the sky, and will generally be closer to being circularly polarised.
 
-For coalescing binary neutron stars of masses
-$m_{1,2} = 1.4\,M_{\odot}$ and  $D_{L}200\,Mpc$, the expected SNR is roughly 8.
-The SNR was determined from $D_{L} = [10,200]\,Mpc$  using a randomly chosen
-sky position which defines the antenna patterns and the GW signal. The range of
-luminosity distance used in this paper represented a high level of SNR up to
-the distance at which the first predicted detectable GW source; $D_{L} =
-200\,Mpc$. The detector limits for the advanced era is a luminosity distance of
-$\sim 450\,Mpc$. \sout{This was not initially tested as the SNR was low enough to
-cause problems with computational time. It was however tested in the random sky
-position simulations, see section~\ref{sec:rndsp}. The initial value for the
-calibration error used in the analysis was chosen randomly.}
+For each source we applied a fixed calibration scale factor of 1.2, 0.8 and 1.3 for H1, L1 and V1 
+respectively.
+
+These simulations allow us to assess how well on average we would be able to constrain the 
+calibration scale for a given CBC-GRB coincidence.
+
+\section{Results}\label{sec:results}
+
+For each of the simulated sources we have calculated the 68\% credible region on the scale factors 
+for each detector (for a Gaussian distribution this is equivalent to the region bounded by 
+$1\sigma$ intervals). For all the signals in each distance increment we have produced the 
+distribution of the half widths (i.e.\ $1\sigma$) of these scale factors. These are shown in 
+figures~\ref{fig:narrowdist} and \ref{fig:widedist} for the case with the inclination distribution 
+with a width of $5^{\circ}$ and $20^{\circ}$ respectively.
 
 \begin{figure}
-  \centering
-  \includegraphics[width = \textwidth]{DvsSNR-mult-det}
-  \caption{The distance and SNR relationship for all three detectors:
-LIGO Hanford, H1, LIGO Livingston, L1, and VIRGO.}
-  \label{fig:distsnr}
+ \begin{center}
+  \includegraphics[width=1.0\textwidth]{scale_factors_5degs_boxplot.png}
+ \end{center}
+ \caption{\label{fig:narrowdist} Add caption!}
 \end{figure}
 
-To begin with, parameter estimation was carried out on the simplest case: a
-single source of GWs using a single detector, LIGO Hanford, at different levels
-of SNR. The aim was to resolve the calibration scale factor with the highest
-degree of certainty.  Before testing the uncertainty resolution using a the
-full range of prior parameters, we fixed all values to known signal values and
-only searched over the scale factor. Results shown for with and without a full
-prior range are given in figures~\ref{fig:sd-empty-D}
-and~\ref{fig:sd-non-empty-D} respectively.
-
 \begin{figure}
-  \centering
-  \includegraphics[width = \textwidth]{SD_empty_D10_200}
-  \caption{\textit{Posterior distribution on the calibration scale factor for
-LIGO Hanford. Simulations carried out using an entirely known source of GWs.}}
-  \label{fig:sd-empty-D}
+ \begin{center}
+  \includegraphics[width=1.0\textwidth]{scale_factors_20degs_boxplot.png}
+ \end{center}
+ \caption{\label{fig:widedist} Add caption!}
 \end{figure}
-
-The figure above is a histogram of the calibration error, with each curve
-representing a different distance (SNR). The black line at $s = 0.8$ is the
-initial scale factor value used to multiply the unanalysed data set. This
-representation of the initial calibration scale factor will be continuous
-throughout the paper.  One would expect with lower SNR (greater $D_{L}$), the
-estimation of the parameter would become less certain. This is clearly shown in
-figure~\ref{fig:sd-empty-D} as the standard deviations of the curves increases
-with smaller SNR/larger luminosity distances. For this entirely known source of
-GWs, we would expect to find the uncertainty in $s$ to be approximately
-$1/$SNR; this is what we see in figure~\ref{fig:sd-empty-D}. The error in
-estimating $s$ for $D_{L} = 200 \,Mpc$ is roughly $\sim 20\%$. This is larger
-than the current method of obtaining the uncertainty in calibration. A full
-test of the new method proposed is carried out in section~\ref{sec:rndsp}.
-
-The next estimation of the calibration scale factor was performed with a full
-range of prior parameters, giving a more realistic system. As before, the
-simulations were carried out for differing levels of SNR and a histogram of the
-calibration error was made, seen in figure~\ref{fig:sd-non-empty-D}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width = \textwidth]{SD_non_empty_D10_200}
-  \caption{\textit{Posterior distribution on the calibration scale factor for
-LIGO Hanford. Simulations carried out with some parameters in the prior.}}
-  \label{fig:sd-non-empty-D}
-\end{figure}
-
-The error in the calibration scale factor, as before, increases with decreasing
-SNR (increasing $D_{L}$). The magnitude of this error however is larger than in
-the previous simulation. With a full prior range given in
-section~\ref{sec:pandl}, the standard deviation increases; at $D_{L} =
-200\,Mpc$ the error in $s$ is $\sim 24\%$. When comparing this new method of
-calibration scale factor estimation with the current, feedback loop system
-method used today, this method is not as accurate. However, it should again be
-noted that this method is an independent estimate of the calibration accuracy.
-
-\sout{A network of detectors are used throughout GW research. Being able to show the
-effect of the calibration error and recovering its uncertainty in the analysis
-would help in giving a detection more confidence. In this section, the
-simulations were carried out using multiple detectors; LIGO Hanford, LIGO
-Livingston, and Virgo. In reality, multiple detectors are used in order to
-determine the (rough) sky position of a signal. Due to their locations, if one
-site detects a signal before the others, then the position of the source can be
-roughly worked out. The aim was, as before, to estimate the calibration scale
-factor with the least amount of uncertainty for differing levels of SNR for a
-single source of GWs.}
-
-For a network of detectors, the overall likelihood function becomes a product
-of each likelihood for all detectors, i.e.
-
-\begin{equation}
-  \label{eq:mult-likeli}
-  p(\vec{d}| \vec{\theta}, \curlH) = \prod \limits_D p(\vec{d}^D| \vec{\theta},
-\curlH)
-\end{equation}
-
-where $D = \{H,L,V\}$. The initial calibration scale factors where again chosen
-at random ($s = [4.6, 2.5, 6.1]$ for LIGO Hanford, LIGO Livingston, and Virgo
-respectively). The simulations were the same as before, using the same
-parameter values. The GRB assumption was again tested in these simulations as a
-handful of parameters were relaxed and given some uncertainty. The results of
-these simulations are shown in figure~\ref{fig:mult-empty-D} and
-figure~\ref{fig:mult-non-D}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width = \textwidth]{MD_empty_D10_200}
-  \caption{Posterior distribution on the calibration scale factor for
-all three detectors. Simulations carried out using an entirely known source of
-GWs.}
-    \label{fig:mult-empty-D}
-\end{figure}
-
-Figure~\ref{fig:mult-empty-D} shows the histograms of the calibration scale
-factor for each detector. These simulations used an entirely known source of
-GWs, at a randomly chosen sky position (the same sky position as before). As
-the figure shows both LIGO sites, Hanford and Livingston, are able to resolve
-the error with a measurable degree of uncertainty. The standard deviation for
-$s$ at $D_{L}=200\,Mpc$ is $24\%$ and $41\%$ of the true scale factor for
-Hanford and Livingston respectively. The estimation for the calibration error
-using VIRGO however, was not as accurate. It can be seen that, for increasing
-luminosity distances, the resolution of $s$ becomes increasingly uncertain
-until the point at which the distribution becomes almost uniform; $D_{L}
-\approx 120\,Mpc$. There are a few reasons why this may be. Firstly, the sky
-position used may not be 'ideal' for generating the antenna patterns. The most
-probable cause of the estimation being very poor will lie in the detector
-itself. The interferometer itself is not as sensitive the two LIGO sites, hence
-the estimation of parameters using this site will not be as precise.
-
-\begin{figure}
-  \centering
-  \includegraphics[width = \textwidth]{MD_non_empty_D10_200}
-  \caption{Posterior distribution on the calibration scale factor for
-all three detectors. Simulations carried out with some prior parameters.}
-  \label{fig:mult-non-D}
-\end{figure}
-
-The histograms shown in figure~\ref{fig:mult-non-D} were made using the full
-range of prior parameters. The uncertainty in the calibration error is larger
-with a less known source of GWs. VIRGO, again, has the least certain
-calibration error resolution with the distribution becoming even more uniform.
-The standard deviation for both LIGO curves become larger with increasing
-luminosity distance as before. For $D_{L} = 200Mpc$,  the standard deviation is
-$\tilde 30\%$ of the true value. When comparing this to the current method,
-this again is not as accurate.
-
-Throughout the duration of this project the antenna patterns were defined by a
-randomly chosen sky position, as seen in equation~\ref{eq:gravsig}. This did
-not give a true representation of the estimation capable by algorithm as this
-position could be a bad one; making parameterisation more difficult and less
-precise. To then test this new method of calibration fully, the entire field of
-view of the detector must be analysed. The sky positions $\alpha$, $\delta$,
-were distributed as: \begin{table}[H] \centering \begin{tabular}{|c|c|} \hline
-Parameter & Distribution \\ \hline $\alpha$       & U($0$,$2\pi$) \\
-$\sin(\delta)$ & U($-1$,$1$)    \\ \hline \end{tabular} \caption{\textit{The
-distribution of sky position. U(a,b) indicates a uniform distribution between a
-and b.}} \label{tab:sp} \end{table}
-
-Figures~\ref{fig:snr-rndsp} and~\ref{fig:std-rndsp} were obtained using a
-luminosity distance $D_{L} = 200 Mpc$ as this is the expected luminosity
-distance predicted to host the first observable GW~\cite{peadvanced}. $100$
-different sky positions were tested using the LIGO Hanford site. The number of
-locations was limited by the length of time these simulations took to carry
-out. Increasing the number of locations would give a more accurate
-representation of the estimation ability; this can be carried out in future.
-Finally, a full range of prior parameters were also used in the simulation
-set-up.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{rand_sp_snr_D200}
-  \caption{Signal to noise ratio for every sky position analysed at
-$D_{L} = 200 \,Mpc$. The vertical black line represents the SNR for the sky
-position used throughout previous simulations}
-  \label{fig:snr-rndsp}
-\end{figure}
-
-Figure~\ref{fig:snr-rndsp} shows the histogram of SNRs for the entire field of
-view for the detector. Predictions state that at a luminosity distance of
-$D_{L}=200\,Mpc$, the SNR will be roughly 8. This is consistent with the
-results obtained. The vertical black line shown in figure~\ref{fig:snr-rndsp}
-represents the SNR used throughout the project using the randomly chosen
-$\alpha$ and $\delta$. One can see that this is a less-than-average SNR and may
-explain the resolution of the calibration error not being too precise. This
-then is a promising result and the full testing of this new calibration method
-can be carried out.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{rand_sp_std_D200}
-  \caption{Standard Deviation (as a $\%$ of the true scale factor
-value) for every sky position analysed at $D_{L}=\,200Mpc$. The vertical black
-line indicates the median of the distribution.}
-  \label{fig:std-rndsp}
-\end{figure}
-
-Figure~\ref{fig:std-rndsp} is the distribution of fractional standard
-deviations on s for each sky position analysed, given as a percentage. The
-vertical black line in this figure shows the \textit{median} of the
-distribution. The results show that the algorithm can resolve the calibration
-error in the data within the range of $\sim 7\%$ to $\sim 14\%$ of the true
-value of $s$, for $D_{L}=200\,Mpc$. Being able to recover the scale factor
-error to this accuracy range confirms that the new method of calibration method
-is very competative with current techniques. These results also confirm that
-the sky position used throughout the project was indeed a bad choice with an
-error estimation of $\sim 24\%$.
-
-Other simulations relating to the randomisation of $\alpha$ and $\delta$ were
-carried out in this project. Different luminosity distances were simulated to
-test different levels of SNR. The expected trend was that the standard
-deviation scales with SNR; with high SNR yielding a better resolution of the
-scale factor error. The results show this trend, given in table
-
-\begin{table}
-  \centering
-  \begin{tabular}{|c|c|c|}
-    \hline
-    Distance ($Mpc$) &$\langle$ SNR $\rangle$ & median(Standard
-Deviation)($\%$) \\
-    \hline
-    $20$ & 74 & 1.5 \\
-    $120$ & 12  & 7 \\
-    $450$ & 3 & 27\\
-    \hline
-  \end{tabular}
-  \caption{\textit{Random sky positions for different luminosity distances;
-$\langle$SNR$\rangle$ is the mean SNR for the distribution. }}
-  \label{tab:dist-randsp}
-\end{table}
-
-The last, and lengthiest, set of simulations carried out used a network of
-detectors. Analysis on a luminosity distance of $200 \,Mpc$ was the only one
-carried out in this paper. This was due to time taken to process these
-simulations. With a full prior range and $100$ random sky positions as before,
-the standard deviations and SNRs were obtained. Results are shown
-in~\ref{tab:mult-rndsp}.
-
-\begin{table}
-\centering
-\begin{tabular}[H]{|c|c|c|}
-\hline
-  Detector   & $\langle$SNR$\rangle$ & median(Standard Deviation)($\%$)\\
-\hline
-LIGO Hanford & 20& 3 \\
-LIGO Livingston & 20 & 10 \\
-Virgo & 5 & 30\\
-\hline
-\end{tabular}
-\caption{\textit{Results from randomising the sky position using a network of
-detectors with a full range of prior parameters at $D_{L} = 200\, Mpc$.}}
-\label{tab:mult-rndsp}
-\end{table}
-
-Table~\ref{tab:mult-rndsp} gives promising results for both LIGO sites. One can
-see that the error estimation for LIGO Hanford is very good, with half of the
-positions analysed having an uncertainty $<3\%$. One would expect the gap in
-estimation accuracy between the two LIGO sites to be smaller as they are built
-to the same sensitivity. This leads to the conclusion that not enough sky
-positions were analysed. Virgo again has a less accurate recovery of the
-calibration scale factor. As mentioned above, the larger uncertainty is
-expected to originate from the sensitivity of the detector itself.
 

### 2015-07-03 src/python/scale_factor_mcmc.py
scale_factor_mcmc:py - various changes:  - allow NS-BH systems to be used (these will have masses drawn from a Gaussian dist with mean of 5 solar masses and standard deviation of one (also used for the prior). The black hole will have a spin aligned with the system orbital angular mommentum, but with the neutron star spin assumed to be negligible (not entirely sure this works properly yet!).  - the scale factor prior has been narrowed to correspond to having 10 and 0.1 being 10 times less probable than have it at 1  - the scale factors will be drawn randomly from some distribution

@@ -65,17 +65,18 @@ Inputs are:
  - fmax - the upper bound on frequency (Hz)
  - dist - the source distance (in Mpc)
  - incl - the source inclination angle (rads)
+ - a1spin - dimensionless spin of the more massive component
 
 This function does not apply the antenna pattern to the output.
 """
-def fdwaveform(phiref, deltaF, m1, m2, fmin, fmax, dist, incl):
+def fdwaveform(phiref, deltaF, m1, m2, fmin, fmax, dist, incl, a1spin):
   ampO = 0   # 0 pN order in amplitude
   phaseO = 7 # 3.5 pN order in phase
   approx = lalsimulation.TaylorF2 # Taylor F2 approximant
   fref = fmin
  
   hptilde, hctilde = lalsimulation.SimInspiralChooseFDWaveform(phiref, deltaF,
-    m1*lal.MSUN_SI, m2*lal.MSUN_SI, 0., 0., 0., 0., 0., 0., fmin,
+    m1*lal.MSUN_SI, m2*lal.MSUN_SI, 0., 0., a1spin, 0., 0., 0., fmin,
     fmax, fref, dist*1e6*lal.PC_SI, incl, 0., 0., None, None, ampO, phaseO, approx)
   
   # return the frequency domain plus and cross waveforms
@@ -136,19 +137,23 @@ def frequency_noise_from_psd(psd, deltaF, seed = None):
 #  - deltaF - the frequency bin size
 #  - resps - the antenna respsonse functions
 #  - asds - the detector noise ASDs
-def lnprob(theta, y, dd, iotaprior, tccentre, dist, fmin, fmax, deltaF, resps, asds): 
-  lp = lnprior(theta, iotaprior, tccentre, len(resps))
+#  - nsbh - set whether using a NS-BH system (this has a different mass prior and includes a spin parameter)
+def lnprob(theta, y, dd, iotaprior, tccentre, dist, fmin, fmax, deltaF, resps, asds, nsbh): 
+  lp = lnprior(theta, iotaprior, tccentre, len(resps), nsbh)
   
   if not np.isfinite(lp):
     return -np.inf
   
-  return lp + lnlike(theta, y, dd, dist, fmin, fmax, deltaF, resps, asds)
+  return lp + lnlike(theta, y, dd, dist, fmin, fmax, deltaF, resps, asds, nsbh)
 
 
 # define the log prior function
[…]

### 2015-07-03 src/python/scale_factor_mcmc.py
scale_factor_mcmc.py: fix mass initialisation for NS-BH signals

@@ -173,7 +173,7 @@ def lnprior(theta, iotawidth, tccentre, ndets, nsbh):
     return -np.inf
   
   if nsbh: # parameter limits for NS-BH system
-    if 0.9 < m2 < 2. and 2.5 < m1 < 9. and  -1. < a1spin < 1.:
+    if 0.9 < m2 < 2. and m1 > 0. and -1. < a1spin < 1.:
       lp = 0.
     else:
       return -np.inf
@@ -618,7 +618,12 @@ if it does not fulfill the SNR criterion.")
     while not -0.5*np.pi < iotaini < 0.5*np.pi:
       iotaini = iotawidth*np.random.randn()
     tcini = -0.01 + 2.*0.01*np.random.rand() + t0 # time of coalescence
-    m1ini = 1.35 + 0.13*np.random.randn() # mass 1
+    if opts.nsbh:
+      m1ini = 5. + 1.*np.random.randn() # distribution for black hole (from http://journals.aps.org/prd/pdf/10.1103/PhysRevD.85.082002)
+      while m1ini < 2.5:
+        m1ini = 5. + 1.*np.random.randn() # make sure mass is over 2.5 solar masses
+    else:
+      m1ini = 1.35 + 0.13*np.random.randn() # mass 1
     m2ini = 1.35 + 0.13*np.random.randn() # mass 2
     while m1ini < m2ini: # m1 must be > m2
       m2ini = 1.35 + 0.13*np.random.randn() # mass 2

### 2015-07-03 src/python/results_script.py
results_script.py: output means and standard deviations to screen

@@ -104,7 +104,6 @@ for i, d in enumerate(dirs):
     relsf.append(vals)
     
   nprelsf = np.array(relsf)
-  print nprelsf.shape
   
   # divide by two to get the half widths and convert to percentage
   nprelsf[:,0] = 100.*nprelsf[:,0]/(2.*info['InjectionParameters']['scales'][0])
@@ -114,6 +113,9 @@ for i, d in enumerate(dirs):
   data.append(nprelsf[:,0])
   data.append(nprelsf[:,1])
   data.append(nprelsf[:,2])
+
+  print np.mean(nprelsf[:,0]), np.mean(nprelsf[:,1]), np.mean(nprelsf[:,2])
+  print np.median(nprelsf[:,0]), np.median(nprelsf[:,1]), np.median(nprelsf[:,2])
 
   # plot output
   #ci1 = credible_interval(nprelsf[:,0], 0.90)

### 2015-07-03 docs/cosmic_cal.tex
Minor updated to draft

@@
 \end{equation}
-where we chose $\sigma = 1$, which in turn means that having the mode at unity gives $\mu = 1$. 
-With these parameters the prior probability density at 0.1 and 10 is $\sim 1/14$ of that at 1. 
-Note that this prior does not provide produce a symmetric amount of probability about unity, but 
+where we chose $\sigma = 1.07$, which in turn means that having the mode at unity gives $\mu = 
+1.15$. With these parameters the prior probability density at 0.1 and 10 is one tenth of that at 
+1. Note that this prior does not provide produce a symmetric amount of probability about unity, but 
 given our simulation criterion described below in Section~\ref{sec:simulations} we find that the 
@@
 
-For each source we applied a fixed calibration scale factor of 1.2, 0.8 and 1.3 for H1, L1 and V1 
-respectively.
+For each simulation the calibration scale factor for each detector has been drawn from a Gaussian 
+distribution with a mean of unity and a standard deviation of 0.125 (equivalent to the detectors 
+having on average a $\pm 10\%$ amplitude calibration error).
 

### 2015-07-03 docs/cosmic_cal.tex
Changed from lower integration freq of 40Hz to 20 Hz

@@
 low}}$ and $i_{f_{\rm high}}$ corresponding to the lower and upper range in frequencies used for 
-our analysis (which were 40\,Hz to 1600\,Hz respectively \MP{Should we actually go down to 20\,Hz -- 
-I imagine for BNS systems it doesn't make much difference?}). The source parameters, and the 
+our analysis, which were 20\,Hz to 1600\,Hz respectively. The source parameters, and the 
 unknown scale factor, have been subsumed into a vector $\vec{\theta}$. For a network of detectors, 

### 2015-07-03 src/python/scale_factor_mcmc.py
scale_factor_mcmc.py: change default fmin to 30 Hz

@@ -365,7 +365,7 @@ a uniform distribution between 0 and 2pi.")
 a uniform distribution between 0 and pi/2.")
 
   parser.add_option("-f", "--fmin", dest="fmin", type="float",
-                    help="Lower frequency bound (Hz) [default: %default]", default=20.)
+                    help="Lower frequency bound (Hz) [default: %default]", default=30.)
   
   parser.add_option("-m", "--fmax", dest="fmax", type="float",
                     help="Upper frequency bound (Hz) [default: %default]", default=1600.)

### 2015-07-03 docs/cosmic_cal.tex
Minor paper changes

@@
 % Describe the signal model (Taylor F2 etc...)
-We use the post-Newtonian Taylor F2 waveform approximation for modelling the
-signal whereby the \ac{GW} phase for 2 point-particles of mass $m_{1}$ and
-$m_{2}$ is given by
-%
-\begin{equation}
-\label{eq:gwphase}
-  \Psi(f) = 2\pi f t_\mathrm{c} - \phi_{\mathrm{c}} - \frac{\pi}{4} +
-\frac{3}{128\eta\nu^{5}}\sum\limits_{k=0}^{7}\alpha_{k}\nu^{k}.
-\end{equation}
-%
-Here $\eta=m_{1}m_{2}/M^{2}$ is the symmetric mass-ratio with $M=m_{1}+m_{2}$
-the total mass.  The post-Newtonian expansion parameter is $\nu=(\pi M
-f)^{1/3}$ and the coefficients are $\alpha$.
+We use the TaylorF2 waveform approximation (see e.g.\ \cite{2009PhRvD..80h4043B} and references 
+therein) with a 3.5 post-Newtonian expansion in phase for modelling our signal (in both simulations 
+and signal recovery).  For \ac{BNS} systems we use non-spinning waveforms under the assumption that 
+spins will be negligible. For \ac{NS-BH} system we use a spinning waveform, but in which only the 
+black hole has non-negligible spin and that its rotational angular momentum is aligned with the 
+systems orbital angular momentum.
 
-The frequency-domain polarisation amplitudes are then defined as
+For a non-spinning system the general form of the frequency-domain polarisation amplitudes are
 %
 \begin{eqnarray}\label{eq:signal}
-  \gws_{+}(f) &=& \frac{1+\cos^{2}\iota}{2D_{L}}
-\frac{\mathcal{M}^{5/6}}{\pi^{2/3}}
-\sqrt{\frac{5}{24}}f^{-7/6}e^{-i\Psi(f)} \nonumber \\
-  \gws_{\times}(f) &=&
- i\frac{\cos^{2}\iota}{D_{L}}\frac{\mathcal{M}^{5/6}}{\pi^{2/3}}
-\sqrt{\frac{5}{24}}f^{-7/6}e^{-i\Psi(f)}
+  \gws_{+}(f) &\propto \frac{1+\cos^{2}\iota}{D_{L}}
+\mathcal{M}^{5/6}f^{-7/6}e^{-i\Psi(f, \nu, t_\mathrm{c}, \phi_{\mathrm{c}})} \nonumber \\
+  \gws_{\times}(f) &\propto
+ \frac{\cos^{2}\iota}{D_{L}}\mathcal{M}^{5/6} f^{-7/6}e^{-i\Psi(f, \nu, t_\mathrm{c}, 
+\phi_{\mathrm{c}})}
 \end{eqnarray}
 %
-where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=M\eta^{3/5}$,
-and $\iota$ is the inclination angle.  The \ac{GW} strain measured
-at the $k$'th detector is then given by
+where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=M\eta^{3/5}$, with 
+$\eta=m_{1}m_{2}/M^{2}$ and $M=m_{1}+m_{2}$, $\iota$ is the inclination angle, and $\nu=(\pi 
+Mf)^{1/3}$.  The \ac{GW} strain measured at the $k^{\textrm{th}}$ detector is then given by
 %
@@
 low}}$ and $i_{f_{\rm high}}$ corresponding to the lower and upper range in frequencies used for 
-our analysis, which were 20\,Hz to 1600\,Hz respectively. The source parameters, and the 
-unknown scale factor, have been subsumed into a vector $\vec{\theta}$. For a network of detectors, 
+our analysis, which were 30\,Hz to 1600\,Hz respectively. The source parameters, and the 
+unknown scale factor, have been subsumed into a vector $\btheta$. For a network of detectors, 
 the overall likelihood function becomes a product of the likelihoods for each detectors, i.e.\

### 2015-07-06 src/python/results_script.py
Results_script.py: added line showing the percentage of the signal parameter prior volume that simulations are drawn from given the SNR contraint

@@ -70,7 +70,7 @@ pl.rc('text', usetex=True)
 pl.rc('font', family='serif')
 pl.rc('font', size=14)
 
-dists = [50, 100, 150, 200, 250, 300, 350, 400, 450]
+dists = [50, 100, 150, 200, 250, 300, 350, 400, 450, 500]
 
 prefix = indir
 
@@ -83,11 +83,13 @@ fnpre = 'info_'
 fig, ax = pl.subplots(figsize=(8,5))
 
 data = []
+meanvfrac = []
 
 for i, d in enumerate(dirs):
   print d
 
   relsf = []
+  vfrac = []
   
   files = [os.path.join(d, f) for f in os.listdir(d) if os.path.isfile(os.path.join(d, f)) and fnpre in f]
   
@@ -102,7 +104,10 @@ for i, d in enumerate(dirs):
       vals.append(info['Results']['Scale68%CredibleInterval'][k][1]-info['Results']['Scale68%CredibleInterval'][k][0])
 
     relsf.append(vals)
-    
+
+    vfrac.append(1./info['Attempts'])
+  
+  meanvfrac.append(np.mean(np.array(vfrac)))
   nprelsf = np.array(relsf)
   
   # divide by two to get the half widths and convert to percentage
@@ -191,12 +196,17 @@ for i in range(numBoxes):
 ax.set_xticklabels(dists)
 ax.set_xticks(dists)
 
-ax.set_xlim((0, 500))
+ax.set_xlim((0, 550))
 ax.set_ylim((0, 100))
 
 pl.legend(hs, ['H1', 'L1', 'V1'], loc='best')
 pl.xlabel('distance (Mpc)')
 pl.ylabel('\% calibration scaling error (1$\sigma$ equivalent)')
+
+ax2 = ax.twinx()
+ax2.plot(dists, 100.*np.array(meanvfrac), 'ko--', markerfacecolor='None', markeredgecolor='k')
+ax2.set_ylabel('\% of total signals observed')
+ax2.set_xlim((0, 550))
 
 try:
   fig.savefig(outfile)

### 2015-07-07 src/python/scale_factor_mcmc.py
scale_factor_mcmc.py: changed to use argparse rather than optparse

@@ -7,7 +7,7 @@ import numpy as np
 import copy
 import sys
 import os
-from optparse import OptionParser
+import argparse
 import json
 
 from matplotlib import pyplot as pl
@@ -259,7 +259,7 @@ def lnlike(theta, y, dd, dist, fmin, fmax, deltaF, resps, asds, nsbh):
 
 # a function to get the credible intervals using a greedy binning method
 def credible_interval(dsamples, ci):
-    n, binedges = np.histogram(dsamples, bins=100)
+    n, binedges = np.histogram(dsamples, bins=250)
     dbins = binedges[1]-binedges[0] # width of a histogram bin
     bins = binedges[0:-1]+dbins/2. # centres of bins
 
@@ -271,6 +271,7 @@ def credible_interval(dsamples, ci):
         frac += float(n[i])/float(len(dsamples))
         j = j+1
         if frac >= ci:
+            #print frac
             break
 
     return (np.min(bins[histIndices[:j]]), np.max(bins[histIndices[:j]]))
@@ -287,196 +288,205 @@ def McQ2Masses(mC, q):
 
 if __name__=='__main__':
   
-  usage = "Usage: %prog [options]"
+  parser = argparse.ArgumentParser( formatter_class=argparse.ArgumentDefaultsHelpFormatter )
 
-  parser = OptionParser( usage = usage )
+  parser.add_argument("-o", "--outpath", dest="outpath", help="The path for \
+the analysis output (a sub-directory based on the pulsar name will be created here that \
+contains the pulsar information)", metavar="DIR")
 
-  parser.add_option("-o", "--outpath", dest="outpath", help="The path for "
-                    "the analysis output (a sub-directory based on the pulsar "
-                    "name will be created here that contains the pulsar "
-                    "information)", metavar="DIR")
-
-  parser.add_option("-O", "--output-samples", dest="outsamps", help="If this flag "
+  parser.add_argument("-O", "--output-samples", dest="outsamps", help="If this flag "
                     "is set then the samples will be output.", default=False, action="store_true",)
 
-  parser.add_option("-g", "--det", dest="dets",
-                    help="An interferometer to be used (multiple interferometers can be set \
-with multiple uses of this flag, e.g. \"-g H1 -g L1 -g V1\") [default: H1]", 
-                    action="append")
+  parser.add_argument("-g", "--det", dest="dets", type=str,
+                      help="An interferometer to be used (multiple interferometers can be set \
[…]

### 2015-07-07 docs/cosmic_cal.tex
minor changes to the paper

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%    INSTITUTE OF PHYSICS PUBLISHING                                   %
-%                                                                      %
-%   `Preparing an article for publication in an Institute of Physics   %
-%    Publishing journal using LaTeX'                                   %
-%                                                                      %
-%    LaTeX source code `ioplau2e.tex' used to generate `author         %
-%    guidelines', the documentation explaining and demonstrating use   %
-%    of the Institute of Physics Publishing LaTeX preprint files       %
-%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
-%                                                                      %
-%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
-%                                                                      %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-% First we have a character check
-%
-% ! exclamation mark    " double quote  
-% # hash                ` opening quote (grave)
-% & ampersand           ' closing quote (acute)
-% $ dollar              % percent       
-% ( open parenthesis    ) close paren.  
-% - hyphen              = equals sign
-% | vertical bar        ~ tilde         
-% @ at sign             _ underscore
-% { open curly brace    } close curly   
-% [ open square         ] close square bracket
-% + plus sign           ; semi-colon    
-% * asterisk            : colon
-% < open angle bracket  > close angle   
-% , comma               . full stop
-% ? question mark       / forward slash 
-% \ backslash           ^ circumflex
-%
-% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
-% abcdefghijklmnopqrstuvwxyz 
-% 1234567890
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
 \documentclass[10pt]{iopart}
@@
 
-%\documentclass[12pt]{iopart}
-%\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
-%Uncomment next line if AMS fonts required
-%\usepackage{iopams}  
 %\usepackage {amsmath}
@@
 \acrodef{BBH}[BBH]{binary black hole}
-\acrodef{NSBH}[NSBH]{neutron star black hole}
+\acrodef{BNS}[BNS]{binary neutron star}
+\acrodef{NSBH}[NSBH]{neutron star-black hole}
 \acrodef{SNR}[SNR]{signal-to-noise ratio}
@@
 % Describe our model for the calibration
-\MP{The gravitational wave strain is measured through the differential arm length, $\Delta L$, changes
-of the interferometer via $h = \Delta L(f) / L$, where $L$ is the full arm length.
+\MP{The gravitational wave strain is measured through the differential arm length, $\Delta L$, 
+changes of the interferometer via $h = \Delta L(f) / L$, where $L$ is the full arm length.
 Calibration is required to relate the actual measured interferometer error signal output
@@
 \begin{equation}
-h(f) = s h_{\rm m}(f) = s \frac{R(f,t) e(f)}{L},
+sh(f) = h_{\rm m}(f) = \frac{R(f,t) e(f)}{L},
 \end{equation}
-where $h(f)$ is the true strain and $h_{\rm m}(f)$ is the measured strain.
-In this analysis we will simplify the situation by assuming that $s$ is a constant with
-respect to frequency, but in future studies that assumption could be dropped and $s$
-could take some functional form, or piece-wise fit, with respect to $f$.}
+where $h(f)$ is the true strain and $h_{\rm m}(f)$ is the measured strain. With this definition it
+means that if $s > 1$ then a signal would appear to have a larger amplitude (e.g.\ be closer) than 
+expected, whereas if $s < 1$ would appear to have a lower amplitude than expected. In this analysis 
+we will simplify the situation by assuming that $s$ is a constant with respect to frequency, but in 
+future studies that assumption could be dropped and $s$ could take some functional form, or 
+piece-wise fit, with respect to $f$.}
 
@@
 
-% Describe what we are going to do in this section
-Let us consider a single \ac{GW} detection of the inspiral stage of a \ac{BNS}
-merger and assume that this signal has been associated with a potential
-\ac{sGRB} detection. We therefore also assume that the inclination angle
-of the binary is consistent with a given \ac{sGRB} beaming angle.  Finally we
-assume that a host galaxy has been identified for the source and that the sky
-position is therefore known to high precision.
-
-\MP{We perform 100 simulations at each of a range of distances from 10\,Mpc out
-to 200\,Mpc. For the simulations at each distance range we have allowed the
-source's sky position to be randomly drawn from a uniform distribution over
-the sky, the polarisation angle $\psi$ to drawn from a uniform distribution in the
-range $[0, \pi/2]$, the coalescence phase to be drawn from a uniform distribution
-in the range $[0, 2\pi]$, and the inclination angle drawn from a Gaussian with range XX?.
-The binary system masses have been fixed as $1.4\,{\rm M}_\odot$ (based on the
-very good assumption that whatever masses are chosen in these cases they will be
-recovered to high accuracy) and the time of coalescence is fixed.}
-
 % Describe the signal model (Taylor F2 etc...)
@@
 and signal recovery).  For \ac{BNS} systems we use non-spinning waveforms under the assumption that 
-spins will be negligible. For \ac{NS-BH} system we use a spinning waveform, but in which only the 
+spins will be negligible. For \ac{NSBH} system we use a spinning waveform, but in which only the 
 black hole has non-negligible spin and that its rotational angular momentum is aligned with the 
@@
 \MP{Our aim is to estimate the marginal posterior probability distribution function (pdf)
-for the unknown calibration scale factor. We use Bayes' theorem for which we need to define
+for the unknown calibration scale factors. We use Bayes' theorem for which we need to define
 a likelihood function and prior pdfs on the signal parameters being estimated. In this
@@
 low}}$ and $i_{f_{\rm high}}$ corresponding to the lower and upper range in frequencies used for 
-our analysis, which were 30\,Hz to 1600\,Hz respectively. The source parameters, and the 
-unknown scale factor, have been subsumed into a vector $\btheta$. For a network of detectors, 
-the overall likelihood function becomes a product of the likelihoods for each detectors, i.e.\
+our analysis, which were 20\,Hz to 400\,Hz respectively. This frequency range was chosen as the 
+vast majority of the \ac{SNR} can be found within this range, so going to lower or higher 
+frequencies provides very little additional information for our analysis. The source parameters, 
+and the unknown scale factor, have been subsumed into a vector $\btheta$. For a network of 
+detectors, the overall likelihood function becomes a product of the likelihoods for each detectors, 
+i.e.\
 \begin{equation}
@@
 
-% describe what parameters we will be considering and their priors
-We assume uniform prior distributions on the unknown nuisance parameters
-$\phi_{\mathrm{c}},\psi$ on the ranges $[0, 2\pi]$ and $[0, \pi/2]$ respectively.
-The time of coalescence is
-restricted to a prior range of $\pm 10$ msec around the true value.  Given our
-assumption of an \ac{sGRB} counterpart and host galaxy, we treat the sky
-position as exactly known. We also consider the cosine of the inclination angle to be drawn
-from a zero-mean Gaussian with standard deviation $0.1$ consistent with an
-uncertainty on \ac{sGRB} opening of XXX~\cm{What was exactly done here?}.  The
-main parameter of interest, the calibration scale factor $s$ is assumed to have
-a uniform prior distribution on the range $(0,2)$.
-
 \subsubsection{Prior ranges}\label{sec:priors}
@@
 would normally be highly correlated with the signal amplitude. The most important of these is that 
-the GRB can provide a very tight constraint on the source distance\footnote{A consequence of our 
-choice of BNS signals is that they will only be observable within the local Universe ($z \sim 
-0.1$), so the relation between the measured GRB redshift and distance is simple and is negligably 
-effected by the choice of cosmology.}. In this analysis we therefore assume that the source distance 
-is known (i.e.\ has a $\delta$-function prior). Another piece of information that we can use from 
-having a coincident GRB is that the system must be close to face-on, which allows us to place prior 
-constraints on the system inclination angle. In our analysis we take two scenarios: the first use a 
+the GRB can provide a very tight constraint on the source distance\footnote{For \ac{BNS} and 
+\ac{NSBH} systems they will be observable only in the relatively nearby Universe ($z \sim 
+0.1$ and $z\sim 0.2$ for \ac{BNS} and \ac{NSBH} systems respectively), so the relation between the 
+measured GRB redshift and distance is simple and our luminosity distance will be negligably 
+effected by the choice of cosmology \MP{Check this?}.}. In this analysis we therefore assume that 
+the source distance is known (i.e.\ has a $\delta$-function prior). Another piece of information 
+that we can use from having a coincident GRB is that the system must be close to face-on, which 
+allows us to place prior constraints on the system inclination angle. In our analysis we take two 
+scenarios: the first use a 
 prior on the beam opening angle (half-width of the beam) with a Gaussian distribution with zero mean 
@@
 
+For the spinning \ac{NSBH} systems we use a uniform prior on the dimensionless spin magnitude of 
+the black hole between $-1$ and 1.
+
 Finally, we require a prior on the calibration scale factors for each detector. We make the 
@@
 To estimate how well we can constrain the calibration scaling each detector in an advanced 
-detector network we have performed simulations with injected BNS signals at a range of distances 
-(50\,Mpc to 450\,Mpc on 50\,Mpc increments). The simulations use the two Advanced LIGO detectors 
-(one at Hanford -- H1 -- and one at Livingston -- L1) and the Advanced Virgo detector (V1). We 
-assume all are expected to be operating at their design sensitivity.
+detector network we have performed simulations with injected \ac{BNS} and \ac{NSBH} signals at a 
+range of distances (50\,Mpc to 500\,Mpc with 50\,Mpc increments for the \ac{BNS} signals and 
+100\,Mpc to 1000\,Mpc with 100\,Mpc increments for the \ac{NSBH} signals). The simulations use the 
+two Advanced LIGO detectors (one at Hanford -- H1 -- and one at Livingston -- L1) and the Advanced 
+Virgo detector (V1). The simulations are all performed in the freqeuncy domain and span a frequency 
+range from 20\,Hz to 400\,Hz. We assume all detectors are expected to be operating at their design
+sensitivity.
 

### 2015-07-08 src/python/results_script.py
results_script.py: changed detectable source distribution calculation

@@ -83,16 +83,17 @@ fnpre = 'info_'
 fig, ax = pl.subplots(figsize=(8,5))
 
 data = []
-meanvfrac = []
+rates = []
 
 for i, d in enumerate(dirs):
   print d
 
   relsf = []
-  vfrac = []
   
   files = [os.path.join(d, f) for f in os.listdir(d) if os.path.isfile(os.path.join(d, f)) and fnpre in f]
   
+  totaltime = 0.
+
   # go through files and extract the relative standard devaition on the scale factors for each detector
   for f in files:
     fo = open(f, 'r')
@@ -105,9 +106,9 @@ for i, d in enumerate(dirs):
 
     relsf.append(vals)
 
-    vfrac.append(1./info['Attempts'])
+    totaltime += info['Attempts']
   
-  meanvfrac.append(np.mean(np.array(vfrac)))
+  rates.append(100.*len(files)/totaltime)
   nprelsf = np.array(relsf)
   
   # divide by two to get the half widths and convert to percentage
@@ -204,8 +205,8 @@ pl.xlabel('distance (Mpc)')
 pl.ylabel('\% calibration scaling error (1$\sigma$ equivalent)')
 
 ax2 = ax.twinx()
-ax2.plot(dists, 100.*np.array(meanvfrac), 'ko--', markerfacecolor='None', markeredgecolor='k')
-ax2.set_ylabel('\% of total signals observed')
+ax2.plot(dists, rates, 'mo--', markerfacecolor='None', markeredgecolor='m')
+ax2.set_ylabel('\% of source distribution detectable')
 ax2.set_xlim((0, 550))
 
 try:

### 2015-07-08 src/python/scale_factor_mcmc.py
simulation_check: made consistent with scale_factor_mcmc in terms of source distribution and added NS-BH

@@ -596,7 +596,7 @@ if it does not fulfill the SNR criterion.")
       htmp[~np.isfinite(htmp)] = 0.
       snr = np.sqrt(4.*deltaF*np.vdot(htmp, htmp).real)
       SNRs.append(snr)
-      print >> sys.stderr, "%s: SNR = %.2f" % (dets[i], snr) 
+      print >> sys.stderr, "%s: SNR = %.2f" % (dets[i], snr)
 
       # create additive noise
       if addnoise:

### 2015-07-08 src/python/simulation_check.py

@@ -63,14 +63,14 @@ Inputs are:
 
 This function does not apply the antenna pattern to the output.
 """
-def fdwaveform(phiref, deltaF, m1, m2, fmin, fmax, dist, incl):
+def fdwaveform(phiref, deltaF, m1, m2, fmin, fmax, dist, incl, a1spin):
   ampO = 0   # 0 pN order in amplitude
   phaseO = 7 # 3.5 pN order in phase
   approx = lalsimulation.TaylorF2 # Taylor F2 approximant
   fref = fmin
  
   hptilde, hctilde = lalsimulation.SimInspiralChooseFDWaveform(phiref, deltaF,
-    m1*lal.MSUN_SI, m2*lal.MSUN_SI, 0., 0., 0., 0., 0., 0., fmin,
+    m1*lal.MSUN_SI, m2*lal.MSUN_SI, 0., 0., a1spin, 0., 0., 0., fmin,
     fmax, fref, dist*1e6*lal.PC_SI, incl, 0., 0., None, None, ampO, phaseO, approx)
   
   # return the frequency domain plus and cross waveforms
@@ -172,7 +172,17 @@ a uniform distribution between 0 and pi/2.")
   parser.add_option("-k", "--psd-noise", dest="psdnoise", default=0, type="int",
                      help="If this flag is non-zero the PSD is estimated as the average of \
 the given number of noisy PSD estimates.")
+  
+  parser.add_option("-s", "--a1spin", dest="a1spin", type="float",
+                    help="If using a NS-BH system this will set the aligned spin magnitude of \
+the black hole (a number between -1 and 1).")
+  
+  parser.add_option("-H", "--nsbh", dest="nsbh", default=False, action="store_true",
+                    help="If this flag is set it will use a neutron star-black hole system \
+for drawing masses and setting spin (a single spin for the black hole will be used).")
  
+  parser.add_option("-w", "--iotawidth", dest="iotawidth", type="float",
+                      help="Width of iota simulation, and prior, distribution (degs)")
 
   # parse input options
   (opts, args) = parser.parse_args()
@@ -205,10 +215,15 @@ the given number of noisy PSD estimates.")
   else:
[…]

### 2015-07-10 src/python/results_script.py
results_script.py: fix bug that was causing relative errors to be wrong

@@ -102,7 +102,8 @@ for i, d in enumerate(dirs):
 
     vals = []
     for k in range(len(info['InjectionParameters']['scales'])):
-      vals.append(info['Results']['Scale68%CredibleInterval'][k][1]-info['Results']['Scale68%CredibleInterval'][k][0])
+      # divide by two to get the half widths and convert to percentage
+      vals.append(100.*(info['Results']['Scale68%CredibleInterval'][k][1]-info['Results']['Scale68%CredibleInterval'][k][0])/(2.*info['InjectionParameters']['scales'][k]))
 
     relsf.append(vals)
 
@@ -110,11 +111,6 @@ for i, d in enumerate(dirs):
   
   rates.append(100.*len(files)/totaltime)
   nprelsf = np.array(relsf)
-  
-  # divide by two to get the half widths and convert to percentage
-  nprelsf[:,0] = 100.*nprelsf[:,0]/(2.*info['InjectionParameters']['scales'][0])
-  nprelsf[:,1] = 100.*nprelsf[:,1]/(2.*info['InjectionParameters']['scales'][1]) 
-  nprelsf[:,2] = 100.*nprelsf[:,2]/(2.*info['InjectionParameters']['scales'][2])
 
   data.append(nprelsf[:,0])
   data.append(nprelsf[:,1])

### 2015-07-10 src/python/scale_factor_mcmc.py
scale_factor_mcmc.py: corrected log normal values

@@ -651,8 +651,8 @@ if it does not fulfill the SNR criterion.")
       a1spinini = -1. + 2.*np.random.rand()
 
     #sfs = np.log(0.1 + (2.-0.5)*np.random.rand(len(scales))) # log scale factors
-    lognormalsigma = 1.
-    lognormalmu = 1.
+    lognormalsigma = 1.0729830131446736
+    lognormalmu = lognormalsigma**2
     # scale factors from a log normal distrbution
     #sfs = np.log(np.random.lognormal(lognormalmu, lognormalsigma, len(scales)))
     sfs = np.random.lognormal(lognormalmu, lognormalsigma, len(scales))

### 2015-07-16 docs/cosmic_cal.tex
Some changes to the prior description

@@
 the source distance is known (i.e.\ has a $\delta$-function prior). Another piece of information 
-that we can use from having a coincident GRB is that the system must be close to face-on, which 
-allows us to place prior constraints on the system inclination angle. In our analysis we take two 
-scenarios: the first use a 
-prior on the beam opening angle (half-width of the beam) with a Gaussian distribution with zero mean 
-and $\sigma = 5^{\circ}$, whilst the second uses $\sigma = 20^{\circ}$. Beam opening angles for 
-short GRBs are relatively poorly constrained as they are based on only a few jet break detection, 
-however \cite{2014ApJ...780..118F} give a median opening angle value of $\sim 10^{\circ}$. If we 
-assume that the beam opening angle is a half-normal distribution, for which the median is given by 
-$\sqrt{2}\erf{}^{-1}(1/2)\sigma \sim 0.67\sigma$, then a median of $\sim 10^{\circ}$ corresponds to 
-$\sigma \sim 14.8^{\circ}$.
+that we can use from having a coincident GRB is that the system is likely to be relatively close to 
+face-on (an therefore nearly circularly polarised), which allows us to place prior constraints on 
+the system inclination angle. Beam opening angles for short GRBs are relatively poorly constrained 
+as they are based on only a few jet break detections, however \cite{2014ApJ...780..118F} give a 
+median opening angle value of $\sim 10^{\circ}$. If we assume that the beam opening angle is a 
+half-normal distribution, for which the median is given by $\sqrt{2}\erf{}^{-1}(1/2)\sigma \sim 
+0.67\sigma$, then a median of $\sim 10^{\circ}$ corresponds to $\sigma \sim 14.8^{\circ}$. We 
+therefore place a Gaussian prior on the inclination angle with zero mean and $\sigma=14.8^{\circ}$.
 
-As we have assumed that the GRB progenitor is a binary neutron star system the prior we use for the 
-system component masses is a Gaussian distribution for both components with means 
-of $1.35\,\textrm{M}_{\odot}$ and standard deviations of $0.13\,\textrm{M}_{\odot}$ 
-\MP{(reference)}. In reality, for all the systems we use in our study, the chirp mass, which 
-we see from equation~\ref{eq:signal} plays a role in the overall signal amplitude, is very well 
-constrained by the systems phase evolution, so our prior could be expanded with minimal effect on 
-the results. In the future neutron star-black hole systems could be included, although this add the 
-complication of having to include the black hole spin (by using BNS systems we have assumed that 
-the spins are small enough that they will introduce only a negligible effect).
+Our prior on the system component masses is dependent on whether we are considering the source 
+being a \ac{BNS} system or a \ac{NSBH}. For the former case the prior we use is a Gaussian 
+distribution for both components with means of $1.35\,\textrm{M}_{\odot}$ and standard deviations 
+of $0.13\,\textrm{M}_{\odot}$ \MP{(reference)}. In reality, for all the systems we use in our 
+study, the chirp mass, which we see from equation~\ref{eq:signal} plays a role in the overall 
+signal amplitude, is very well constrained by the systems phase evolution, so our prior could be 
+expanded with minimal effect on the results. In the case of \ac{NSBH} systems we use the same 
+prior as above for the neutron star, but for the black hole we use a prior based on the 
+canonical masss distribution used for the rate results in \cite{2012PhRvD..85h2002A} with a mean of 
+$5\,\textrm{M}_{\odot}$ and a standard deviation of $1\,\textrm{M}_{\odot}$. For the \ac{NSBH}
+systems we also require a prior on the black hole spin (for \ac{BNS} systems we have assumed that 
+the spins are small enough that they will be negligible). We use a uniform prior on the 
+normalised aligned spin magnitude between $-1$ and $1$.
 

### 2015-07-16 docs/cosmic_cal.tex
Written some more results

@@
 \acrodef{EM}[EM]{electro-magnetic}
+\acrodef{aLIGO}[aLIGO]{Advanced LIGO}
+\acrodef{AdV}[AdV]{Advanced Virgo}
 
@@
 overview of the calibration method currently used, and is an extensive study in
-itself, see \cite{Vitale:2012} and references therein. For advanced LIGO and Virgo,
+itself, see \cite{Vitale:2012} and references therein. For \ac{aLIGO} and \ac{AdV},
 at the frequency range used in this project, the error in the calibration using
@@
 $\eta=m_{1}m_{2}/M^{2}$ and $M=m_{1}+m_{2}$, $\iota$ is the inclination angle, and $\nu=(\pi 
-Mf)^{1/3}$.  The \ac{GW} strain measured at the $k^{\textrm{th}}$ detector is then given by
+Mf)^{1/3}$.  The \ac{GW} strain measured at the $k^{\rm{th}}$ detector is then given by
 %
@@
 
-To estimate how well we can constrain the calibration scaling each detector in an advanced 
+To estimate how well we can constrain the calibration scaling for each detector in an advanced 
 detector network we have performed simulations with injected \ac{BNS} and \ac{NSBH} signals at a 
@@
 range from 20\,Hz to 400\,Hz. We assume all detectors are expected to be operating at their design
-sensitivity.
+sensitivity (as given by Figure 1 of \cite{2013arXiv1304.0670L}). As such we add noise to each 
+simulation based on the sensitivity curve, but scaled with the associated calibration scale factor.
 
 % Describe the ensemble of simulations used
-As described in Section~\ref{sec:priors} we had two different scenarios for the prior distribution 
-of source system inclination angles, so we have produced separate simulations for both cases. In 
-each case 1000 sources were created with distributions on the inclination, component masses,
-reference phase and polarisation angle drawn from their respective prior distributions. For the 
-time of coalescence the times were held fixed within the centre of the prior window. When drawing 
-the sky position points were drawn uniformly over the sky. However, with all these an additional 
-constraint was imposed based on the fact that we require a distributions sources that could 
-actually have been detected. We therefore include the requirement from \cite{2012PhRvD..85h2002A} 
-that their be at least two detectors in the network for which the SNR is $\gtrsim 5.5$ (as these 
-would be signals coincident with a GRB we do not include the often used further constraint that 
-the total coherent network SNR is greater than 12). This generally means that as we go to larger
-distances our population of sources that fulfil this criterion will not be uniformly distributed
-over the sky, and will generally be closer to being circularly polarised.
-
-For each simulation the calibration scale factor for each detector has been drawn from a Gaussian 
-distribution with a mean of unity and a standard deviation of 0.125 (equivalent to the detectors 
-having on average a $\pm 10\%$ amplitude calibration error).
+Separately for the \ac{BNS} and \ac{NSBH} systems we have performed 2000 simulations at each 
+distance value. For proposing injections for each simulation the source parameters were randomly 
+drawn from the prior distributions given in Section~\ref{sec:priors}, with the sky position being 
+drawn randomly from a uniform distribution, and with the scaling factors for each detector drawn 
+from a Gaussian distribution with a mean of one and standard deviation of $0.125$ (equivalent to a 
+mean fractional offset of 10\%). However, in accepting a proposed injection as one to be analysed 
+we introduced a criteria that the signal be ``detectable'', based on it having an SNR of $\geq 5.5$ 
+in at least two detectors as used in \cite{2012PhRvD..85h2002A} (as these would be signals 
+coincident with a GRB we do not include the often used further constraint that the total coherent 
+network SNR is greater than 12). The coalescence times for all injections were held fixed 
+within the centre of the prior window. This criteria means that as we go to larger distances our 
+population of injected sources will not be uniformly distributed over the sky, and will generally 
+be closer to being circularly polarised. For the \ac{NSBH} systems it also means that at larger 
+distances we will have more sources with larger mass ratios \MP{Check this}.
 
@@
 
-For each of the simulated sources we have calculated the 68\% credible region on the scale factors 
-for each detector (for a Gaussian distribution this is equivalent to the region bounded by 
-$1\sigma$ intervals). For all the signals in each distance increment we have produced the 
-distribution of the half widths (i.e.\ $1\sigma$) of these scale factors. These are shown in 
-figures~\ref{fig:narrowdist} and \ref{fig:widedist} for the case with the inclination distribution 
-with a width of $5^{\circ}$ and $20^{\circ}$ respectively.
+For each of the simulated sources we have used the {\tt emcee} python MCMC package to perform 
+parameter estimation over the unknown source parameters using the prior as discussed in 
+Section~\ref{sec:priors}. In each case we use an estimate of the noise PSD based on the design 
+sensitivity, but simulated as if estimated by averaging 64 separate PSD estimates and scaled with 
+the same factor as applied to the injection and noise\footnote{We do not account for there being 
+a difference between the estimated PSD and the actual PSD of the analysed section of data. This 
+difference would be very much correlated with the scale factor, so in reality our estimate of the 
+scale factor would be a combination of the calibration offset and the difference in the PSD. As 
+such our results on real data would be an upper limit on the calibration scale factor.}. This has 
+provided posterior probability distributions on the calibration scale factors for each detector 
+\MP{Show an example for both source types}. From these we have calculated the 68\% credible region 
+on the scale factors for each detector (if these were Gaussian distributions this is equivalent to 
+the region bounded by $1\sigma$ intervals). For all the signals at each distance increment we have 
+produced the distribution of the fractional half widths (i.e.\ $1\sigma$) of these scale factors 
+confidence intervals compared to the actual value. These are shown as boxplots in 
+figures~\ref{fig:bnsresults} and \ref{fig:nsbhresults} for the \ac{BNS} and \ac{NSBH} systems 
+respectively. The boxes show the extent from the lower to upper quartile of the values, whilst the 
+whiskers represent range of values. The black line within each box gives the median value and the 
+star gives the mean value. Also shown on each plot as the dashed magenta line is the percentage of 
+proposed injections that actually fulfilled the detection criterion.
 
@@
  \begin{center}
-  \includegraphics[width=1.0\textwidth]{scale_factors_5degs_boxplot.png}
+  \includegraphics[width=1.0\textwidth]{scale_factor.pdf}
  \end{center}
- \caption{\label{fig:narrowdist} Add caption!}
+ \caption{\label{fig:bnsresults} Distribution of the percentage accuracy at which the calibration 
+scale factor can be determined for a three detector network if assuming \ac{BNS} systems (provided 
+a coincident GRB is observed and can yield a distance estimate). The boxplots span the lower to 
+upper quartile range of the distributions, with the median value shown as a horizontal line within 
+the box and the mean shown as a star. The dashed magenta line shows the percentage of source drawn 
+from the prior distribution that would be detectable at each distance value.}
 \end{figure}
@@
  \begin{center}
-  \includegraphics[width=1.0\textwidth]{scale_factors_20degs_boxplot.png}
+  \includegraphics[width=1.0\textwidth]{scale_factor_nsbh.pdf}
  \end{center}
- \caption{\label{fig:widedist} Add caption!}
+ \caption{\label{fig:nsbhresults} Distribution of the percentage accuracy at which the calibration 
+scale factor can be determined for a three detector network if assuming \ac{NSBH} systems (provided 
+a coincident GRB is observed and can yield a distance estimate). The plot contents are the same as 
+in figure~\ref{fig:bnsresults}.}
 \end{figure}
+
+In figure~\ref{fig:bnsresults} we see that unsurprisingly on average the scale factor can be 
+recovered to equivalent precision in both \ac{aLIGO} detectors, with uncertainties generally within 
+the 10\% range for sources at 100\,Mpc. An interesting feature that we see is that for distances 
+$\gtrsim 250$\,Mpc the upper extent of the uncertainty hits a maximum at $\sim 25\mbox{--}30\%$, 
+whilst the width of boxes narrows. This is due to our ``detectability'' criteria, in that for all 
+distances we will only be seeing those sources with high enough SNR that we would consider them 
+detectable. There will therefore be fewer sources with SNR higher than this criteria at large 
+distances, giving us a narrower range, and we automatically exclude those with SNR that are too 
+small thus truncating our uncertainty distribution at the upper range. However, this does show that 
+on average for sources that are detectable out to 450\,Mpc we would be able to constrain the 
+calibration scale factor uncertainty to $\sim 20\%$. As the detection criteria will generally come 
+from the SNR being high enough within the two \ac{aLIGO} instruments this means that the SNR in 
+\ac{AdV} can be small and thus the ability to constrain the scale factor for it becomes poor 
+(although it still provides information that the calibration is not grossly inaccurate).
 
@@
 simulations, it was shown that this new method of calibration can reach and
-surpass the current method in terms of accuracy. The results in
-figure~\ref{fig:std-rndsp} show that for half of the randomly chosen sky
-positions relate to an estimation of $s$ within a range comparable with the
-current methods.
+surpass the current method in terms of accuracy.
 

### 2015-07-16 docs/cosmic_cal.tex
Added some more to the results section, but it really needs to be checked that what I say is true

@@
 
+In figure~\ref{fig:bnsresults} we see slightly different results. Despite the generally higher SNR 
+for the \ac{NSBH} systems we find that on averge that the scale factor uncertainty is still roughly 
+20\% at distances to 400--500\,Mpc. The range of uncertainty values is also far broader at all 
+distances and there is not the plateau and upper limit seen for the \ac{BNS} systems. \MP{This 
+needs checked, but ... This broader range of uncertainties appears to be down to the inclusion of 
+the spin parameter and correlations this has with the mass ratio $q$ - show posterior examples. The 
+lack of an upper plateau seems to be due to the detection criteria having a different effect on the 
+distribution of sources that could be seen, in particular it biases the observable sources towards 
+having a higher mass ratio (again check this) and thus having more correlation with the spin.}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-07-21 docs/cosmic_cal.tex
Added a posterior probability figure and ended a line mid-sentence!

@@
 provided posterior probability distributions on the calibration scale factors for each detector 
-\MP{Show an example for both source types}. From these we have calculated the 68\% credible region 
+\MP{Show an example for both source types}. An example of the marginalised posterior probablil From 
+these we have calculated the 68\% credible region 
 on the scale factors for each detector (if these were Gaussian distributions this is equivalent to 
@@
 proposed injections that actually fulfilled the detection criterion.
+
+\begin{figure}
+ \begin{center}
+  \includegraphics[width=1.0\textwidth]{bns_post_fig.pdf}
+ \end{center}
+ \caption{\label{fig:bnspost} The marginalised posterior probability distributions for the unknown
+ parameters of a \ac{BNS} system including calibration scaling factors for the three detectors (H1, 
+L1 and V1). The injected signal had \ac{SNR} of 8.7, 10.9 and 4.1 and percentage uncertainties in 
+the calibration scale factors of 14\%, 11\% and 27\% for each of the detectors respectively.}
+\end{figure}
 

### 2015-07-21 src/python/scale_factor_mcmc.py

@@ -778,12 +778,15 @@ if it does not fulfill the SNR criterion.")
     mC = (m1inj*m2inj)**(3./5.) / (m1inj+m2inj)**(1./5.)
     q = m2inj/m1inj
 
+    # set t0 to 0
+    samples[:,3] = samples[:,3] - t0
+
     if args.nsbh:
       labels = ["$\psi$", "$\phi_0$", "$\iota$", "$t_c$", "$\mathcal{M}$", "$q$", "$a_1$"]
-      truths = [psi, phi0, iota, t0, mC, q, a1spin]
+      truths = [psi, phi0, iota, 0.0, mC, q, a1spin]
     else:
       labels = ["$\psi$", "$\phi_0$", "$\iota$", "$t_c$", "$\mathcal{M}$", "$q$"]
-      truths = [psi, phi0, iota, t0, mC, q]
+      truths = [psi, phi0, iota, 0.0, mC, q]
     for i in range(len(dets)):
       labels.append("$s_{\mathrm{%s}}}$" % dets[i])
       truths.append(scales[i])
@@ -793,5 +796,5 @@ if it does not fulfill the SNR criterion.")
 
     fig = triangle.corner(samples, labels=labels, truths=truths, data_kwargs={'color': 'darkblue', 'ms': 2}, plot_density=True, no_fill_contours=False, plot_contours=True, levels=levels)
 
-    plotfile = os.path.join(outpath, 'posterior_plot_'+intseed+'.png')
+    plotfile = os.path.join(outpath, 'posterior_plot_'+intseed+'.pdf')
     fig.savefig(plotfile)

### 2015-07-24 docs/cosmic_cal.tex
Added posterior plot showing example of NS-BH system, and added more text

@@
 
+A simple assumption that one could make would be that the uncertainties on the calibration scale 
+factors (if they are independent of other parameters and have a Gaussian probability distribtuion) 
+should be given by $\sim 1/{\rm SNR}$ for each detector. As we will show below this assumption is 
+reasonable, but correlations do exist between parameters meaning that it does not completely hold, 
+especially in the case on \ac{NSBH} systems with a spinning component.
+
 For each of the simulated sources we have used the {\tt emcee} python MCMC package to perform 
@@
 provided posterior probability distributions on the calibration scale factors for each detector 
-\MP{Show an example for both source types}. An example of the marginalised posterior probablil From 
-these we have calculated the 68\% credible region 
-on the scale factors for each detector (if these were Gaussian distributions this is equivalent to 
-the region bounded by $1\sigma$ intervals). For all the signals at each distance increment we have 
+\MP{Show an example for both source types}. Examples of the marginalised posterior probabilities
+for a \ac{BNS} system and a \ac{NSBH} system observed with the three detector network are shown in 
+figures~\ref{fig:bnspost} and \ref{fig:nsbhpost} respectively.
+
+From these posterior distributions we have calculated the 68\% credible region on the scale factors 
+for each detector (if these were Gaussian distributions this is equivalent to the region bounded by 
+$1\sigma$ intervals). For all the signals at each distance increment we have 
 produced the distribution of the fractional half widths (i.e.\ $1\sigma$) of these scale factors 
@@
 respectively. The boxes show the extent from the lower to upper quartile of the values, whilst the 
-whiskers represent range of values. The black line within each box gives the median value and the 
-star gives the mean value. Also shown on each plot as the dashed magenta line is the percentage of 
-proposed injections that actually fulfilled the detection criterion.
+whiskers represent the range of values. The black line within each box gives the median value and 
+the star gives the mean value. Also shown on each plot as the dashed magenta line is the percentage 
+of proposed injections that actually fulfilled the detection criterion.
 
@@
  parameters of a \ac{BNS} system including calibration scaling factors for the three detectors (H1, 
-L1 and V1). The injected signal had \ac{SNR} of 8.7, 10.9 and 4.1 and percentage uncertainties in 
-the calibration scale factors of 14\%, 11\% and 27\% for each of the detectors respectively.}
+L1 and V1). The injected signal was at a distance of 250\,Mpc, had \ac{SNR} of 8.7, 10.9 and 4.1 
+and percentage uncertainties in the calibration scale factors of 14\%, 11\% and 27\% for each of 
+the detectors respectively.}
+\end{figure}
+
+\begin{figure}
+ \begin{center}
+  \includegraphics[width=1.0\textwidth]{nsbh_post_fig.pdf}
+ \end{center}
+ \caption{\label{fig:nsbhpost} The marginalised posterior probability distributions for the unknown
+ parameters of a \ac{NSBH} system including calibration scaling factors for the three detectors 
+(H1, L1 and V1). The injected signal was at a distance of 450\,Mpc, had \ac{SNR} of 7.7, 10.9 and 
+4.2 and percentage uncertainties in the calibration scale factors of 17\%, 14\% and 34\% for each of 
+the detectors respectively.}
 \end{figure}
@@
 \ac{AdV} can be small and thus the ability to constrain the scale factor for it becomes poor 
-(although it still provides information that the calibration is not grossly inaccurate).
+(although it still provides information that the calibration is not grossly inaccurate). We also 
+see that true uncertainties achievable for recovering the calibration scale factors are indeed very 
+similar to the simple assumption that they would be given by $sim 1/{\rm SNR}$. Although, they are 
+probably slightly poorer due to their still be correlations with $\iota$ even within the 
+constrained prior range, and slight correlation between the scale factors for H1 and L1 as seen in 
+figure~\ref{fig:bnspost}.
 
-In figure~\ref{fig:bnsresults} we see slightly different results. Despite the generally higher SNR 
+In figure~\ref{fig:nsbhresults} we see slightly different results. Despite the generally higher SNR 
 for the \ac{NSBH} systems we find that on averge that the scale factor uncertainty is still roughly 
@@
 \cm{What did we learn.}
-
-This paper has shown that for differing levels of SNR, the error in the
-calibration is resolved within a certain accuracy range. Initial simulations
-used a bad sky position, resulting a low SNR which created larger uncertainty
-in resolving $s$. It was shown that for this sky position, the calibration
-scale factor estimation was recovered with an accuracy of roughly a $27\%$ for
-$D_{L} = 200\,Mpc$ for LIGO Hanford. The resolution of $s$ for a network of
-detectors was also shown.
-
-The main results obtained in this project were from analysing randomly chosen
-sky positions from the entire field of view for the detector. Through these
-simulations, it was shown that this new method of calibration can reach and
-surpass the current method in terms of accuracy.
-
-These results are reassuring for parameter estimation of the defining GW signal
-and would help in reducing the uncertainty in a direct detection. This project
-was; however, time constrained. There are many, many things that could be done
-with this project. A true noise realisation with each frequency bin having a
-random Gaussian noise would be the next step in calibration error estimation.
 

### 2015-07-27 src/python/results_script_nsbh.py
Change NSBH results script to only use converged posterior chains

@@ -70,7 +70,7 @@ pl.rc('text', usetex=True)
 pl.rc('font', family='serif')
 pl.rc('font', size=14)
 
-dists = [100, 200, 300, 400, 500, 600, 700, 800, 900, 1000]
+dists = [100, 200, 300, 400, 500, 600, 700, 800, 900] #, 1000]
 
 prefix = indir
 
@@ -94,6 +94,7 @@ for i, d in enumerate(dirs):
   files = [os.path.join(d, f) for f in os.listdir(d) if os.path.isfile(os.path.join(d, f)) and fnpre in f]
   
   # go through files and extract the relative standard devaition on the scale factors for each detector
+  count = 0
   for f in files:
     fo = open(f, 'r')
     info = json.load(fo)
@@ -101,10 +102,20 @@ for i, d in enumerate(dirs):
 
     vals = []
     for k in range(len(info['InjectionParameters']['scales'])):
-      # divide by two to get the half widths and convert to percentage
-      vals.append(100.*(info['Results']['Scale68%CredibleInterval'][k][1]-info['Results']['Scale68%CredibleInterval'][k][0])/(2.*info['InjectionParameters']['scales'][k]))
+      # check whether chain looks like it's converged with a simple check of the histogram
+      histd = info['Results']['ScaleHist'][k]
+      histd = np.array(histd)
+      # get number of nonzero values
+      nonzero = np.zeros(len(histd[0]))[np.array(histd[0]) > 0]
 
-    relsf.append(vals)
+      # say converged chain must have more than 75% of posterior points being non-zero
+      # and the standard deviation of the points must be within a sixth of the width of the histogram
+      if float(len(nonzero))/float(len(histd[0])) > 0.75 and info['Results']['ScaleSigma'][k] < (histd[1][-1]-histd[1][0])/6.:
+        # divide by two to get the half widths and convert to percentage
+        vals.append(100.*(info['Results']['Scale68%CredibleInterval'][k][1]-info['Results']['Scale68%CredibleInterval'][k][0])/(2.*info['InjectionParameters']['scales'][k]))
+
+    if len(vals) == 3:
+      relsf.append(vals)
     totaltime += info['Attempts']   
 
   rates.append(100.*len(files)/totaltime)
@@ -200,6 +211,7 @@ pl.ylabel('\% calibration scaling error (1$\sigma$ equivalent)')
 ax2 = ax.twinx()
[…]

### 2015-07-27 src/python/results_script.py
results_script.py: change axis label colour

@@ -203,6 +203,7 @@ pl.ylabel('\% calibration scaling error (1$\sigma$ equivalent)')
 ax2 = ax.twinx()
 ax2.plot(dists, rates, 'mo--', markerfacecolor='None', markeredgecolor='m')
 ax2.set_ylabel('\% of source distribution detectable')
+ax2.yaxis.label.set_color('magenta')
 ax2.set_xlim((0, 550))
 
 try:

### 2015-07-27 src/python/results_script.py
results_scripts.py: change this to also only us chains that look like they have converged

@@ -102,10 +102,20 @@ for i, d in enumerate(dirs):
 
     vals = []
     for k in range(len(info['InjectionParameters']['scales'])):
-      # divide by two to get the half widths and convert to percentage
-      vals.append(100.*(info['Results']['Scale68%CredibleInterval'][k][1]-info['Results']['Scale68%CredibleInterval'][k][0])/(2.*info['InjectionParameters']['scales'][k]))
+      # check whether chain looks like it's converged with a simple check of the histogram
+      histd = info['Results']['ScaleHist'][k]
+      histd = np.array(histd)
+      # get number of nonzero values
+      nonzero = np.zeros(len(histd[0]))[np.array(histd[0]) > 0]
+      
+      # say converged chain must have more than 75% of posterior points being non-zero
+      # and the standard deviation of the points must be within a sixth of the width of the histogram
+      if float(len(nonzero))/float(len(histd[0])) > 0.75 and info['Results']['ScaleSigma'][k] < (histd[1][-1]-histd[1][0])/6.:
+        # divide by two to get the half widths and convert to percentage
+        vals.append(100.*(info['Results']['Scale68%CredibleInterval'][k][1]-info['Results']['Scale68%CredibleInterval'][k][0])/(2.*info['InjectionParameters']['scales'][k]))
 
-    relsf.append(vals)
+    if len(vals) == 3:
+      relsf.append(vals)
 
     totaltime += info['Attempts']
   
@@ -118,28 +128,6 @@ for i, d in enumerate(dirs):
 
   print np.mean(nprelsf[:,0]), np.mean(nprelsf[:,1]), np.mean(nprelsf[:,2])
   print np.median(nprelsf[:,0]), np.median(nprelsf[:,1]), np.median(nprelsf[:,2])
-
-  # plot output
-  #ci1 = credible_interval(nprelsf[:,0], 0.90)
-  #ci2 = credible_interval(nprelsf[:,1], 0.90)
-  #ci3 = credible_interval(nprelsf[:,2], 0.90)
-  
-  #pl.plot(dists[i]-7.5, np.mean(nprelsf[:,0]), 'bo', lw=2, ms=6)
[…]

### 2015-07-27 src/python/results_script_nsbh.py

@@ -94,7 +94,6 @@ for i, d in enumerate(dirs):
   files = [os.path.join(d, f) for f in os.listdir(d) if os.path.isfile(os.path.join(d, f)) and fnpre in f]
   
   # go through files and extract the relative standard devaition on the scale factors for each detector
-  count = 0
   for f in files:
     fo = open(f, 'r')
     info = json.load(fo)
@@ -116,6 +115,7 @@ for i, d in enumerate(dirs):
 
     if len(vals) == 3:
       relsf.append(vals)
+
     totaltime += info['Attempts']   
 
   rates.append(100.*len(files)/totaltime)
@@ -127,28 +127,6 @@ for i, d in enumerate(dirs):
 
   print np.mean(nprelsf[:,0]), np.mean(nprelsf[:,1]), np.mean(nprelsf[:,2])
   print np.median(nprelsf[:,0]), np.median(nprelsf[:,1]), np.median(nprelsf[:,2])
-
-  # plot output
-  #ci1 = credible_interval(nprelsf[:,0], 0.90)
-  #ci2 = credible_interval(nprelsf[:,1], 0.90)
-  #ci3 = credible_interval(nprelsf[:,2], 0.90)
-  
-  #pl.plot(dists[i]-7.5, np.mean(nprelsf[:,0]), 'bo', lw=2, ms=6)
-  #pl.plot(dists[i], np.mean(nprelsf[:,1]), 'ro', lw=2, ms=6)
-  #pl.plot(dists[i]+7.5, np.mean(nprelsf[:,2]), 'go', lw=2, ms=6)
-  
-  #pl.plot(dists[i]-7.5, np.median(nprelsf[:,0]), 'bx', lw=2, ms=6)
-  #pl.plot(dists[i], np.median(nprelsf[:,1]), 'rx', lw=2, ms=6)
-  #pl.plot(dists[i]+7.5, np.median(nprelsf[:,2]), 'gx', lw=2, ms=6)
-  
-  #if i == 0:
-  #  pl.plot([dists[i]-7.5, dists[i]-7.5], ci1, 'b', label='H1', lw=2)
-  #  pl.plot([dists[i], dists[i]], ci2, 'r', label='L1', lw=2)
-  #  pl.plot([dists[i]+7.5, dists[i]+7.5], ci3, 'g', label='V1', lw=2)
-  #else:
[…]

### 2015-07-28 docs/cosmic_cal.tex
cosmic_cal.tex: added comment
Some minor updates to the paper

@@
 \acrodef{AdV}[AdV]{Advanced Virgo}
+\acrodef{PSD}[PSD]{power spectral density}
 
@@
 range of distances (50\,Mpc to 500\,Mpc with 50\,Mpc increments for the \ac{BNS} signals and 
-100\,Mpc to 1000\,Mpc with 100\,Mpc increments for the \ac{NSBH} signals). The simulations use the 
+100\,Mpc to 900\,Mpc with 100\,Mpc increments for the \ac{NSBH} signals). The simulations use the 
 two Advanced LIGO detectors (one at Hanford -- H1 -- and one at Livingston -- L1) and the Advanced 
@@
 % Describe the ensemble of simulations used
-Separately for the \ac{BNS} and \ac{NSBH} systems we have performed 2000 simulations at each 
-distance value. For proposing injections for each simulation the source parameters were randomly 
+Separately for the \ac{BNS} and \ac{NSBH} systems we have performed up to 2000 simulations at each 
+distance value\footnote{Simulations on which our MCMC did not produce converged chains within our 
+analysis time were discarded and not re-run, although there were always at least 1000 simulations 
+that completed.}. For proposing injections for each simulation the source parameters were randomly 
 drawn from the prior distributions given in Section~\ref{sec:priors}, with the sky position being 
@@
 A simple assumption that one could make would be that the uncertainties on the calibration scale 
-factors (if they are independent of other parameters and have a Gaussian probability distribtuion) 
+factors (if they are independent of other parameters and have a roughly Gaussian probability 
+distribtuion)
 should be given by $\sim 1/{\rm SNR}$ for each detector. As we will show below this assumption is 
-reasonable, but correlations do exist between parameters meaning that it does not completely hold, 
-especially in the case on \ac{NSBH} systems with a spinning component.
+reasonable, but small correlations do exist between parameters meaning that it does not completely 
+hold.
 
-For each of the simulated sources we have used the {\tt emcee} python MCMC package to perform 
-parameter estimation over the unknown source parameters using the prior as discussed in 
-Section~\ref{sec:priors}. In each case we use an estimate of the noise PSD based on the design 
-sensitivity, but simulated as if estimated by averaging 64 separate PSD estimates and scaled with 
-the same factor as applied to the injection and noise\footnote{We do not account for there being 
-a difference between the estimated PSD and the actual PSD of the analysed section of data. This 
-difference would be very much correlated with the scale factor, so in reality our estimate of the 
-scale factor would be a combination of the calibration offset and the difference in the PSD. As 
-such our results on real data would be an upper limit on the calibration scale factor.}. This has 
-provided posterior probability distributions on the calibration scale factors for each detector 
-\MP{Show an example for both source types}. Examples of the marginalised posterior probabilities
-for a \ac{BNS} system and a \ac{NSBH} system observed with the three detector network are shown in 
-figures~\ref{fig:bnspost} and \ref{fig:nsbhpost} respectively.
+For each of the simulated sources we have used the {\tt emcee} python MCMC package 
+\cite{2013PASP..125..306F} to perform parameter estimation over the unknown source parameters using 
+the priors as discussed in Section~\ref{sec:priors}. In each case when calculating the likelihood 
+we used an estimate of the noise \ac{PSD} based on the advanced detector design sensitivities 
+(using those given in \cite{2013arXiv1304.0670L}), but simulated as if calculated by averaging 64 
+separate noisy \ac{PSD} estimates and scaled with the same factor as applied to the injection and 
+noise\footnote{We do not account for there being a difference between the estimated \ac{PSD} and the 
+actual \ac{PSD} of the analysed section of data. This difference would be very much correlated with 
+the scale factor, so in reality our estimate of the scale factor would be a combination of the 
+calibration offset and the difference in the \ac{PSD}. As such our results on real data would be an 
+upper limit on the calibration scale factor.}. This has provided posterior probability distributions 
+on the calibration scale factors for each detector. Examples of the marginalised posterior 
+probabilities for a \ac{BNS} system and a \ac{NSBH} system observed with the three detector network 
+are shown in figures~\ref{fig:bnspost} and \ref{fig:nsbhpost} respectively.
 
-From these posterior distributions we have calculated the 68\% credible region on the scale factors 
-for each detector (if these were Gaussian distributions this is equivalent to the region bounded by 
-$1\sigma$ intervals). For all the signals at each distance increment we have 
-produced the distribution of the fractional half widths (i.e.\ $1\sigma$) of these scale factors 
-confidence intervals compared to the actual value. These are shown as boxplots in 
-figures~\ref{fig:bnsresults} and \ref{fig:nsbhresults} for the \ac{BNS} and \ac{NSBH} systems 
+From these posterior distributions we have calculated the 68\% credible region on the 
+calibration scale factors for each detector (if these were Gaussian distributions this is 
+equivalent to the region bounded by $1\sigma$ intervals). For all the signals at each distance 
+increment we have produced the distribution of the fractional half widths (i.e.\ $1\sigma$) of 
+these scale factor's confidence intervals compared to the actual value. These are shown as boxplots 
+in figures~\ref{fig:bnsresults} and \ref{fig:nsbhresults} for the \ac{BNS} and \ac{NSBH} systems 
 respectively. The boxes show the extent from the lower to upper quartile of the values, whilst the 
@@
 the star gives the mean value. Also shown on each plot as the dashed magenta line is the percentage 
-of proposed injections that actually fulfilled the detection criterion.
+of signal drawn from the prior distribution that would fulfill the detection criterion.
 
@@
  \caption{\label{fig:bnspost} The marginalised posterior probability distributions for the unknown
- parameters of a \ac{BNS} system including calibration scaling factors for the three detectors (H1, 
-L1 and V1). The injected signal was at a distance of 250\,Mpc, had \ac{SNR} of 8.7, 10.9 and 4.1 
-and percentage uncertainties in the calibration scale factors of 14\%, 11\% and 27\% for each of 
-the detectors respectively.}
+ parameters of a \ac{BNS} system, including calibration scaling factors for the three detectors 
+(H1, L1 and V1). The injected signal was at a distance of 250\,Mpc, had \ac{SNR} of 8.7, 10.9 and 
+4.1 and percentage uncertainties in the calibration scale factors of 14\%, 11\% and 27\% for each 
+of the detectors respectively.}
 \end{figure}
@@
  \caption{\label{fig:nsbhpost} The marginalised posterior probability distributions for the unknown
- parameters of a \ac{NSBH} system including calibration scaling factors for the three detectors 
+ parameters of a \ac{NSBH} system, including calibration scaling factors for the three detectors 
 (H1, L1 and V1). The injected signal was at a distance of 450\,Mpc, had \ac{SNR} of 7.7, 10.9 and 
@@
  \caption{\label{fig:bnsresults} Distribution of the percentage accuracy at which the calibration 
-scale factor can be determined for a three detector network if assuming \ac{BNS} systems (provided 
+scale factors can be determined for a three detector network if assuming \ac{BNS} systems (provided 
 a coincident GRB is observed and can yield a distance estimate). The boxplots span the lower to 
@@
  \caption{\label{fig:nsbhresults} Distribution of the percentage accuracy at which the calibration 
-scale factor can be determined for a three detector network if assuming \ac{NSBH} systems (provided 
-a coincident GRB is observed and can yield a distance estimate). The plot contents are the same as 
+scale factors can be determined for a three detector network if assuming \ac{NSBH} systems 
+(provided a coincident GRB is observed and can yield a distance estimate). The plot contents are 
+the same as 
 in figure~\ref{fig:bnsresults}.}
@@
 
-In figure~\ref{fig:nsbhresults} we see slightly different results. Despite the generally higher SNR 
-for the \ac{NSBH} systems we find that on averge that the scale factor uncertainty is still roughly 
-20\% at distances to 400--500\,Mpc. The range of uncertainty values is also far broader at all 
-distances and there is not the plateau and upper limit seen for the \ac{BNS} systems. \MP{This 
-needs checked, but ... This broader range of uncertainties appears to be down to the inclusion of 
-the spin parameter and correlations this has with the mass ratio $q$ - show posterior examples. The 
-lack of an upper plateau seems to be due to the detection criteria having a different effect on the 
-distribution of sources that could be seen, in particular it biases the observable sources towards 
-having a higher mass ratio (again check this) and thus having more correlation with the spin.}
+In figure~\ref{fig:nsbhresults} we see very similar results although the higher \ac{SNR} of the 
+signals means that we can provide $\sim 20\mbox{--}25\%$ uncertainties on the calibration scale 
+factors for H1 and L1 out to greater distances.
 

### 2015-07-29 src/python/results_script_nsbh.py
scale_factor_nsbh.pdf: updated figure to shrink the x-axis range

@@ -190,7 +190,8 @@ ax2 = ax.twinx()
 ax2.plot(dists, rates, 'mo--', markerfacecolor='None', markeredgecolor='m')
 ax2.set_ylabel('\% of source distribution detectable')
 ax2.yaxis.label.set_color('magenta')
-ax2.set_xlim((0, 1100))
+#ax2.set_xlim((0, 1100))
+ax2.set_xlim((0, 1000))
 
 try:
   fig.savefig(outfile)

### 2015-08-05 src/python/results_script.py
Changed the criterion for assessing chain convergence to make it stricter

@@ -93,6 +93,7 @@ for i, d in enumerate(dirs):
   files = [os.path.join(d, f) for f in os.listdir(d) if os.path.isfile(os.path.join(d, f)) and fnpre in f]
   
   totaltime = 0.
+  accepted = 0
 
   # go through files and extract the relative standard devaition on the scale factors for each detector
   for f in files:
@@ -108,14 +109,15 @@ for i, d in enumerate(dirs):
       # get number of nonzero values
       nonzero = np.zeros(len(histd[0]))[np.array(histd[0]) > 0]
       
-      # say converged chain must have more than 75% of posterior points being non-zero
+      # say converged chain must have more than 90% of posterior points being non-zero
       # and the standard deviation of the points must be within a sixth of the width of the histogram
-      if float(len(nonzero))/float(len(histd[0])) > 0.75 and info['Results']['ScaleSigma'][k] < (histd[1][-1]-histd[1][0])/6.:
+      if float(len(nonzero))/float(len(histd[0])) > 0.9 and info['Results']['ScaleSigma'][k] < (histd[1][-1]-histd[1][0])/6.:
         # divide by two to get the half widths and convert to percentage
         vals.append(100.*(info['Results']['Scale68%CredibleInterval'][k][1]-info['Results']['Scale68%CredibleInterval'][k][0])/(2.*info['InjectionParameters']['scales'][k]))
 
     if len(vals) == 3:
       relsf.append(vals)
+      accepted += 1
 
     totaltime += info['Attempts']
   
@@ -126,6 +128,7 @@ for i, d in enumerate(dirs):
   data.append(nprelsf[:,1])
   data.append(nprelsf[:,2])
 
+  print accepted
   print np.mean(nprelsf[:,0]), np.mean(nprelsf[:,1]), np.mean(nprelsf[:,2])
   print np.median(nprelsf[:,0]), np.median(nprelsf[:,1]), np.median(nprelsf[:,2])
 

### 2015-08-05 src/python/results_script_nsbh.py

@@ -94,6 +94,7 @@ for i, d in enumerate(dirs):
   files = [os.path.join(d, f) for f in os.listdir(d) if os.path.isfile(os.path.join(d, f)) and fnpre in f]
   
   # go through files and extract the relative standard devaition on the scale factors for each detector
+  acc = 0
   for f in files:
     fo = open(f, 'r')
     info = json.load(fo)
@@ -107,14 +108,15 @@ for i, d in enumerate(dirs):
       # get number of nonzero values
       nonzero = np.zeros(len(histd[0]))[np.array(histd[0]) > 0]
 
-      # say converged chain must have more than 75% of posterior points being non-zero
+      # say converged chain must have more than 90% of posterior points being non-zero
       # and the standard deviation of the points must be within a sixth of the width of the histogram
-      if float(len(nonzero))/float(len(histd[0])) > 0.75 and info['Results']['ScaleSigma'][k] < (histd[1][-1]-histd[1][0])/6.:
+      if float(len(nonzero))/float(len(histd[0])) > 0.9 and info['Results']['ScaleSigma'][k] < (histd[1][-1]-histd[1][0])/6.:
         # divide by two to get the half widths and convert to percentage
         vals.append(100.*(info['Results']['Scale68%CredibleInterval'][k][1]-info['Results']['Scale68%CredibleInterval'][k][0])/(2.*info['InjectionParameters']['scales'][k]))
 
     if len(vals) == 3:
       relsf.append(vals)
+      acc += 1
 
     totaltime += info['Attempts']   
 
@@ -125,6 +127,7 @@ for i, d in enumerate(dirs):
   data.append(nprelsf[:,1])
   data.append(nprelsf[:,2])
 
+  print acc
   print np.mean(nprelsf[:,0]), np.mean(nprelsf[:,1]), np.mean(nprelsf[:,2])
   print np.median(nprelsf[:,0]), np.median(nprelsf[:,1]), np.median(nprelsf[:,2])
 

### 2015-08-05 docs/cosmic_cal.tex
Some large chunks of text removed

@@
 
-The technique used throughout GW research in calibrating the detectors is
-carried out through a complicated system of physical manipulations of the
-Fabret-Perot Michelson interferometer; where an elaborate feedback system is
-used to sustain a defined measurement in arm length difference between the
-moving mirrors~\cite{LIGOCal}. In general, the interferometers are set to move
-at a frequency corresponding to a GW signal. An error signal is then found
-through the Pound-Driver-Hall technique~\cite{Black} for mirror stabilisation
-which is proportional to the difference in mirror separation of the
-Fabret-Perot Michelson interferometer. The Pound-Drever-Hall technique is
-essentially a control loop feedback system in which the difference output (in
-GW detectors, this is the distance the mirrors move due to the GW) and the the
-output set is reduced to the highest degree of certainty. This is just a brief
-overview of the calibration method currently used, and is an extensive study in
-itself, see \cite{Vitale:2012} and references therein. For \ac{aLIGO} and \ac{AdV},
-at the frequency range used in this project, the error in the calibration using
-the `hardware injection method' is roughly $10\%$ \cite{Vitale:2012}. This is a
-benchmark set for the error estimation using the proposed new method in this
-project.~\cm{Clean this up and expand.}
+The technique used throughout \ac{GW} research in calibrating the detectors is carried out through 
+a complicated system of physical manipulations of the Fabret-Perot Michelson interferometer; where 
+an elaborate feedback system is used to sustain a defined measurement in arm length difference 
+between the moving mirrors. A comprehensive description of the calibration procedure (in particular 
+for the LIGO detectors during their fifth science run) can be found in \cite{2010NIMPA.624..223A}, 
+and a brief description is giving in \cite{Vitale:2012} and references therein. For \ac{aLIGO} 
+and \ac{AdV}, at the frequency range used in this project, the error in the amplitude calibration 
+using the `hardware injection method' is roughly $10\%$ \cite{Vitale:2012}. This is a benchmark set 
+for the error estimation using the proposed new method in this project.~\cm{Clean this up and 
+expand.}
 
 % Describe our model for the calibration
-\MP{The gravitational wave strain is measured through the differential arm length, $\Delta L$, 
-changes of the interferometer via $h = \Delta L(f) / L$, where $L$ is the full arm length.
-Calibration is required to relate the actual measured interferometer error signal output
-$e(f)$ to $\Delta L$. This relation is known as the length response function, $R(f,t)$,
-defined such that
+The \ac{GW} strain is measured through the differential arm length, $\Delta L$, changes of the 
+interferometer via $h(f,t) = \Delta L(f,t) / L$, where $L$ is the full arm length. Calibration is 
+required to relate the actual measured interferometer error signal output $e(f)$ to $\Delta L$. 
+This relation is known as the length response function, $R(f)$, defined such that
 \begin{equation}
-\Delta L(f) = R(f,t) e(f),
+\Delta L(f,t) = R(f) e(f, t),
 \end{equation}
-where $R$ varies slowly in time (in comparison to transient signal timescales).
+where where we assume $R$ varies only slowly in time (in comparison to transient signal timescales).
 Calculation of $R$ requires the measurement of various functions within a control feedback loop
-(see \cite{Vitale:2012} for a brief overview of this), which are subject to
-uncertainties. In this study we will assume an estimate of $R$ is available (although
-in theory we could take on the role of estimating $R$ itself), but that it differs from the
-truth through some unknown scale factor, $s$, so that
+(see \cite{Vitale:2012} for a brief overview of this), which are subject to uncertainties. In this 
+study we will assume an estimate of $R$ is available (although in theory we could take on the role 
+of estimating $R$ itself), but that it differs from the truth through some unknown scale factor, 
+$s$, so that
 \begin{equation}
@@
 means that if $s > 1$ then a signal would appear to have a larger amplitude (e.g.\ be closer) than 
-expected, whereas if $s < 1$ would appear to have a lower amplitude than expected. In this analysis 
-we will simplify the situation by assuming that $s$ is a constant with respect to frequency, but in 
-future studies that assumption could be dropped and $s$ could take some functional form, or 
-piece-wise fit, with respect to $f$.}
+in reality\footnote{However, the \ac{SNR} would be the correct value as the signal and the noise 
+will be contain the scale factor}, whereas if $s < 1$ it would appear to have a lower amplitude 
+than reality. In this analysis we will simplify the situation by assuming that $s$ is a constant 
+with respect to frequency, but in future studies that assumption could be dropped and $s$ could 
+take some functional form, or piece-wise fit, with respect to $f$.
 
@@
 % observing a coincident event
-The most likely scenario in which a coincident \ac{GW}--\ac{sGRB} event would
-be identified is through the targeted follow-up of an observed \ac{sGRB} or by
-post-facto matching of \ac{GW} trigger lists with known \ac{sGRB} events.  The
-likelihood of being able to follow-up \ac{GW} events using gamma-ray telescopes
-with low enough latency to catch a \ac{sGRB} is low.  Compounding this issue is
-the relatively large \ac{GW} sky error-box giving a field-of-view for the
-\ac{EM} observatories to search spanning $\sim 100$'s of square
-degrees~\cite{grb}.  For the \ac{GW} follow-up of \ac{sGRB} scenario the merger
-time for \ac{BNS} systems will be estimated from the \ac{sGRB} to within a few
-seconds~\cite{grb}.  This makes the follow-up search less computationally
-expensive since it is performed over a smaller range of data using potentially
-fewer numbers of waveform templates.  This computational saving enables the use
-of a more computationally expensive multi-detector coherent scheme rather than
-the cheaper coincidence methods used in the untargeted searches.
+The most likely scenario in which a coincident \ac{GW}--\ac{sGRB} event would be identified is 
+through the targeted follow-up of an observed \ac{sGRB} or by post-facto matching of \ac{GW} 
+trigger lists with known \ac{sGRB} events.  The likelihood of being able to follow-up \ac{GW} 
+events using gamma-ray telescopes with low enough latency to catch a \ac{sGRB} is low.  Compounding 
+this issue is the relatively large \ac{GW} sky error-box giving a field-of-view for the \ac{EM} 
+observatories to search spanning $\sim 100$'s of square degrees~\cite{grb}.  For the \ac{GW} 
+follow-up of \ac{sGRB} scenario the merger time for \ac{BNS} systems will be estimated from the 
+\ac{sGRB} to within a few seconds~\cite{grb}.  This makes the follow-up search less computationally
+expensive since it is performed over a smaller range of data using potentially fewer numbers of 
+waveform templates.  This computational saving enables the use of a more computationally expensive 
+multi-detector coherent scheme rather than the cheaper coincidence methods used in the untargeted 
+searches.
 
 % discuss the inclination angle issue
-A fortunate consequence of a joint \ac{GW}--\ac{sGRB} observation will be the
-fact that in order for such an event to be observed, the \ac{BNS} system must
-have had its orbital angular momentum vector pointing towards (or away) from
-the detector.  The actual inclination angle of the system, defined as the angle
-between the orbital angular momentum vector and the line of sight, must be $<$
-half of the \ac{sGRB} beaming angle.  This prerequisite property limits us to
-systems that are approximately ``face-on'' and therefore biases us to higher
-\ac{SNR} signals.  However, as discussed earlier, the property of beaming
-severely impacts the probable rate of such joint observations.
+A fortunate consequence of a joint \ac{GW}--\ac{sGRB} observation will be the fact that in order 
+for such an event to be observed, the \ac{BNS} system must have had its orbital angular momentum 
+vector pointing towards (or away) from the detector.  The actual inclination angle of the system, 
+defined as the angle between the orbital angular momentum vector and the line of sight, must be $<$
+half of the \ac{sGRB} beaming angle.  This prerequisite property limits us to systems that are 
+approximately ``face-on'' and therefore biases us to higher \ac{SNR} signals.  However, as 
+discussed earlier, the property of beaming severely impacts the probable rate of such joint 
+observations.
 
@@
 
-\MP{We want to assess how well a calibration scale factor can be estimated
-from a single observed \ac{GW} associated with a particular \ac{sGRB}. As we {\it a priori}
-have no knowledge of the likely location and distance of such an event we have performed
-simulations of multiple events to see the expected distribution the accuracy of the
-calibration scale factor recovery.}
+We want to assess how well the calibration scale factor can be estimated from a single observed 
+\ac{GW} associated with a particular \ac{sGRB}. As we {\it a priori} have no knowledge of the 
+likely location and distance of such an event we have performed simulations of multiple events to 
+see the expected distribution the accuracy of the calibration scale factor recovery.
 
@@
 black hole has non-negligible spin and that its rotational angular momentum is aligned with the 
-systems orbital angular momentum.
+system's orbital angular momentum.
 
@@
    \gws^{k}(f) = \big[ F_{+}^{k}(\alpha, \delta, \psi)\gws_{+}(f) +
-F_{\times}^D(\alpha, \delta, \psi)\gws_{\times}(f)\big]
+F_{\times}^k(\alpha, \delta, \psi)\gws_{\times}(f)\big]
 \end{equation}
 %
-where $F_{+},F_{\times}$ are the antenna response functions which are dependent
-upon the polarisation angle $\psi$ and the sky position of the source
-$\alpha,\delta$.
+where $F_{+},F_{\times}$ are the antenna response functions which are dependent upon the 
+polarisation angle $\psi$ and the sky position of the source $\alpha$ and $\delta$.
 
 % Define the detectors and the noise epoch we will be using
-For this analysis we will consider advanced generation instruments operating
-at their design sensitivity, \MP{with noise power spectral densities taken from
-\cite{2013arXiv1304.0670L}}. The network we consider is the 3 detector H1--L1--V1 configuration.
+For this analysis we will consider the advanced generation LIGO and Virgo detectors operating
+at their design sensitivity, with noise power spectral densities taken from
+\cite{2013arXiv1304.0670L}. The network we consider is the two \ac{aLIGO} detectors (H1 and L1) and 
+the \ac{AdV} detector (V1).
 
@@
 %
-where $\tilde{n}(f)$ is the Fourier transform of Gaussian noise drawn from a
-distribution consistent with the power-spectral-density of the advanced
-detector noise.
+where $\tilde{n}(f)$ is the Fourier transform of Gaussian noise drawn from a distribution 
+consistent with the \ac{PSD} of the advanced detector noise.
 
@@
 
-\MP{Our aim is to estimate the marginal posterior probability distribution function (pdf)
-for the unknown calibration scale factors. We use Bayes' theorem for which we need to define
-a likelihood function and prior pdfs on the signal parameters being estimated. In this
-analysis we use the Markov chain Monte Carlo code {\tt emcee} \cite{2013PASP..125..306F} to sample 
-the posterior distribution.} We use a Gaussian likelihood function given by
+To assess the ability to estimate the unknown calibration scale factors we want to calculate their 
+marginal posterior probability distribution functions (pdf) given the data contains a signal. We 
+use Bayes' theorem for which we need to define a likelihood function and prior pdfs on the signal 
+parameters being estimated. In this analysis we use the Markov chain Monte Carlo code {\tt emcee} 
+\cite{2013PASP..125..306F} to sample the posterior distribution. We use a Gaussian likelihood 
+function given by
 %
@@
 % 
-where $\Delta f$ is the frequency resolution of the data, $S_k$ is the $k^{\rm th}$ detector's one 
-sided noise power spectral density, and the $i$ indices increment over frequency with $i_{f_{\rm 
-low}}$ and $i_{f_{\rm high}}$ corresponding to the lower and upper range in frequencies used for 
-our analysis, which were 20\,Hz to 400\,Hz respectively. This frequency range was chosen as the 
-vast majority of the \ac{SNR} can be found within this range, so going to lower or higher 
-frequencies provides very little additional information for our analysis. The source parameters, 
-and the unknown scale factor, have been subsumed into a vector $\btheta$. For a network of 
-detectors, the overall likelihood function becomes a product of the likelihoods for each detectors, 
-i.e.\
-\begin{equation}
-  \label{eq:mult-likeli}
-  p(\bi{d}| \btheta, \curlH) = \prod \limits_D p(\bi{d}^D| \vec{\theta}, \curlH)
-\end{equation}
-where $D = \{H,L,V\}$.
+where $\bi{d}$ is an array containing the data for all detectors, $\Delta f$ is the frequency 
+resolution, $S_k$ is the $k^{\rm th}$ detector's one sided noise \ac{PSD}, and the $i$ indices 
+increment over frequency with $i_{f_{\rm low}}$ and $i_{f_{\rm high}}$ corresponding to the lower 
+and upper range in frequencies used for our analysis, which were 20\,Hz to 400\,Hz respectively. 
+This frequency range was chosen as the vast majority of the \ac{SNR} for the inspiralling signals 
+we use can be found within this range, so going to lower or higher frequencies provides very little 
+additional information. The source parameters, and the unknown scale factors, have been subsumed 
+into a vector $\btheta$.
 
@@
 
-The primary reason why we can use coincident observations with a GRB as a potential check of 
-detector calibration is that the GRB allows us to constrain our priors on various parameters that 
-would normally be highly correlated with the signal amplitude. The most important of these is that 
-the GRB can provide a very tight constraint on the source distance\footnote{For \ac{BNS} and 
-\ac{NSBH} systems they will be observable only in the relatively nearby Universe ($z \sim 
+The primary reason why we can use coincident observations with a \ac{sGRB} as a potential check of 
+detector calibration is that the \ac{GRB} allows us to constrain the priors on various parameters 
+that would normally be highly correlated with the signal amplitude. The most important of these is 
+that the \ac{GRB} can provide a very tight constraint on the source distance\footnote{For \ac{BNS} 
+and \ac{NSBH} systems they will be observable only in the relatively nearby Universe ($z \sim 
 0.1$ and $z\sim 0.2$ for \ac{BNS} and \ac{NSBH} systems respectively), so the relation between the 
-measured GRB redshift and distance is simple and our luminosity distance will be negligably 
+measured GRB redshift and distance is simple and our luminosity distance will be negligibly 
 effected by the choice of cosmology \MP{Check this?}.}. In this analysis we therefore assume that 
 the source distance is known (i.e.\ has a $\delta$-function prior). Another piece of information 
-that we can use from having a coincident GRB is that the system is likely to be relatively close to 
-face-on (an therefore nearly circularly polarised), which allows us to place prior constraints on 
-the system inclination angle. Beam opening angles for short GRBs are relatively poorly constrained 
-as they are based on only a few jet break detections, however \cite{2014ApJ...780..118F} give a 
-median opening angle value of $\sim 10^{\circ}$. If we assume that the beam opening angle is a 
-half-normal distribution, for which the median is given by $\sqrt{2}\erf{}^{-1}(1/2)\sigma \sim 
-0.67\sigma$, then a median of $\sim 10^{\circ}$ corresponds to $\sigma \sim 14.8^{\circ}$. We 
-therefore place a Gaussian prior on the inclination angle with zero mean and $\sigma=14.8^{\circ}$.
+that we can use from having a coincident \ac{GRB} is that the system is likely to be relatively 
+close to face-on (an therefore nearly circularly polarised), which allows us to place prior 
+constraints on the system inclination angle. Beam opening angles for \acp{sGRB} are relatively 
+poorly constrained as they are based on only a few jet break detections, however 
+\cite{2014ApJ...780..118F} give a median opening angle value of $\sim 10^{\circ}$. If we assume 
+that the beam opening angle is a half-normal distribution, for which the median is given by 
+$\sqrt{2}\erf{}^{-1}(1/2)\sigma \sim 0.67\sigma$, then a median of $\sim 10^{\circ}$ corresponds to 
+$\sigma \sim 14.8^{\circ}$. We therefore place a Gaussian prior on the inclination angle with zero 
+mean and $\sigma=14.8^{\circ}$.
 
@@
 study, the chirp mass, which we see from equation~\ref{eq:signal} plays a role in the overall 
-signal amplitude, is very well constrained by the systems phase evolution, so our prior could be 
+signal amplitude, is very well constrained by the system's phase evolution, so our prior could be 
 expanded with minimal effect on the results. In the case of \ac{NSBH} systems we use the same 
@@
 
-For the spinning \ac{NSBH} systems we use a uniform prior on the dimensionless spin magnitude of 
-the black hole between $-1$ and 1.
-
 Finally, we require a prior on the calibration scale factors for each detector. We make the 
-assumption that in general the calibration will be correct, so want to use a prior that is peaked 
-at unity. We also assume that the probability density that the calibration scale factor is either 
-larger or smaller than unity by an equivalent factor, e.g.\ $s=10$ or $s=0.1$, is the same. For this 
-we therefore use a log-normal distribution as our prior
+assumption that in general the calibration applied to the detector data will be correct, so want to 
+use a prior that is peaked at unity. We also assume that the probability density that the 
+calibration scale factor is either larger or smaller than unity by an equivalent factor, e.g.\ 
+$s=10$ or $s=0.1$, is the same. For the scale factor for each detector we we therefore use a 
+log-normal distribution as our prior
 \begin{equation}
@@
 1.15$. With these parameters the prior probability density at 0.1 and 10 is one tenth of that at 
-1. Note that this prior does not provide produce a symmetric amount of probability about unity, but 
-given our simulation criterion described below in Section~\ref{sec:simulations} we find that the 
-likelihood generally overwhelms the prior. As $s$ is a scale factor we also have the further 
-constraint that it must be positive.
+1. Note that this prior does not give a symmetric amount of probability about unity, but given our 
+simulation criterion described below in Section~\ref{sec:simulations} we find that the likelihood 
+generally overwhelms the prior. As $s$ is a scale factor we also have the further constraint that 
+it must be positive.
 
@@
 range of distances (50\,Mpc to 500\,Mpc with 50\,Mpc increments for the \ac{BNS} signals and 
-100\,Mpc to 900\,Mpc with 100\,Mpc increments for the \ac{NSBH} signals). The simulations use the 
-two Advanced LIGO detectors (one at Hanford -- H1 -- and one at Livingston -- L1) and the Advanced 
-Virgo detector (V1). The simulations are all performed in the freqeuncy domain and span a frequency 
-range from 20\,Hz to 400\,Hz. We assume all detectors are expected to be operating at their design
-sensitivity (as given by Figure 1 of \cite{2013arXiv1304.0670L}). As such we add noise to each 
-simulation based on the sensitivity curve, but scaled with the associated calibration scale factor.
+100\,Mpc to 900\,Mpc with 100\,Mpc increments for the \ac{NSBH} signals). The simulations used the 
+two \ac{aLIGO} detectors (H1 and L1) and the \ac{AdV} detector (V1). The simulations were all 
+performed in the frequency domain and spanned a frequency range from 20\,Hz to 400\,Hz. We assumed
+all detectors were operating at their design sensitivity (as given by Figure~1 of 
+\cite{2013arXiv1304.0670L}). As such we added noise to each simulation based on the sensitivity 
+curve, but scaled with the associated calibration scale factor.
 
@@
 drawn from the prior distributions given in Section~\ref{sec:priors}, with the sky position being 
-drawn randomly from a uniform distribution, and with the scaling factors for each detector drawn 
-from a Gaussian distribution with a mean of one and standard deviation of $0.125$ (equivalent to a 
-mean fractional offset of 10\%). However, in accepting a proposed injection as one to be analysed 
-we introduced a criteria that the signal be ``detectable'', based on it having an SNR of $\geq 5.5$ 
-in at least two detectors as used in \cite{2012PhRvD..85h2002A} (as these would be signals 
-coincident with a GRB we do not include the often used further constraint that the total coherent 
-network SNR is greater than 12). The coalescence times for all injections were held fixed 
-within the centre of the prior window. This criteria means that as we go to larger distances our 
-population of injected sources will not be uniformly distributed over the sky, and will generally 
-be closer to being circularly polarised. For the \ac{NSBH} systems it also means that at larger 
-distances we will have more sources with larger mass ratios \MP{Check this}.
+drawn randomly from a uniform distribution, the coalescence times held fixed within the centre of 
+the prior window, and with the calibration scaling factors for each detector drawn from a Gaussian 
+distribution with a mean of one and standard deviation of $0.125$ (equivalent to a mean fractional 
+offset of 10\%). However, in accepting a proposed injection as one to be analysed we introduced a 
+criteria that the signal be ``detectable'', based on that used in \cite{2012PhRvD..85h2002A} of it 
+having an \ac{SNR} of $\geq 5.5$ in at least two detectors (as these would be signals coincident 
+with a \ac{GRB} we do not include the often used further constraint that the total coherent network 
+\ac{SNR} is greater than 12).  This criteria means that at larger distances our population of 
+injected sources was not be uniformly distributed over the sky, and were generally closer to being 
+circularly polarised.
 
-These simulations allow us to assess how well on average we would be able to constrain the 
-calibration scale for a given CBC-GRB coincidence.
+These simulations have allowed us to assess how well on average we would be able to constrain the 
+calibration scale for a given \ac{CBC}-\ac{GRB} coincidence.
 
@@
 factors (if they are independent of other parameters and have a roughly Gaussian probability 
-distribtuion)
-should be given by $\sim 1/{\rm SNR}$ for each detector. As we will show below this assumption is 
-reasonable, but small correlations do exist between parameters meaning that it does not completely 
-hold.
+distribution) should be given by $\sim 1/{\rm SNR}$ for each detector. As we will show below this 
+assumption is reasonable, but small correlations do exist between parameters meaning that it does 
+not completely hold.
 
@@
 scale factors can be determined for a three detector network if assuming \ac{BNS} systems (provided 
-a coincident GRB is observed and can yield a distance estimate). The boxplots span the lower to 
-upper quartile range of the distributions, with the median value shown as a horizontal line within 
-the box and the mean shown as a star. The dashed magenta line shows the percentage of source drawn 
-from the prior distribution that would be detectable at each distance value.}
+a coincident \ac{GRB} is observed and can yield a distance estimate). The boxplots span the lower 
+to upper quartile range of the distributions, with the median value shown as a horizontal line 
+within the box and the mean shown as a star. The dashed magenta line shows the percentage of source 
+drawn from the prior distribution that would be detectable at each distance value.}
 \end{figure}
@@
 scale factors can be determined for a three detector network if assuming \ac{NSBH} systems 
-(provided a coincident GRB is observed and can yield a distance estimate). The plot contents are 
-the same as 
-in figure~\ref{fig:bnsresults}.}
+(provided a coincident \ac{GRB} is observed and can yield a distance estimate). The plot contents 
+are the same as in figure~\ref{fig:bnsresults}.}
 \end{figure}
 
-In figure~\ref{fig:bnsresults} we see that unsurprisingly on average the scale factor can be 
-recovered to equivalent precision in both \ac{aLIGO} detectors, with uncertainties generally within 
-the 10\% range for sources at 100\,Mpc. An interesting feature that we see is that for distances 
-$\gtrsim 250$\,Mpc the upper extent of the uncertainty hits a maximum at $\sim 25\mbox{--}30\%$, 
-whilst the width of boxes narrows. This is due to our ``detectability'' criteria, in that for all 
-distances we will only be seeing those sources with high enough SNR that we would consider them 
-detectable. There will therefore be fewer sources with SNR higher than this criteria at large 
-distances, giving us a narrower range, and we automatically exclude those with SNR that are too 
-small thus truncating our uncertainty distribution at the upper range. However, this does show that 
-on average for sources that are detectable out to 450\,Mpc we would be able to constrain the 
-calibration scale factor uncertainty to $\sim 20\%$. As the detection criteria will generally come 
-from the SNR being high enough within the two \ac{aLIGO} instruments this means that the SNR in 
-\ac{AdV} can be small and thus the ability to constrain the scale factor for it becomes poor 
-(although it still provides information that the calibration is not grossly inaccurate). We also 
-see that true uncertainties achievable for recovering the calibration scale factors are indeed very 
-similar to the simple assumption that they would be given by $sim 1/{\rm SNR}$. Although, they are 
-probably slightly poorer due to their still be correlations with $\iota$ even within the 
-constrained prior range, and slight correlation between the scale factors for H1 and L1 as seen in 
-figure~\ref{fig:bnspost}.
+In figure~\ref{fig:bnsresults} we see that on average the scale factor can be recovered to 
+equivalent precision in both \ac{aLIGO} detectors, as would be expected, with uncertainties 
+generally within the 10\% range for sources at 100\,Mpc. This is comparable to previous estimates 
+of the calibration error in the initial LIGO detectors. An interesting feature is that for 
+distances $\gtrsim 250$\,Mpc the upper extent of the uncertainty hits a maximum at $\sim 
+25\mbox{--}30\%$, whilst the width of boxes narrows. This is due to our ``detectability'' criteria, 
+in that for all distances we will only be seeing those sources with a high enough \ac{SNR} that we 
+would consider them detectable. There will therefore be fewer sources with \ac{SNR} higher than 
+this criteria at large distances, giving us a narrower range, and we automatically exclude those 
+with \ac{SNR} that are too small thus truncating our uncertainty distribution at the upper end. 
+However, this does show that on average for sources that are detectable out to 450\,Mpc we would be 
+able to constrain the calibration scale factor uncertainty for the \ac{aLIGO} detectors to $\sim 
+20\%$. As the detection criteria will generally come from the \ac{SNR} being high enough within the 
+two \ac{aLIGO} instruments this means that the SNR in \ac{AdV} can be small and thus the ability to 
+constrain the scale factor for it becomes poor (although it still provides information that the 
+calibration is not grossly inaccurate). We also see that true uncertainties achievable for 
+recovering the calibration scale factors are indeed very similar to the simple assumption that they 
+would be given by $sim 1/{\rm SNR}$, i.e.\ at the upper end of the distribution the \ac{SNR} in 
+the two \ac{aLIGO} detectors will be $\sim 5.5$, which would be expected to produce uncertainties 
+of $\lesssim 20\%$. Our results are probably slightly poorer than the \ac{SNR} expectation due to 
+their still be correlations with $\iota$ even within the constrained prior range, and slight 
+correlations between the scale factors for H1 and L1 as seen in figure~\ref{fig:bnspost}.
 
-In figure~\ref{fig:nsbhresults} we see very similar results although the higher \ac{SNR} of the 
-signals means that we can provide $\sim 20\mbox{--}25\%$ uncertainties on the calibration scale 
-factors for H1 and L1 out to greater distances.
+In figure~\ref{fig:nsbhresults} we see very similar results for the \ac{NSBH} systems although the 
+higher \ac{SNR} of the signals means that we can provide $\sim 20\mbox{--}25\%$ uncertainties on 
+the calibration scale factors for H1 and L1 out to greater distances.
+
+\MP{Here's my current thinking regarding why the distributions for the NSBH systems are broader
+that those for the BNS systems. At distances where the distributions in SNR for each type of system 
+are equivalent for the NSBH systems the ``detectable'' systems can be produced
+by a broader range of parameter combinations, in particular chirp mass and $\iota$, this means
+that the particular correlations between these are the scale factors will mean that as a whole
+the distribution will be broader. I need to convince myself of this.}
 
@@
 \cm{What did we learn.}
-
-This project is only the beginning in testing this new method of calibration.
-More rigorous testing has to be carried out for a number of different systems.
-As mentioned above, a larger number of random sky positions have to be analysed
-to give better understanding of the resolution capabilities of the method. The
-current technique of calibration defines the error in the detector to be a
-complex function of frequency and phase of the GW. Instead of the simple scale
-factor used so far, a more realistic error can be tested to ensure that the
-estimation in its value is still comparable with current methods.
 

### 2015-08-05 src/python/simulation_check.py
simulation_check.py: allow script to just use sources that would have been detected

@@ -185,6 +185,10 @@ for drawing masses and setting spin (a single spin for the black hole will be us
   parser.add_option("-w", "--iotawidth", dest="iotawidth", type="float",
                       help="Width of iota simulation, and prior, distribution (degs)")
 
+  parser.add_option("-A", "--above-threshold", dest="abovethresh", default=False,
+                    action="store_true", help="If this flag is set only above detection \
+threshold events will be used")
+
   # parse input options
   (opts, args) = parser.parse_args()
 
@@ -260,30 +264,9 @@ for drawing masses and setting spin (a single spin for the black hole will be us
     m1mean = 1.35
     m1sigma = 0.13
 
-  m1list = []
-  m2list = []
-  for i in range(Nsim):
-    m1tmp = m1mean + m1sigma*np.random.randn()
-    if opts.nsbh:
-      while m1tmp < 2.5:
-        m1tmp = m1mean + m1sigma*np.random.randn()
-    m2tmp = m2mean + m2sigma*np.random.randn()
-    while m2tmp > m1tmp:
-      m2tmp = m2mean + m2sigma*np.random.randn()
-    m1list.append(m1tmp)
-    m2list.append(m2tmp)
-
-  m1 = np.array(m1list)
-  m2 = np.array(m2list)
-
-  indSNRs = []
-  netSNRs = []
-
   asds = []
   # create frequency series for PSDs
   psd = lal.CreateREAL8FrequencySeries('name', t0, 0., deltaF, lal.Unit(), 1+int(fmax/deltaF))
-
-  accepted = 0. # values above threshold
 
   for i in range(len(dets)):
     if dets[i] in ['H1', 'L1']:
@@ -308,6 +291,33 @@ for drawing masses and setting spin (a single spin for the black hole will be us
     else:
       asds.append(np.sqrt(psd.data.data))
 
+  m1list = []
+  m2list = []
+  for i in range(Nsim):
+    m1tmp = m1mean + m1sigma*np.random.randn()
+    if opts.nsbh:
+      while m1tmp < 2.5:
+        m1tmp = m1mean + m1sigma*np.random.randn()
+    m2tmp = m2mean + m2sigma*np.random.randn()
+    while m2tmp > m1tmp:
[…]

### 2015-08-06 src/python/results_script.py
Changed boxplots to extend from the 5th to 95th percentile

@@ -139,7 +139,7 @@ for dist in dists:
   positions.append(dist+10)
  
 #bp = pl.boxplot(data, notch=0, sym='+', positions=positions, widths=8)
-bp = pl.boxplot(data, notch=0, sym='', positions=positions, widths=8) 
+bp = pl.boxplot(data, whis=[5, 95], notch=0, sym='', positions=positions, widths=8) 
 
 pl.setp(bp['boxes'], color='black')
 pl.setp(bp['whiskers'], color='black')

### 2015-08-06 src/python/results_script_nsbh.py

@@ -137,7 +137,7 @@ for dist in dists:
   positions.append(dist)
   positions.append(dist+20)
 
-bp = pl.boxplot(data, notch=0, sym='', positions=positions, widths=16) 
+bp = pl.boxplot(data, whis=[5, 95], notch=0, sym='', positions=positions, widths=16) 
 #bp = pl.boxplot(data, notch=0, sym='x', positions=positions, widths=16)
 
 pl.setp(bp['boxes'], color='black')

### 2015-08-06 docs/cosmic_cal.tex
Added more to the document - still need to go through some more bits and check equation 6 is ok

@@
 \begin{abstract}
-  We present an investigation into the potential of calibrating \MP{change to: ``assessing the validity of the calibration''?}
-  gravitational wave detector outputs through the use of standard
+  We present an investigation into the potential of assessing the validity of the 
+calibration of gravitational wave detector outputs through the use of standard
   sirens. Such signals, as measured via gravitational wave
@@
 % introduce GWs and BNS systems
-It is expected that the advanced generation of interferometric \ac{GW}
-detectors will detect waves emitted from $O(10s)$ of compact binary
-coalescences.  One such class of these cataclysmic events, the inspiral and
-merger of \ac{BNS} systems will be detected out to a maximum range of $\approx
-450$ Mpc. Assuming the current best estimates for the cosmological parameters,
-this is equivalent to a redshift $z\approx 0.1$. As noted by \cite{1986Natur.323..310S} \ac{CBC}
-systems can be used as cosmological distance markers, otherwise known as
-``standard sirens'' (analogous to the \ac{EM} standard candles). Standard sirens
-allow ...
+It is expected that the advanced generation of interferometric \ac{GW} detectors will detect waves 
+emitted from $O(10s)$ of \acp{CBC}. One such class of these cataclysmic events, the inspiral and
+merger of \ac{BNS} systems will be detected out to a maximum range of $\approx 450$ Mpc. Assuming 
+the current best estimates for the cosmological parameters, this is equivalent to a redshift 
+$z\approx 0.1$. As noted by \cite{1986Natur.323..310S} \ac{CBC} systems can be used as cosmological 
+distance markers, otherwise known as ``standard sirens'' (analogous to the \ac{EM} standard 
+candles). Standard sirens allow ...
 
 % BNS systems are GRBs
-It is considered likely that the merger of \ac{BNS} systems, in addition to
-emitting detectable \acp{GW}, is also the mechanism for producing \acp{sGRB}.
-In this scenario these \ac{EM} events are produced through the ... and result
-in tightly beamed emission parallel to the orbital angular momentum vector of
-the \ac{BNS} system. Additional evidence for the coincidence of \acp{sGRB} with
-\ac{GW} events are the estimated astrophysical rates of both phenomena.
-Observations of \acp{sGRB} give a rate of XXX Mpc$^{-3}$yr$^{-1}$ which can be
-compared to estimates of \ac{BNS} merger rates of XXX-XXX Mpc$^{-3}$yr$^{-1}$.
-This latter range is obtained from population synthesis models together with
-knowledge of the X known \ac{BNS} systems in our galaxy.
+It is considered likely that the merger of \ac{BNS} systems, in addition to emitting detectable 
+\acp{GW}, is also the mechanism for producing \acp{sGRB}. In this scenario these \ac{EM} events are 
+produced through the ... and result in tightly beamed emission parallel to the orbital angular 
+momentum vector of the \ac{BNS} system. Additional evidence for the coincidence of \acp{sGRB} with
+\ac{GW} events are the estimated astrophysical rates of both phenomena. Observations of \acp{sGRB} 
+give a rate of XXX Mpc$^{-3}$yr$^{-1}$ which can be compared to estimates of \ac{BNS} merger rates 
+of XXX-XXX Mpc$^{-3}$yr$^{-1}$. This latter range is obtained from population synthesis models 
+together with knowledge of the X known \ac{BNS} systems in our galaxy.
 
 % talk about the calibration idea
-If a single \ac{GW} event is observed in coincidence with a \ac{sGRB} it will \MP{may?}
-be possible to identify the host galaxy of the \ac{BNS} source.  With this
-information a spectroscopic redshift can be very accurately obtained. With
-knowledge of the redshift, using current best estimates for the Hubble constant
-and other cosmological parameters, the true luminosity distance can be
-estimated to $\sim 1\%$ accuracy.  The \ac{GW} measurement acts as a standard
-siren also giving us a direct measurement of the luminosity distance to the source.
-The accuracy of such a measurement depends on a number of factors including the
-accuracy with which the \ac{GW} detector has been calibrated.  Hence comparison
-with the distance estimate from the \ac{sGRB} we can re-calibrate (or validate)
-the existing experimentally obtained calibration.
+If a single \ac{GW} event is observed in coincidence with a \ac{sGRB} it may be possible to 
+identify the host galaxy of the \ac{BNS} source.  With this information a spectroscopic redshift 
+can be very accurately obtained. With knowledge of the redshift, using current best estimates for 
+the Hubble constant and other cosmological parameters, the true luminosity distance can be
+estimated to $\sim 1\%$ accuracy.  The \ac{GW} measurement acts as a standard siren also giving us 
+a direct measurement of the luminosity distance to the source. The accuracy of such a measurement 
+depends on a number of factors including the accuracy with which the \ac{GW} detector has been 
+calibrated.  Hence comparison with the distance estimate from the \ac{sGRB} we can re-calibrate (or 
+validate) the existing experimentally obtained calibration.
 
 % What do we do in this paper
-In this paper we investigate the feasibility of this approach for single
-coincident \ac{GW}--\ac{sGRB} events and establish the validation power of such
-a calibration technique.  In Sec.~\ref{sec:calibration} we briefly summarize
-the existing experimental technique for \ac{GW} detector calibration and its
-expected accuracy.  We then review the concept of \ac{GW} standard sirens and
-their proposed coincident \ac{EM} signatures, the \acp{sGRB} in
-Secs.~\ref{sec:sirens} and~\ref{sec:GRB}. In Sec.~\ref{sec:single} we present
-the results of Monte-Carlo simulations for the purpose of luminosity distance
-estimation using \ac{GW} events. In Sec.~\ref{sec:cosmo} we show the likely
-accuracy of \ac{sGRB} distances obtained from their host galaxy redshift and in
-Sec.~\ref{sec:discussion} we conclude.
+In this paper we investigate the feasibility of this approach for single coincident 
+\ac{GW}--\ac{sGRB} events and establish the validation power of such a calibration technique.  In 
+Sec.~\ref{sec:calibration} we briefly summarize the existing experimental technique for \ac{GW} 
+detector calibration and its expected accuracy.  We then review the concept of \ac{GW} standard 
+sirens and their proposed coincident \ac{EM} signatures, the \acp{sGRB} in
+Secs.~\ref{sec:sirens} and~\ref{sec:GRB}. In Sec.~\ref{sec:cosmo} we show the likely accuracy of 
+\ac{sGRB} distances obtained from their host galaxy redshift and in Sec.~\ref{sec:discussion} we 
+conclude.
 
@@
 \section{Gravitational wave detector calibration\label{sec:calibration}}
-
-\cm{We need to define our model for the calibration, i.e. are we talking
-about amplitude only and are we breaking this up over frequency.}
-
-\MP{In this paper I think we stick to just talking about a single overall amplitude calibration scale.}
 
@@
 using the `hardware injection method' is roughly $10\%$ \cite{Vitale:2012}. This is a benchmark set 
-for the error estimation using the proposed new method in this project.~\cm{Clean this up and 
-expand.}
+for the error estimation using the proposed new method in this project.
 
@@
 \end{equation}
-where where we assume $R$ varies only slowly in time (in comparison to transient signal timescales).
-Calculation of $R$ requires the measurement of various functions within a control feedback loop
-(see \cite{Vitale:2012} for a brief overview of this), which are subject to uncertainties. In this 
+where where we assume $R$ varies only slowly in time (in comparison to transient signal 
+time-scales). Calculation of $R$ requires the measurement of various functions within a control 
+feedback loop (see \cite{2010NIMPA.624..223A}, which are subject to uncertainties. In this 
 study we will assume an estimate of $R$ is available (although in theory we could take on the role 
@@
 $s$, so that
-\begin{equation}
+\begin{equation}\label{eq:scalefactor}
 sh(f) = h_{\rm m}(f) = \frac{R(f,t) e(f)}{L},
@@
 in reality\footnote{However, the \ac{SNR} would be the correct value as the signal and the noise 
-will be contain the scale factor}, whereas if $s < 1$ it would appear to have a lower amplitude 
+will both contain the scale factor}, whereas if $s < 1$ it would appear to have a lower amplitude 
 than reality. In this analysis we will simplify the situation by assuming that $s$ is a constant 
 with respect to frequency, but in future studies that assumption could be dropped and $s$ could 
-take some functional form, or piece-wise fit, with respect to $f$.
+take some functional form, or piece-wise fit, with respect to $f$ (e.g.\ in a similar way to the 
+method of \cite{2013PhRvD..88h4044L} used for fitting differences in \ac{PSD} estimates).
 
@@
 The most likely scenario in which a coincident \ac{GW}--\ac{sGRB} event would be identified is 
-through the targeted follow-up of an observed \ac{sGRB} or by post-facto matching of \ac{GW} 
+through the targeted follow-up of an observed \ac{GW} or by post-facto matching of \ac{GW} 
 trigger lists with known \ac{sGRB} events.  The likelihood of being able to follow-up \ac{GW} 
@@
 
-\cm{This should be the simplest case we can do.  We need to cut a lot of this.}
-
-We want to assess how well the calibration scale factor can be estimated from a single observed 
-\ac{GW} associated with a particular \ac{sGRB}. As we {\it a priori} have no knowledge of the 
-likely location and distance of such an event we have performed simulations of multiple events to 
-see the expected distribution the accuracy of the calibration scale factor recovery.
+We want to assess how well the calibration scale factor defined in Eqn.~\ref{eq:scalefactor} can be 
+estimated from a single observed \ac{GW} associated with a particular \ac{sGRB}. As we {\it a 
+priori} have no knowledge of the likely location and distance of such an event we have performed 
+simulations of multiple events to see the expected distribution in the accuracy of the calibration 
+scale factor recovery.
 
@@
 %
-where $F_{+},F_{\times}$ are the antenna response functions which are dependent upon the 
+where $F_{+}$ and $F_{\times}$ are the antenna response functions which are dependent upon the 
 polarisation angle $\psi$ and the sky position of the source $\alpha$ and $\delta$.
@@
 % Define the detectors and the noise epoch we will be using
-For this analysis we will consider the advanced generation LIGO and Virgo detectors operating
-at their design sensitivity, with noise power spectral densities taken from
+For this analysis we will consider the advanced generation \ac{aLIGO} and \ac{AdV} detectors 
+operating at their design sensitivity, with noise power spectral densities taken from
 \cite{2013arXiv1304.0670L}. The network we consider is the two \ac{aLIGO} detectors (H1 and L1) and 
@@
 \begin{equation}
-  \tilde{d}_{k}(f)=\tilde{n}_{k}(f)+s\gws_{k}(f,\btheta)
+  \tilde{d}_{k}(f)=s(\tilde{n}_{k}(f)+\gws_{k}(f,\btheta))
 \end{equation}
 %
-where $\tilde{n}(f)$ is the Fourier transform of Gaussian noise drawn from a distribution 
-consistent with the \ac{PSD} of the advanced detector noise.
+where $\tilde{n}(f)$ is the Fourier transform of the true (Gaussian) strain noise drawn from a 
+distribution consistent with the \ac{PSD} of the advanced detector noise and $\btheta$ is the set 
+of waveform parameters.
 
@@
 To assess the ability to estimate the unknown calibration scale factors we want to calculate their 
-marginal posterior probability distribution functions (pdf) given the data contains a signal. We 
-use Bayes' theorem for which we need to define a likelihood function and prior pdfs on the signal 
-parameters being estimated. In this analysis we use the Markov chain Monte Carlo code {\tt emcee} 
-\cite{2013PASP..125..306F} to sample the posterior distribution. We use a Gaussian likelihood 
-function given by
+marginal posterior probability distribution functions (pdf). We use Bayes' theorem for which we 
+need to define a likelihood function and prior pdfs on the signal parameters being estimated. In 
+this analysis we use the Markov chain Monte Carlo code {\tt emcee} \cite{2013PASP..125..306F} to 
+sample the posterior distribution. We use a Gaussian likelihood function given by
 %
-%\begin{equation}
-%  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp\Bigg[ 2\Delta
-%f\sum_{k>0}\frac{|\tilde{d}(f_{k}) - \tilde{h}(f_{k})|^{2}}{S(f_{k})}\Bigg]
-%\end{equation}
 \begin{equation}
-\fl p(\bi{d} | \btheta, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm det}}
+\fl p(\bi{d} | \btheta, \bi{s}, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm det}}
   \sum_{i=i_{f_{\rm low}}}^{i_{f_{\rm high}}}\frac{\Re{\left\{ \tilde{d}^{*}_{k,i} 
@@
 canonical masss distribution used for the rate results in \cite{2012PhRvD..85h2002A} with a mean of 
-$5\,\textrm{M}_{\odot}$ and a standard deviation of $1\,\textrm{M}_{\odot}$. For the \ac{NSBH}
-systems we also require a prior on the black hole spin (for \ac{BNS} systems we have assumed that 
-the spins are small enough that they will be negligible). We use a uniform prior on the 
-normalised aligned spin magnitude between $-1$ and $1$.
+$5\,\textrm{M}_{\odot}$ and a standard deviation of $1\,\textrm{M}_{\odot}$. We note that the mass 
+range of black holes could be quite different from this, but use this range as an example for this 
+type of system. For the \ac{NSBH} systems we also require a prior on the black hole spin (for 
+\ac{BNS} systems we have assumed that the spins are small enough that they will be negligible). We 
+use a uniform prior on the normalised aligned spin magnitude between $-1$ and $1$.
 
@@
 % Describe the ensemble of simulations used
-Separately for the \ac{BNS} and \ac{NSBH} systems we have performed up to 2000 simulations at each 
-distance value\footnote{Simulations on which our MCMC did not produce converged chains within our 
-analysis time were discarded and not re-run, although there were always at least 1000 simulations 
-that completed.}. For proposing injections for each simulation the source parameters were randomly 
-drawn from the prior distributions given in Section~\ref{sec:priors}, with the sky position being 
-drawn randomly from a uniform distribution, the coalescence times held fixed within the centre of 
-the prior window, and with the calibration scaling factors for each detector drawn from a Gaussian 
-distribution with a mean of one and standard deviation of $0.125$ (equivalent to a mean fractional 
-offset of 10\%). However, in accepting a proposed injection as one to be analysed we introduced a 
-criteria that the signal be ``detectable'', based on that used in \cite{2012PhRvD..85h2002A} of it 
-having an \ac{SNR} of $\geq 5.5$ in at least two detectors (as these would be signals coincident 
-with a \ac{GRB} we do not include the often used further constraint that the total coherent network 
-\ac{SNR} is greater than 12).  This criteria means that at larger distances our population of 
-injected sources was not be uniformly distributed over the sky, and were generally closer to being 
-circularly polarised.
+Separately for the \ac{BNS} and \ac{NSBH} systems we have performed a large number simulations at 
+each distance value. For proposing injections for each simulation the source parameters were 
+randomly drawn from the prior distributions given in Section~\ref{sec:priors}, with the sky 
+position being drawn randomly from a uniform distribution, the coalescence times held fixed within 
+the centre of the prior window, and with the calibration scaling factors for each detector drawn 
+from a Gaussian distribution with a mean of one and standard deviation of $0.125$ (equivalent to a 
+mean fractional offset of 10\%). However, in accepting a proposed injection as one to be analysed 
+we introduced a criteria that the signal be ``detectable'', based on that used in 
+\cite{2012PhRvD..85h2002A} of it having an \ac{SNR} of $\geq 5.5$ in at least two detectors (as 
+these would be signals coincident with a \ac{sGRB} we do not include the often used further 
+constraint that the total coherent network \ac{SNR} is greater than 12). This criteria means that 
+at larger distances our population of injected sources was not uniformly distributed over the sky, 
+and were generally closer to being circularly polarised.
 
@@
 separate noisy \ac{PSD} estimates and scaled with the same factor as applied to the injection and 
-noise\footnote{We do not account for there being a difference between the estimated \ac{PSD} and the 
-actual \ac{PSD} of the analysed section of data. This difference would be very much correlated with 
-the scale factor, so in reality our estimate of the scale factor would be a combination of the 
+noise\footnote{We do not account for there being a potential difference between the estimated 
+\ac{PSD} and the actual \ac{PSD} of the analysed section of data as described in e.g.\ 
+\cite{2013PhRvD..88h4044L}. This difference would be very highly correlated with the calibration 
+scale factor, so in reality our estimate of the scale factor would be a combination of the
 calibration offset and the difference in the \ac{PSD}. As such our results on real data would be an 
@@
 
-From these posterior distributions we have calculated the 68\% credible region on the 
-calibration scale factors for each detector (if these were Gaussian distributions this is 
-equivalent to the region bounded by $1\sigma$ intervals). For all the signals at each distance 
-increment we have produced the distribution of the fractional half widths (i.e.\ $1\sigma$) of 
-these scale factor's confidence intervals compared to the actual value. These are shown as boxplots 
-in figures~\ref{fig:bnsresults} and \ref{fig:nsbhresults} for the \ac{BNS} and \ac{NSBH} systems 
-respectively. The boxes show the extent from the lower to upper quartile of the values, whilst the 
-whiskers represent the range of values. The black line within each box gives the median value and 
-the star gives the mean value. Also shown on each plot as the dashed magenta line is the percentage 
-of signal drawn from the prior distribution that would fulfill the detection criterion.
+From these posterior distributions we have calculated the 68\% credible region on the calibration 
+scale factors for each detector (if these were Gaussian distributions this is equivalent to the 
+region either side of the mean bounded by the $1\sigma$ intervals). For all the signals at each 
+distance increment we have produced the distribution of the fractional half widths (i.e.\ $1\sigma$) 
+of these scale factors' confidence intervals compared to the actual value. These are shown as 
+boxplots in figures~\ref{fig:bnsresults} and \ref{fig:nsbhresults} for the \ac{BNS} and \ac{NSBH} 
+systems respectively. The boxes show the extent from the lower to upper quartile of the values, 
+whilst the whiskers extend from the 5$^{\rm th}$ to 95$^{\rm th}$ percentile. The black line 
+within each box gives the median value and the star gives the mean value. Also shown on each plot as 
+the dashed magenta line is the percentage of signals drawn from the prior distribution that fulfill 
+the detection criterion.
 
@@
  \end{center}
- \caption{\label{fig:bnsresults} Distribution of the percentage accuracy at which the calibration 
+ \caption{\label{fig:bnsresults} Distributions of the percentage accuracy at which the calibration 
 scale factors can be determined for a three detector network if assuming \ac{BNS} systems (provided 
@@
  \end{center}
- \caption{\label{fig:nsbhresults} Distribution of the percentage accuracy at which the calibration 
+ \caption{\label{fig:nsbhresults} Distributions of the percentage accuracy at which the calibration 
 scale factors can be determined for a three detector network if assuming \ac{NSBH} systems 
@@
 of the calibration error in the initial LIGO detectors. An interesting feature is that for 
-distances $\gtrsim 250$\,Mpc the upper extent of the uncertainty hits a maximum at $\sim 
-25\mbox{--}30\%$, whilst the width of boxes narrows. This is due to our ``detectability'' criteria, 
+distances $\gtrsim 250$\,Mpc the upper extent of the uncertainty for H1 and L1 hits a maximum at 
+$\sim 25\%$, whilst the width of boxes narrows. This is due to our ``detectability'' criteria, 
 in that for all distances we will only be seeing those sources with a high enough \ac{SNR} that we 
@@
 higher \ac{SNR} of the signals means that we can provide $\sim 20\mbox{--}25\%$ uncertainties on 
-the calibration scale factors for H1 and L1 out to greater distances.
+the calibration scale factors for H1 and L1 out to greater distances. The distributions generally 
+appear slightly broader than those for the \ac{BNS} systems. This may be entirely explicable 
+through two factors. The first factor down to the fact that including the spin parameter leads to 
+string correlations between the chirp mass, mass ratio and spin. These strong correlations make 
+convergence of the MCMC take longer and means we have fewer independent samples to estimate the 
+posterior distributions with. This leads to a larger statistical fluctiation on the results. The 
+second factor is that there appears to be a modest effect due to the population of sources that 
+are detectable. For distances at which the \ac{BNS} and \ac{NSBH} systems would give the same 
+percentage of obsverable sources there is a slight increase in the mean and standard distribution of 
+\acp{SNR} of the \ac{NSBH} systems over the \ac{BNS} systems.
 
-\MP{Here's my current thinking regarding why the distributions for the NSBH systems are broader
-that those for the BNS systems. At distances where the distributions in SNR for each type of system 
-are equivalent for the NSBH systems the ``detectable'' systems can be produced
-by a broader range of parameter combinations, in particular chirp mass and $\iota$, this means
-that the particular correlations between these are the scale factors will mean that as a whole
-the distribution will be broader. I need to convince myself of this.}
+As noted previously the mass distribution for the black holes could be quite different from the one 
+we used, with masses extending to $\gtrsim 10$\,M$_{\odot}$. We expect these would produce similar 
+results to those we see in figure~\ref{fig:nsbhresults}, but again extending to further distances. 
+Introducing the merger and ring-down phase, which we currently ignore, may also help increase 
+\ac{SNR}.
 
@@
 
-\cm{What did we learn.}
+We have attempted to see how well we can assess the detector calibration for \ac{aLIGO} and 
+\ac{AdV} using astrophysical sources. To do this we require that a \ac{sGRB}, for which it has been 
+possible to measure the distance, is observed in coincidence with a \ac{CBC} signal in the \ac{GW} 
+detectors. This enables us to assume a known distance to the source, and also limit the inclination 
+of the source, which in turn allows us to test the consistency of the detector calibration. We do 
+this by including an unknown scale factor on the true signal detector data, which we estimate given 
+the data. We find that for detectable \ac{BNS} sources the uncertainty of on the calibration scale 
+factor could on average be determined to $\lesssim 10\%$ of its true value for the \ac{aLIGO} 
+and \ac{AdV} detectors out to 100\,Mpc. This is comparible to the hoped for level of calibration 
+accuracy of the detectors. For sources at the standard single detector \ac{BNS} horizon distance of 
+$\sim 450$\,Mpc the scale factor could on average be determined to within $\lesssim 20\%$ of its 
+true value for the \ac{aLIGO} detectors. Similar results were found for \ac{NSBH} sources, although 
+sources could be observed out to higher distances.
 
-For applications of this calibration method, hypothesis testing could be
-carried out. Testing different hypothesis is the technique used in determining
-a GW signal in a data-set when compared to a noise-only data-set. Being able to
-show with confidence that the data contains a signal is the primary focus of GW
-research at the moment. Therefore if uncertainty can be removed from the
-analysis this will ensure a more confident detection. Using this new method is
-resolving and removing the uncertainty is an area which has to be explored.
+The requirement of a coincident \ac{sGRB} with a known distance means that there will be 
+considerable latency for this calibration assessment, but importantly this method would provide 
+semi-independent consistency check that the calibration is reasonably accurate. We do note that this 
+method has to make its own assumptions, including that the waveforms we use accurately reflect the 
+signal and the calibration, the phase calibration does not change to much over frequency and that 
+the overall calibration does not change considerably over the length of a signal. These are studies 
+that could be incorporated in the future.
 
-\MP{We emphasise that our work has assumed that the PSD used in the likelihood function
+We imagine that a test for consistency of calibration could be performed with an ensemble of 
+\ac{CBC} sources, although such a method may have to marginalise over highly uncertain population 
+models and may give quite weak bounds on the overall calibration.
+
+We emphasise that our work has assumed that the PSD used in the likelihood function
 when estimating the calibration scale factor is an accurate representation of the
@@
 with any uncertainty in the PSD estimate, but would still offer an upper limit
-on the calibration uncertainty.}
+on the calibration uncertainty.
 
-\MP{The BayesLine algorithm \cite{2015PhRvD..91h4034L} may provide a natural
-way to perform this analysis in particular when accounting for a frequency
-dependent calibration.}
+A further extension of this work would be for the calibration scale factor to be estimated as a 
+function of frequency. This would require the frequency series to be broken up into a set of 
+sections and the calibration scale factor estimated for each. However, the scale factor uncertainty 
+will obviously be related to the \ac{SNR} of a signal within a particular section. It could 
+therefore be a job of Bayesian model selection to work out the optimal number of frequency 
+sections to estimate the scale factor for. The BayesLine algorithm \cite{2015PhRvD..91h4034L}, or 
+the method described in \cite{2013PhRvD..88h4044L}, may provide a natural way to perform such an
+analysis. Indeed these methods, and those of \cite{Vitale:2012} are already being used for 
+marginalising over calibration uncertainties, but do not attempt to estimate it.
 

### 2015-08-07 docs/cosmic_cal.tex
Some more updates to the document

@@
 \newcommand{\gws}{\tilde{h}}
+\newcommand{\scf}{\ensuremath{\mathcal{C}}}
 
@@
 \acrodef{PSD}[PSD]{power spectral density}
+\acrodef{MCMC}[MCMC]{Markov chain Monte Carlo}
 
@@
 of estimating $R$ itself), but that it differs from the truth through some unknown scale factor, 
-$s$, so that
+$\scf$, so that
 \begin{equation}\label{eq:scalefactor}
-sh(f) = h_{\rm m}(f) = \frac{R(f,t) e(f)}{L},
+\scf h(f) = h_{\rm m}(f) = \frac{R(f,t) e(f)}{L},
 \end{equation}
 where $h(f)$ is the true strain and $h_{\rm m}(f)$ is the measured strain. With this definition it
-means that if $s > 1$ then a signal would appear to have a larger amplitude (e.g.\ be closer) than 
-in reality\footnote{However, the \ac{SNR} would be the correct value as the signal and the noise 
-will both contain the scale factor}, whereas if $s < 1$ it would appear to have a lower amplitude 
-than reality. In this analysis we will simplify the situation by assuming that $s$ is a constant 
-with respect to frequency, but in future studies that assumption could be dropped and $s$ could 
-take some functional form, or piece-wise fit, with respect to $f$ (e.g.\ in a similar way to the 
-method of \cite{2013PhRvD..88h4044L} used for fitting differences in \ac{PSD} estimates).
+means that if $\scf > 1$ then a signal would appear to have a larger amplitude (e.g.\ be 
+closer) than in reality\footnote{However, the \ac{SNR} would be the correct value as the signal and 
+the noise will both contain the scale factor}, whereas if $\scf < 1$ it would appear to have a 
+lower amplitude than reality. In this analysis we will simplify the situation by assuming that 
+$\scf$ is a constant with respect to frequency, but in future studies that assumption could be 
+dropped and $\scf$ could take some functional form, or piece-wise fit, with respect to $f$ (e.g.\ 
+in a similar way to the method of \cite{2013PhRvD..88h4044L} used for fitting differences in 
+\ac{PSD} estimates).
 
@@
 \begin{equation}
-  \tilde{d}_{k}(f)=s(\tilde{n}_{k}(f)+\gws_{k}(f,\btheta))
+  \tilde{d}_{k}(f)=\scf_{k}(\tilde{n}_{k}(f)+\gws_{k}(f,\btheta))
 \end{equation}
 %
-where $\tilde{n}(f)$ is the Fourier transform of the true (Gaussian) strain noise drawn from a 
-distribution consistent with the \ac{PSD} of the advanced detector noise and $\btheta$ is the set 
-of waveform parameters.
+where $\tilde{n}(f)$ is the Fourier transform of the true (in our case Gaussian) strain noise drawn 
+from a, $\btheta$ is the set of waveform parameters and $\scf_k$ is the calibration scale factor for 
+the $k^{\rm th}$ detector.
 
@@
 \begin{equation}
-\fl p(\bi{d} | \btheta, \bi{s}, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm det}}
+\fl p(\bi{d} | \btheta, \vec{\scf}, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm 
+det}}
   \sum_{i=i_{f_{\rm low}}}^{i_{f_{\rm high}}}\frac{\Re{\left\{ \tilde{d}^{*}_{k,i} 
-\gws_{k,i}(\btheta) - \frac{1}{2}\left(\gws^*_{k,i}(\btheta)\gws_{k,i}(\btheta) + 
+\scf_{k}\gws_{k,i}(\btheta) - \frac{1}{2}\left(\scf_k^2 \gws^*_{k,i}(\btheta)\gws_{k,i}(\btheta) + 
 \tilde{d}^*_{k,i}\tilde{d}_{k,i}\right) \right\}}}{S_{k,i}}\right]}
@@
 where $\bi{d}$ is an array containing the data for all detectors, $\Delta f$ is the frequency 
-resolution, $S_k$ is the $k^{\rm th}$ detector's one sided noise \ac{PSD}, and the $i$ indices 
-increment over frequency with $i_{f_{\rm low}}$ and $i_{f_{\rm high}}$ corresponding to the lower 
+resolution, $S_k$ is the $k^{\rm th}$ detector's measured one sided noise \ac{PSD} (and as such 
+will contain the effect of the calibration scale factor), and the $i$ indices increment over 
+frequency with $i_{f_{\rm low}}$ and $i_{f_{\rm high}}$ corresponding to the lower 
 and upper range in frequencies used for our analysis, which were 20\,Hz to 400\,Hz respectively. 
@@
 we use can be found within this range, so going to lower or higher frequencies provides very little 
-additional information. The source parameters, and the unknown scale factors, have been subsumed 
-into a vector $\btheta$.
+additional information. The source parameters have been subsumed into a vector $\btheta$ and the 
+calibration scale factors are within the vector $\vec{\scf}$.
 
@@
 
-The primary reason why we can use coincident observations with a \ac{sGRB} as a potential check of 
-detector calibration is that the \ac{GRB} allows us to constrain the priors on various parameters 
-that would normally be highly correlated with the signal amplitude. The most important of these is 
-that the \ac{GRB} can provide a very tight constraint on the source distance\footnote{For \ac{BNS} 
+The primary reason why we can use coincident observations with a \ac{sGRB} as a check of detector 
+calibration is that the \ac{sGRB} allows us to constrain the priors on various parameters that 
+would normally be highly correlated with the signal amplitude. The most important of these is 
+that the \ac{sGRB} can provide a very tight constraint on the source distance\footnote{For \ac{BNS} 
 and \ac{NSBH} systems they will be observable only in the relatively nearby Universe ($z \sim 
 0.1$ and $z\sim 0.2$ for \ac{BNS} and \ac{NSBH} systems respectively), so the relation between the 
-measured GRB redshift and distance is simple and our luminosity distance will be negligibly 
+measured \ac{sGRB} redshift and distance is simple and our luminosity distance will be negligibly 
 effected by the choice of cosmology \MP{Check this?}.}. In this analysis we therefore assume that 
 the source distance is known (i.e.\ has a $\delta$-function prior). Another piece of information 
-that we can use from having a coincident \ac{GRB} is that the system is likely to be relatively 
-close to face-on (an therefore nearly circularly polarised), which allows us to place prior 
+that we can use from having a coincident \ac{sGRB} is that the system is likely to be relatively 
+close to face-on (and therefore nearly circularly polarised), which allows us to place prior 
 constraints on the system inclination angle. Beam opening angles for \acp{sGRB} are relatively 
@@
 
-The final important piece of information that we can make use of from the GRB observation is the 
-sky position of the source, which we take to be known precisely.
+The final important piece of information that we can make use of from the \ac{sGRB} observation is 
+the sky position of the source, which we take to be known precisely.
 
@@
 recorded time of the signal that would be returned by the detection pipeline. For the reference 
-phase a uniform prior between 0 and $2\pi$ is used, and for the polarisation angle a uniform prior 
+phase a uniform prior between 0 and $\pi$ is used, and for the polarisation angle a uniform prior 
 between 0 and $\pi/2$ is used. As the signals we use are generally close to being circularly 
@@
 \begin{equation}
- p(s|I) = \frac{1}{s\sigma\sqrt{2\pi}}\exp{\left( -\frac{(\ln{s} - \mu)^2}{2\sigma^2} \right)},
+ p(\scf|I) = \frac{1}{\scf\sigma\sqrt{2\pi}}\exp{\left( -\frac{(\ln{\scf} - \mu)^2}{2\sigma^2} 
+\right)},
 \end{equation}
@@
 simulation criterion described below in Section~\ref{sec:simulations} we find that the likelihood 
-generally overwhelms the prior. As $s$ is a scale factor we also have the further constraint that 
-it must be positive.
+generally overwhelms the prior. As $\scf$ is a scale factor we also have the further constraint 
+that it must be positive.
 
@@
 Separately for the \ac{BNS} and \ac{NSBH} systems we have performed a large number simulations at 
-each distance value. For proposing injections for each simulation the source parameters were 
+each distance value. When proposing injections for each simulation the source parameters were 
 randomly drawn from the prior distributions given in Section~\ref{sec:priors}, with the sky 
-position being drawn randomly from a uniform distribution, the coalescence times held fixed within 
-the centre of the prior window, and with the calibration scaling factors for each detector drawn 
-from a Gaussian distribution with a mean of one and standard deviation of $0.125$ (equivalent to a 
-mean fractional offset of 10\%). However, in accepting a proposed injection as one to be analysed 
+position being drawn randomly from a uniform distribution on the sky, the coalescence times held 
+fixed within the centre of the prior window, and with the calibration scaling factors for each 
+detector drawn from a Gaussian distribution with a mean of one and standard deviation of $0.125$ 
+(equivalent to a mean fractional offset of 10\%). However, in accepting a proposed injection as one 
+to be analysed 
 we introduced a criteria that the signal be ``detectable'', based on that used in 
@@
 These simulations have allowed us to assess how well on average we would be able to constrain the 
-calibration scale for a given \ac{CBC}-\ac{GRB} coincidence.
+calibration scale for a given \ac{CBC}-\ac{sGRB} coincidence.
 
@@
 
-For each of the simulated sources we have used the {\tt emcee} python MCMC package 
+For each of the simulated sources we have used the {\tt emcee} python \ac{MCMC} package 
 \cite{2013PASP..125..306F} to perform parameter estimation over the unknown source parameters using 
@@
 
-From these posterior distributions we have calculated the 68\% credible region on the calibration 
-scale factors for each detector (if these were Gaussian distributions this is equivalent to the 
-region either side of the mean bounded by the $1\sigma$ intervals). For all the signals at each 
-distance increment we have produced the distribution of the fractional half widths (i.e.\ $1\sigma$) 
-of these scale factors' confidence intervals compared to the actual value. These are shown as 
-boxplots in figures~\ref{fig:bnsresults} and \ref{fig:nsbhresults} for the \ac{BNS} and \ac{NSBH} 
-systems respectively. The boxes show the extent from the lower to upper quartile of the values, 
-whilst the whiskers extend from the 5$^{\rm th}$ to 95$^{\rm th}$ percentile. The black line 
-within each box gives the median value and the star gives the mean value. Also shown on each plot as 
-the dashed magenta line is the percentage of signals drawn from the prior distribution that fulfill 
-the detection criterion.
+From these posterior distributions we have calculated the minimal 68\% credible region on the 
+calibration scale factors for each detector (if these were Gaussian distributions this is 
+equivalent to the region either side of the mean bounded by the $1\sigma$ intervals). For all the 
+signals at each distance increment we have produced the distribution of the fractional half widths 
+(i.e.\ $1\sigma$) of these scale factors' confidence intervals compared to the actual value. These 
+are shown as boxplots in figures~\ref{fig:bnsresults} and \ref{fig:nsbhresults} for the \ac{BNS} 
+and \ac{NSBH} systems respectively. The boxes show the extent from the lower to upper quartile of 
+the values, whilst the whiskers extend from the 5$^{\rm th}$ to 95$^{\rm th}$ percentile. The black 
+line within each box gives the median value and the star gives the mean value. Also shown on each 
+plot as the dashed magenta line is the percentage of signals drawn from the prior distribution that 
+fulfil the detection criterion.
 
@@
 recovering the calibration scale factors are indeed very similar to the simple assumption that they 
-would be given by $sim 1/{\rm SNR}$, i.e.\ at the upper end of the distribution the \ac{SNR} in 
+would be given by $\sim 1/{\rm SNR}$, i.e.\ at the upper end of the distribution the \ac{SNR} in 
 the two \ac{aLIGO} detectors will be $\sim 5.5$, which would be expected to produce uncertainties 
@@
 appear slightly broader than those for the \ac{BNS} systems. This may be entirely explicable 
-through two factors. The first factor down to the fact that including the spin parameter leads to 
-string correlations between the chirp mass, mass ratio and spin. These strong correlations make 
-convergence of the MCMC take longer and means we have fewer independent samples to estimate the 
-posterior distributions with. This leads to a larger statistical fluctiation on the results. The 
-second factor is that there appears to be a modest effect due to the population of sources that 
+through two factors. The first factor is from the fact that including the spin parameter leads to 
+strong correlations between the chirp mass, mass ratio and spin. These strong correlations make 
+convergence of the \ac{MCMC} take longer and means we have fewer independent samples to estimate 
+the posterior distributions with. This leads to a larger statistical fluctuation on the results. 
+The second factor is that there appears to be a modest effect due to the population of sources that 
 are detectable. For distances at which the \ac{BNS} and \ac{NSBH} systems would give the same 
-percentage of obsverable sources there is a slight increase in the mean and standard distribution of 
-\acp{SNR} of the \ac{NSBH} systems over the \ac{BNS} systems.
+percentage of observable sources there is a slight increase in the mean and standard distribution 
+of \acp{SNR} of the \ac{NSBH} systems over the \ac{BNS} systems.
 
@@
 results to those we see in figure~\ref{fig:nsbhresults}, but again extending to further distances. 
-Introducing the merger and ring-down phase, which we currently ignore, may also help increase 
+Including the merger and ring-down phase, which we currently ignore, may also help increase 
 \ac{SNR}.
@@
 We have attempted to see how well we can assess the detector calibration for \ac{aLIGO} and 
-\ac{AdV} using astrophysical sources. To do this we require that a \ac{sGRB}, for which it has been 
-possible to measure the distance, is observed in coincidence with a \ac{CBC} signal in the \ac{GW} 
-detectors. This enables us to assume a known distance to the source, and also limit the inclination 
-of the source, which in turn allows us to test the consistency of the detector calibration. We do 
-this by including an unknown scale factor on the true signal detector data, which we estimate given 
-the data. We find that for detectable \ac{BNS} sources the uncertainty of on the calibration scale 
-factor could on average be determined to $\lesssim 10\%$ of its true value for the \ac{aLIGO} 
-and \ac{AdV} detectors out to 100\,Mpc. This is comparible to the hoped for level of calibration 
-accuracy of the detectors. For sources at the standard single detector \ac{BNS} horizon distance of 
-$\sim 450$\,Mpc the scale factor could on average be determined to within $\lesssim 20\%$ of its 
-true value for the \ac{aLIGO} detectors. Similar results were found for \ac{NSBH} sources, although 
-sources could be observed out to higher distances.
+\ac{AdV} using astrophysical sources. To do this we require that a \ac{sGRB}, for which it has 
+been possible to measure the distance, is observed in coincidence with a \ac{CBC} signal in the 
+\ac{GW} detectors. This enables us to assume a known distance to the source, and also limit the 
+inclination of the source, which in turn allows us to test the consistency of the detector 
+calibration. We do this by including an unknown scale factor on the true signal and noise, which we 
+estimate given the data. We find that for detectable \ac{BNS} sources the uncertainty of on the 
+calibration scale factor could on average be determined to $\lesssim 10\%$ of its true value for the 
+\ac{aLIGO} and \ac{AdV} detectors out to 100\,Mpc. This is comparable to the hoped for level of 
+calibration accuracy of the detectors. For sources at the standard single detector \ac{BNS} horizon 
+distance of $\sim 450$\,Mpc the scale factor could on average be determined to within $\lesssim 
+20\%$ of its true value for the \ac{aLIGO} detectors. Similar results were found for \ac{NSBH} 
+sources, although sources could be observed out to higher distances.
 
 The requirement of a coincident \ac{sGRB} with a known distance means that there will be 
-considerable latency for this calibration assessment, but importantly this method would provide 
-semi-independent consistency check that the calibration is reasonably accurate. We do note that this 
-method has to make its own assumptions, including that the waveforms we use accurately reflect the 
-signal and the calibration, the phase calibration does not change to much over frequency and that 
+considerable latency for this method of calibration assessment, but importantly this method would 
+provide semi-independent consistency check that the calibration is reasonably accurate. We do note 
+that this method has to make its own assumptions, including that the waveforms we use accurately 
+reflect the signal, that the phase calibration does not change to much over frequency and that 
 the overall calibration does not change considerably over the length of a signal. These are studies 
@@
 We imagine that a test for consistency of calibration could be performed with an ensemble of 
-\ac{CBC} sources, although such a method may have to marginalise over highly uncertain population 
-models and may give quite weak bounds on the overall calibration.
+\ac{CBC} sources even without \ac{sGRB} counterparts, although such a method would have to 
+marginalise over highly uncertain population models and may give quite weak bounds on the overall 
+calibration uncertainty.
 
-We emphasise that our work has assumed that the PSD used in the likelihood function
-when estimating the calibration scale factor is an accurate representation of the
-PSD at the time of the observed CBC signal. In reality the PSD may be estimated from
-a period close to, but not overlapping, the signal time and therefore may be slightly
-different \cite{2013PhRvD..88h4044L}. Our result would therefore be completely correlated
-with any uncertainty in the PSD estimate, but would still offer an upper limit
-on the calibration uncertainty.
+We emphasise that our work has assumed that the \ac{PSD} used in the likelihood function when 
+estimating the calibration scale factor is an accurate representation of the \ac{PSD} at the time 
+of the observed \ac{CBC} signal. In reality the \ac{PSD} may be estimated from a period close to, 
+but not overlapping, the signal time and therefore may be slightly different 
+\cite{2013PhRvD..88h4044L}. Our result would therefore be highly correlated with any 
+uncertainty in the \ac{PSD} estimate, but would still offer an upper limit on the calibration 
+uncertainty. However, future plans for \ac{CBC} parameter estimation include estimating the 
+\ac{PSD} with the signal using BayesLine \cite{2015PhRvD..91h4034L}.
 
 A further extension of this work would be for the calibration scale factor to be estimated as a 
-function of frequency. This would require the frequency series to be broken up into a set of 
-sections and the calibration scale factor estimated for each. However, the scale factor uncertainty 
-will obviously be related to the \ac{SNR} of a signal within a particular section. It could 
+function of frequency. This would require the frequency series to be broken up into sections and 
+the calibration scale factor estimated for each. However, the scale factor uncertainty will 
+obviously be related to the \ac{SNR} of a signal within a particular section. It could 
 therefore be a job of Bayesian model selection to work out the optimal number of frequency 
@@
 
-We would like to acknowledge the useful discussions with a whole bunch
-of people. LW was part funded for this work through a Royal Astronomical
-Society Undergraduate Research Bursary.
+We would like to acknowledge the useful discussions with a whole bunch of people. LW was part 
+funded for this work through a Royal Astronomical Society Undergraduate Research Bursary.
 MP is funded by the STFC under grant number ST/L000946/1.

### 2015-08-07 src/python/scale_factor_mcmc.py
scale_factor_mcmc.py: changes axes labels for the calibration scale factors

@@ -817,7 +817,7 @@ if it does not fulfill the SNR criterion.")
       labels = ["$\psi$", "$\phi_0$", "$\iota$", "$t_c$", "$\mathcal{M}$", "$q$"]
       truths = [psi, phi0, iota, 0.0, mC, q]
     for i in range(len(dets)):
-      labels.append("$s_{\mathrm{%s}}}$" % dets[i])
+      labels.append("$\mathcal{C}_{\mathrm{%s}}}$" % dets[i])
       truths.append(scales[i])
 
     # plot 1, 2 and 3 sigma contours

### 2015-09-22 docs/cosmic_cal.tex
cosmic_cal.tex: made some minor changes to use revtex style for PRD

@@
-\documentclass[10pt]{iopart}
+\IfFileExists{revtex4-1.cls}{\documentclass[prd, twocolumn, lengthcheck,
+superscriptaddress, showpacs, letterpaper, nofootinbib]{revtex4-1}}{
+\IfFileExists{revtex4.cls}{\documentclass[prd, twocolumn, lengthcheck,
+superscriptaddress, showpacs, letterpaper, nofootinbib]{revtex4}}{}
+}
+
 \usepackage{latexsym}
 \usepackage{graphicx}
-\usepackage{iopams}
+
 \usepackage{color}
 
-%\usepackage {amsmath}
-%\usepackage {amssymb}
-%\usepackage {amsfonts}
-%\usepackage {amsthm}
-%\usepackage {mathrsfs}
-%\usepackage {natbib}
-%\usepackage {latexsym}
-%\usepackage {graphicx}
-%\usepackage {dsfont}
-%\usepackage {times}
-%\usepackage {txfonts}
-%\usepackage {rotating}
-%\usepackage {wasysym}
-%\usepackage {multirow}
-%\usepackage {hhline}
+\usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{hyperref}
-%\usepackage {color}
 \usepackage{bm}
-%\usepackage{appendix}
 \usepackage{acronym}
-%\usepackage{dcolumn}   % needed for some tables
 \usepackage{url}
@@
 
-%\usepackage{harvard}
 \usepackage{amsopn} % for using DeclareMathOperator
@@
 \newcommand{\lam}{\bm{\lambda}}
-%\newcommand{\data}{\bm{d}}
-%\newcommand{\sigpar}{\bm{\theta}}
-%\newcommand{\nuispar}{\bm{\lambda}}
-%\newcommand{\globpar}{\bm{\gamma}}
-%\newcommand{\model}{\mathcal{H}}
 
@@
 \newcommand{\scf}{\ensuremath{\mathcal{C}}}
+
+\newcommand{\btheta}{\mathbf{\theta}}
 
@@
 
+% define macro for PASP from aas_macros.sty
+\newcommand\pasp{PASP}
+
 \begin{document}
@@
 
-\author{L.~Wright$^1$, M.~Pitkin$^1$ \& C.~Messenger$^1$}
-\address{$^1$ SUPA, School of Physics and Astronomy, University of
+\author{L.~Wright}\affiliation{SUPA, School of Physics and Astronomy, University of
   Glasgow, Glasgow G12 8QQ, United Kingdom}
-\eads{\mailto{matthew.pitkin@glasgow.ac.uk}, \mailto{christopher.messenger@glasgow.ac.uk}}
+\author{M.~Pitkin}\affiliation{SUPA, School of Physics and Astronomy, University of
+  Glasgow, Glasgow G12 8QQ, United Kingdom}
+\author{C.~Messenger}\affiliation{SUPA, School of Physics and Astronomy, University of
+  Glasgow, Glasgow G12 8QQ, United Kingdom}
+
+\email{matthew.pitkin@glasgow.ac.uk}
+\email{christopher.messenger@glasgow.ac.uk}
+
+\date{\today}
 
@@
 \end{abstract}
+
+\pacs{04.80.Nn, 95.55.Ym}
+\preprint{LIGO-P15XXXXX}
+
+\maketitle
 
@@
 \acrodef{MCMC}[MCMC]{Markov chain Monte Carlo}
-
-%Uncomment for PACS numbers title message
-%\pacs{00.00, 20.00, 42.10}
-% Keywords required only for MST, PB, PMB, PM, JOA, JOB? 
-%\vspace{2pc}
-%\noindent{\it Keywords}: Article preparation, IOP journals
-% Uncomment for Submitted to journal title message
-%\submitto{\JPA}
-% Comment out if separate title page not required
-\maketitle
 
@@
 sample the posterior distribution. We use a Gaussian likelihood function given by
-%
+\begin{widetext}
 \begin{equation}
-\fl p(\bi{d} | \btheta, \vec{\scf}, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm 
+p(\bm{d} | \btheta, \vec{\scf}, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm 
 det}}
@@
 \end{equation}
-% 
-where $\bi{d}$ is an array containing the data for all detectors, $\Delta f$ is the frequency 
+\end{widetext}
+where $\bm{d}$ is an array containing the data for all detectors, $\Delta f$ is the frequency 
 resolution, $S_k$ is the $k^{\rm th}$ detector's measured one sided noise \ac{PSD} (and as such 
@@
 
-\begin{figure}
+\begin{figure*}
  \begin{center}
@@
 of the detectors respectively.}
-\end{figure}
+\end{figure*}
 
-\begin{figure}
+\begin{figure*}
  \begin{center}
@@
 the detectors respectively.}
-\end{figure}
+\end{figure*}
 
-\begin{figure}
+\begin{figure*}
  \begin{center}
@@
 drawn from the prior distribution that would be detectable at each distance value.}
-\end{figure}
+\end{figure*}
 
-\begin{figure}
+\begin{figure*}
  \begin{center}
@@
 are the same as in figure~\ref{fig:bnsresults}.}
-\end{figure}
+\end{figure*}
 
@@
 
-\ack
+\acknowledgements
 
@@
 
-\section*{References}
-
-%\biaxbliographystyle{jphysicsB}
-\bibliographystyle{iopart-num}
 \bibliography{masterbib}

### 2015-09-22 docs/cosmic_cal.tex
Added text to the standard siren and GRB sections plus refs.

--- 2015-08-07 docs/cosmic_cal.tex
@@
 \acrodef{GW}[GW]{gravitational wave}
+\acrodef{NS}[NS]{neutron star}
 \acrodef{BNS}[BNS]{binary neutron star}
@@
 \acrodef{MCMC}[MCMC]{Markov chain Monte Carlo}
+\acrodef{ET}[ET]{Einstein Telescope}
 
@@
 
-We should explain what this means starting with reference
-to~\cite{1986Natur.323..310S}.
+% Standard siren basics
+The idea of \ac{GW} standard sirens is directly analogous to the concept of
+standard candles in \ac{EM} astronomy and was first proposed
+in~\cite{1986Natur.323..310S}. Unlike the primary \ac{EM} standard candle
+event, Type 1a Supernovae, the measured luminosity of a \ac{CBC} in \acp{GW} is
+not only a function of distance. It also depends upon the chirp mass (a
+function of the component masses) and the binary orientation with respect to
+the detector network. However, measurement of the phase evolution of such an
+event allows accurate determination of the chirp mass~\footnote{In reality it
+is the redshifted chirp mass that is measured.}. Timing information from the
+different signal arrival times at each detector in the network allows
+reasonably accurate sky position determination.  Finally, amplitude variation
+between differently oriented interferometers allows some (weaker) level of
+determination of the binary system orientation.  A great help to this final
+stage for this project is the assumption that the source has an \ac{sGRB}
+counterpart.  It follows that we therefore have a well constrained prior belief
+on the inclination of the source orbit with respect to the line-of-sight based
+on the expected level of beaming in the \ac{sGRB} emission.
+
+% The classic use of sirens
+Standard sirens are clearly a very powerful tool for \ac{GW} cosmology since
+they give us a direct measure of the absolute luminosty distance to sources.
+This is distinct from Type 1a Supernovae standard candles that only provide
+relative luminosty distance measures and require calibration via other methods
+as part of the cosmological distance ladder. At first sight the missing
+component for standard sirens is a complementary measurement of the source
+redshift which would allow each event to inform the distance-redshift
+relationship and hence obtain cosmological parameter estimates.  However, it
+was not thought possible (until recently) that redshift measurements could be
+obtained from \ac{GW} observations. In~\cite{DelPozzo:2011uf} (building on the
+original idea proposed in~\cite{1986Natur.323..310S} the idea of identifying
+host galaxies within the sky position uncertainty region was developed. Each
+galaxy would provide a redshift measure and would contribute to a disparate mix
+of possible cosmological parameters. However, by combining multiple (${\sim}20$)
+detections, the true set of cosmological parameters emerges with an expected
+${\sim}5\%$ uncertainty on the Hubble constant.
+
+% continued use of sirens
+The use of \ac{GW} events with \ac{sGRB} counterparts was investigated
+in~\cite{2010CQGra..27u5006S} and~\cite{2011PhRvD..83b3005Z} with respect to
+the 3rd generation \ac{GW} interferometer, the \ac{ET}. It was
+assumed in both studies that 1000 coincident events would be observed and
+cosmological parameter estimation ability was compared with current \ac{EM}
+results with focus on dark energy parameters. Other methods for cosmological
+inference (see~\cite{1996PhRvD..53.2878F},~\cite{2012PhRvD..85b3535T},
+and~\cite{2012PhRvD..86b3502T}) have been proposed for \ac{BNS} systems that do
+not rely on any redshift measurements and instead use the distribution of
+measured \acp{SNR} combined with some assumptions regarding the \ac{NS} mass
+and spatial distributions. Finally we mention that for \ac{BNS} systems it has
+been shown in~\cite{Messenger:2011ux} and~\cite{2013arXiv1312.1862M} that features
+of the tidal and post-merger hyper-massive \ac{NS} stages of the \ac{GW}
+waveform can allow the measurement of the source redshift if the \ac{NS}
+equation-of-state is sufficiently well known. This development, and seperately, the method
+based on \ac{SNR} distributions will allow \ac{GW} cosmological inference
+without the need for complementary \ac{EM} observations.
+
+% put this in context with calibration
+To put the existing work on \ac{GW} cosmology in context with the study
+described in this manuscript we are essentially turning the standard
+cosmological problem upside-down. We assume that the existing \ac{EM}
+cosmological parameter estimation (specifically for the Hubble constant) is
+accurately determined and therefore a known quantity. We then assume that the
+unknown in our analysis is the absolute amplitude calibration of the
+interferometers in the global \ac{GW} detector network. The issue of
+calibration has largely been ignored in \ac{GW} cosmological inference studies
+and in Sec.~\ref{sec:discussion} we discuss some of the implications of
+combining these issues in future analyses.
 
@@
 % what are sGRBs 
-\cm{What are sGRBs?}\\
+It is believed that \acp{sGRB} (those \acp{GRB} with duration ${<}2$ sec) are
+emitted during the merger of
+\acp{BNS}~\cite{1989Natur.340..126E},~\cite{1992ApJ...395L..83N}. The emmission
+from these events is highly beamed along the binary rotation axis and hence
+only potentially observable for a fraction of \ac{BNS} mergers. The distances
+to these objects are obtained via identification of
+
+The observed redshift range is from about 0.2 to 2 for short GRBs (sGRBs), with
+a mean of about 0.4. The XRT commonly observes sGRBs up to z ≃ 0.5 but in 10.5
+yr of operation Swift has not seen one with a measured redshift z < 0.1, i.e.,
+the NS-NS merger aLIGO target GW horizon ∼400 Mpc, suggesting that such events
+may be quite rare.
+Estimates for jet beaming are θj ∼ 5◦ for lGRBs and θj ∼ 5 − 15◦ for sGRBs
+(Burrows et al. 2006, Grupe et al. 2006, Fong et al. 2012). Beaming angles for
+sGRBs are still highly uncertain. The beaming factors fb = 1 − cos θj ≃ θj2/2
+are roughly 1/300 for lGRBs and 1/30 for sGRBs. Based on the observed rate of
+sGRBs by Swift, Coward et al. (2012) estimate a LVC detection rateof∼3−30yr−1
+forθj ≃ 15◦. Chen&Holz(2013)claim3−7yr−1 forGRBGW+EM detections. Kelley et al.
+(2013) estimate the rate of Swift or Fermi observations joint with LVC
+detections to be ∼0.07 yr−1. Siellez, Boe ̈r, & Gendre (2014) consider current
+and future high energy missions and estimate a rate of simultaneous GW+EM
+detections of ∼0.1 − 4 yr−1 in the LVC era. Wanderman & Piran (2015) estimate a
+co-detection rate LVC+Fermi of 0.1 − 1 yr−1 and LVC+Swift of 0.02 − 0.14 yr−1.
 

### 2015-09-22 docs/cosmic_cal.tex

@@
 \acrodef{ET}[ET]{Einstein Telescope}
-
-%Uncomment for PACS numbers title message
-%\pacs{00.00, 20.00, 42.10}
-% Keywords required only for MST, PB, PMB, PM, JOA, JOB? 
-%\vspace{2pc}
-%\noindent{\it Keywords}: Article preparation, IOP journals
-% Uncomment for Submitted to journal title message
-%\submitto{\JPA}
-% Comment out if separate title page not required
-\maketitle
 

### 2015-09-22 docs/cosmic_cal.tex
A bit more edits and fixed the compile.

@@
 \acrodef{ET}[ET]{Einstein Telescope}
+\acrodef{XRT}[XRT]{x-ray telescope}
 
@@
 from these events is highly beamed along the binary rotation axis and hence
-only potentially observable for a fraction of \ac{BNS} mergers. The distances
-to these objects are obtained via identification of
+only potentially observable for a fraction of \ac{BNS} mergers. If the event
+exhibits an optical afterglow then the host galaxy can be identified from which
+a redshift can be obtained.  The fraction of events with associated redshifts
+is ${\sim}1/3$. The range to which the \emph{Swift} \ac{XRT} has detected
+\acp{sGRB} is $z{\sim}2$ and the nearest event with associated redshift is at
+$z{sim}0.1$~\cite{2015GCN..17278...1C} which is approximately equal to the
+horizon range of the advanced \ac{GW} detector network.
 
-The observed redshift range is from about 0.2 to 2 for short GRBs (sGRBs), with
-a mean of about 0.4. The XRT commonly observes sGRBs up to z ≃ 0.5 but in 10.5
-yr of operation Swift has not seen one with a measured redshift z < 0.1, i.e.,
-the NS-NS merger aLIGO target GW horizon ∼400 Mpc, suggesting that such events
-may be quite rare.
-Estimates for jet beaming are θj ∼ 5◦ for lGRBs and θj ∼ 5 − 15◦ for sGRBs
-(Burrows et al. 2006, Grupe et al. 2006, Fong et al. 2012). Beaming angles for
-sGRBs are still highly uncertain. The beaming factors fb = 1 − cos θj ≃ θj2/2
-are roughly 1/300 for lGRBs and 1/30 for sGRBs. Based on the observed rate of
-sGRBs by Swift, Coward et al. (2012) estimate a LVC detection rateof∼3−30yr−1
-forθj ≃ 15◦. Chen&Holz(2013)claim3−7yr−1 forGRBGW+EM detections. Kelley et al.
-(2013) estimate the rate of Swift or Fermi observations joint with LVC
-detections to be ∼0.07 yr−1. Siellez, Boe ̈r, & Gendre (2014) consider current
-and future high energy missions and estimate a rate of simultaneous GW+EM
-detections of ∼0.1 − 4 yr−1 in the LVC era. Wanderman & Piran (2015) estimate a
-co-detection rate LVC+Fermi of 0.1 − 1 yr−1 and LVC+Swift of 0.02 − 0.14 yr−1.
+% this is copied text for reference !! 
+%Estimates for jet beaming are θj ∼ 5◦ for lGRBs and θj ∼ 5 − 15◦ for sGRBs
+%(Burrows et al. 2006, Grupe et al. 2006, Fong et al. 2012). Beaming angles for
+%sGRBs are still highly uncertain. The beaming factors fb = 1 − cos θj ≃ θj2/2
+%are roughly 1/300 for lGRBs and 1/30 for sGRBs. Based on the observed rate of
+%sGRBs by Swift, Coward et al. (2012) estimate a LVC detection rateof∼3−30yr−1
+%forθj ≃ 15◦. Chen&Holz(2013)claim3−7yr−1 forGRBGW+EM detections. Kelley et al.
+%(2013) estimate the rate of Swift or Fermi observations joint with LVC
+%detections to be ∼0.07 yr−1. Siellez, Boe ̈r, & Gendre (2014) consider current
+%and future high energy missions and estimate a rate of simultaneous GW+EM
+%detections of ∼0.1 − 4 yr−1 in the LVC era. Wanderman & Piran (2015) estimate a
+%co-detection rate LVC+Fermi of 0.1 − 1 yr−1 and LVC+Swift of 0.02 − 0.14 yr−1.
 

### 2015-09-25 docs/cosmic_cal.tex
Some minor edits and fixes of some refs

@@
 
-\newcommand{\btheta}{\mathbf{\theta}}
+\newcommand{\btheta}{\bm{\theta}}
 
@@
 sirens and their proposed coincident \ac{EM} signatures, the \acp{sGRB} in
-Secs.~\ref{sec:sirens} and~\ref{sec:GRB}. In Sec.~\ref{sec:cosmo} we show the likely accuracy of 
-\ac{sGRB} distances obtained from their host galaxy redshift and in Sec.~\ref{sec:discussion} we 
+Secs.~\ref{sec:sirens} and~\ref{sec:GRB}. In Secs.~\ref{sec:analysis} and \ref{sec:results} we 
+discuss out analysis method and the results, and in Sec.~\ref{sec:discussion} we 
 conclude.
@@
 The technique used throughout \ac{GW} research in calibrating the detectors is carried out through 
-a complicated system of physical manipulations of the Fabret-Perot Michelson interferometer; where 
+a complicated system of physical manipulations of the Fabry-P\'{e}rot Michelson interferometer; 
+where 
 an elaborate feedback system is used to sustain a defined measurement in arm length difference 
@@
 using the `hardware injection method' is roughly $10\%$ \cite{Vitale:2012}. This is a benchmark set 
-for the error estimation using the proposed new method in this project.
+for the error estimation using the proposed method in this project.
 
@@
 time-scales). Calculation of $R$ requires the measurement of various functions within a control 
-feedback loop (see \cite{2010NIMPA.624..223A}, which are subject to uncertainties. In this 
+feedback loop (see \cite{2010NIMPA.624..223A}), which are subject to uncertainties. In this 
 study we will assume an estimate of $R$ is available (although in theory we could take on the role 
@@
 closer) than in reality\footnote{However, the \ac{SNR} would be the correct value as the signal and 
-the noise will both contain the scale factor}, whereas if $\scf < 1$ it would appear to have a 
+the noise will both contain the scale factor.}, whereas if $\scf < 1$ it would appear to have a 
 lower amplitude than reality. In this analysis we will simplify the situation by assuming that 
@@
 in~\cite{1986Natur.323..310S}. Unlike the primary \ac{EM} standard candle
-event, Type 1a Supernovae, the measured luminosity of a \ac{CBC} in \acp{GW} is
+event, Type Ia Supernovae, the measured luminosity of a \ac{CBC} in \acp{GW} is
 not only a function of distance. It also depends upon the chirp mass (a
@@
 they give us a direct measure of the absolute luminosty distance to sources.
-This is distinct from Type 1a Supernovae standard candles that only provide
+This is distinct from Type Ia Supernovae standard candles that only provide
 relative luminosty distance measures and require calibration via other methods
@@
 obtained from \ac{GW} observations. In~\cite{DelPozzo:2011uf} (building on the
-original idea proposed in~\cite{1986Natur.323..310S} the idea of identifying
+original idea proposed in~\cite{1986Natur.323..310S}) the idea of identifying
 host galaxies within the sky position uncertainty region was developed. Each
@@
 emitted during the merger of
-\acp{BNS}~\cite{1989Natur.340..126E},~\cite{1992ApJ...395L..83N}. The emmission
+\ac{BNS} or \ac{NSBH} systems ~\cite{1989Natur.340..126E,1992ApJ...395L..83N}. The emission
 from these events is highly beamed along the binary rotation axis and hence
-only potentially observable for a fraction of \ac{BNS} mergers. If the event
+only potentially observable for a fraction of the mergers. If the event
 exhibits an optical afterglow then the host galaxy can be identified from which
@@
 \acp{sGRB} is $z{\sim}2$ and the nearest event with associated redshift is at
-$z{sim}0.1$~\cite{2015GCN..17278...1C} which is approximately equal to the
+$z{\sim}0.1$~\cite{2015GCN..17278...1C} which is approximately equal to the
 horizon range of the advanced \ac{GW} detector network.
@@
 observatories to search spanning $\sim 100$'s of square degrees~\cite{grb}.  For the \ac{GW} 
-follow-up of \ac{sGRB} scenario the merger time for \ac{BNS} systems will be estimated from the 
-\ac{sGRB} to within a few seconds~\cite{grb}.  This makes the follow-up search less computationally
-expensive since it is performed over a smaller range of data using potentially fewer numbers of 
-waveform templates.  This computational saving enables the use of a more computationally expensive 
-multi-detector coherent scheme rather than the cheaper coincidence methods used in the untargeted 
-searches.
+follow-up of \ac{sGRB} scenario the merger time for \ac{BNS}/\ac{NSBH} systems will be estimated 
+from the \ac{sGRB} to within a few seconds~\cite{grb}.  This makes the follow-up search less 
+computationally expensive since it is performed over a smaller range of data using potentially 
+fewer numbers of waveform templates.  This computational saving enables the use of a more 
+computationally expensive multi-detector coherent scheme rather than the cheaper coincidence 
+methods used in the untargeted searches.
 
@@
 A fortunate consequence of a joint \ac{GW}--\ac{sGRB} observation will be the fact that in order 
-for such an event to be observed, the \ac{BNS} system must have had its orbital angular momentum 
-vector pointing towards (or away) from the detector.  The actual inclination angle of the system, 
-defined as the angle between the orbital angular momentum vector and the line of sight, must be $<$
+for such an event to be observed, the \ac{BNS}/\ac{NSBH} system must have had its orbital angular 
+momentum vector pointing towards (or away) from the detector.  The actual inclination angle of the 
+system, defined as the angle between the orbital angular momentum vector and the line of sight, 
+must be $<$
 half of the \ac{sGRB} beaming angle.  This prerequisite property limits us to systems that are 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis}
+\section{Analysis}\label{sec:analysis}
 
@@
 %
-where $\tilde{n}(f)$ is the Fourier transform of the true (in our case Gaussian) strain noise drawn 
-from a, $\btheta$ is the set of waveform parameters and $\scf_k$ is the calibration scale factor for 
-the $k^{\rm th}$ detector.
+where $\tilde{n}(f)$ is the Fourier transform of the true (in our case Gaussian) strain noise, 
+$\btheta$ is the set of waveform parameters and $\scf_k$ is the calibration scale factor for 
+the $k^{\rm th}$ detector, which we want to estimate.
 
@@
 \begin{equation}
-p(\bm{d} | \btheta, \vec{\scf}, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm 
+p(\bm{d} | \btheta, \bm{\scf}, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm 
 det}}
@@
 additional information. The source parameters have been subsumed into a vector $\btheta$ and the 
-calibration scale factors are within the vector $\vec{\scf}$.
+calibration scale factors are within the vector $\bm{\scf}$.
 
@@
 calibration scale factor is either larger or smaller than unity by an equivalent factor, e.g.\ 
-$s=10$ or $s=0.1$, is the same. For the scale factor for each detector we we therefore use a 
+$\scf=10$ or $\scf=0.1$, is the same. For the scale factor for each detector we we therefore use a 
 log-normal distribution as our prior
@@
 all detectors were operating at their design sensitivity (as given by Figure~1 of 
-\cite{2013arXiv1304.0670L}). As such we added noise to each simulation based on the sensitivity 
-curve, but scaled with the associated calibration scale factor.
+\cite{2013arXiv1304.0670L}). As such we added coloured Gaussian noise to each simulation based on 
+the sensitivity curve, but scaled with the associated calibration scale factor.
 
@@
 
-From these posterior distributions we have calculated the minimal 68\% credible region on the 
+From these posterior distributions we have calculated the minimal 68\% credible region for the 
 calibration scale factors for each detector (if these were Gaussian distributions this is 
@@
 signals at each distance increment we have produced the distribution of the fractional half widths 
-(i.e.\ $1\sigma$) of these scale factors' confidence intervals compared to the actual value. These 
+(i.e.\ $1\sigma$) of these scale factors' confidence intervals compared to the true value. These 
 are shown as boxplots in figures~\ref{fig:bnsresults} and \ref{fig:nsbhresults} for the \ac{BNS} 
@@
 of $\lesssim 20\%$. Our results are probably slightly poorer than the \ac{SNR} expectation due to 
-their still be correlations with $\iota$ even within the constrained prior range, and slight 
+their still being correlations with $\iota$ even within the constrained prior range, and slight 
 correlations between the scale factors for H1 and L1 as seen in figure~\ref{fig:bnspost}.
@@
 are detectable. For distances at which the \ac{BNS} and \ac{NSBH} systems would give the same 
-percentage of observable sources there is a slight increase in the mean and standard distribution 
+percentage of observable sources there is a slight increase in the mean and standard deviation 
 of \acp{SNR} of the \ac{NSBH} systems over the \ac{BNS} systems.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Distance estimates from \acp{sGRB}\label{sec:cosmo}}
+%\section{Distance estimates from \acp{sGRB}\label{sec:cosmo}}
 
-~\cm{Describe the cosmological analysis used to obtain the distance from the
-GRB}
+%~\cm{Describe the cosmological analysis used to obtain the distance from the GRB}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Multiple events without counterpart\label{sec:multiple}}
+%\section{Multiple events without counterpart\label{sec:multiple}}
 
-\cm{A discussion of the issues regarding multiple events
-without EM counterparts.}
+%\cm{A discussion of the issues regarding multiple events
+%without EM counterparts.}
 
@@
 been possible to measure the distance, is observed in coincidence with a \ac{CBC} signal in the 
-\ac{GW} detectors. This enables us to assume a known distance to the source, and also limit the 
-inclination of the source, which in turn allows us to test the consistency of the detector 
-calibration. We do this by including an unknown scale factor on the true signal and noise, which we 
-estimate given the data. We find that for detectable \ac{BNS} sources the uncertainty of on the 
-calibration scale factor could on average be determined to $\lesssim 10\%$ of its true value for the 
-\ac{aLIGO} and \ac{AdV} detectors out to 100\,Mpc. This is comparable to the hoped for level of 
-calibration accuracy of the detectors. For sources at the standard single detector \ac{BNS} horizon 
-distance of $\sim 450$\,Mpc the scale factor could on average be determined to within $\lesssim 
-20\%$ of its true value for the \ac{aLIGO} detectors. Similar results were found for \ac{NSBH} 
-sources, although sources could be observed out to higher distances.
+\ac{GW} detectors. This enables us to assume a known distance and sky position for the source, and 
+also limit the inclination of the source, which in turn allows us to test the consistency of the 
+detector calibration. We do this by including an unknown scale factor on the true signal and noise, 
+which we estimate given the data. We find that for detectable \ac{BNS} sources the uncertainty 
+on the calibration scale factor could on average be determined to $\lesssim 10\%$ of its true value 
+for the \ac{aLIGO} and \ac{AdV} detectors out to 100\,Mpc. This is comparable to the hoped for 
+level of calibration accuracy of the detectors. For sources at the standard single detector 
+\ac{BNS} horizon distance of $\sim 450$\,Mpc the scale factor could on average be determined to 
+within $\lesssim 20\%$ of its true value for the \ac{aLIGO} detectors. Similar results were found 
+for \ac{NSBH} sources, although sources could be observed out to higher distances.
 

### 2015-09-28 docs/cosmic_cal.tex
Minor changes I forgot to commit last week.

--- 2015-09-22 docs/cosmic_cal.tex
@@
 
-% this is copied text for reference !! 
-%Estimates for jet beaming are θj ∼ 5◦ for lGRBs and θj ∼ 5 − 15◦ for sGRBs
-%(Burrows et al. 2006, Grupe et al. 2006, Fong et al. 2012). Beaming angles for
-%sGRBs are still highly uncertain. The beaming factors fb = 1 − cos θj ≃ θj2/2
-%are roughly 1/300 for lGRBs and 1/30 for sGRBs. Based on the observed rate of
-%sGRBs by Swift, Coward et al. (2012) estimate a LVC detection rateof∼3−30yr−1
-%forθj ≃ 15◦. Chen&Holz(2013)claim3−7yr−1 forGRBGW+EM detections. Kelley et al.
-%(2013) estimate the rate of Swift or Fermi observations joint with LVC
-%detections to be ∼0.07 yr−1. Siellez, Boe ̈r, & Gendre (2014) consider current
-%and future high energy missions and estimate a rate of simultaneous GW+EM
-%detections of ∼0.1 − 4 yr−1 in the LVC era. Wanderman & Piran (2015) estimate a
-%co-detection rate LVC+Fermi of 0.1 − 1 yr−1 and LVC+Swift of 0.02 − 0.14 yr−1.
-
 % discuss joint-event rates
-\cm{Discuss the expected rate of joint detections}\\
+The predicted detection rates of \ac{BNS} events (irrespective of an \ac{sGRB}
+counterpart) are derived from three main sources, population sythesis models,
+extrapolations based on known galactic \ac{BNS} systems, and from observations
+of \ac{sGRB} events. Various studies using the first two methods have been
+compiled into an overall detection rate of $0.2$--$200$ \ac{BNS} events per
+year~\cite{2013arXiv1304.0670L}~\cite{2010CQGra..27q3001A}~\footnote{We should
+briefly mention why the rates differ by a factor of 2.  I can't remember right
+now.}. In~\cite{2012MNRAS.425.2668C} using the rate of observed \acp{sGRB} and
+assuming a beaming angle of $15^{o}$ a similar event rate of $1$--$180$ is
+obtained. The beaming angle is the primary factor in determining the fraction
+of detected \ac{GW} \ac{BNS} signals with an \ac{EM} \ac{sGRB} counterpart and
+is unfortunately highly uncertain.  Various authors have provided estimates on
+the rate of joint detections to be in the range $0.02$--$7$ per
+year~\cite{2013PhRvL.111r1101C},~\cite{Kelley:2012fl},~\cite{2014arXiv1405.2254S},~\cite{2015MNRAS.448.3026W}.
 

### 2015-09-28 docs/cosmic_cal.tex
I got all the way through it.

@@
 \acrodef{PSD}[PSD]{power spectral density}
+\acrodef{PDF}[PDF]{probability distribution function}
 \acrodef{MCMC}[MCMC]{Markov chain Monte Carlo}
@@
 
-We want to assess how well the calibration scale factor defined in Eqn.~\ref{eq:scalefactor} can be 
-estimated from a single observed \ac{GW} associated with a particular \ac{sGRB}. As we {\it a 
-priori} have no knowledge of the likely location and distance of such an event we have performed 
-simulations of multiple events to see the expected distribution in the accuracy of the calibration 
-scale factor recovery.
+% the main aim of the analysis
+The main aim of this work is to assess how well the calibration scale factor
+defined in Eqn.~\ref{eq:scalefactor} can be estimated from a single observed
+\ac{GW} associated with a particular \ac{sGRB}. As we {\it a priori} have
+little knowledge of the likely location and distance of such an event we have
+performed simulations of multiple events to see the expected distribution in
+the accuracy of the calibration scale factor recovery.
 
 % Describe the signal model (Taylor F2 etc...)
-We use the TaylorF2 waveform approximation (see e.g.\ \cite{2009PhRvD..80h4043B} and references 
-therein) with a 3.5 post-Newtonian expansion in phase for modelling our signal (in both simulations 
-and signal recovery).  For \ac{BNS} systems we use non-spinning waveforms under the assumption that 
-spins will be negligible. For \ac{NSBH} system we use a spinning waveform, but in which only the 
-black hole has non-negligible spin and that its rotational angular momentum is aligned with the 
-system's orbital angular momentum.
+We use the TaylorF2 waveform approximation (see e.g.~\cite{2009PhRvD..80h4043B}
+and references therein) with a 3.5 post-Newtonian expansion in phase for
+modelling our signal (in both simulations and signal recovery).  For \ac{BNS}
+systems we use non-spinning waveforms under the assumption that spins will be
+negligible~\cm{ref?}. For \ac{NSBH} system we use a spinning waveform, but in which only
+the black hole has non-negligible spin and that its rotational angular momentum
+is aligned with the system's orbital angular momentum.
 
-For a non-spinning system the general form of the frequency-domain polarisation amplitudes are
+For a non-spinning system the general form of the frequency-domain polarisation
+amplitudes are
 %
-\begin{eqnarray}\label{eq:signal}
-  \gws_{+}(f) &\propto \frac{1+\cos^{2}\iota}{D_{L}}
-\mathcal{M}^{5/6}f^{-7/6}e^{-i\Psi(f, \nu, t_\mathrm{c}, \phi_{\mathrm{c}})} \nonumber \\
-  \gws_{\times}(f) &\propto
- \frac{\cos^{2}\iota}{D_{L}}\mathcal{M}^{5/6} f^{-7/6}e^{-i\Psi(f, \nu, t_\mathrm{c}, 
-\phi_{\mathrm{c}})}
-\end{eqnarray}
+\begin{align}\label{eq:signal} 
+\gws_{+}(f) &\propto
+\left(1+\cos^{2}\iota\right)D_{L}^{-1} \mathcal{M}^{5/6}f^{-7/6}e^{-i\Psi(f,
+\nu, t_\mathrm{c}, \phi_{\mathrm{c}})} \nonumber \\ \gws_{\times}(f) &\propto
+\left(\cos^{2}\iota\right)D_{L}^{-1}\mathcal{M}^{5/6} f^{-7/6}e^{-i\Psi(f, \nu,
+t_\mathrm{c}, \phi_{\mathrm{c}})} 
+\end{align}
 %
-where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=M\eta^{3/5}$, with 
-$\eta=m_{1}m_{2}/M^{2}$ and $M=m_{1}+m_{2}$, $\iota$ is the inclination angle, and $\nu=(\pi 
-Mf)^{1/3}$.  The \ac{GW} strain measured at the $k^{\rm{th}}$ detector is then given by
+where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=M\eta^{3/5}$,
+with the symmetric mass ratio $\eta=m_{1}m_{2}/M^{2}$.  The total mass
+$M=m_{1}+m_{2}$ where $m_{1}$ and $m_{2}$ are the component masses, $\iota$ is
+the inclination angle, and $\nu=(\pi Mf)^{1/3}$.  The \ac{GW} strain measured
+at the $k^{\rm{th}}$ detector is then given by
 %
-\begin{equation}
-  \label{eq:gravsig}
-   \gws^{k}(f) = \big[ F_{+}^{k}(\alpha, \delta, \psi)\gws_{+}(f) +
-F_{\times}^k(\alpha, \delta, \psi)\gws_{\times}(f)\big]
+\begin{equation} \label{eq:gravsig} 
+\gws^{k}(f) = \big[ F_{+}^{k}(\alpha,
+\delta, \psi)\gws_{+}(f) + F_{\times}^k(\alpha, \delta,
+\psi)\gws_{\times}(f)\big] 
 \end{equation}
 %
-where $F_{+}$ and $F_{\times}$ are the antenna response functions which are dependent upon the 
-polarisation angle $\psi$ and the sky position of the source $\alpha$ and $\delta$.
+where $F^{k}_{+}$ and $F^{k}_{\times}$ are the antenna response functions which
+are dependent upon the polarisation angle $\psi$ and the sky position of the
+source $\alpha$ and $\delta$.
 
 % Define the detectors and the noise epoch we will be using
-For this analysis we will consider the advanced generation \ac{aLIGO} and \ac{AdV} detectors 
-operating at their design sensitivity, with noise power spectral densities taken from
-\cite{2013arXiv1304.0670L}. The network we consider is the two \ac{aLIGO} detectors (H1 and L1) and 
+For this analysis we will consider the advanced generation \ac{aLIGO} and
+\ac{AdV} detectors operating at their design sensitivity, with noise power
+spectral densities taken from \cite{2013arXiv1304.0670L}. The detector network
+we consider consists of the two \ac{aLIGO} detectors (H1 and L1) and 
 the \ac{AdV} detector (V1).
@@
 %
-where $\tilde{n}(f)$ is the Fourier transform of the true (in our case Gaussian) strain noise, 
-$\btheta$ is the set of waveform parameters and $\scf_k$ is the calibration scale factor for 
-the $k^{\rm th}$ detector, which we want to estimate.
+where $\tilde{n}(f)$ is the Fourier transform of the true (in this case
+Gaussian) strain noise, $\btheta$ is the set of waveform parameters and
+$\scf_k$ is the calibration scale factor for the $k^{\rm th}$ detector, the
+parameters we are interested in estimating.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Method}
 
-To assess the ability to estimate the unknown calibration scale factors we want to calculate their 
-marginal posterior probability distribution functions (pdf). We use Bayes' theorem for which we 
-need to define a likelihood function and prior pdfs on the signal parameters being estimated. In 
-this analysis we use the Markov chain Monte Carlo code {\tt emcee} \cite{2013PASP..125..306F} to 
-sample the posterior distribution. We use a Gaussian likelihood function given by
+% Introduce the Bayesian approach and the likelihood
+To assess the ability to estimate the unknown calibration scale factors we
+calculate their marginal posterior \acp{PDF}.  We use Bayes' theorem for which
+we need to define a likelihood function and prior \acp{PDF} on the signal
+parameters and the calibration scale factors. In this analysis we use the
+\ac{MCMC} code {\tt emcee}~\cite{2013PASP..125..306F} to sample
+the posterior distribution. We use a Gaussian likelihood function given by
+%
 \begin{widetext}
@@
 det}}
-  \sum_{i=i_{f_{\rm low}}}^{i_{f_{\rm high}}}\frac{\Re{\left\{ \tilde{d}^{*}_{k,i} 
+  \sum_{i=i_{\rm low}}^{i_{\rm high}}\frac{\Re{\left\{ \tilde{d}^{*}_{k,i} 
 \scf_{k}\gws_{k,i}(\btheta) - \frac{1}{2}\left(\scf_k^2 \gws^*_{k,i}(\btheta)\gws_{k,i}(\btheta) + 
@@
 \end{widetext}
-where $\bm{d}$ is an array containing the data for all detectors, $\Delta f$ is the frequency 
-resolution, $S_k$ is the $k^{\rm th}$ detector's measured one sided noise \ac{PSD} (and as such 
-will contain the effect of the calibration scale factor), and the $i$ indices increment over 
-frequency with $i_{f_{\rm low}}$ and $i_{f_{\rm high}}$ corresponding to the lower 
-and upper range in frequencies used for our analysis, which were 20\,Hz to 400\,Hz respectively. 
-This frequency range was chosen as the vast majority of the \ac{SNR} for the inspiralling signals 
-we use can be found within this range, so going to lower or higher frequencies provides very little 
-additional information. The source parameters have been subsumed into a vector $\btheta$ and the 
-calibration scale factors are within the vector $\bm{\scf}$.
+%
+where $\bm{d}$ is an array containing the data for all detectors, $\Delta f$ is
+the frequency resolution, $S_k$ is the $k^{\rm th}$ detector's estimated
+one-sided noise \ac{PSD} (and as such will contain the effect of the
+calibration scale factor), and the $i$ indices increment over frequency with
+$i_{\rm low}$ and $i_{\rm high}$ corresponding to the lower and upper range in
+frequencies used for our analysis, which were 20\,Hz to 400\,Hz respectively.
+This frequency range was chosen as the vast majority of the \ac{SNR} for the
+inspiralling signals we use can be found within this range, so going to lower
+or higher frequencies provides very little additional information. The source
+parameters have been subsumed into a vector $\btheta$ and the calibration scale
+factors are within the vector $\bm{\scf}$.
 
@@
 
-The primary reason why we can use coincident observations with a \ac{sGRB} as a check of detector 
-calibration is that the \ac{sGRB} allows us to constrain the priors on various parameters that 
-would normally be highly correlated with the signal amplitude. The most important of these is 
-that the \ac{sGRB} can provide a very tight constraint on the source distance\footnote{For \ac{BNS} 
-and \ac{NSBH} systems they will be observable only in the relatively nearby Universe ($z \sim 
-0.1$ and $z\sim 0.2$ for \ac{BNS} and \ac{NSBH} systems respectively), so the relation between the 
-measured \ac{sGRB} redshift and distance is simple and our luminosity distance will be negligibly 
-effected by the choice of cosmology \MP{Check this?}.}. In this analysis we therefore assume that 
-the source distance is known (i.e.\ has a $\delta$-function prior). Another piece of information 
-that we can use from having a coincident \ac{sGRB} is that the system is likely to be relatively 
-close to face-on (and therefore nearly circularly polarised), which allows us to place prior 
-constraints on the system inclination angle. Beam opening angles for \acp{sGRB} are relatively 
-poorly constrained as they are based on only a few jet break detections, however 
-\cite{2014ApJ...780..118F} give a median opening angle value of $\sim 10^{\circ}$. If we assume 
-that the beam opening angle is a half-normal distribution, for which the median is given by 
-$\sqrt{2}\erf{}^{-1}(1/2)\sigma \sim 0.67\sigma$, then a median of $\sim 10^{\circ}$ corresponds to 
-$\sigma \sim 14.8^{\circ}$. We therefore place a Gaussian prior on the inclination angle with zero 
-mean and $\sigma=14.8^{\circ}$.
+% mention of distance delta function prior
+The primary reason why coincident observations with a \ac{sGRB} can be used as
+a check of detector calibration is that the \ac{sGRB} allows us to constrain
+the priors on various parameters that would normally be highly correlated with
+the signal amplitude. The most important of these is that the \ac{sGRB} can
+provide a very tight constraint on the source distance\footnote{For \ac{BNS}
+and \ac{NSBH} systems they will be observable only in the relatively nearby
+Universe ($z \sim 0.1$ and $z\sim 0.2$ for \ac{BNS} and \ac{NSBH} systems
+respectively), so the relation between the measured \ac{sGRB} redshift and
+distance is simple and our luminosity distance will be negligibly effected by
+the choice of cosmology \MP{Check this?}.}. In this analysis we therefore
+assume that the source distance is known (i.e.\ has a $\delta$-function
+prior).~\cm{Should we not say a little bit more about this? A really accurate
+distance from a really accurate redshift is only as good as the Hubble constant
+which currently has $\sim$ few percent error?  We should also say something
+about peculiar motion.}
 
-Our prior on the system component masses is dependent on whether we are considering the source 
-being a \ac{BNS} system or a \ac{NSBH}. For the former case the prior we use is a Gaussian 
-distribution for both components with means of $1.35\,\textrm{M}_{\odot}$ and standard deviations 
-of $0.13\,\textrm{M}_{\odot}$ \MP{(reference)}. In reality, for all the systems we use in our 
-study, the chirp mass, which we see from equation~\ref{eq:signal} plays a role in the overall 
-signal amplitude, is very well constrained by the system's phase evolution, so our prior could be 
-expanded with minimal effect on the results. In the case of \ac{NSBH} systems we use the same 
-prior as above for the neutron star, but for the black hole we use a prior based on the 
-canonical masss distribution used for the rate results in \cite{2012PhRvD..85h2002A} with a mean of 
-$5\,\textrm{M}_{\odot}$ and a standard deviation of $1\,\textrm{M}_{\odot}$. We note that the mass 
-range of black holes could be quite different from this, but use this range as an example for this 
-type of system. For the \ac{NSBH} systems we also require a prior on the black hole spin (for 
-\ac{BNS} systems we have assumed that the spins are small enough that they will be negligible). We 
-use a uniform prior on the normalised aligned spin magnitude between $-1$ and $1$.
+% discuss the inclination and opening angle
+Other information relevant for a coincident \ac{sGRB} is that the system is
+likely to be relatively close to face-on (and therefore nearly circularly
+polarised), which allows us to place prior constraints on the system
+inclination angle. Beam opening angles for \acp{sGRB} are relatively poorly
+constrained as they are based on only a few jet-break detections. A comparative
+study of estimates and lower-limits of opening angles in 13 \acp{sGRB} given
+in~\cite{2014ApJ...780..118F} provide a median opening angle value of $\sim
+10^{\circ}$. We assume that the beam opening angle is a half-normal
+distribution, for which the median is given by $\sqrt{2}\erf{}^{-1}(1/2)\sigma
+\sim 0.67\sigma$. It follows that the median of a $\sim 10^{\circ}$ opening
+angle corresponds to $\sigma \sim 14.8^{\circ}$. We therefore place a Gaussian
+prior on the inclination angle with zero mean and a standard deviation
+$\sigma=14.8^{\circ}$ truncated at values below zero.
 
-The final important piece of information that we can make use of from the \ac{sGRB} observation is 
-the sky position of the source, which we take to be known precisely.
+% discuss the component mass and spin priors 
+The prior on the system component masses is dependent on whether we are
+considering the source being a \ac{BNS} system or a \ac{NSBH}. For the former
+case the prior we use is a Gaussian distribution for both components with means
+of $1.35\,\textrm{M}_{\odot}$ and standard deviations of
+$0.13\,\textrm{M}_{\odot}$ \MP{(reference)}. In reality, for all the systems we
+use in our study, the chirp mass, which we see from Eq.~\ref{eq:signal}
+plays a role in the overall signal amplitude, is very well constrained by the
+system's phase evolution, so our prior could be expanded with minimal effect on
+the results. In the case of \ac{NSBH} systems we use the same prior as above
+for the neutron star, but for the black hole we use a prior based on the
+canonical masss distribution used for the rate results in
+\cite{2012PhRvD..85h2002A} with a mean of $5\,\textrm{M}_{\odot}$ and a
+standard deviation of $1\,\textrm{M}_{\odot}$. We note that the mass range of
+black holes could be quite different from this, but use this range as an
+example for this type of system. For the \ac{NSBH} systems we also require a
+prior on the black hole spin (for \ac{BNS} systems we have assumed that the
+spins are small enough that they will be negligible). We use a uniform prior on
+the normalised aligned spin magnitude between $-1$ and $1$.
 
-For the time of coalescence we use a uniform prior of spanning $\pm0.01$\,seconds around the 
-recorded time of the signal that would be returned by the detection pipeline. For the reference 
-phase a uniform prior between 0 and $\pi$ is used, and for the polarisation angle a uniform prior 
-between 0 and $\pi/2$ is used. As the signals we use are generally close to being circularly 
-polarised (face-on) the phase and polarisation angle will be degenerate meaning the exact 
-combination of them will have very little effect of the result.
+% discuss the known sky position
+The final important piece of information that we can make use of from the
+\ac{sGRB} observation is the sky position of the source. We assume that this is 
+known precisely and has been obtained, in conjunction with the source redshift,
+from the identification of the host galaxy.
 
-Finally, we require a prior on the calibration scale factors for each detector. We make the 
-assumption that in general the calibration applied to the detector data will be correct, so want to 
-use a prior that is peaked at unity. We also assume that the probability density that the 
-calibration scale factor is either larger or smaller than unity by an equivalent factor, e.g.\ 
-$\scf=10$ or $\scf=0.1$, is the same. For the scale factor for each detector we we therefore use a 
-log-normal distribution as our prior
-\begin{equation}
+% discuss the nuisance parameters
+For the time of coalescence $t_{\mathrm{c}}$ we use a uniform prior spanning
+$\pm0.01$\,seconds around the recorded time of the signal that would be
+returned by the \ac{GW} detection pipeline. For the reference phase
+$\phi_{\mathrm{c}}$ a uniform prior on the range $(0,\pi]$ is used, and for the
+polarisation angle we use a uniform prior on the range $(0,\pi/2]$. As the
+signals we use are generally close to being circularly polarised (face-on) the
+phase and polarisation angle will be largely degenerate and their exact
+combination will have little effect on the result.
+
+% the scale factor prior
+Finally, we require a prior on the calibration scale factors for each detector.
+We make the assumption that in general the calibration applied to the detector
+data will be correct, so want to use a prior that is peaked at $\scf_k=1$. We
+also assume that the prior \ac{PDF} of the calibration scale factor is either
+larger or smaller than unity by an equivalent factor, e.g.\ $p(\scf|I)=10$ is
+equivelent to $p(\scf|I)=0.1$. For the scale factor in each detector we 
+therefore use the following log-normal distribution as our prior
+%
+\begin{equation}\label{eq:scale_prior}
  p(\scf|I) = \frac{1}{\scf\sigma\sqrt{2\pi}}\exp{\left( -\frac{(\ln{\scf} - \mu)^2}{2\sigma^2} 
@@
 \end{equation}
-where we chose $\sigma = 1.07$, which in turn means that having the mode at unity gives $\mu = 
-1.15$. With these parameters the prior probability density at 0.1 and 10 is one tenth of that at 
-1. Note that this prior does not give a symmetric amount of probability about unity, but given our 
-simulation criterion described below in Section~\ref{sec:simulations} we find that the likelihood 
-generally overwhelms the prior. As $\scf$ is a scale factor we also have the further constraint 
-that it must be positive.
+%
+where we chose $\sigma = 1.07$, which together with the requirement that the
+mode is at unity gives $\mu = 1.15$. With these parameters the prior
+probability density at 0.1 and 10 is one tenth of that at 1. Note that this
+prior does not give a symmetric amount of probability about unity, but given
+our simulation criterion described in Section~\ref{sec:simulations} we find
+that the likelihood generally overwhelms the prior~\cm{I'm not sure about this
+sentence.  Why does the symmetry matter?}. As $\scf$ is a scale factor we also
+have the further constraint that it must be positive and hence the log-normal
+prior is a natural choice.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Simulations}\label{sec:simulations}
 
-To estimate how well we can constrain the calibration scaling for each detector in an advanced 
-detector network we have performed simulations with injected \ac{BNS} and \ac{NSBH} signals at a 
-range of distances (50\,Mpc to 500\,Mpc with 50\,Mpc increments for the \ac{BNS} signals and 
-100\,Mpc to 900\,Mpc with 100\,Mpc increments for the \ac{NSBH} signals). The simulations used the 
-two \ac{aLIGO} detectors (H1 and L1) and the \ac{AdV} detector (V1). The simulations were all 
-performed in the frequency domain and spanned a frequency range from 20\,Hz to 400\,Hz. We assumed
-all detectors were operating at their design sensitivity (as given by Figure~1 of 
-\cite{2013arXiv1304.0670L}). As such we added coloured Gaussian noise to each simulation based on 
-the sensitivity curve, but scaled with the associated calibration scale factor.
+% a general description of the simulations 
+To estimate how well the calibration scaling can be constrained for each
+detector in an advanced detector network we have performed simulations with
+injected \ac{BNS} and \ac{NSBH} signals. These signal injections were simulated
+at a range of distances: 50\,Mpc to 500\,Mpc with 50\,Mpc increments for the
+\ac{BNS} signals and 100\,Mpc to 900\,Mpc with 100\,Mpc increments for the
+\ac{NSBH} signals. The simulations used the two \ac{aLIGO} detectors (H1 and
+L1) and the \ac{AdV} detector (V1) and were all performed in the frequency
+domain and spanned a frequency range from 20\,Hz to 400\,Hz~\cm{can you remind
+me and maybe the reader of why a 400 Hz cutoff is OK?}. We assumed all
+detectors were operating at their design sensitivity (as given by Figure~1 of
+\cite{2013arXiv1304.0670L}). As such we added coloured Gaussian noise to each
+simulation based on the sensitivity curve, but scaled with the associated
+calibration scale factor.
 
 % Describe the ensemble of simulations used
-Separately for the \ac{BNS} and \ac{NSBH} systems we have performed a large number simulations at 
-each distance value. When proposing injections for each simulation the source parameters were 
-randomly drawn from the prior distributions given in Section~\ref{sec:priors}, with the sky 
-position being drawn randomly from a uniform distribution on the sky, the coalescence times held 
-fixed within the centre of the prior window, and with the calibration scaling factors for each 
-detector drawn from a Gaussian distribution with a mean of one and standard deviation of $0.125$ 
-(equivalent to a mean fractional offset of 10\%). However, in accepting a proposed injection as one 
-to be analysed 
-we introduced a criteria that the signal be ``detectable'', based on that used in 
-\cite{2012PhRvD..85h2002A} of it having an \ac{SNR} of $\geq 5.5$ in at least two detectors (as 
-these would be signals coincident with a \ac{sGRB} we do not include the often used further 
-constraint that the total coherent network \ac{SNR} is greater than 12). This criteria means that 
-at larger distances our population of injected sources was not uniformly distributed over the sky, 
-and were generally closer to being circularly polarised.
+Separately for the \ac{BNS} and \ac{NSBH} systems we have performed a large
+number~\cm{how many exactly/roughly?} simulations at each distance value. When
+proposing injections for each simulation the source parameters were randomly
+drawn from the prior distributions described in Section~\ref{sec:priors}, with
+the sky position being drawn randomly from a uniform distribution on the sky,
+the coalescence times held fixed within the centre of the prior window, and
+with the calibration scaling factors for each detector drawn from a Gaussian
+distribution with a mean of one and standard deviation of $0.125$ (equivalent
+to a mean fractional offset of 10\%)~\cm{should we explicitly acknowledge that
+we do not simulate from the same priors that we assume in the recovery?}.
+However, in accepting a proposed injection as one to be analysed we introduced
+a criteria that the signal be ``detectable'', based on that used in
+\cite{2012PhRvD..85h2002A}. We therefore only analysed injections that had an
+\ac{SNR} of $\geq 5.5$ in at least two detectors. Since we assume that these
+signals would be coincident with a \ac{sGRB} we do not include the often used
+further constraint that the total coherent network \ac{SNR} is greater than 12.
+For injections close to the \ac{SNR} threshold, the application of the
+threshold resulted in the preferential selection of better oriented systems.
+Therefore, at larger distances the population of injected sources was not
+uniformly distributed over the sky and instead favoured locations with better
+antenna response.  Similarly, despite already constrained by the prior to small
+inclination angles, the accepted injections show preference for sources very
+close to circularly polarised.
 
-These simulations have allowed us to assess how well on average we would be able to constrain the 
-calibration scale for a given \ac{CBC}-\ac{sGRB} coincidence.
+% final clarification on the simulation distribution
+These simulations have allowed us to assess how well on average we would be
+able to constrain the calibration scale for a given detected \ac{CBC}-\ac{sGRB}
+coincidence.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}\label{sec:results}
 
-A simple assumption that one could make would be that the uncertainties on the calibration scale 
-factors (if they are independent of other parameters and have a roughly Gaussian probability 
-distribution) should be given by $\sim 1/{\rm SNR}$ for each detector. As we will show below this 
-assumption is reasonable, but small correlations do exist between parameters meaning that it does 
-not completely hold.
+% a general and simple results statement 
+A simple assumption that one could make would be that the uncertainties on the
+calibration scale factors (if they are independent of other parameters and have
+a roughly Gaussian probability distribution) should be given by $\sim 1/{\rm
+SNR}$ for each detector. As we will show below this assumption is reasonable,
+but small correlations do exist between parameters meaning that it does not
+completely hold.
 
-For each of the simulated sources we have used the {\tt emcee} python \ac{MCMC} package 
-\cite{2013PASP..125..306F} to perform parameter estimation over the unknown source parameters using 
-the priors as discussed in Section~\ref{sec:priors}. In each case when calculating the likelihood 
-we used an estimate of the noise \ac{PSD} based on the advanced detector design sensitivities 
-(using those given in \cite{2013arXiv1304.0670L}), but simulated as if calculated by averaging 64 
-separate noisy \ac{PSD} estimates and scaled with the same factor as applied to the injection and 
-noise\footnote{We do not account for there being a potential difference between the estimated 
-\ac{PSD} and the actual \ac{PSD} of the analysed section of data as described in e.g.\ 
-\cite{2013PhRvD..88h4044L}. This difference would be very highly correlated with the calibration 
-scale factor, so in reality our estimate of the scale factor would be a combination of the
-calibration offset and the difference in the \ac{PSD}. As such our results on real data would be an 
-upper limit on the calibration scale factor.}. This has provided posterior probability distributions 
-on the calibration scale factors for each detector. Examples of the marginalised posterior 
-probabilities for a \ac{BNS} system and a \ac{NSBH} system observed with the three detector network 
-are shown in figures~\ref{fig:bnspost} and \ref{fig:nsbhpost} respectively.
+% a discussion of the PSD estimate
+For each of the simulated sources we have used the {\tt emcee} python \ac{MCMC}
+package~\cite{2013PASP..125..306F} to perform parameter estimation over the
+unknown source parameters using the priors as discussed in
+Section~\ref{sec:priors}. In each case when calculating the likelihood we used
+an estimate of the noise \ac{PSD} based on the advanced detector design
+sensitivities (using those given in~\cite{2013arXiv1304.0670L}), but simulated
+as if calculated by averaging 64 separate noisy \ac{PSD} estimates and scaled
+with the same factor as applied to the injection and noise~\cm{Do you estimate
+the PSD using 64 segments? This sentence might be a bit confusing to the
+reader.}\footnote{We do not account for there being a potential difference
+between the estimated \ac{PSD} and the actual \ac{PSD} of the analysed section
+of data as described in e.g.\ \cite{2013PhRvD..88h4044L}. This difference would
+be very highly correlated with the calibration scale factor, so in reality our
+estimate of the scale factor would be a combination of the calibration offset
+and the difference in the \ac{PSD}. As such our results on real data would be
+an upper limit on the calibration scale factor.~\cm{Even though we're hiding
+this in a footnote is it really this bad. How big is the statistical error on
+PSD estimation using 64 segments?}}. This has provided posterior probability
+distributions on the calibration scale factors for each detector.  Examples of
+the marginalised posterior probabilities for a \ac{BNS} system and a \ac{NSBH}
+system observed with the three detector network are shown in
+figures~\ref{fig:bnspost} and \ref{fig:nsbhpost} respectively.
 
-From these posterior distributions we have calculated the minimal 68\% credible region for the 
-calibration scale factors for each detector (if these were Gaussian distributions this is 
-equivalent to the region either side of the mean bounded by the $1\sigma$ intervals). For all the 
-signals at each distance increment we have produced the distribution of the fractional half widths 
-(i.e.\ $1\sigma$) of these scale factors' confidence intervals compared to the true value. These 
-are shown as boxplots in figures~\ref{fig:bnsresults} and \ref{fig:nsbhresults} for the \ac{BNS} 
-and \ac{NSBH} systems respectively. The boxes show the extent from the lower to upper quartile of 
-the values, whilst the whiskers extend from the 5$^{\rm th}$ to 95$^{\rm th}$ percentile. The black 
-line within each box gives the median value and the star gives the mean value. Also shown on each 
-plot as the dashed magenta line is the percentage of signals drawn from the prior distribution that 
-fulfil the detection criterion.
+% describe the main boxplot results
+From these posterior distributions we have calculated the minimal 68\% credible
+region for the calibration scale factors for each detector (if these were
+Gaussian distributions this is equivalent to the region either side of the mean
+bounded by the $1\sigma$ intervals). For all the signals at each distance
+increment we have produced the distribution of the fractional half widths
+(i.e.\ $1\sigma$) of these scale factors' confidence intervals compared to the
+true value. These are shown as boxplots in figures~\ref{fig:bnsresults} and
+\ref{fig:nsbhresults} for the \ac{BNS} and \ac{NSBH} systems respectively. The
+boxes show the extent from the lower to upper quartile of the values, whilst
+the whiskers extend from the 5$^{\rm th}$ to 95$^{\rm th}$ percentile. The
+black line within each box gives the median value and the star gives the mean
+value. Also shown on each plot as the dashed magenta line is the percentage of
+signals drawn from the prior distribution that fulfil the detection criterion.
 
@@
 4.1 and percentage uncertainties in the calibration scale factors of 14\%, 11\% and 27\% for each 
-of the detectors respectively.}
+of the detectors respectively.~\cm{The fonts are a bit small.  Maybe also
+highlight somehow the boxes refereing to the scale factors?}}
 \end{figure*}
@@
 4.2 and percentage uncertainties in the calibration scale factors of 17\%, 14\% and 34\% for each of 
-the detectors respectively.}
+the detectors respectively.~\cm{same things about the fonts and highlighting
+the scale factors.}}
 \end{figure*}
@@
  \caption{\label{fig:bnsresults} Distributions of the percentage accuracy at which the calibration 
-scale factors can be determined for a three detector network if assuming \ac{BNS} systems (provided 
-a coincident \ac{GRB} is observed and can yield a distance estimate). The boxplots span the lower 
-to upper quartile range of the distributions, with the median value shown as a horizontal line 
-within the box and the mean shown as a star. The dashed magenta line shows the percentage of source 
-drawn from the prior distribution that would be detectable at each distance value.}
+scale factors can be determined for a three detector network if assuming
+\ac{BNS} systems (provided a coincident \ac{GRB} is observed and can yield a
+distance estimate). The boxplots span the lower to upper quartile range of the
+distributions, with the median value shown as a horizontal line within the box
+and the mean shown as a white star. The dashed magenta line shows the
+percentage of sources drawn from the prior distribution that would be detectable
+at each distance value.} 
 \end{figure*}
@@
 
-In figure~\ref{fig:bnsresults} we see that on average the scale factor can be recovered to 
-equivalent precision in both \ac{aLIGO} detectors, as would be expected, with uncertainties 
-generally within the 10\% range for sources at 100\,Mpc. This is comparable to previous estimates 
-of the calibration error in the initial LIGO detectors. An interesting feature is that for 
-distances $\gtrsim 250$\,Mpc the upper extent of the uncertainty for H1 and L1 hits a maximum at 
-$\sim 25\%$, whilst the width of boxes narrows. This is due to our ``detectability'' criteria, 
-in that for all distances we will only be seeing those sources with a high enough \ac{SNR} that we 
-would consider them detectable. There will therefore be fewer sources with \ac{SNR} higher than 
-this criteria at large distances, giving us a narrower range, and we automatically exclude those 
-with \ac{SNR} that are too small thus truncating our uncertainty distribution at the upper end. 
-However, this does show that on average for sources that are detectable out to 450\,Mpc we would be 
-able to constrain the calibration scale factor uncertainty for the \ac{aLIGO} detectors to $\sim 
-20\%$. As the detection criteria will generally come from the \ac{SNR} being high enough within the 
-two \ac{aLIGO} instruments this means that the SNR in \ac{AdV} can be small and thus the ability to 
-constrain the scale factor for it becomes poor (although it still provides information that the 
-calibration is not grossly inaccurate). We also see that true uncertainties achievable for 
-recovering the calibration scale factors are indeed very similar to the simple assumption that they 
-would be given by $\sim 1/{\rm SNR}$, i.e.\ at the upper end of the distribution the \ac{SNR} in 
-the two \ac{aLIGO} detectors will be $\sim 5.5$, which would be expected to produce uncertainties 
-of $\lesssim 20\%$. Our results are probably slightly poorer than the \ac{SNR} expectation due to 
-their still being correlations with $\iota$ even within the constrained prior range, and slight 
-correlations between the scale factors for H1 and L1 as seen in figure~\ref{fig:bnspost}.
+% discuss the BNS plots - general behaviour
+In figure~\ref{fig:bnsresults} we see that on average the scale factor can be
+recovered to equivalent precision in both \ac{aLIGO} detectors, as would be
+expected, with uncertainties generally within the 10\% range for sources at
+100\,Mpc. This is comparable to previous estimates of the calibration error in
+the initial LIGO detectors. An interesting feature is that for distances
+$\gtrsim 250$\,Mpc the upper extent of the uncertainty for H1 and L1 hits a
+maximum at $\sim 25\%$, whilst the width of boxes narrows. This is due to our
+``detectability'' criteria, where for all distances we will only be seeing
+those sources with a high enough \ac{SNR} that we would consider them
+detectable. There will therefore be fewer sources with \ac{SNR} higher than
+this criteria at large distances, giving us a narrower range, and we
+automatically exclude those with \ac{SNR} that are too small thus truncating
+our uncertainty distribution at the upper end.  However, this does show that on
+average for sources that are detectable out to 450\,Mpc we would be able to
+constrain the calibration scale factor uncertainty for the \ac{aLIGO} detectors
+to $\sim 20\%$.
 
-In figure~\ref{fig:nsbhresults} we see very similar results for the \ac{NSBH} systems although the 
-higher \ac{SNR} of the signals means that we can provide $\sim 20\mbox{--}25\%$ uncertainties on 
-the calibration scale factors for H1 and L1 out to greater distances. The distributions generally 
-appear slightly broader than those for the \ac{BNS} systems. This may be entirely explicable 
-through two factors. The first factor is from the fact that including the spin parameter leads to 
-strong correlations between the chirp mass, mass ratio and spin. These strong correlations make 
-convergence of the \ac{MCMC} take longer and means we have fewer independent samples to estimate 
-the posterior distributions with. This leads to a larger statistical fluctuation on the results. 
-The second factor is that there appears to be a modest effect due to the population of sources that 
-are detectable. For distances at which the \ac{BNS} and \ac{NSBH} systems would give the same 
-percentage of observable sources there is a slight increase in the mean and standard deviation 
-of \acp{SNR} of the \ac{NSBH} systems over the \ac{BNS} systems.
+% discuss the Virgo contribution and the simple estimate
+The largest \ac{SNR} contribution will generally come from either of the
+\ac{aLIGO} detectors and thus the detection criteria (\ac{SNR} threshold) will
+not apply to to\ac{adV} result. Hence, the \ac{SNR} in \ac{AdV} can be small
+and thus the ability to constrain its scale factor becomes poor (although it
+still provides information that the calibration is not grossly inaccurate). We
+also see that true uncertainties achievable for recovering the calibration
+scale factors are indeed very similar to the simple assumption that they would
+be given by $\sim 1/{\rm SNR}$, i.e.\ at the upper end of the distribution the
+\ac{SNR} in the two \ac{aLIGO} detectors will be $\sim 5.5$, which would be
+expected to produce uncertainties of $\lesssim 20\%$. Our results are slightly
+poorer than estimates based on this simple assumption due to correlations with
+$\iota$ even within the constrained prior range, and slight correlations
+between the scale factors for H1 and L1 as seen in figure~\ref{fig:bnspost}.
 
-As noted previously the mass distribution for the black holes could be quite different from the one 
-we used, with masses extending to $\gtrsim 10$\,M$_{\odot}$. We expect these would produce similar 
-results to those we see in figure~\ref{fig:nsbhresults}, but again extending to further distances. 
-Including the merger and ring-down phase, which we currently ignore, may also help increase 
-\ac{SNR}.
+% discuss the NSBH sources
+In figure~\ref{fig:nsbhresults} we see very similar results for the \ac{NSBH}
+systems although the higher \ac{SNR} of the signals means that we can provide
+$\sim 20\mbox{--}25\%$ uncertainties on the calibration scale factors for H1
+and L1 out to greater distances. The distributions generally appear slightly
+broader than those for the \ac{BNS} systems. This may be explained
+by two factors. The first is from the fact that including the spin
+parameter leads to strong correlations between the chirp mass, mass ratio and
+spin. These strong correlations make convergence of the \ac{MCMC} take longer
+and means we have fewer independent samples with which to estimate the posterior
+distributions. This leads to a larger statistical fluctuation on the
+results.  The second factor is that there appears to be a modest effect due to
+the population of sources that are detectable. For distances at which the
+\ac{BNS} and \ac{NSBH} systems would give the same percentage of observable
+sources there is a slight increase in the mean and standard deviation of
+\acp{SNR} of the \ac{NSBH} systems over the \ac{BNS} systems~\cm{can we go any
+further with this?}.
+
+% reiterate the BH mass distribution
+As noted previously, the mass distribution for black holes could be quite
+different from the one used here, with masses extending to $\gtrsim
+10$\,M$_{\odot}$. We expect these would produce similar results to those we see
+in figure~\ref{fig:nsbhresults}, but again extending to further distances.
+Including the merger and ring-down phase, which we currently ignore, will also
+increase \ac{SNR} and act to reduce the calibration scale uncertainties.
 
@@
 
-We have attempted to see how well we can assess the detector calibration for \ac{aLIGO} and 
-\ac{AdV} using astrophysical sources. To do this we require that a \ac{sGRB}, for which it has 
-been possible to measure the distance, is observed in coincidence with a \ac{CBC} signal in the 
-\ac{GW} detectors. This enables us to assume a known distance and sky position for the source, and 
-also limit the inclination of the source, which in turn allows us to test the consistency of the 
-detector calibration. We do this by including an unknown scale factor on the true signal and noise, 
-which we estimate given the data. We find that for detectable \ac{BNS} sources the uncertainty 
-on the calibration scale factor could on average be determined to $\lesssim 10\%$ of its true value 
-for the \ac{aLIGO} and \ac{AdV} detectors out to 100\,Mpc. This is comparable to the hoped for 
-level of calibration accuracy of the detectors. For sources at the standard single detector 
-\ac{BNS} horizon distance of $\sim 450$\,Mpc the scale factor could on average be determined to 
-within $\lesssim 20\%$ of its true value for the \ac{aLIGO} detectors. Similar results were found 
-for \ac{NSBH} sources, although sources could be observed out to higher distances.
+% summary of main results
+We have established how well we can assess the detector calibration for
+\ac{aLIGO} and \ac{AdV} using astrophysical sources as standard sirens. To do
+this we require that a \ac{sGRB}, for which it has been possible to measure the
+distance, is observed in coincidence with a \ac{CBC} signal in the \ac{GW}
+detectors. This enables us to assume a known distance and sky position for the
+source, and also limit the inclination of the source, which in turn allows us
+to test the consistency of the detector calibration. We do this by including an
+unknown scale factor on the true signal and noise, which we estimate given the
+data from the 3 inteferometer advanced detector network. We find that for
+detectable \ac{BNS} sources the uncertainty on the calibration scale factor
+could on average be determined to $\lesssim 10\%$ of its true value for the
+\ac{aLIGO} and \ac{AdV} detectors out to 100\,Mpc. This is comparable to the
+proposed accuracy level of the hardware calibration of the detectors. For
+sources at the standard single detector \ac{BNS} horizon distance of $\sim
+450$\,Mpc the scale factor could on average be determined to within $\lesssim
+20\%$ of its true value for the \ac{aLIGO} detectors. Similar results were
+found for \ac{NSBH} sources, although sources could be observed out to higher
+distances.
 
-The requirement of a coincident \ac{sGRB} with a known distance means that there will be 
-considerable latency for this method of calibration assessment, but importantly this method would 
-provide semi-independent consistency check that the calibration is reasonably accurate. We do note 
-that this method has to make its own assumptions, including that the waveforms we use accurately 
-reflect the signal, that the phase calibration does not change to much over frequency and that 
-the overall calibration does not change considerably over the length of a signal. These are studies 
-that could be incorporated in the future.
+% mention caveats and future work
+The requirement of a coincident \ac{sGRB} with a known distance means that
+there will likely be considerable delay (beyond initial \ac{GW} detection) for
+the implementation of this method of calibration assessment.  Importantly
+however, this method would provide a semi-independent consistency check that
+the calibration is reasonably accurate.  We acknowledge that we have made a
+number of simplifying assumptions, including that the waveforms we use
+accurately model the true signal, and that the overall calibration remains
+constant over the duration of a signal and its \ac{PSD} estimation. A further
+major simplification made here is that the absolute calibration is constant
+with frequency.  Other authors~\cm{Ref: Cornish? and others?} have addressed
+this issue in the context of robust parameter estimation for astrophysical
+sources and in future work we intend to apply related methods for astrophysical
+calibration.
 
-We imagine that a test for consistency of calibration could be performed with an ensemble of 
-\ac{CBC} sources even without \ac{sGRB} counterparts, although such a method would have to 
-marginalise over highly uncertain population models and may give quite weak bounds on the overall 
-calibration uncertainty.
+% comparitive calibration
+We imagine that a test for consistency of calibration between detectors could be
+performed with an ensemble of \ac{CBC} sources without relying on \ac{EM}
+counterparts. Such a method would have to marginalise over highly
+uncertain population models and may give quite weak bounds on the overall
+calibration uncertainty.~\cm{I'm not sure what you mean here? I want to add a
+paragraph on relative calibration after I understand what you mean.}
 
-We emphasise that our work has assumed that the \ac{PSD} used in the likelihood function when 
-estimating the calibration scale factor is an accurate representation of the \ac{PSD} at the time 
-of the observed \ac{CBC} signal. In reality the \ac{PSD} may be estimated from a period close to, 
-but not overlapping, the signal time and therefore may be slightly different 
-\cite{2013PhRvD..88h4044L}. Our result would therefore be highly correlated with any 
-uncertainty in the \ac{PSD} estimate, but would still offer an upper limit on the calibration 
-uncertainty. However, future plans for \ac{CBC} parameter estimation include estimating the 
-\ac{PSD} with the signal using BayesLine \cite{2015PhRvD..91h4034L}.
+% discus PSD uncertainty
+We emphasise that our work has assumed that the \ac{PSD} used in the likelihood
+function is an accurate representation of the \ac{PSD} at the time of the
+observed \ac{CBC} signal. In reality the \ac{PSD} may be estimated from a
+period close to, but not overlapping, the signal time and therefore may be
+slightly different \cite{2013PhRvD..88h4044L}. Our result would therefore be
+highly correlated with any uncertainty in the \ac{PSD} estimate, but would
+still offer an upper limit on the calibration uncertainty. However, future
+plans for \ac{CBC} parameter estimation include estimating the \ac{PSD} with
+the signal using BayesLine \cite{2015PhRvD..91h4034L}~\cm{I'm nit sure that
+this last point really helps us.}.
 
-A further extension of this work would be for the calibration scale factor to be estimated as a 
-function of frequency. This would require the frequency series to be broken up into sections and 
-the calibration scale factor estimated for each. However, the scale factor uncertainty will 
-obviously be related to the \ac{SNR} of a signal within a particular section. It could 
-therefore be a job of Bayesian model selection to work out the optimal number of frequency 
-sections to estimate the scale factor for. The BayesLine algorithm \cite{2015PhRvD..91h4034L}, or 
-the method described in \cite{2013PhRvD..88h4044L}, may provide a natural way to perform such an
-analysis. Indeed these methods, and those of \cite{Vitale:2012} are already being used for 
-marginalising over calibration uncertainties, but do not attempt to estimate it.
+% frequency dependence
+A further extension of this work would be for the calibration scale factor to
+be estimated as a function of frequency. This may require the frequency series
+to be divided into sections and the calibration scale factor estimated for
+each. However, since the scale factor uncertainty is related to the \ac{SNR}
+then the \ac{SNR} contribution in each frequency section will be of relevance.
+In this case Bayesian model selection may be applicabl for selecting the
+optimal number of frequency sections with which to estimate the scale factor.
+The BayesLine algorithm \cite{2015PhRvD..91h4034L}, or the method described in
+\cite{2013PhRvD..88h4044L}, may provide a natural way to perform such an
+analysis. Indeed these methods, and those of \cite{Vitale:2012} are already
+being used for marginalising over calibration uncertainties, but do not yet
+attempt to estimate it.
 
@@
 funded for this work through a Royal Astronomical Society Undergraduate Research Bursary.
-MP is funded by the STFC under grant number ST/L000946/1.
+MP is funded by the STFC under grant number ST/L000946/1 and CM is funded via
+XXX.
 

### 2015-09-29 docs/cosmic_cal.tex
Some minor updates and further references

@@
 exhibits an optical afterglow then the host galaxy can be identified from which
-a redshift can be obtained.  The fraction of events with associated redshifts
-is ${\sim}1/3$. The range to which the \emph{Swift} \ac{XRT} has detected
+a redshift can be obtained (e.g.~\cite{2005Natur.437..851G}).  The fraction of events with 
+associated redshifts is ${\sim}1/3$. The range to which the \emph{Swift} \ac{XRT} has detected
 \acp{sGRB} is $z{\sim}2$ and the nearest event with associated redshift is at
@@
 compiled into an overall detection rate of $0.2$--$200$ \ac{BNS} events per
-year~\cite{2013arXiv1304.0670L}~\cite{2010CQGra..27q3001A}~\footnote{We should
+year~\cite{2013arXiv1304.0670L,2010CQGra..27q3001A}~\footnote{We should
 briefly mention why the rates differ by a factor of 2.  I can't remember right
@@
 the rate of joint detections to be in the range $0.02$--$7$ per
-year~\cite{2013PhRvL.111r1101C},~\cite{Kelley:2012fl},~\cite{2014arXiv1405.2254S},~\cite{2015MNRAS.448.3026W}.
+year~\cite{2013PhRvL.111r1101C,Kelley:2012fl,2014arXiv1405.2254S,2015MNRAS.448.3026W}.
 
@@
 this issue is the relatively large \ac{GW} sky error-box giving a field-of-view for the \ac{EM} 
-observatories to search spanning $\sim 100$'s of square degrees~\cite{grb}.  For the \ac{GW} 
+observatories to search spanning $\sim 100$'s of square degrees (e.g.\ 
+\cite{2011CQGra..28j5021F,2013arXiv1304.0670L,2014ApJ...795..105S}).  For the \ac{GW} 
 follow-up of \ac{sGRB} scenario the merger time for \ac{BNS}/\ac{NSBH} systems will be estimated 
-from the \ac{sGRB} to within a few seconds~\cite{grb}.  This makes the follow-up search less 
+from the \ac{sGRB} to within a few seconds (see e.g. the discussion in Sec.\ 2.2 of 
+\cite{2012ApJ...760...12A}). This makes the follow-up search less 
 computationally expensive since it is performed over a smaller range of data using potentially 
@@
 of $1.35\,\textrm{M}_{\odot}$ and standard deviations of
-$0.13\,\textrm{M}_{\odot}$ \MP{(reference)}. In reality, for all the systems we
-use in our study, the chirp mass, which we see from Eq.~\ref{eq:signal}
+$0.13\,\textrm{M}_{\odot}$, which is roughly consistent with the double neutron star system
+masses found in \cite{2012ApJ...757...55O} albeit with a wider distribution. In reality, for all 
+the systems we use in our study, the chirp mass, which we see from Eq.~\ref{eq:signal}
 plays a role in the overall signal amplitude, is very well constrained by the

### 2015-09-30 src/python/scale_factor_mcmc.py
Slight fix to plotting script

@@ -757,7 +757,7 @@ if it does not fulfill the SNR criterion.")
     commandline += arg + ' '
   outdict['CommandLine'] = commandline
 
-  # get scale factor standard devaitions and 95% credible intervals
+  # get scale factor standard deviations and 95% credible intervals
   scstds = []
   cis95 = []
   cis90 = []
@@ -799,13 +799,37 @@ if it does not fulfill the SNR criterion.")
 
   if args.plot:
     try:
-      import triangle
+      from corner import corner
     except:
-      print >> sys.stderr, "Can't load triangle.py, so no plot will be produced"
+      print >> sys.stderr, "Can't load corner.py, so no plot will be produced"
       sys.exit(0)
 
     mC = (m1inj*m2inj)**(3./5.) / (m1inj+m2inj)**(1./5.)
     q = m2inj/m1inj
+
+    # make sure psi and phi0 are in their correct range (they don't seem to be returned correctly through theta)
+    i = 0
+    for psival, phival in samples[:,0:2]:
+      nmod = 0.
+      if psival > np.pi/2.:
+        nmod = np.floor(psival/(np.pi/2.))
+        psival = np.fmod(psival, np.pi/2.)
+      elif psival < 0.:
+        nmod = np.ceil(-psival/(np.pi/2.))
+        psival += nmod*(np.pi/2.)
+ 
+      phival += nmod*(np.pi/2.)
+ 
+      # wrap phi0 in the 0->pi range
+      if phival < 0.:
+        nmod = np.ceil(-phival/np.pi)
+        phival += nmod*np.pi
+      elif phival > np.pi:
+        phival = np.fmod(phival, np.pi)
+
+      samples[i,0] = psival
+      samples[i,1] = phival
+      i = i+1
 
     # set t0 to 0
     samples[:,3] = samples[:,3] - t0
@@ -823,7 +847,20 @@ if it does not fulfill the SNR criterion.")
     # plot 1, 2 and 3 sigma contours
     levels = 1.-np.exp(-0.5*np.array([1., 2., 3.])**2)
 
-    fig = triangle.corner(samples, labels=labels, truths=truths, data_kwargs={'color': 'darkblue', 'ms': 2}, plot_density=True, no_fill_contours=False, plot_contours=True, levels=levels)
[…]

### 2015-09-30 docs/cosmic_cal.tex
Minor changes

@@
 
-\author{L.~Wright}\affiliation{SUPA, School of Physics and Astronomy, University of
+\author{M.~Pitkin}
+\email{matthew.pitkin@glasgow.ac.uk}
+\affiliation{SUPA, School of Physics and Astronomy, University of
   Glasgow, Glasgow G12 8QQ, United Kingdom}
-\author{M.~Pitkin}\affiliation{SUPA, School of Physics and Astronomy, University of
+\author{C.~Messenger}
+\email{christopher.messenger@glasgow.ac.uk}
+\affiliation{SUPA, School of Physics and Astronomy, University of
   Glasgow, Glasgow G12 8QQ, United Kingdom}
-\author{C.~Messenger}\affiliation{SUPA, School of Physics and Astronomy, University of
+\author{L.~Wright}
+\altaffiliation{Now at: School of Mathematics and Physics, Queen's University Belfast, Belfast BT7 
+1NN, United Kingdom}
+\affiliation{SUPA, School of Physics and Astronomy, University of
   Glasgow, Glasgow G12 8QQ, United Kingdom}
-
-\email{matthew.pitkin@glasgow.ac.uk}
-\email{christopher.messenger@glasgow.ac.uk}
 
@@
 % BNS systems are GRBs
-It is considered likely that the merger of \ac{BNS} systems, in addition to emitting detectable 
-\acp{GW}, is also the mechanism for producing \acp{sGRB}. In this scenario these \ac{EM} events are 
-produced through the ... and result in tightly beamed emission parallel to the orbital angular 
+It is considered likely that the merger of \ac{BNS} and/or \ac{NSBH} systems, in addition to 
+emitting detectable \acp{GW}, is also the mechanism for producing \acp{sGRB} 
+\cite{1992ApJ...395L..83N}. In this scenario these events produce tightly beamed \ac{EM} emission 
+parallel to the orbital angular 
 momentum vector of the \ac{BNS} system. Additional evidence for the coincidence of \acp{sGRB} with
@@
 give a rate of XXX Mpc$^{-3}$yr$^{-1}$ which can be compared to estimates of \ac{BNS} merger rates 
-of XXX-XXX Mpc$^{-3}$yr$^{-1}$. This latter range is obtained from population synthesis models 
-together with knowledge of the X known \ac{BNS} systems in our galaxy.
+of XXX-XXX Mpc$^{-3}$yr$^{-1}$, or \ac{NSBH} merger rates. The ranges of rates for \acp{BNS} 
+is obtained from population synthesis models together with knowledge of the X known \ac{BNS} systems 
+in our galaxy, whilst the range for \acp{NSBH} is entirely from population synthesis as not such 
+systems have been observed.
 
@@
  \caption{\label{fig:bnspost} The marginalised posterior probability distributions for the unknown
- parameters of a \ac{BNS} system, including calibration scaling factors for the three detectors 
-(H1, L1 and V1). The injected signal was at a distance of 250\,Mpc, had \ac{SNR} of 8.7, 10.9 and 
-4.1 and percentage uncertainties in the calibration scale factors of 14\%, 11\% and 27\% for each 
-of the detectors respectively.~\cm{The fonts are a bit small.  Maybe also
-highlight somehow the boxes refereing to the scale factors?}}
+ parameters of a \ac{BNS} system, including calibration scaling factors (surrounded by the red 
+borders) for the three detectors (H1, L1 and V1). The injected signal was at a distance of 
+250\,Mpc, had \ac{SNR} of 8.7, 10.9 and 4.1 and percentage uncertainties in the calibration scale 
+factors of 14\%, 11\% and 27\% for each of the detectors respectively.}
 \end{figure*}
@@
 4.2 and percentage uncertainties in the calibration scale factors of 17\%, 14\% and 34\% for each of 
-the detectors respectively.~\cm{same things about the fonts and highlighting
-the scale factors.}}
+the detectors respectively.}
 \end{figure*}

### 2015-10-01 docs/cosmic_cal.tex
Finished another pass through the paper - almost there

@@
 \ac{GW} events are the estimated astrophysical rates of both phenomena. Observations of \acp{sGRB} 
-give a rate of XXX Mpc$^{-3}$yr$^{-1}$ which can be compared to estimates of \ac{BNS} merger rates 
-of XXX-XXX Mpc$^{-3}$yr$^{-1}$, or \ac{NSBH} merger rates. The ranges of rates for \acp{BNS} 
-is obtained from population synthesis models together with knowledge of the X known \ac{BNS} systems 
-in our galaxy, whilst the range for \acp{NSBH} is entirely from population synthesis as not such 
-systems have been observed.
+give a rate of XXX\,Mpc$^{-3}$\,yr$^{-1}$ which can be compared to estimates of \ac{BNS} merger 
+rates of $10^{-8}$--$10^{-5}$\,Mpc$^{-3}$\,yr$^{-1}$ \cite{2010CQGra..27q3001A}, or \ac{NSBH} merger 
+rates of $6\,\times\,10^{-10}$--$1\,\times\,10^{-6}$\,Mpc$^{-3}$\,yr$^{-1}$ 
+\cite{2010CQGra..27q3001A}. However, the recent work of \cite{2015MNRAS.448..928K}, which studied 
+known galactic \ac{BNS} systems in more detail suggest a rate approximately five times lower. The 
+estimates for \ac{NSBH} systems come entirely from population synthesis models as no such systems 
+have been observed.
+%The ranges of rates for \acp{BNS} is obtained from population synthesis models together with 
+%knowledge of the X known \ac{BNS} systems in our galaxy, whilst the range for \acp{NSBH} is 
+%entirely from population synthesis as not such systems have been observed.
 
-% talk about the calibration idea
 If a single \ac{GW} event is observed in coincidence with a \ac{sGRB} it may be possible to 
 identify the host galaxy of the \ac{BNS} source.  With this information a spectroscopic redshift 
-can be very accurately obtained. With knowledge of the redshift, using current best estimates for 
-the Hubble constant and other cosmological parameters, the true luminosity distance can be
-estimated to $\sim 1\%$ accuracy.  The \ac{GW} measurement acts as a standard siren also giving us 
-a direct measurement of the luminosity distance to the source. The accuracy of such a measurement 
-depends on a number of factors including the accuracy with which the \ac{GW} detector has been 
-calibrated.  Hence comparison with the distance estimate from the \ac{sGRB} we can re-calibrate (or 
-validate) the existing experimentally obtained calibration.
+can be very accurately obtained (see values quoted in e.g. \cite{0004-637X-664-2-1000} and 
+references therein). With knowledge of the redshift, using current best 
+estimates for the Hubble constant and other cosmological parameters, the true luminosity distance 
+can be estimated to $\sim 1\%$ accuracy.  The \ac{GW} measurement acts as a standard siren also 
+giving us a direct measurement of the luminosity distance to the source. The accuracy of such a 
+measurement depends on a number of factors including the accuracy with which the \ac{GW} detector 
+has been calibrated.  Hence, by comparison with the distance estimate from the \ac{sGRB} we can 
+re-calibrate (or validate) the existing experimentally obtained calibration.
 
@@
 a complicated system of physical manipulations of the Fabry-P\'{e}rot Michelson interferometer; 
-where 
-an elaborate feedback system is used to sustain a defined measurement in arm length difference 
-between the moving mirrors. A comprehensive description of the calibration procedure (in particular 
-for the LIGO detectors during their fifth science run) can be found in \cite{2010NIMPA.624..223A}, 
-and a brief description is giving in \cite{Vitale:2012} and references therein. For \ac{aLIGO} 
-and \ac{AdV}, at the frequency range used in this project, the error in the amplitude calibration 
-using the `hardware injection method' is roughly $10\%$ \cite{Vitale:2012}. This is a benchmark set 
-for the error estimation using the proposed method in this project.
+where an elaborate feedback system is used to sustain a defined measurement in arm length 
+difference between the moving mirrors. A comprehensive description of the calibration procedure (in 
+particular for the LIGO detectors during their fifth science run) can be found in 
+\cite{2010NIMPA.624..223A}, and a brief description is giving in \cite{Vitale:2012} and references 
+therein. For \ac{aLIGO} and \ac{AdV}, at the frequency range used in this project, the error in the 
+amplitude calibration is roughly $10\%$ \cite{Vitale:2012}. This is a benchmark set for the error 
+estimation using the proposed method in this project.
 
@@
 of estimating $R$ itself), but that it differs from the truth through some unknown scale factor, 
-$\scf$, so that
+$\scf$, so that at a particular time we have
 \begin{equation}\label{eq:scalefactor}
-\scf h(f) = h_{\rm m}(f) = \frac{R(f,t) e(f)}{L},
+\scf h(f) = h_{\rm m}(f) = \frac{R(f) e(f)}{L},
 \end{equation}
@@
 between differently oriented interferometers allows some (weaker) level of
-determination of the binary system orientation.  A great help to this final
-stage for this project is the assumption that the source has an \ac{sGRB}
-counterpart.  It follows that we therefore have a well constrained prior belief
-on the inclination of the source orbit with respect to the line-of-sight based
-on the expected level of beaming in the \ac{sGRB} emission.
+determination of the binary system orientation.
+%A great help to this final
+%stage for this project is the assumption that the source has an \ac{sGRB}
+%counterpart.  It follows that we therefore have a well constrained prior belief
+%on the inclination of the source orbit with respect to the line-of-sight based
+%on the expected level of beaming in the \ac{sGRB} emission.
 
@@
 Standard sirens are clearly a very powerful tool for \ac{GW} cosmology since
-they give us a direct measure of the absolute luminosty distance to sources.
+they give us a direct measure of the absolute luminosity distance to sources.
 This is distinct from Type Ia Supernovae standard candles that only provide
-relative luminosty distance measures and require calibration via other methods
+relative luminosity distance measures and require calibration via other methods
 as part of the cosmological distance ladder. At first sight the missing
@@
 % continued use of sirens
-The use of \ac{GW} events with \ac{sGRB} counterparts was investigated
-in~\cite{2010CQGra..27u5006S} and~\cite{2011PhRvD..83b3005Z} with respect to
+The use of \ac{GW} events with \ac{sGRB} counterparts for use as a cosmological tool was 
+investigated in~\cite{2010CQGra..27u5006S} and~\cite{2011PhRvD..83b3005Z} with respect to
 the 3rd generation \ac{GW} interferometer, the \ac{ET}. It was
@@
 results with focus on dark energy parameters. Other methods for cosmological
-inference (see~\cite{1996PhRvD..53.2878F},~\cite{2012PhRvD..85b3535T},
-and~\cite{2012PhRvD..86b3502T}) have been proposed for \ac{BNS} systems that do
+inference (see~\cite{1996PhRvD..53.2878F,2012PhRvD..85b3535T, 2012PhRvD..86b3502T}) have 
+been proposed for \ac{BNS} systems that do
 not rely on any redshift measurements and instead use the distribution of
 measured \acp{SNR} combined with some assumptions regarding the \ac{NS} mass
-and spatial distributions. Finally we mention that for \ac{BNS} systems it has
+and spatial distributions. Finally, we mention that for \ac{BNS} systems it has
 been shown in~\cite{Messenger:2011ux} and~\cite{2013arXiv1312.1862M} that features
@@
 waveform can allow the measurement of the source redshift if the \ac{NS}
-equation-of-state is sufficiently well known. This development, and seperately, the method
+equation-of-state is sufficiently well known. This development, and separately, the method
 based on \ac{SNR} distributions will allow \ac{GW} cosmological inference
@@
 $z{\sim}0.1$~\cite{2015GCN..17278...1C} which is approximately equal to the
-horizon range of the advanced \ac{GW} detector network.
+horizon range of the advanced \ac{GW} detector network for \ac{BNS} systems.
 
@@
 The predicted detection rates of \ac{BNS} events (irrespective of an \ac{sGRB}
-counterpart) are derived from three main sources, population sythesis models,
+counterpart) are derived from three main sources, population synthesis models,
 extrapolations based on known galactic \ac{BNS} systems, and from observations
@@
 compiled into an overall detection rate of $0.2$--$200$ \ac{BNS} events per
-year~\cite{2013arXiv1304.0670L,2010CQGra..27q3001A}~\footnote{We should
-briefly mention why the rates differ by a factor of 2.  I can't remember right
-now.}. In~\cite{2012MNRAS.425.2668C} using the rate of observed \acp{sGRB} and
+year~\cite{2013arXiv1304.0670L,2010CQGra..27q3001A}\footnote{We note that this rate is that given 
+in Table 1 of \cite{2013arXiv1304.0670L} for an \ac{aLIGO} and \ac{AdV} network consisting to two 
+\ac{aLIGO} detectors operating at design sensitivity. However, a higher rate of $0.4$--$400$ is 
+given in \cite{2010CQGra..27q3001A}, which may differ due to the use of a third \ac{aLIGO} 
+detector, but the discrepancy may also have contributions from slightly different threshold 
+definitions or design sensitivity curves used. We also note the study mentioned in 
+Sec.~\ref{sec:intro} that potentially lowers the expected rate by a factor of five.}. 
+In~\cite{2012MNRAS.425.2668C} using the rate of observed \acp{sGRB} and
 assuming a beaming angle of $15^{o}$ a similar event rate of $1$--$180$ is
@@
 little knowledge of the likely location and distance of such an event we have
-performed simulations of multiple events to see the expected distribution in
-the accuracy of the calibration scale factor recovery.
+performed simulations of multiple events to see the distribution in the accuracy of the calibration 
+scale factor recovery.
 
 % Describe the signal model (Taylor F2 etc...)
-We use the TaylorF2 waveform approximation (see e.g.~\cite{2009PhRvD..80h4043B}
+For our signals we use the TaylorF2 waveform approximation (see e.g.~\cite{2009PhRvD..80h4043B}
 and references therein) with a 3.5 post-Newtonian expansion in phase for
 modelling our signal (in both simulations and signal recovery).  For \ac{BNS}
-systems we use non-spinning waveforms under the assumption that spins will be
-negligible~\cm{ref?}. For \ac{NSBH} system we use a spinning waveform, but in which only
-the black hole has non-negligible spin and that its rotational angular momentum
-is aligned with the system's orbital angular momentum.
+systems we use non-spinning waveforms under the assumption that the component spins will be small 
+and have a negligible effect on the analysis\footnote{In future studies we could create our 
+simulations including small spins and confirm that it does not have a significant effect on our 
+results if recovery is with non-spinning templates.} (see discussion in \cite{2014ApJ...795..105S} 
+and studies in \cite{2012PhRvD..86h4017B}). For \ac{NSBH} system we use a spinning waveform, but in 
+which only the black hole has non-negligible spin and its rotational angular momentum is aligned 
+with the system's orbital angular momentum.
 
@@
 \left(1+\cos^{2}\iota\right)D_{L}^{-1} \mathcal{M}^{5/6}f^{-7/6}e^{-i\Psi(f,
-\nu, t_\mathrm{c}, \phi_{\mathrm{c}})} \nonumber \\ \gws_{\times}(f) &\propto
-\left(\cos^{2}\iota\right)D_{L}^{-1}\mathcal{M}^{5/6} f^{-7/6}e^{-i\Psi(f, \nu,
+\mathcal{M}, t_\mathrm{c}, \phi_{\mathrm{c}})} \nonumber \\ \gws_{\times}(f) &\propto
+\left(\cos^{2}\iota\right)D_{L}^{-1}\mathcal{M}^{5/6} f^{-7/6}e^{-i\Psi(f, \mathcal{M},
 t_\mathrm{c}, \phi_{\mathrm{c}})} 
@@
 $M=m_{1}+m_{2}$ where $m_{1}$ and $m_{2}$ are the component masses, $\iota$ is
-the inclination angle, and $\nu=(\pi Mf)^{1/3}$.  The \ac{GW} strain measured
+the inclination angle.  The \ac{GW} strain measured
 at the $k^{\rm{th}}$ detector is then given by
@@
 % Define the detectors and the noise epoch we will be using
-For this analysis we will consider the advanced generation \ac{aLIGO} and
+For this analysis we consider the advanced generation \ac{aLIGO} and
 \ac{AdV} detectors operating at their design sensitivity, with noise power
@@
 Gaussian) strain noise, $\btheta$ is the set of waveform parameters and
-$\scf_k$ is the calibration scale factor for the $k^{\rm th}$ detector, the
+$\scf_k$ is the calibration scale factor for the $k^{\rm th}$ detector, which are the
 parameters we are interested in estimating.
@@
 the signal amplitude. The most important of these is that the \ac{sGRB} can
-provide a very tight constraint on the source distance\footnote{For \ac{BNS}
-and \ac{NSBH} systems they will be observable only in the relatively nearby
-Universe ($z \sim 0.1$ and $z\sim 0.2$ for \ac{BNS} and \ac{NSBH} systems
-respectively), so the relation between the measured \ac{sGRB} redshift and
-distance is simple and our luminosity distance will be negligibly effected by
-the choice of cosmology \MP{Check this?}.}. In this analysis we therefore
+provide a very tight constraint on the source distance. In this analysis we therefore
 assume that the source distance is known (i.e.\ has a $\delta$-function
@@
 angle corresponds to $\sigma \sim 14.8^{\circ}$. We therefore place a Gaussian
-prior on the inclination angle with zero mean and a standard deviation
-$\sigma=14.8^{\circ}$ truncated at values below zero.
+prior on the inclination angle $\iota$ with zero mean and a standard deviation
+$\sigma=14.8^{\circ}$.
 
@@
 larger or smaller than unity by an equivalent factor, e.g.\ $p(\scf|I)=10$ is
-equivelent to $p(\scf|I)=0.1$. For the scale factor in each detector we 
+equivalent to $p(\scf|I)=0.1$. For the scale factor in each detector we 
 therefore use the following log-normal distribution as our prior
@@
 probability density at 0.1 and 10 is one tenth of that at 1. Note that this
-prior does not give a symmetric amount of probability about unity, but given
+prior does not give a symmetric amount of probability about unity (so out prior favour \scf\xspace 
+values greater than 1), but given
 our simulation criterion described in Section~\ref{sec:simulations} we find
@@
 L1) and the \ac{AdV} detector (V1) and were all performed in the frequency
-domain and spanned a frequency range from 20\,Hz to 400\,Hz~\cm{can you remind
-me and maybe the reader of why a 400 Hz cutoff is OK?}. We assumed all
+domain and spanned a frequency range from 20\,Hz to 400\,Hz, which is a range that contains the 
+vast majority of the \ac{SNR} and will therefore be all that is required to provide good 
+constraints on \scf. We assumed all
 detectors were operating at their design sensitivity (as given by Figure~1 of
@@
 Separately for the \ac{BNS} and \ac{NSBH} systems we have performed a large
-number~\cm{how many exactly/roughly?} simulations at each distance value. When
+number (several hundred up to $O(1000)$) simulations at each distance value. When
 proposing injections for each simulation the source parameters were randomly
@@
 uniformly distributed over the sky and instead favoured locations with better
-antenna response.  Similarly, despite already constrained by the prior to small
+antenna response.  Similarly, despite being already constrained by the prior to small
 inclination angles, the accepted injections show preference for sources very
@@
 an estimate of the noise \ac{PSD} based on the advanced detector design
-sensitivities (using those given in~\cite{2013arXiv1304.0670L}), but simulated
-as if calculated by averaging 64 separate noisy \ac{PSD} estimates and scaled
-with the same factor as applied to the injection and noise~\cm{Do you estimate
-the PSD using 64 segments? This sentence might be a bit confusing to the
-reader.}\footnote{We do not account for there being a potential difference
+sensitivities (using those given in~\cite{2013arXiv1304.0670L}), but calculated by averaging 
+64 separate noisy realisations of the \ac{PSD} and scaled
+with the same factor as applied to the injection and noise\footnote{We do not account for there 
+being a potential difference
 between the estimated \ac{PSD} and the actual \ac{PSD} of the analysed section
@@
 % discuss the Virgo contribution and the simple estimate
-The largest \ac{SNR} contribution will generally come from either of the
+The largest \ac{SNR} contribution will generally come from the two
 \ac{aLIGO} detectors and thus the detection criteria (\ac{SNR} threshold) will
-not apply to to\ac{adV} result. Hence, the \ac{SNR} in \ac{AdV} can be small
+not apply to \ac{AdV} result. Hence, the \ac{SNR} in \ac{AdV} can be small
 and thus the ability to constrain its scale factor becomes poor (although it
@@
 10$\,M$_{\odot}$. We expect these would produce similar results to those we see
-in figure~\ref{fig:nsbhresults}, but again extending to further distances.
+in figure~\ref{fig:nsbhresults}, but again extending to further distances. However, at higher 
+masses we would encounter the problem that the chirp mass will be less well constrained due to 
+there being fewer observable cycles during the inspiral stage.
 Including the merger and ring-down phase, which we currently ignore, will also
@@
 major simplification made here is that the absolute calibration is constant
-with frequency.  Other authors~\cm{Ref: Cornish? and others?} have addressed
+with frequency.  Other authors (e.g.\ \cite{Vitale:2012,2013PhRvD..88h4044L} have addressed
 this issue in the context of robust parameter estimation for astrophysical
@@
 % comparitive calibration
-We imagine that a test for consistency of calibration between detectors could be
-performed with an ensemble of \ac{CBC} sources without relying on \ac{EM}
-counterparts. Such a method would have to marginalise over highly
-uncertain population models and may give quite weak bounds on the overall
-calibration uncertainty.~\cm{I'm not sure what you mean here? I want to add a
-paragraph on relative calibration after I understand what you mean.}
+%We imagine that a test for consistency of calibration between detectors could be
+%performed with an ensemble of \ac{CBC} sources without relying on \ac{EM}
+%counterparts. Such a method would have to marginalise over highly
+%uncertain population models and may give quite weak bounds on the overall
+%calibration uncertainty.~\cm{I'm not sure what you mean here? I want to add a
+%paragraph on relative calibration after I understand what you mean.}
+
+\MP{Chris: feel free to add a short paragraph on relative calibration here.}
 
@@
 highly correlated with any uncertainty in the \ac{PSD} estimate, but would
-still offer an upper limit on the calibration uncertainty. However, future
-plans for \ac{CBC} parameter estimation include estimating the \ac{PSD} with
-the signal using BayesLine \cite{2015PhRvD..91h4034L}~\cm{I'm nit sure that
-this last point really helps us.}.
+still offer an upper limit on the calibration uncertainty. 
+%However, future
+%plans for \ac{CBC} parameter estimation include estimating the \ac{PSD} with
+%the signal using BayesLine \cite{2015PhRvD..91h4034L}~\cm{I'm nit sure that
+%this last point really helps us.}.
 

### 2015-10-07 docs/cosmic_cal.tex
Added short section on LISA verification binaries

@@
 
-\newcommand{\dcc}{LIGO-PXXXXXXXX}
+\newcommand{\dcc}{LIGO-P1500198}
 \newcommand{\cm}[1]{\textcolor{red}{CM: #1}}
@@
 re-calibrate (or validate) the existing experimentally obtained calibration.
+
+For future planned space-based detectors such as eLISA there are several known galactic binary 
+systems known as ``verification'' binaries \cite{2006CQGra..23S.809S}, in that their masses and 
+orbital parameters allow them to be guarenteed sources for such detectors. These too could be used 
+for assessing calibiration for such detectors. However, generally for these verification binaries 
+the inclination of the system is not known and distances are fairly uncertain, so an accurate 
+calibration assessment may not be possible. Although there is at least one (J0651) currently known 
+eclipsing white dwarf binary \cite{2041-8205-737-1-L23}, which does provide the system inclination 
+\cite{2012ApJ...757L..21H, 2013ASPC..467...47K} and has an approximately 10\% distance uncertainty 
+\cite{2041-8205-737-1-L23}, which make it an excellent candiate for an astrophysical calibrator at 
+its frequency.
 
@@
 through the targeted follow-up of an observed \ac{GW} or by post-facto matching of \ac{GW} 
-trigger lists with known \ac{sGRB} events.  The likelihood of being able to follow-up \ac{GW} 
+trigger lists with known (or subsequently found \cite{2015ApJS..217....8B}) \ac{sGRB} events.  
+The likelihood of being able to follow-up \ac{GW} 
 events using gamma-ray telescopes with low enough latency to catch a \ac{sGRB} is low.  Compounding 
@@
 MP is funded by the STFC under grant number ST/L000946/1 and CM is funded via
-XXX.
+XXX. This document has LIGO DCC number \dcc.
 

### 2015-10-07 docs/cosmic_cal.tex
Added commnets from meeting.

@@
 \begin{abstract}
-  We present an investigation into the potential of assessing the validity of the 
-calibration of gravitational wave detector outputs through the use of standard
+  We present an investigation assessing the validity of the 
+of gravitational wave detector calibration through the use of gamma-ray
+bursts as standard
   sirens. Such signals, as measured via gravitational wave
@@
   galaxy is identified for a given source then its redshift can be
-  obtained and combined with current knowledge of the cosmological
-  parameters.  This will yield the true luminosity distance and allow
-  the direct comparison with the estimated value.  Discrepancies can
-  then be used to correct the original calibration.~\cm{Update at end}
+  combined with current knowledge of the cosmological
+  parameters yielding the true luminosity distance.  This will then allow
+  a direct comparison with the estimated value and discrepancies can
+  then be used to correct the original calibration.~\MP{Update with results.
+End with a statement comparing to existing method but stressing its independent.}
 \end{abstract}
@@
 distance markers, otherwise known as ``standard sirens'' (analogous to the \ac{EM} standard 
-candles). Standard sirens allow ...
+candles). \cm{Standard sirens allow ... fix me}
 
@@
 \ac{GW} events are the estimated astrophysical rates of both phenomena. Observations of \acp{sGRB} 
-give a rate of XXX\,Mpc$^{-3}$\,yr$^{-1}$ which can be compared to estimates of \ac{BNS} merger 
+give a rate of \cm{XXX}\,Mpc$^{-3}$\,yr$^{-1}$ which can be compared to estimates of \ac{BNS} merger 
 rates of $10^{-8}$--$10^{-5}$\,Mpc$^{-3}$\,yr$^{-1}$ \cite{2010CQGra..27q3001A}, or \ac{NSBH} merger 
@@
 estimates for \ac{NSBH} systems come entirely from population synthesis models as no such systems 
-have been observed.
+have yet been observed.
 %The ranges of rates for \acp{BNS} is obtained from population synthesis models together with 
@@
 
-For future planned space-based detectors such as eLISA there are several known galactic binary 
+\MP{Move this paragraph to conclusions} For future planned space-based detectors such as eLISA there are several known galactic binary 
 systems known as ``verification'' binaries \cite{2006CQGra..23S.809S}, in that their masses and 
@@
 % continued use of sirens
-The use of \ac{GW} events with \ac{sGRB} counterparts for use as a cosmological tool was 
+\cm{Lose, trim or merge this paragraph with previous one.} The use of \ac{GW} events with \ac{sGRB} counterparts for use as a cosmological tool was 
 investigated in~\cite{2010CQGra..27u5006S} and~\cite{2011PhRvD..83b3005Z} with respect to
@@
 compiled into an overall detection rate of $0.2$--$200$ \ac{BNS} events per
-year~\cite{2013arXiv1304.0670L,2010CQGra..27q3001A}\footnote{We note that this rate is that given 
+year~\cite{2013arXiv1304.0670L,2010CQGra..27q3001A}\footnote{\cm{Email Patrick
+S about this} We note that this rate is that given 
 in Table 1 of \cite{2013arXiv1304.0670L} for an \ac{aLIGO} and \ac{AdV} network consisting to two 
@@
 prior on the inclination angle $\iota$ with zero mean and a standard deviation
-$\sigma=14.8^{\circ}$.
+$\sigma=14.8^{\circ}$.~\MP{Check regarding the face/arse orientation issue.}
 
@@
 probability density at 0.1 and 10 is one tenth of that at 1. Note that this
-prior does not give a symmetric amount of probability about unity (so out prior favour \scf\xspace 
-values greater than 1), but given
-our simulation criterion described in Section~\ref{sec:simulations} we find
-that the likelihood generally overwhelms the prior~\cm{I'm not sure about this
-sentence.  Why does the symmetry matter?}. As $\scf$ is a scale factor we also
-have the further constraint that it must be positive and hence the log-normal
-prior is a natural choice.
+prior does not give a symmetric amount of probability about unity (so our prior
+favour \scf\xspace values greater than 1), but given our simulation criterion
+described in Section~\ref{sec:simulations} we find that the likelihood
+generally overwhelms the prior. As $\scf$ is a scale factor we also have the
+further constraint that it must be positive and hence the log-normal prior is a
+natural choice.
 
@@
 and the difference in the \ac{PSD}. As such our results on real data would be
-an upper limit on the calibration scale factor.~\cm{Even though we're hiding
-this in a footnote is it really this bad. How big is the statistical error on
-PSD estimation using 64 segments?}}. This has provided posterior probability
+an upper limit on the calibration scale factor.}. This has provided posterior probability
 distributions on the calibration scale factors for each detector.  Examples of
@@
 sources there is a slight increase in the mean and standard deviation of
-\acp{SNR} of the \ac{NSBH} systems over the \ac{BNS} systems~\cm{can we go any
-further with this?}.
+\acp{SNR} of the \ac{NSBH} systems over the \ac{BNS} systems.
 
@@
 the implementation of this method of calibration assessment.  Importantly
-however, this method would provide a semi-independent consistency check that
+however, this method would provide a \MP{semi-independent} consistency check that
 the calibration is reasonably accurate.  We acknowledge that we have made a
@@
 
-\MP{Chris: feel free to add a short paragraph on relative calibration here.}
+\cm{Chris: feel free to add a short paragraph on relative calibration here.}
 

### 2015-10-08 docs/cosmic_cal.tex
Finished my final edits before LVC circulation.

@@
 % introduce GWs and BNS systems
-It is expected that the advanced generation of interferometric \ac{GW} detectors will detect waves 
-emitted from $O(10s)$ of \acp{CBC}. One such class of these cataclysmic events, the inspiral and
-merger of \ac{BNS} systems will be detected out to a maximum range of $\approx 450$ Mpc. Assuming 
-the current best estimates for the cosmological parameters, this is equivalent to a redshift 
-$z\approx 0.1$. As noted by \cite{1986Natur.323..310S} \ac{CBC} systems can be used as cosmological 
-distance markers, otherwise known as ``standard sirens'' (analogous to the \ac{EM} standard 
-candles). \cm{Standard sirens allow ... fix me}
+It is expected that the advanced generation of interferometric \ac{GW}
+detectors will detect waves emitted from $O(10s)$ of \acp{CBC} per year. One
+such class of these cataclysmic events, the inspiral and merger of \ac{BNS}
+systems will be detected out to a maximum range of $\approx 450$ Mpc. Assuming
+the current best estimates for the cosmological parameters, this is equivalent
+to a redshift $z\approx 0.1$. As noted by~\citet{1986Natur.323..310S},\ac{CBC}
+systems can be used as cosmological distance markers, otherwise known as
+``standard sirens'' (analogous to \ac{EM} standard candles). The nature of
+these sources will allow us to estimate, amongst other parameters, their
+luminosity distance without the need for calibration against the cosmic
+distance ladder.
 
 % BNS systems are GRBs
-It is considered likely that the merger of \ac{BNS} and/or \ac{NSBH} systems, in addition to 
-emitting detectable \acp{GW}, is also the mechanism for producing \acp{sGRB} 
-\cite{1992ApJ...395L..83N}. In this scenario these events produce tightly beamed \ac{EM} emission 
-parallel to the orbital angular 
-momentum vector of the \ac{BNS} system. Additional evidence for the coincidence of \acp{sGRB} with
-\ac{GW} events are the estimated astrophysical rates of both phenomena. Observations of \acp{sGRB} 
-give a rate of \cm{XXX}\,Mpc$^{-3}$\,yr$^{-1}$ which can be compared to estimates of \ac{BNS} merger 
-rates of $10^{-8}$--$10^{-5}$\,Mpc$^{-3}$\,yr$^{-1}$ \cite{2010CQGra..27q3001A}, or \ac{NSBH} merger 
-rates of $6\,\times\,10^{-10}$--$1\,\times\,10^{-6}$\,Mpc$^{-3}$\,yr$^{-1}$ 
-\cite{2010CQGra..27q3001A}. However, the recent work of \cite{2015MNRAS.448..928K}, which studied 
-known galactic \ac{BNS} systems in more detail suggest a rate approximately five times lower. The 
-estimates for \ac{NSBH} systems come entirely from population synthesis models as no such systems 
-have yet been observed.
-%The ranges of rates for \acp{BNS} is obtained from population synthesis models together with 
-%knowledge of the X known \ac{BNS} systems in our galaxy, whilst the range for \acp{NSBH} is 
-%entirely from population synthesis as not such systems have been observed.
+It is considered likely that the merger of \ac{BNS} and/or \ac{NSBH} systems,
+in addition to emitting detectable \acp{GW}, is also the mechanism for
+producing \acp{sGRB} \cite{1992ApJ...395L..83N}. In this scenario these events
+produce tightly beamed \ac{EM} emission parallel to the orbital angular
+momentum vector of the \ac{BNS} system. Additional evidence for the coincidence
+of \acp{sGRB} with \ac{GW} events are the estimated astrophysical rates of both
+phenomena. Observations of \acp{sGRB} give an inferred rate of \ac{sGRB}
+producing \ac{CBC} mergers of $8\times 10^{-9}$--$1.1\times
+10^{-6}$\,Mpc$^{-3}$\,yr$^{-1}$~\cite{2012MNRAS.425.2668C} which can be
+compared to independently obtained estimates of \ac{BNS} merger rates of
+$10^{-8}$--$10^{-5}$\,Mpc$^{-3}$\,yr$^{-1}$ \cite{2010CQGra..27q3001A}, or
+\ac{NSBH} merger rates of
+$6\,\times\,10^{-10}$--$1\,\times\,10^{-6}$\,Mpc$^{-3}$\,yr$^{-1}$
+\cite{2010CQGra..27q3001A}. However, the recent work of
+\cite{2015MNRAS.448..928K}, which studied known galactic \ac{BNS} systems in
+more detail suggest a rate approximately five times lower. The estimates for
+\ac{NSBH} systems come entirely from population synthesis models as no such
+systems have yet been observed.
+%The ranges of rates for \acp{BNS} is obtained from population synthesis models
+%together with knowledge of the X known \ac{BNS} systems in our galaxy, whilst
+%the range for \acp{NSBH} is entirely from population synthesis as not such
+%systems have been observed.
 
-If a single \ac{GW} event is observed in coincidence with a \ac{sGRB} it may be possible to 
-identify the host galaxy of the \ac{BNS} source.  With this information a spectroscopic redshift 
-can be very accurately obtained (see values quoted in e.g. \cite{0004-637X-664-2-1000} and 
-references therein). With knowledge of the redshift, using current best 
-estimates for the Hubble constant and other cosmological parameters, the true luminosity distance 
-can be estimated to $\sim 1\%$ accuracy.  The \ac{GW} measurement acts as a standard siren also 
-giving us a direct measurement of the luminosity distance to the source. The accuracy of such a 
-measurement depends on a number of factors including the accuracy with which the \ac{GW} detector 
-has been calibrated.  Hence, by comparison with the distance estimate from the \ac{sGRB} we can 
+If a single \ac{GW} event is observed in coincidence with a \ac{sGRB} it may be
+possible to identify the host galaxy of the \ac{BNS} source.  With this
+information a spectroscopic redshift can be very accurately obtained (see
+values quoted in e.g. \cite{0004-637X-664-2-1000} and references therein). With
+knowledge of the redshift, using current best estimates for the Hubble constant
+and other cosmological parameters, the true luminosity distance can be
+estimated to $\sim 1\%$ accuracy.  The \ac{GW} measurement acts as a standard
+siren also giving us a direct measurement of the luminosity distance to the
+source. The accuracy of such a measurement depends on a number of factors
+including the accuracy with which the \ac{GW} detector has been calibrated.
+Hence, by comparison with the distance estimate from the \ac{sGRB} we can
 re-calibrate (or validate) the existing experimentally obtained calibration.
@@
 systems known as ``verification'' binaries \cite{2006CQGra..23S.809S}, in that their masses and 
-orbital parameters allow them to be guarenteed sources for such detectors. These too could be used 
-for assessing calibiration for such detectors. However, generally for these verification binaries 
+orbital parameters allow them to be guaranteed sources for such detectors. These too could be used 
+for assessing calibration for such detectors. However, generally for these verification binaries 
 the inclination of the system is not known and distances are fairly uncertain, so an accurate 
@@
 \cite{2012ApJ...757L..21H, 2013ASPC..467...47K} and has an approximately 10\% distance uncertainty 
-\cite{2041-8205-737-1-L23}, which make it an excellent candiate for an astrophysical calibrator at 
+\cite{2041-8205-737-1-L23}, which make it an excellent candidate for an astrophysical calibrator at 
 its frequency.
@@
 different signal arrival times at each detector in the network allows
-reasonably accurate sky position determination.  Finally, amplitude variation
-between differently oriented interferometers allows some (weaker) level of
-determination of the binary system orientation.
-%A great help to this final
-%stage for this project is the assumption that the source has an \ac{sGRB}
-%counterpart.  It follows that we therefore have a well constrained prior belief
-%on the inclination of the source orbit with respect to the line-of-sight based
-%on the expected level of beaming in the \ac{sGRB} emission.
+reasonably accurate sky position determination~\cite{2014ApJ...795..105S}.
+Finally, amplitude variation between differently oriented interferometers
+allows some (weaker) level of determination of the binary system orientation.
+%A great help to this final stage for this project is the assumption that the
+%source has an \ac{sGRB} counterpart.  It follows that we therefore have a well
+%constrained prior belief on the inclination of the source orbit with respect
+%to the line-of-sight based on the expected level of beaming in the \ac{sGRB}
+%emission.
 
@@
 relative luminosity distance measures and require calibration via other methods
-as part of the cosmological distance ladder. At first sight the missing
-component for standard sirens is a complementary measurement of the source
-redshift which would allow each event to inform the distance-redshift
-relationship and hence obtain cosmological parameter estimates.  However, it
-was not thought possible (until recently) that redshift measurements could be
-obtained from \ac{GW} observations. In~\cite{DelPozzo:2011uf} (building on the
-original idea proposed in~\cite{1986Natur.323..310S}) the idea of identifying
-host galaxies within the sky position uncertainty region was developed. Each
-galaxy would provide a redshift measure and would contribute to a disparate mix
-of possible cosmological parameters. However, by combining multiple (${\sim}20$)
-detections, the true set of cosmological parameters emerges with an expected
-${\sim}5\%$ uncertainty on the Hubble constant.
-
-% continued use of sirens
-\cm{Lose, trim or merge this paragraph with previous one.} The use of \ac{GW} events with \ac{sGRB} counterparts for use as a cosmological tool was 
-investigated in~\cite{2010CQGra..27u5006S} and~\cite{2011PhRvD..83b3005Z} with respect to
-the 3rd generation \ac{GW} interferometer, the \ac{ET}. It was
-assumed in both studies that 1000 coincident events would be observed and
-cosmological parameter estimation ability was compared with current \ac{EM}
-results with focus on dark energy parameters. Other methods for cosmological
-inference (see~\cite{1996PhRvD..53.2878F,2012PhRvD..85b3535T, 2012PhRvD..86b3502T}) have 
-been proposed for \ac{BNS} systems that do
-not rely on any redshift measurements and instead use the distribution of
-measured \acp{SNR} combined with some assumptions regarding the \ac{NS} mass
-and spatial distributions. Finally, we mention that for \ac{BNS} systems it has
-been shown in~\cite{Messenger:2011ux} and~\cite{2013arXiv1312.1862M} that features
-of the tidal and post-merger hyper-massive \ac{NS} stages of the \ac{GW}
-waveform can allow the measurement of the source redshift if the \ac{NS}
-equation-of-state is sufficiently well known. This development, and separately, the method
-based on \ac{SNR} distributions will allow \ac{GW} cosmological inference
-without the need for complementary \ac{EM} observations.
+as part of the cosmological distance ladder. The use of \ac{GW} events with
+\ac{sGRB} counterparts for use as a cosmological tool was investigated
+in~\cite{2010CQGra..27u5006S} and~\cite{2011PhRvD..83b3005Z} with respect to
+the 3rd generation \ac{GW} interferometer, the \ac{ET}. In this case, the
+redshift information obtained from the \ac{sGRB} host galaxy, provides a
+complementary measurement allowing each event to inform the distance-redshift
+relationship and hence obtain cosmological parameter estimates. Other methods
+for cosmological inference (see~\cite{1996PhRvD..53.2878F,2012PhRvD..85b3535T,
+2012PhRvD..86b3502T,Messenger:2011ux,2013arXiv1312.1862M}) have been proposed
+for \ac{BNS} systems that do not rely on any redshift measurements. These
+instead use either the distribution of measured \acp{SNR} and assumed \ac{NS}
+mass and spatial distributions or exploit the features of the tidal and
+post-merger hyper-massive \ac{NS} stages of the \ac{GW} waveform.
 
@@
 compiled into an overall detection rate of $0.2$--$200$ \ac{BNS} events per
-year~\cite{2013arXiv1304.0670L,2010CQGra..27q3001A}\footnote{\cm{Email Patrick
-S about this} We note that this rate is that given 
-in Table 1 of \cite{2013arXiv1304.0670L} for an \ac{aLIGO} and \ac{AdV} network consisting to two 
-\ac{aLIGO} detectors operating at design sensitivity. However, a higher rate of $0.4$--$400$ is 
-given in \cite{2010CQGra..27q3001A}, which may differ due to the use of a third \ac{aLIGO} 
-detector, but the discrepancy may also have contributions from slightly different threshold 
-definitions or design sensitivity curves used. We also note the study mentioned in 
-Sec.~\ref{sec:intro} that potentially lowers the expected rate by a factor of five.}. 
-In~\cite{2012MNRAS.425.2668C} using the rate of observed \acp{sGRB} and
-assuming a beaming angle of $15^{o}$ a similar event rate of $1$--$180$ is
-obtained. The beaming angle is the primary factor in determining the fraction
-of detected \ac{GW} \ac{BNS} signals with an \ac{EM} \ac{sGRB} counterpart and
-is unfortunately highly uncertain.  Various authors have provided estimates on
-the rate of joint detections to be in the range $0.02$--$7$ per
+year~\cite{2013arXiv1304.0670L,2010CQGra..27q3001A}\footnote{We note that this
+rate is that given in Table 1 of \cite{2013arXiv1304.0670L} for an \ac{aLIGO}
+and \ac{AdV} network consisting to two \ac{aLIGO} detectors operating at design
+sensitivity. However, a higher rate of $0.4$--$400$ is given in
+\cite{2010CQGra..27q3001A}, which may differ due to the use of a third
+\ac{aLIGO} detector, but the discrepancy may also have contributions from
+slightly different threshold definitions or design sensitivity curves used. We
+also note the study mentioned in Sec.~\ref{sec:intro} that potentially lowers
+the expected rate by a factor of five.}.  In~\cite{2012MNRAS.425.2668C} using
+the rate of observed \acp{sGRB} and assuming a beaming angle of $15^{o}$ a
+similar event rate of $1$--$180$ is obtained. The beaming angle is the primary
+factor in determining the fraction of detected \ac{GW} \ac{BNS} signals with an
+\ac{EM} \ac{sGRB} counterpart and is unfortunately highly uncertain.  Various
+authors have provided estimates on the rate of joint detections to be in the
+range $0.02$--$7$ per
 year~\cite{2013PhRvL.111r1101C,Kelley:2012fl,2014arXiv1405.2254S,2015MNRAS.448.3026W}.
@@
 The main aim of this work is to assess how well the calibration scale factor
-defined in Eqn.~\ref{eq:scalefactor} can be estimated from a single observed
+defined in Eq.~\ref{eq:scalefactor} can be estimated from a single observed
 \ac{GW} associated with a particular \ac{sGRB}. As we {\it a priori} have
@@
 assume that the source distance is known (i.e.\ has a $\delta$-function
-prior).~\cm{Should we not say a little bit more about this? A really accurate
-distance from a really accurate redshift is only as good as the Hubble constant
-which currently has $\sim$ few percent error?  We should also say something
-about peculiar motion.}
+prior) whereas in reality the error on this distance will be dependent
+upon the uncertainties in both the redshift measurement and the Hubble
+constant. For spectroscopic redshifts this uncertainty will be dominated by the
+Hubble constant and will be of $O(1\%)$.
 
@@
 for the neutron star, but for the black hole we use a prior based on the
-canonical masss distribution used for the rate results in
+canonical mass distribution used for the rate results in
 \cite{2012PhRvD..85h2002A} with a mean of $5\,\textrm{M}_{\odot}$ and a
@@
 Separately for the \ac{BNS} and \ac{NSBH} systems we have performed a large
-number (several hundred up to $O(1000)$) simulations at each distance value. When
-proposing injections for each simulation the source parameters were randomly
-drawn from the prior distributions described in Section~\ref{sec:priors}, with
-the sky position being drawn randomly from a uniform distribution on the sky,
-the coalescence times held fixed within the centre of the prior window, and
-with the calibration scaling factors for each detector drawn from a Gaussian
-distribution with a mean of one and standard deviation of $0.125$ (equivalent
-to a mean fractional offset of 10\%)~\cm{should we explicitly acknowledge that
-we do not simulate from the same priors that we assume in the recovery?}.
-However, in accepting a proposed injection as one to be analysed we introduced
-a criteria that the signal be ``detectable'', based on that used in
+number (several hundred up to $O(1000)$) of simulations at each distance value.
+When proposing injections for each simulation the source parameters were
+randomly drawn from the prior distributions described in
+Section~\ref{sec:priors}, with the sky position being drawn randomly from a
+uniform distribution on the sky, the coalescence times held fixed within the
+centre of the prior window, and with the calibration scaling factors for each
+detector drawn from a Gaussian distribution with a mean of one and standard
+deviation of $0.125$ (equivalent to a mean fractional offset of 10\%).  Note
+that this distribution is broadly consistent with the scale of calibration
+uncertainties that are expected via experimental methods but not the same as
+the prior that we assume on $\mathcal{C}$ (see Eq.~\ref{eq:scale_prior}) when
+performing our analysis.
+
+% describe the detection cut 
+In accepting a proposed injection as one to be analysed we introduced a
+criteria that the signal be ``detectable'', based on that used in
 \cite{2012PhRvD..85h2002A}. We therefore only analysed injections that had an
@@
 uniformly distributed over the sky and instead favoured locations with better
-antenna response.  Similarly, despite being already constrained by the prior to small
-inclination angles, the accepted injections show preference for sources very
-close to circularly polarised.
+antenna response.  Similarly, despite being already constrained by the prior to
+small inclination angles, the accepted injections show preference for sources
+very close to circularly polarised.
 
@@
 value. Also shown on each plot as the dashed magenta line is the percentage of
-signals drawn from the prior distribution that fulfil the detection criterion.
+signals drawn from the prior distribution that fulfill the detection criterion.
 
@@
 maximum at $\sim 25\%$, whilst the width of boxes narrows. This is due to our
-``detectability'' criteria, where for all distances we will only be seeing
+``detectibility'' criteria, where for all distances we will only be seeing
 those sources with a high enough \ac{SNR} that we would consider them
@@
 unknown scale factor on the true signal and noise, which we estimate given the
-data from the 3 inteferometer advanced detector network. We find that for
+data from the 3 interferometer advanced detector network. We find that for
 detectable \ac{BNS} sources the uncertainty on the calibration scale factor
@@
 
-% comparitive calibration
+% comparative calibration
 %We imagine that a test for consistency of calibration between detectors could be
@@
 
-\cm{Chris: feel free to add a short paragraph on relative calibration here.}
+% comparative calibration
+A further application of astrophysical calibration is the relative calibration
+of \ac{GW} detectors.  In this case we would be interested in obtaining
+posterior distributions on the ratios of calibration scale factors between
+detectors. For \ac{CBC} events, in this scenario there would be no explicit
+requirement for a redshift measurement and therefore a cosmologically inferred
+distance. There would however be a requirement for an electromagnetically
+inferred sky position of $O(\rm{degrees})$ accuracy. A more suitable source for
+relative calibration estimation would be that of continuous \ac{GW} emission.
+In that case, sky position and orientation information would be obtained to a
+high level of accuracy from single detector analyses alone. A primary use for
+relative calibration information would be to account for potential biases in
+the sky position estimates for \ac{CBC} and burst sources. In this case,
+amplitude information is a secondary, but important factor, in breaking
+degeneracies on sky position estimates.
 
-% discus PSD uncertainty
+% discuss PSD uncertainty
 We emphasise that our work has assumed that the \ac{PSD} used in the likelihood
@@
 then the \ac{SNR} contribution in each frequency section will be of relevance.
-In this case Bayesian model selection may be applicabl for selecting the
+In this case Bayesian model selection may be applicable for selecting the
 optimal number of frequency sections with which to estimate the scale factor.
@@
 
-We would like to acknowledge the useful discussions with a whole bunch of people. LW was part 
-funded for this work through a Royal Astronomical Society Undergraduate Research Bursary.
-MP is funded by the STFC under grant number ST/L000946/1 and CM is funded via
-XXX. This document has LIGO DCC number \dcc.
+We would like to acknowledge useful discussions within the LIGO-Virgo
+Collaboration and specifically with Salvatore Vitale. LW was part funded for
+this work through a Royal Astronomical Society Undergraduate Research Bursary.
+MP and CM are funded by the Science and Technology Research Council under grant
+number ST/L000946/1. CM is also supported by a Glasgow University Lord Kelvin
+Adam Smith Fellowship. This document has LIGO DCC number \dcc.
 

### 2015-10-09 docs/cosmic_cal.tex
Some final changes

@@
 \begin{abstract}
-  We present an investigation assessing the validity of the 
-of gravitational wave detector calibration through the use of gamma-ray
-bursts as standard
+  We investigate a method to assess the validity of gravitational wave detector calibration 
+through the use of gamma-ray bursts as standard
   sirens. Such signals, as measured via gravitational wave
-  observations provide an estimated luminosity distance which is
+  observations, provide an estimated luminosity distance that is
   subject to uncertainties in the calibration of the data.  If a host
@@
   parameters yielding the true luminosity distance.  This will then allow
-  a direct comparison with the estimated value and discrepancies can
-  then be used to correct the original calibration.~\MP{Update with results.
-End with a statement comparing to existing method but stressing its independent.}
+  a direct comparison with the estimated value and can validate the accuracy of the original 
+calibration. We use simulations of individual detectable gravitational wave signals from binary 
+neutron star (BNS) or neutron star-black hole (NSBH) systems, which we assume to be found in 
+coincidence with short gamma-ray bursts, to estimate any discrepancy in the overall scaling of the 
+calibration for detectors in the Advanced LIGO and Advanced Virgo network. We find that the 
+amplitude scaling of the calibration for the LIGO instruments could on average be confirmed to 
+within $\sim 10\%$ for a BNS source within 100\,Mpc. This result is largely independent of the 
+current detector calibration method and gives an uncertainty that is competitive with that expected 
+in the current calibration procedure. Confirmation of the calibration accuracy to within 
+$\sim 20\%$ can be found with BNS sources out to $\sim 500$\,Mpc.
 \end{abstract}
@@
 \pacs{04.80.Nn, 95.55.Ym}
-\preprint{LIGO-P15XXXXX}
+\preprint{\dcc}
 
@@
 produce tightly beamed \ac{EM} emission parallel to the orbital angular
-momentum vector of the \ac{BNS} system. Additional evidence for the coincidence
+momentum vector of the system. Additional evidence for the coincidence
 of \acp{sGRB} with \ac{GW} events are the estimated astrophysical rates of both
@@
 $6\,\times\,10^{-10}$--$1\,\times\,10^{-6}$\,Mpc$^{-3}$\,yr$^{-1}$
-\cite{2010CQGra..27q3001A}. However, the recent work of
+\cite{2010CQGra..27q3001A}\footnote{However, the recent work of
 \cite{2015MNRAS.448..928K}, which studied known galactic \ac{BNS} systems in
-more detail suggest a rate approximately five times lower. The estimates for
-\ac{NSBH} systems come entirely from population synthesis models as no such
-systems have yet been observed.
+more detail suggest a rate approximately five times lower.}.
+%The estimates for
+%\ac{NSBH} systems come entirely from population synthesis models as no such
+%systems have yet been observed.
 %The ranges of rates for \acp{BNS} is obtained from population synthesis models
@@
 If a single \ac{GW} event is observed in coincidence with a \ac{sGRB} it may be
-possible to identify the host galaxy of the \ac{BNS} source.  With this
+possible to identify the host galaxy of the source.  With this
 information a spectroscopic redshift can be very accurately obtained (see
-values quoted in e.g. \cite{0004-637X-664-2-1000} and references therein). With
+values quoted in e.g.\ \cite{0004-637X-664-2-1000} and references therein). With
 knowledge of the redshift, using current best estimates for the Hubble constant
@@
 
-\MP{Move this paragraph to conclusions} For future planned space-based detectors such as eLISA there are several known galactic binary 
-systems known as ``verification'' binaries \cite{2006CQGra..23S.809S}, in that their masses and 
-orbital parameters allow them to be guaranteed sources for such detectors. These too could be used 
-for assessing calibration for such detectors. However, generally for these verification binaries 
-the inclination of the system is not known and distances are fairly uncertain, so an accurate 
-calibration assessment may not be possible. Although there is at least one (J0651) currently known 
-eclipsing white dwarf binary \cite{2041-8205-737-1-L23}, which does provide the system inclination 
-\cite{2012ApJ...757L..21H, 2013ASPC..467...47K} and has an approximately 10\% distance uncertainty 
-\cite{2041-8205-737-1-L23}, which make it an excellent candidate for an astrophysical calibrator at 
-its frequency.
-
 % What do we do in this paper
@@
 Secs.~\ref{sec:sirens} and~\ref{sec:GRB}. In Secs.~\ref{sec:analysis} and \ref{sec:results} we 
-discuss out analysis method and the results, and in Sec.~\ref{sec:discussion} we 
+discuss our analysis method and the results, and in Sec.~\ref{sec:discussion} we 
 conclude.
@@
 therein. For \ac{aLIGO} and \ac{AdV}, at the frequency range used in this project, the error in the 
-amplitude calibration is roughly $10\%$ \cite{Vitale:2012}. This is a benchmark set for the error 
-estimation using the proposed method in this project.
+amplitude calibration is expected to be roughly $10\%$ \cite{Vitale:2012}. This is a benchmark set 
+for the error estimation using the proposed method in this project.
 
@@
 \end{equation}
-where where we assume $R$ varies only slowly in time (in comparison to transient signal 
+where we assume $R$ varies only slowly in time (in comparison to transient signal 
 time-scales). Calculation of $R$ requires the measurement of various functions within a control 
@@
 % what are sGRBs 
-It is believed that \acp{sGRB} (those \acp{GRB} with duration ${<}2$ sec) are
+It is believed that \acp{sGRB} (those gamma-ray bursts with duration ${<}2$ sec) are
 emitted during the merger of
@@
 the expected rate by a factor of five.}.  In~\cite{2012MNRAS.425.2668C} using
-the rate of observed \acp{sGRB} and assuming a beaming angle of $15^{o}$ a
+the rate of observed \acp{sGRB} and assuming a beaming angle of $15^{\circ}$ a
 similar event rate of $1$--$180$ is obtained. The beaming angle is the primary
@@
 For a non-spinning system the general form of the frequency-domain polarisation
-amplitudes are
+amplitudes is
 %
@@
 \left(1+\cos^{2}\iota\right)D_{L}^{-1} \mathcal{M}^{5/6}f^{-7/6}e^{-i\Psi(f,
-\mathcal{M}, t_\mathrm{c}, \phi_{\mathrm{c}})} \nonumber \\ \gws_{\times}(f) &\propto
+\mathcal{M}, t_\mathrm{c}, \phi_{\mathrm{c}})}, \nonumber \\
+\gws_{\times}(f) &\propto
 \left(\cos^{2}\iota\right)D_{L}^{-1}\mathcal{M}^{5/6} f^{-7/6}e^{-i\Psi(f, \mathcal{M},
@@
 where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=M\eta^{3/5}$,
-with the symmetric mass ratio $\eta=m_{1}m_{2}/M^{2}$.  The total mass
-$M=m_{1}+m_{2}$ where $m_{1}$ and $m_{2}$ are the component masses, $\iota$ is
-the inclination angle.  The \ac{GW} strain measured
+with the symmetric mass ratio $\eta=m_{1}m_{2}/M^{2}$, total mass
+$M=m_{1}+m_{2}$ (where $m_{1}$ and $m_{2}$ are the component masses), and $\iota$ is
+the inclination angle.  The Fourier transform of the \ac{GW} strain measured
 at the $k^{\rm{th}}$ detector is then given by
@@
 constrained as they are based on only a few jet-break detections. A comparative
-study of estimates and lower-limits of opening angles in 13 \acp{sGRB} given
-in~\cite{2014ApJ...780..118F} provide a median opening angle value of $\sim
+study of estimates and lower-limits of opening angles for 13 \acp{sGRB} given
+in~\cite{2014ApJ...780..118F} provides a median opening angle value of $\sim
 10^{\circ}$. We assume that the beam opening angle is a half-normal
@@
 prior on the inclination angle $\iota$ with zero mean and a standard deviation
-$\sigma=14.8^{\circ}$.~\MP{Check regarding the face/arse orientation issue.}
+$\sigma=14.8^{\circ}$. This, and the choice of polarisation angle prior range given below, 
+explicitly fixes the handedness of the system's rotation (or alternatively whether it is face-on or 
+back-on). If we were to allow the inclination to have two modes (one for face-on and one 
+for back-on), or alternatively to double the range of the polarisation angle, it would have no 
+effect on our ability to estimate the signal amplitude as the two modes should be identical with 
+regards to the amplitude parameter probability distributions. However, we are consistent between the 
+range of simulated signals discussed in Sec.~\ref{sec:simulations} and the initial ranges stated.
 
@@
 known precisely and has been obtained, in conjunction with the source redshift,
-from the identification of the host galaxy.
+from the identification of the host galaxy. This fixes the time delay between detectors to a known 
+value and, along with the polarisation angle, defines the antenna response patterns.
 
@@
 We make the assumption that in general the calibration applied to the detector
-data will be correct, so want to use a prior that is peaked at $\scf_k=1$. We
+data will be close to being correct, so want to use a prior that is peaked at $\scf_k=1$. We
 also assume that the prior \ac{PDF} of the calibration scale factor is either
@@
 prior does not give a symmetric amount of probability about unity (so our prior
-favour \scf\xspace values greater than 1), but given our simulation criterion
+favours \scf\xspace values greater than 1), but given our simulation criterion
 described in Section~\ref{sec:simulations} we find that the likelihood
@@
 that this distribution is broadly consistent with the scale of calibration
-uncertainties that are expected via experimental methods but not the same as
-the prior that we assume on $\mathcal{C}$ (see Eq.~\ref{eq:scale_prior}) when
-performing our analysis.
+uncertainties that are expected via experimental methods, but is not the same as
+the prior that we assume on $\mathcal{C}$ (see Eq.~\ref{eq:scale_prior}) when recovering 
+signals.
 
@@
 uniformly distributed over the sky and instead favoured locations with better
-antenna response.  Similarly, despite being already constrained by the prior to
+antenna response.  Similarly, despite already being constrained by the prior to
 small inclination angles, the accepted injections show preference for sources
@@
 These simulations have allowed us to assess how well on average we would be
-able to constrain the calibration scale for a given detected \ac{CBC}-\ac{sGRB}
+able to constrain the calibration scale for a detected \ac{CBC}-\ac{sGRB}
 coincidence.
@@
 the implementation of this method of calibration assessment.  Importantly
-however, this method would provide a \MP{semi-independent} consistency check that
-the calibration is reasonably accurate.  We acknowledge that we have made a
-number of simplifying assumptions, including that the waveforms we use
-accurately model the true signal, and that the overall calibration remains
-constant over the duration of a signal and its \ac{PSD} estimation. A further
+however, this method would provide an independent\footnote{There are caveats to this independence. 
+We rely on a source having been detected, which most likely implies that calibration is not grossly 
+inaccurate and is fairly consistent between detectors.} consistency check that the calibration is 
+reasonably accurate.  We acknowledge that we have made a number of simplifying assumptions, 
+including that the waveforms we use accurately model the true signal, and that the overall 
+calibration remains constant over the duration of a signal and its \ac{PSD} estimation. A further
 major simplification made here is that the absolute calibration is constant
-with frequency.  Other authors (e.g.\ \cite{Vitale:2012,2013PhRvD..88h4044L} have addressed
+with frequency.  Other authors (e.g.\ \cite{Vitale:2012,2013PhRvD..88h4044L}) have addressed
 this issue in the context of robust parameter estimation for astrophysical
@@
 
+It is also worth noting that astrophysical calibration may be possible for future planned 
+space-based detectors such as eLISA \cite{2013GWN.....6....4A}. For such detectors there are 
+several 
+galactic binary systems known as ``verification'' binaries \cite{2006CQGra..23S.809S}, in that 
+their masses and orbital parameters allow them to be guaranteed sources. These too could be used 
+for assessing detector calibration. However, generally for these verification binaries 
+the inclination of the system is not known and distances are fairly uncertain, so a precise 
+calibration assessment may not be possible. Although there is at least one currently known 
+eclipsing white dwarf binary (J0651) \cite{2041-8205-737-1-L23}, which does provide the system 
+inclination \cite{2012ApJ...757L..21H, 2013ASPC..467...47K} and has an approximately 10\% distance 
+uncertainty \cite{2041-8205-737-1-L23}, which makes it an excellent candidate for an astrophysical 
+calibrator.
+
 \acknowledgements

### 2015-11-06 src/python/simulation_check.py
change triangle to corner plot

@@ -11,7 +11,7 @@ from optparse import OptionParser
 import json
 
 from matplotlib import pyplot as pl
-import triangle
+import corner
 
 # swig lal modules for signal generation
 import lal

### 2015-11-06 src/python/results_script.py
Add P-P plots for 50 and 500 Mpc to results script

@@ -3,14 +3,16 @@
 """
 Script to go through a set of directories and gather up information from all the
 info_* files to produce a histrogram of how well the calibration scale factor
-can be recovered for different distance.
+can be recovered for different distance and the P-P plots of injected versus
+recovered confidence intervals.
 """
 
 import json
 import sys
 import os
 import sys
-
+from scipy.interpolate import interp1d
+from scipy.stats import beta
 from optparse import OptionParser
 
 import numpy as np
@@ -48,22 +50,16 @@ else:
       parser.print_help()
       sys.exit(0)
 
-# a function to get the credible intervals using a greedy binning method
-def credible_interval(dsamples, ci):
-  n, binedges = np.histogram(dsamples, bins=100)
-  dbins = binedges[1]-binedges[0] # width of a histogram bin
-  bins = binedges[0:-1]+dbins/2. # centres of bins
 
-  histIndices=np.argsort(n)[::-1]  # indices of the points of the histogram in decreasing order
-  frac = 0.0
-  j = 0
-  for i in histIndices:
-    frac += float(n[i])/float(len(dsamples))
-    j = j+1
-    if frac >= ci:
-      break
+# get the credible region in which the injected value is found (bounded from zero for convenience)
+def credible_inj(histbins, histvals, injval)
+  # normalise histogram and get cumulative sum
+  cumvals = np.cumsum(histvals/np.sum(histvals))
 
-  return (np.min(bins[histIndices[:j]]), np.max(bins[histIndices[:j]]))
+  if injval <= histbins[0] or injval >= histbins[-1]:
+    return 1. # injection outside posterior
+  else:
+    return cumvals[(np.abs(histbins-injval)).argmin()]
 
 
 pl.rc('text', usetex=True)
@@ -85,6 +81,9 @@ fig, ax = pl.subplots(figsize=(8,5))
 data = []
 rates = []
 
+pp50 = []
+pp500 = []
+
 for i, d in enumerate(dirs):
   print d
 
@@ -95,11 +94,17 @@ for i, d in enumerate(dirs):
   totaltime = 0.
   accepted = 0
 
+  # get info for p-p plots for first and last distance
[…]

### 2015-11-06 src/python/scale_factor_mcmc.py

@@ -320,7 +320,7 @@ the analysis output (a sub-directory based on the pulsar name will be created he
 contains the pulsar information)", metavar="DIR")
 
   parser.add_argument("-O", "--output-samples", dest="outsamps", help="If this flag "
-                    "is set then the samples will be output.", default=False, action="store_true",)
+                    "is set then the samples will be output.", default=False, action="store_true")
 
   parser.add_argument("-g", "--det", dest="dets", type=str,
                       help="An interferometer to be used (multiple interferometers can be set \

### 2015-11-06 docs/cosmic_cal.tex
Started addressing Salvos comments

@@
 interferometer via $h(f,t) = \Delta L(f,t) / L$, where $L$ is the full arm length. Calibration is 
-required to relate the actual measured interferometer error signal output $e(f)$ to $\Delta L$. 
+required to relate the actual measured interferometer error signal output $e(f, t)$ to $\Delta L$. 
 This relation is known as the length response function, $R(f)$, defined such that
@@
 cosmological problem upside-down. We assume that the existing \ac{EM}
-cosmological parameter estimation (specifically for the Hubble constant) is
+cosmological parameter estimation (specifically for the Hubble constant, which is has 
+uncertainties of $O(1\%)$) is
 accurately determined and therefore a known quantity. We then assume that the
@@
 % Describe the signal model (Taylor F2 etc...)
-For our signals we use the TaylorF2 waveform approximation (see e.g.~\cite{2009PhRvD..80h4043B}
+For all our signals, for both simulation and recovery, we use the TaylorF2 waveform approximation (see e.g.~\cite{2009PhRvD..80h4043B}
 and references therein) with a 3.5 post-Newtonian expansion in phase for
@@
 %
-where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=M\eta^{3/5}$,
-with the symmetric mass ratio $\eta=m_{1}m_{2}/M^{2}$, total mass
+where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=Mq^{3/5}$,
+with the symmetric mass ratio $q=m_{1}m_{2}/M^{2}$, total mass
 $M=m_{1}+m_{2}$ (where $m_{1}$ and $m_{2}$ are the component masses), and $\iota$ is
@@
 This frequency range was chosen as the vast majority of the \ac{SNR} for the
-inspiralling signals we use can be found within this range, so going to lower
-or higher frequencies provides very little additional information. The source
+inspiralling signals we use can be found within it (for a typical \ac{BNS}
+system of two $1.4\,{\rm M_{odot}}$ stars, or a \ac{NSBH} system with component masses of
+$5\,{\rm M_{\odot}}$ and $1.4\,{\rm M_{\odot}}$, only $\sim 3-4\%$ more \ac{SNR} is gained
+by using a 10 to 1500\,Hz frequency range), so going to lower
+or higher frequencies provides very little additional information whilst increasing
+the computational cost\footnote{In a real analysis, rather than this case study, the additional
+computational cost of using a larger frequency range may be worthwhile, but in this case we had
+to run many simulations and the computational cost was a limiting factor.}. The source
 parameters have been subsumed into a vector $\btheta$ and the calibration scale
@@
 data will be close to being correct, so want to use a prior that is peaked at $\scf_k=1$. We
-also assume that the prior \ac{PDF} of the calibration scale factor is either
+also (na\"{i}vely) assume that the prior \ac{PDF} of the calibration scale factor is either
 larger or smaller than unity by an equivalent factor, e.g.\ $p(\scf|I)=10$ is
@@
 mode is at unity gives $\mu = 1.15$. With these parameters the prior
-probability density at 0.1 and 10 is one tenth of that at 1. Note that this
+probability density at 0.1 and 10 is one tenth of that at 1 making this a fairly
+conservative prior designed to have little impact on our results. Note that this
 prior does not give a symmetric amount of probability about unity (so our prior
@@
 described in Section~\ref{sec:simulations} we find that the likelihood
-generally overwhelms the prior. As $\scf$ is a scale factor we also have the
-further constraint that it must be positive and hence the log-normal prior is a
-natural choice.
+generally overwhelms the prior and our na\"{i}ve assumptions have little effect.
 
@@
 the prior that we assume on $\mathcal{C}$ (see Eq.~\ref{eq:scale_prior}) when recovering 
-signals.
+signals\footnote{The reason behind this is that we wanted to use a fairly na\"{i}ve and
+conservative prior for signal recovery that had little impact in the results, i.e.\
+to produce results domintaed by the likelihood for these parameters.}.
 
@@
 threshold resulted in the preferential selection of better oriented systems.
-Therefore, at larger distances the population of injected sources was not
-uniformly distributed over the sky and instead favoured locations with better
+Therefore, at larger distances the population of injected
+sources was not uniformly distributed over the sky and instead favoured locations with better
 antenna response.  Similarly, despite already being constrained by the prior to
@@
 borders) for the three detectors (H1, L1 and V1). The injected signal was at a distance of 
-250\,Mpc, had \ac{SNR} of 8.7, 10.9 and 4.1 and percentage uncertainties in the calibration scale 
+250\,Mpc, had \ac{SNR} of 8.7, 10.9 and 4.1 and percentage relative uncertainties in the calibration scale 
 factors of 14\%, 11\% and 27\% for each of the detectors respectively.}
@@
 (H1, L1 and V1). The injected signal was at a distance of 450\,Mpc, had \ac{SNR} of 7.7, 10.9 and 
-4.2 and percentage uncertainties in the calibration scale factors of 17\%, 14\% and 34\% for each of 
+4.2 and percentage relative uncertainties in the calibration scale factors of 17\%, 14\% and 34\% for each of 
 the detectors respectively.}

### 2015-11-06 src/python/results_script.py
results_script.py: fixed some bugs in P-P plot creation

@@ -52,9 +52,9 @@ else:
 
 
 # get the credible region in which the injected value is found (bounded from zero for convenience)
-def credible_inj(histbins, histvals, injval)
+def credible_inj(histbins, histvals, injval):
   # normalise histogram and get cumulative sum
-  cumvals = np.cumsum(histvals/np.sum(histvals))
+  cumvals = np.cumsum(histvals.astype(float)/np.sum(histvals))
 
   if injval <= histbins[0] or injval >= histbins[-1]:
     return 1. # injection outside posterior
@@ -120,7 +120,7 @@ for i, d in enumerate(dirs):
         # divide by two to get the half widths and convert to percentage
         vals.append(100.*(info['Results']['Scale68%CredibleInterval'][k][1]-info['Results']['Scale68%CredibleInterval'][k][0])/(2.*info['InjectionParameters']['scales'][k]))
         if i == 0 or i == len(dirs)-1:
-          ppss.append(credible_inj(histd[1], histd[0], info['InjectionParameters']['scales'][k]))
+          ppss.append(credible_inj(np.array(histd[1]), np.array(histd[0]), info['InjectionParameters']['scales'][k]))
 
     if len(vals) == 3:
       relsf.append(vals)
@@ -231,6 +231,7 @@ alpha = 1.-0.95
 
 for i in range(3):
   # get cumulative histogram of found regions
+  if i == 0: print pp50[:,i]
   ax.hist(pp50[:,i], bins=len(pp50[:,i]), cumulative=True, normed=True, histtype='step', color=boxColors[i])
 
 ax.set_xlabel('Credible interval (CI)')
@@ -259,9 +260,9 @@ for i, v in enumerate(cs):
   errortop[i] = intf(1.-alpha/2.)
   errorbottom[i] = intf(alpha/2.)
 
-pl.fill_between(bins, errorbottom, errortop, alpha=0.25, facecolor='grey', edgecolor'grey')
+pl.fill_between(bins, errorbottom, errortop, alpha=0.25, facecolor='grey', edgecolor='grey')
 
-fig.savefig('pp50Mpc_'+outfile)
+fig.savefig(os.path.join(os.path.dirname(outfile), 'pp50Mpc_'+os.path.basename(outfile)))
 
 fig.clf()
 pl.close(fig)
@@ -295,7 +296,6 @@ for i, v in enumerate(cs):
[…]

### 2015-11-06 src/python/results_script.py
Updated plots to have thin grids

@@ -65,6 +65,9 @@ def credible_inj(histbins, histvals, injval):
 pl.rc('text', usetex=True)
 pl.rc('font', family='serif')
 pl.rc('font', size=14)
+pl.rc('grid', linestyle=':')
+pl.rc('grid', alpha=0.5)
+pl.rc('grid', linewidth=0.5)
 
 dists = [50, 100, 150, 200, 250, 300, 350, 400, 450, 500]
 
@@ -204,6 +207,7 @@ ax.set_ylim((0, 100))
 pl.legend(hs, ['H1', 'L1', 'V1'], loc='best')
 pl.xlabel('distance (Mpc)')
 pl.ylabel('\% calibration scaling error (1$\sigma$ equivalent)')
+ax.grid(True)
 
 ax2 = ax.twinx()
 ax2.plot(dists, rates, 'mo--', markerfacecolor='None', markeredgecolor='m')
@@ -239,7 +243,7 @@ ax.set_ylabel('Cumulative fraction of true values within CI')
 ax.set_xlim((0., 1.))
 ax.set_ylim((0., 1.))
 pl.legend(['H1', 'L1', 'V1'], loc='best')
-
+ax.grid(True)
 ax.plot([0., 1.], [0., 1.], 'k--') # plot diagonal
 
 # plot target confidence band
@@ -262,7 +266,7 @@ for i, v in enumerate(cs):
 
 pl.fill_between(bins, errorbottom, errortop, alpha=0.25, facecolor='grey', edgecolor='grey')
 
-fig.savefig(os.path.join(os.path.dirname(outfile), 'pp50Mpc_'+os.path.basename(outfile)))
+fig.savefig(os.path.join(os.path.dirname(outfile), 'pp50Mpc.pdf'))
 
 fig.clf()
 pl.close(fig)
@@ -278,7 +282,7 @@ ax.set_ylabel('Cumulative fraction of true values within CI')
 ax.set_xlim((0., 1.))
 ax.set_ylim((0., 1.))
 pl.legend(['H1', 'L1', 'V1'], loc='best')
-
+ax.grid(True)
 ax.plot([0., 1.], [0., 1.], 'k--') # plot diagonal
 
 # plot target confidence band
@@ -298,4 +302,4 @@ for i, v in enumerate(cs):
 
 pl.fill_between(bins, errorbottom, errortop, alpha=0.25, facecolor='grey', edgecolor='grey')
 
[…]

### 2015-11-06 src/python/results_script_nsbh.py

@@ -69,6 +69,9 @@ def credible_interval(dsamples, ci):
 pl.rc('text', usetex=True)
 pl.rc('font', family='serif')
 pl.rc('font', size=14)
+pl.rc('grid', linestyle=':')
+pl.rc('grid', alpha=0.5)
+pl.rc('grid', linewidth=0.5)
 
 dists = [100, 200, 300, 400, 500, 600, 700, 800, 900] #, 1000]
 
@@ -185,6 +188,7 @@ ax.set_xticks(dists)
 
 ax.set_xlim((0, 1100))
 ax.set_ylim((0, 100))
+ax.grid(True)
 
 pl.legend(hs, ['H1', 'L1', 'V1'], loc='best')
 pl.xlabel('distance (Mpc)')

### 2015-11-06 src/python/results_script.py
Updated pp500Mpc plot

@@ -235,7 +235,6 @@ alpha = 1.-0.95
 
 for i in range(3):
   # get cumulative histogram of found regions
-  if i == 0: print pp50[:,i]
   ax.hist(pp50[:,i], bins=len(pp50[:,i]), cumulative=True, normed=True, histtype='step', color=boxColors[i])
 
 ax.set_xlabel('Credible interval (CI)')
@@ -275,7 +274,7 @@ fig, ax = pl.subplots(figsize=(6,5))
 
 for i in range(3):
   # get cumulative histogram of found regions
-  ax.hist(pp500[:,i], bins=100, cumulative=True, normed=True, histtype='step', color=boxColors[i])
+  ax.hist(pp500[:,i], bins=len(pp500[:,i]), cumulative=True, normed=True, histtype='step', color=boxColors[i])
 
 ax.set_xlabel('Credible interval (CI)')
 ax.set_ylabel('Cumulative fraction of true values within CI')

### 2015-11-06 docs/cosmic_cal.tex
Added discussion of p-p plots

@@
 \usepackage{xspace}
+\usepackage{subcaption}
 
@@
 \newcommand{\scf}{\ensuremath{\mathcal{C}}}
+\newcommand{\scft}{\scf\xspace}
 
@@
 
+% discuss any bias on the actual recovered values
+\subsection{Parameter biases}
+
+It is also interested to see if our analysis has any bias on the recovered \scft \acp{PDF}, or whether they
+are recovered in a way that is consistent with the simulated values.
+To do this we have produced plots showing the cumulative fraction of the true (simulated)
+\scft values recovered within given posterior credible intervals (see e.g.\ \cite{2014PhRvD..89h4060S})
+for the \ac{BNS} systems at 50 and 500\,Mpc (see figure~\ref{fig:ppplots}).
+Self-consistent \acp{PDF} should show that a fraction of injected values falls within the corresponding credible interval,
+and therefore the cumulative fraction should lie along the diagonal of the plot. Any
+biases would show up as significant deviations from the diagonal.
+
+Figure~\ref{subfig:pp50} shows
+that for sources at 50\,Mpc there is not a very significant bias on the recovery of \scft for any of
+the detectors. However, in figure~\ref{subfig:pp500}, with sources at 500\,Mpc, whilst
+H1 and L1 follow the diagonal well, V1 goes significantly above the diagonal.
+This shows that the \acp{PDF} on \scft for V1 are too broad. The reason for this is that at these
+larger distances the \acp{PDF} on \scft for V1 are not well constrained and start to become dominated
+by the prior. As we have previously noted this prior is not that same as the distribution used
+to generate the \scf values, but is instead far broader and more conservative. So, this bias is
+showing up due to the simulated values of \scft being generated from a much narrow range.
+
+Despite the inconsistency in V1 it is satisfying to see that for the detectors for which the
+\scft values are reasonably constrained (i.e.\ H1 and L1 in figure~\ref{fig:bnsresults}) the
+actual value is recovered consistently, with little effect from the prior.
+
+\begin{figure*}
+\begin{center}
+
+\begin{tabular}{c c}
+\subcaptionbox{\label{subfig:pp50}}{
+ \includegraphics[width=0.475\textwidth]{pp50Mpc}}
+
+&
+
+ \subcaptionbox{\label{subfig:pp500}}{
+ \includegraphics[width=0.475\textwidth]{pp500Mpc}}
+\end{tabular}
+\end{center}
+\caption{\label{fig:ppplots} Each figure shows the cumulative fraction of true \scft values
+found within a
+given fractional credible interval versus fractional credible interval for each detector for
+the \ac{BNS} simulations at (\subref{subfig:pp50}) 50\,Mpc and (\subref{subfig:pp500})
+500\,Mpc. The grey shaded region is a 95\% credible band for the expected deviations from diagonal.}
+\end{figure*}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 this work through a Royal Astronomical Society Undergraduate Research Bursary.
-MP and CM are funded by the Science and Technology Research Council under grant
+MP and CM are funded by the Science and Technology Facilities Council (STFC) under grant
 number ST/L000946/1. CM is also supported by a Glasgow University Lord Kelvin
-Adam Smith Fellowship. This document has LIGO DCC number \dcc.
+Adam Smith Fellowship. We are grateful for computational resources provided by
+Cardiff University, and funded by STFC grant ST/I006285/1 supporting UK Involvement in
+the operation of Advanced LIGO. This document has LIGO DCC number \dcc.
 

### 2015-11-06 docs/cosmic_cal.tex
Added in prep reference for CW work

@@
 relative calibration information would be to account for potential biases in
-the sky position estimates for \ac{CBC} and burst sources. In this case,
+the sky position estimates for \ac{CBC} and burst sources (Vitale, Messenger \& Pitkin, in preparation). In this case,
 amplitude information is a secondary, but important factor, in breaking

### 2015-11-09 src/python/results_script.py
Updated PP plots to not use matplotlib histogram and instead use step to get rid of bad looking plots

@@ -234,8 +234,12 @@ ps = np.linspace(0., 1., 1000)
 alpha = 1.-0.95
 
 for i in range(3):
+  nb, binedges = np.histogram(pp50[:,i], bins=len(pp50[:,i]))
+  cn = np.cumsum(nb)/float(len(pp50)) 
+
   # get cumulative histogram of found regions
-  ax.hist(pp50[:,i], bins=len(pp50[:,i]), cumulative=True, normed=True, histtype='step', color=boxColors[i])
+  #ax.hist(pp50[:,i], bins=len(pp50[:,i]), cumulative=True, normed=True, histtype='step', color=boxColors[i])
+  ax.step(binedges[:-1], cn, color=boxColors[i])
 
 ax.set_xlabel('Credible interval (CI)')
 ax.set_ylabel('Cumulative fraction of true values within CI')
@@ -273,8 +277,12 @@ pl.close(fig)
 fig, ax = pl.subplots(figsize=(6,5))
 
 for i in range(3):
+  nb, binedges = np.histogram(pp500[:,i], bins=len(pp500[:,i]))              
+  cn = np.cumsum(nb)/float(len(pp500))
+
   # get cumulative histogram of found regions
-  ax.hist(pp500[:,i], bins=len(pp500[:,i]), cumulative=True, normed=True, histtype='step', color=boxColors[i])
+  # ax.hist(pp500[:,i], bins=len(pp500[:,i]), cumulative=True, normed=True, histtype='step', color=boxColors[i])
+  ax.step(binedges[:-1], cn, color=boxColors[i])
 
 ax.set_xlabel('Credible interval (CI)')
 ax.set_ylabel('Cumulative fraction of true values within CI')

### 2015-11-09 docs/cosmic_cal.tex
Added some last minor fixes before submission to arXiv

@@
 
-\newcommand{\curlH}{\mathcal{H}}
 \newcommand{\gws}{\tilde{h}}
@@
 \begin{abstract}
-  We investigate a method to assess the validity of gravitational wave detector calibration 
+  We investigate a method to assess the validity of gravitational-wave detector calibration 
 through the use of gamma-ray bursts as standard
-  sirens. Such signals, as measured via gravitational wave
+  sirens. Such signals, as measured via gravitational-wave
   observations, provide an estimated luminosity distance that is
@@
   a direct comparison with the estimated value and can validate the accuracy of the original 
-calibration. We use simulations of individual detectable gravitational wave signals from binary 
+calibration. We use simulations of individual detectable gravitational-wave signals from binary 
 neutron star (BNS) or neutron star-black hole (NSBH) systems, which we assume to be found in 
@@
 % acronym definitions
-\acrodef{GW}[GW]{gravitational wave}
+\acrodef{GW}[GW]{gravitational-wave}
 \acrodef{NS}[NS]{neutron star}
@@
 the current best estimates for the cosmological parameters, this is equivalent
-to a redshift $z\approx 0.1$. As noted by~\citet{1986Natur.323..310S},\ac{CBC}
+to a redshift $z\approx 0.1$. As noted by~\citet{1986Natur.323..310S}, \ac{CBC}
 systems can be used as cosmological distance markers, otherwise known as
@@
 more detail suggest a rate approximately five times lower.}.
-%The estimates for
-%\ac{NSBH} systems come entirely from population synthesis models as no such
-%systems have yet been observed.
-%The ranges of rates for \acp{BNS} is obtained from population synthesis models
-%together with knowledge of the X known \ac{BNS} systems in our galaxy, whilst
-%the range for \acp{NSBH} is entirely from population synthesis as not such
-%systems have been observed.
 
@@
 particular for the LIGO detectors during their fifth science run) can be found in 
-\cite{2010NIMPA.624..223A}, and a brief description is giving in \cite{Vitale:2012} and references 
+\cite{2010NIMPA.624..223A}, and a brief description is given in \cite{Vitale:2012} and references 
 therein. For \ac{aLIGO} and \ac{AdV}, at the frequency range used in this project, the error in the 
@@
 closer) than in reality\footnote{However, the \ac{SNR} would be the correct value as the signal and 
-the noise will both contain the scale factor.}, whereas if $\scf < 1$ it would appear to have a 
+the noise will both contain the same scale factor.}, whereas if $\scf < 1$ it would appear to have a 
 lower amplitude than reality. In this analysis we will simplify the situation by assuming that 
-$\scf$ is a constant with respect to frequency, but in future studies that assumption could be 
+$\scf$ is a constant with respect to frequency (and is real, so has no phase component), but in future studies that assumption could be 
 dropped and $\scf$ could take some functional form, or piece-wise fit, with respect to $f$ (e.g.\ 
@@
 the detector network. However, measurement of the phase evolution of such an
-event allows accurate determination of the chirp mass~\footnote{In reality it
+event allows accurate determination of the chirp mass\footnote{In reality it
 is the redshifted chirp mass that is measured.}. Timing information from the
 different signal arrival times at each detector in the network allows
-reasonably accurate sky position determination~\cite{2014ApJ...795..105S}.
+some level of sky position determination~\cite{2014ApJ...795..105S}.
 Finally, amplitude variation between differently oriented interferometers
 allows some (weaker) level of determination of the binary system orientation.
-%A great help to this final stage for this project is the assumption that the
-%source has an \ac{sGRB} counterpart.  It follows that we therefore have a well
-%constrained prior belief on the inclination of the source orbit with respect
-%to the line-of-sight based on the expected level of beaming in the \ac{sGRB}
-%emission.
 
@@
 unknown in our analysis is the absolute amplitude calibration of the
-interferometers in the global \ac{GW} detector network. The issue of
-calibration has largely been ignored in \ac{GW} cosmological inference studies
-and in Sec.~\ref{sec:discussion} we discuss some of the implications of
-combining these issues in future analyses.
+interferometers in the global \ac{GW} detector network.
 
@@
 only potentially observable for a fraction of the mergers. If the event
-exhibits an optical afterglow then the host galaxy can be identified from which
+exhibits an optical afterglow then the host galaxy can be identified and from which
 a redshift can be obtained (e.g.~\cite{2005Natur.437..851G}).  The fraction of events with 
@@
 \acp{sGRB} is $z{\sim}2$ and the nearest event with associated redshift is at
-$z{\sim}0.1$~\cite{2015GCN..17278...1C} which is approximately equal to the
+$z{\sim}0.1$~\cite{2015GCN..17278...1C}, which is approximately equal to the
 horizon range of the advanced \ac{GW} detector network for \ac{BNS} systems.
@@
 follow-up of \ac{sGRB} scenario the merger time for \ac{BNS}/\ac{NSBH} systems will be estimated 
-from the \ac{sGRB} to within a few seconds (see e.g. the discussion in Sec.\ 2.2 of 
+from the \ac{sGRB} to within a few seconds (see e.g.\ the discussion in Sec.\ 2.2 of 
 \cite{2012ApJ...760...12A}). This makes the follow-up search less 
@@
 % Describe the signal model (Taylor F2 etc...)
-For all our signals, for both simulation and recovery, we use the TaylorF2 waveform approximation (see e.g.~\cite{2009PhRvD..80h4043B}
+For all our signals we use the TaylorF2 waveform approximation (see e.g.~\cite{2009PhRvD..80h4043B}
 and references therein) with a 3.5 post-Newtonian expansion in phase for
@@
 results if recovery is with non-spinning templates.} (see discussion in \cite{2014ApJ...795..105S} 
-and studies in \cite{2012PhRvD..86h4017B}). For \ac{NSBH} system we use a spinning waveform, but in 
+and studies in \cite{2012PhRvD..86h4017B}). For \ac{NSBH} systems we use a spinning waveform, but in 
 which only the black hole has non-negligible spin and its rotational angular momentum is aligned 
@@
 $M=m_{1}+m_{2}$ (where $m_{1}$ and $m_{2}$ are the component masses), and $\iota$ is
-the inclination angle.  The Fourier transform of the \ac{GW} strain measured
+the binary inclination angle.  The Fourier transform of the \ac{GW} strain measured
 at the $k^{\rm{th}}$ detector is then given by
@@
 \begin{equation} \label{eq:gravsig} 
-\gws^{k}(f) = \big[ F_{+}^{k}(\alpha,
+\gws^{k}(f) = F_{+}^{k}(\alpha,
 \delta, \psi)\gws_{+}(f) + F_{\times}^k(\alpha, \delta,
-\psi)\gws_{\times}(f)\big] 
+\psi)\gws_{\times}(f)
 \end{equation}
@@
 \begin{equation}
-p(\bm{d} | \btheta, \bm{\scf}, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm 
+p(\bm{d} | \btheta, \bm{\scf}, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm 
 det}}
@@
 inspiralling signals we use can be found within it (for a typical \ac{BNS}
-system of two $1.4\,{\rm M_{odot}}$ stars, or a \ac{NSBH} system with component masses of
+system of two $1.4\,{\rm M_{\odot}}$ stars, or a \ac{NSBH} system with component masses of
 $5\,{\rm M_{\odot}}$ and $1.4\,{\rm M_{\odot}}$, only $\sim 3-4\%$ more \ac{SNR} is gained
@@
 distribution, for which the median is given by $\sqrt{2}\erf{}^{-1}(1/2)\sigma
-\sim 0.67\sigma$. It follows that the median of a $\sim 10^{\circ}$ opening
-angle corresponds to $\sigma \sim 14.8^{\circ}$. We therefore place a Gaussian
+\sim 0.67\sigma$. It follows that a median opening angle of a $\sim 10^{\circ}$
+corresponds to $\sigma \sim 14.8^{\circ}$. We therefore place a Gaussian
 prior on the inclination angle $\iota$ with zero mean and a standard deviation
@@
 To estimate how well the calibration scaling can be constrained for each
-detector in an advanced detector network we have performed simulations with
-injected \ac{BNS} and \ac{NSBH} signals. These signal injections were simulated
+detector in an advanced detector network we have performed analyses with simulated
+\ac{BNS} and \ac{NSBH} signals. These were simulated
 at a range of distances: 50\,Mpc to 500\,Mpc with 50\,Mpc increments for the
@@
 % describe the detection cut 
-In accepting a proposed injection as one to be analysed we introduced a
+In accepting a proposed simulation as one to be analysed we introduced a
 criteria that the signal be ``detectable'', based on that used in
-\cite{2012PhRvD..85h2002A}. We therefore only analysed injections that had an
+\cite{2012PhRvD..85h2002A}. We therefore only analysed simulations that had an
 \ac{SNR} of $\geq 5.5$ in at least two detectors. Since we assume that these
@@
 further constraint that the total coherent network \ac{SNR} is greater than 12.
-For injections close to the \ac{SNR} threshold, the application of the
+For simulations close to the \ac{SNR} threshold, the application of the
 threshold resulted in the preferential selection of better oriented systems.
@@
 antenna response.  Similarly, despite already being constrained by the prior to
-small inclination angles, the accepted injections show preference for sources
+small inclination angles, the accepted simulations show preference for sources
 very close to circularly polarised.
@@
 64 separate noisy realisations of the \ac{PSD} and scaled
-with the same factor as applied to the injection and noise\footnote{We do not account for there 
+with the same calibraion scale factor as applied to the injection and noise\footnote{We do not account for there 
 being a potential difference
@@
  parameters of a \ac{BNS} system, including calibration scaling factors (surrounded by the red 
-borders) for the three detectors (H1, L1 and V1). The injected signal was at a distance of 
+borders) for the three detectors (H1, L1 and V1). The simulated signal was at a distance of 
 250\,Mpc, had \ac{SNR} of 8.7, 10.9 and 4.1 and percentage relative uncertainties in the calibration scale 
@@
  parameters of a \ac{NSBH} system, including calibration scaling factors for the three detectors 
-(H1, L1 and V1). The injected signal was at a distance of 450\,Mpc, had \ac{SNR} of 7.7, 10.9 and 
+(H1, L1 and V1). The simulated signal was at a distance of 450\,Mpc, had \ac{SNR} of 7.7, 10.9 and 
 4.2 and percentage relative uncertainties in the calibration scale factors of 17\%, 14\% and 34\% for each of 
@@
  \caption{\label{fig:bnsresults} Distributions of the percentage accuracy at which the calibration 
-scale factors can be determined for a three detector network if assuming
+scale factors can be determined for a three detector network if using
 \ac{BNS} systems (provided a coincident \ac{GRB} is observed and can yield a
@@
  \caption{\label{fig:nsbhresults} Distributions of the percentage accuracy at which the calibration 
-scale factors can be determined for a three detector network if assuming \ac{NSBH} systems 
+scale factors can be determined for a three detector network if using \ac{NSBH} systems 
 (provided a coincident \ac{GRB} is observed and can yield a distance estimate). The plot contents 
@@
 \ac{aLIGO} detectors and thus the detection criteria (\ac{SNR} threshold) will
-not apply to \ac{AdV} result. Hence, the \ac{SNR} in \ac{AdV} can be small
+not apply to the \ac{AdV} result. Hence, the \ac{SNR} in \ac{AdV} can be small
 and thus the ability to constrain its scale factor becomes poor (although it
@@
 Including the merger and ring-down phase, which we currently ignore, will also
-increase \ac{SNR} and act to reduce the calibration scale uncertainties.
+increase \ac{SNR} and act to reduce the calibration scale uncertainties, provided there is
+sufficient \ac{SNR} in the inspiral to constrain the chirp mass.
 
@@
 To do this we have produced plots showing the cumulative fraction of the true (simulated)
-\scft values recovered within given posterior credible intervals (see e.g.\ \cite{2014PhRvD..89h4060S})
+\scft values recovered within given posterior credible intervals (see e.g.\ \cite{2014PhRvD..89h4060S} for the initial use
+of this type of plot to check consistency of \ac{GW} analyes)
 for the \ac{BNS} systems at 50 and 500\,Mpc (see figure~\ref{fig:ppplots}).
@@
 Figure~\ref{subfig:pp50} shows
-that for sources at 50\,Mpc there is not a very significant bias on the recovery of \scft for any of
+that for sources at 50\,Mpc there is no highly significant bias on the recovery of \scft for any of
 the detectors. However, in figure~\ref{subfig:pp500}, with sources at 500\,Mpc, whilst
@@
 by the prior. As we have previously noted this prior is not that same as the distribution used
-to generate the \scf values, but is instead far broader and more conservative. So, this bias is
+to generate the \scft values, but is instead far broader and more conservative. So, this bias is
 showing up due to the simulated values of \scft being generated from a much narrow range.
@@
 \end{center}
-\caption{\label{fig:ppplots} Each figure shows the cumulative fraction of true \scft values
+\caption{\label{fig:ppplots} Both figures shows the cumulative fraction of true \scft values
 found within a
@@
 \end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Distance estimates from \acp{sGRB}\label{sec:cosmo}}
-
-%~\cm{Describe the cosmological analysis used to obtain the distance from the GRB}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Multiple events without counterpart\label{sec:multiple}}
-
-%\cm{A discussion of the issues regarding multiple events
-%without EM counterparts.}
 
@@
 unknown scale factor on the true signal and noise, which we estimate given the
-data from the 3 interferometer advanced detector network. We find that for
+data from the three interferometer advanced detector network. We find that for
 detectable \ac{BNS} sources the uncertainty on the calibration scale factor
@@
 inaccurate and is fairly consistent between detectors.} consistency check that the calibration is 
-reasonably accurate.  We acknowledge that we have made a number of simplifying assumptions, 
+reasonably accurate in addition to the existing calbiration methods \cite{2010NIMPA.624..223A}.  We acknowledge that we have made a number of simplifying assumptions, 
 including that the waveforms we use accurately model the true signal, and that the overall 
@@
 sources and in future work we intend to apply related methods for astrophysical
-calibration.
-
-% comparative calibration
-%We imagine that a test for consistency of calibration between detectors could be
-%performed with an ensemble of \ac{CBC} sources without relying on \ac{EM}
-%counterparts. Such a method would have to marginalise over highly
-%uncertain population models and may give quite weak bounds on the overall
-%calibration uncertainty.~\cm{I'm not sure what you mean here? I want to add a
-%paragraph on relative calibration after I understand what you mean.}
+calibration.  We also have neglected any phase uncertainty in the calibration.
 
@@
 of \ac{GW} detectors.  In this case we would be interested in obtaining
-posterior distributions on the ratios of calibration scale factors between
+posterior distributions on the ratios of calibration scale factors and phases between
 detectors. For \ac{CBC} events, in this scenario there would be no explicit
@@
 being used for marginalising over calibration uncertainties, but do not yet
-attempt to estimate it.
+attempt to estimate them.
 

### 2015-11-09 docs/cosmic_cal.tex
Changes for submission to arXiv

@@
-\IfFileExists{revtex4-1.cls}{\documentclass[prd, twocolumn, lengthcheck,
-superscriptaddress, showpacs, letterpaper, nofootinbib]{revtex4-1}}{
-\IfFileExists{revtex4.cls}{\documentclass[prd, twocolumn, lengthcheck,
-superscriptaddress, showpacs, letterpaper, nofootinbib]{revtex4}}{}
-}
+%\IfFileExists{revtex4-1.cls}{\documentclass[prd, twocolumn, lengthcheck,
+%superscriptaddress, showpacs, letterpaper, nofootinbib]{revtex4-1}}{
+%\IfFileExists{revtex4.cls}{\documentclass[prd, twocolumn, lengthcheck,
+%superscriptaddress, showpacs, letterpaper, nofootinbib]{revtex4}}{}
+%}
+\documentclass[prd, twocolumn, lengthcheck, superscriptaddress, showpacs, letterpaper, nofootinbib]{revtex4-1}
 
@@
 %% ----- input git-version tag
-\input{tag.tex}
+%\input{tag.tex}
 

### 2015-11-12 docs/cosmic_cal.tex
Changes to Figs 1 and 2 - remove the psi, phi and tc parameters to make the other plots larger and conform to PRDs request that the axes labels are bigger

@@
  \end{center}
- \caption{\label{fig:bnspost} The marginalised posterior probability distributions for the unknown
+ \caption{\label{fig:bnspost} The marginalised posterior probability distributions for some of the unknown
  parameters of a \ac{BNS} system, including calibration scaling factors (surrounded by the red 
 borders) for the three detectors (H1, L1 and V1). The simulated signal was at a distance of 
-250\,Mpc, had \ac{SNR} of 8.7, 10.9 and 4.1 and percentage relative uncertainties in the calibration scale 
-factors of 14\%, 11\% and 27\% for each of the detectors respectively.}
+250\,Mpc, had \ac{SNR} of 7.9, 9.2 and 5.2 and percentage relative uncertainties in the calibration scale 
+factors of 17\%, 15\% and 24\% for each of the detectors respectively.}
 \end{figure*}
@@
  \end{center}
- \caption{\label{fig:nsbhpost} The marginalised posterior probability distributions for the unknown
+ \caption{\label{fig:nsbhpost} The marginalised posterior probability distributions for some of the unknown
  parameters of a \ac{NSBH} system, including calibration scaling factors for the three detectors 
-(H1, L1 and V1). The simulated signal was at a distance of 450\,Mpc, had \ac{SNR} of 7.7, 10.9 and 
-4.2 and percentage relative uncertainties in the calibration scale factors of 17\%, 14\% and 34\% for each of 
+(H1, L1 and V1). The simulated signal was at a distance of 450\,Mpc, had \ac{SNR} of 9.8, 10.3 and 
+3.8 and percentage relative uncertainties in the calibration scale factors of 15\%, 14\% and 36\% for each of 
 the detectors respectively.}

### 2015-11-12 src/python/scale_factor_mcmc.py

@@ -834,12 +834,24 @@ if it does not fulfill the SNR criterion.")
     # set t0 to 0
     samples[:,3] = samples[:,3] - t0
 
+    # set Mc to be centred at 0 and in 1/1000 solar mass units
+    samples[:,4] = (samples[:,4] - mC)*1e3
+
+    # for the moment ignore psi, phi0 and t_c in output plots
+    samplesnew = np.delete(samples, 0, 1) # delete psi
+    samplesnew = np.delete(samplesnew, 0, 1) # delete phi0
+    samplesnew = np.delete(samplesnew, 1, 1) # delete t_c
+
     if args.nsbh:
-      labels = ["$\psi$", "$\phi_0$", "$\iota$", "$t_c$", "$\mathcal{M}$", "$q$", "$a_1$"]
-      truths = [psi, phi0, iota, 0.0, mC, q, a1spin]
+      #labels = ["$\psi$", "$\phi_0$", "$\iota$", "$t_c$", "$\mathcal{M}$", "$q$", "$a_1$"]
+      #truths = [psi, phi0, iota, 0.0, mC, q, a1spin]
+      labels = ["$\iota\,\mathrm{rads}$", "$\mathcal{M}-\mathcal{M}_{\mathrm{true}}\,10^{-3}\mathrm{M}_{\odot}$", "$q$", "$a_1$"]
+      truths = [iota, mC-mC, q, a1spin]
     else:
-      labels = ["$\psi$", "$\phi_0$", "$\iota$", "$t_c$", "$\mathcal{M}$", "$q$"]
-      truths = [psi, phi0, iota, 0.0, mC, q]
+      #labels = ["$\psi$", "$\phi_0$", "$\iota$", "$t_c$", "$\mathcal{M}$", "$q$"]
+      #truths = [psi, phi0, iota, 0.0, mC, q]
+      labels = ["$\iota \,\mathrm{rads}$", "$\mathcal{M}-\mathcal{M}_{\mathrm{true}}\,10^{-3}\mathrm{M}_{\odot}$", "$q$"]
+      truths = [iota, mC-mC, q]
     for i in range(len(dets)):
       labels.append("$\mathcal{C}_{\mathrm{%s}}}$" % dets[i])
       truths.append(scales[i])
@@ -847,7 +859,14 @@ if it does not fulfill the SNR criterion.")
     # plot 1, 2 and 3 sigma contours
     levels = 1.-np.exp(-0.5*np.array([1., 2., 3.])**2)
 
[…]

### 2016-02-02 docs/response_to_referee.txt
Added copy of referees report for editing to include responses

----------------------------------------------------------------------
Report of the Referee -- DL11679/Pitkin
----------------------------------------------------------------------

This article proposes a novel method of calibrating ground-based
gravitational-wave detectors by employing compact-binary systems as
standard sirens. I find it to be scientifically interesting, novel,
and well explained. Provided that the authors can address my following
points, I suggest this article for publication subject to minor
revision.

** MAJOR COMMENTS

-- In section 5, I have a problem with the assumption of the sGRB jet
opening angle being modelled as a half-normal distribution. Figure 3
in Fong et al. (2014) shows that the long GRB distribution may be
modelled by something like a Rayleigh distribution (or something
similar to give a tail to high values), but with little weight at
small angles. If the same is true of the sGRB population (and from
Figure 3 the long tail seems to be present here too) then a
half-normal distribution for the opening angle is not a good match to
the data. Did you try other priors on the inclination angle which may
better represent the true distribution? I ask that you try to use a
better prior for the inclination angle, or alternatively provide me
with a better justification of your current choice.

** MINOR COMMENTS

INTRODUCTION

-- The first sentence should be followed by a reference to Abadie et
al. (2010), or a similar updated article on predicted detection rates
if available.

-- "will be detected out to a maximum range of ≈ 450 Mpc." -> clarify
whether this is the horizon distance or sensemon distance.

-- "Observations of sGRBs give an inferred rate of sGRB producing CBC
mergers" -> mention that the sGRB rate must be corrected for beaming
effects to map to CBC merger rate.

-- "may be possible to identify the host galaxy of the source" -> is a
host galaxy necessary for redshift identification? Can the redshift
not be attained from spectroscopic analysis of the sGRB light curve?

-- "With knowledge of the redshift, using current best estimates for
the Hubble constant and other cosmological parameters, the true
luminosity distance can be estimated to ∼ 1% accuracy." -> reference
needed for current estimates of cosmological parameters.

SECTION II

-- Semi-colon in first sentence can be simply replaced by a comma.

-- "at the frequency range used in this project" --> explicitly
mention the frequency range.

-- "error in the amplitude calibration" -> is this the strain
amplitude or the error signal output. Clarify either way.

-- "within a control feedback loop (see [7]), which are subject to
uncertainties." -> This is quite a vague sentence. What kind of
uncertainties?

-- Even though the math is quite simple, I would advocate explicitly
stating the relationship between the true length response function and
the estimated one via the calibration scale factor.

-- "larger amplitude (e.g. be closer)" -> it would only appear to be
closer if we had a handle on the other parameters affecting the
amplitude e.g. the chirp mass. We obviously get the latter from the
signal's phase evolution, but this point should be clarified.

-- How realistic is the assumption of a constant calibration scale
factor over the full LIGO band? What physical mechanism would lead us
to make a poor estimation of the length response function, and why
would other calibration tests fail to mitigate the effect dealt with
in this article?

SECTION III

-- "different signal arrival times at each detector in the network" ->
the concept of a network has not been mentioned or discussed before. A
simple sentence explaining the purpose and composition of the kind of
detector network expected in the advanced era is necessary.

-- A reference for the Einstein Telescope is necessary, whether this
be the original design study, or possibly studies of arm orientation
preferences.

-- "The use of GW events with sGRB counterparts for use as a
cosmological tool was investigated in [11] and [12] with respect to
the 3rd generation GW interferometer, the Einstein Telescope (ET)." ->
the concept was also explored for advanced era detectors in a later
section of Taylor, Gair, Mandel (2012).

-- "These instead use either the distribution [...] assumed neutron
star (NS) mass" -> to clarify, Taylor, Gair, Mandel (2012) assume the
*form* of the NS mass distribution, but search over the parameters of
the distribution as hyperparameters in a hierarchical modeling scheme.

-- "specifically for the Hubble constant, which is has uncertainties
of O(1%)" -> reference to Planck results or other appropriate sources.

SECTION IV

-- "If the event exhibits an optical afterglow then the host galaxy
can be identified and from which a redshift can be obtained" -> I
reiterate my question from earlier: why can the redshift not be
determined from the sGRB light curve itself?

-- "The fraction of events with associated redshifts is ∼1/3" -> do
you have a reference for this?

-- "equal to the horizon range of the advanced GW detector network for
BNS systems." -> what is the network composition in this case?

-- "derived from three main sources, population synthesis models..."
-> "derived from three main sources: population synthesis models..."

-- "In [3] using the rate of observed sGRBs and assuming a beaming
angle of 15◦ a similar event rate of 1–180 is obtained." -> This
sentence was a bit difficult to parse. I suggest restructuring it to
with something like "A similar event rate of 1-180 was obtained in
[3], where the rate of observed sGRBs was converted to the rate of BNS
events by assuming a beaming angle of 15 degrees."

-- Is the beaming angle here the full opening angle, or the half
opening angle?

-- "For the GW followup of sGRB scenario the merger time for BNS/NSBH
systems will be estimated from the sGRB to within a few seconds" ->
This sentence was difficult to understand. Suggest changing it to "For
the scenario of a GW-followup of an observed sGRB, the merger time for
BNS/NSBH systems will be estimated from the sGRB to within a few
seconds."

-- "However, as discussed earlier, the property of beaming severely
impacts the probable rate of such joint observations." -> So am I to
understand from this final paragraph that although the probability of
such joint observations is low, the SNR distribution of any detected
signals will be comparatively higher than the regular
GW-detection-only SNR distribution?

SECTION V

-- In equation 3, I believe the inclination term in hx should simply
be cosine(iota) rather than the square. Please proof check all
equations to ensure they match your intent.

-- "the sky position of the source α and δ." -> I assume these
correspond to RA and DEC, but this should be made clear.

-- See my comment on the modeling of the inclination angle
distribution above.

-- "for the black hole we use a prior based on the canonical mass
distribution used for the rate results in [34] with a mean of 5 M and
a standard deviation of 1 M." -> you mention that this distribution is
only an example for this type of system, but how would you do this in
reality? Would you make the mean and standard deviation further
parameters in your search in a hierarchical modelling scheme? How
would your current choices bias calibration attempts if the population
you are targeting is significantly different than your assumptions
here?

-- In equation 7, do your choices mean p(C=0.1) = p(C=10) or
p(lnC=ln(0.1)) = p(lnC=ln(10))?

-- "uniform distribution on the sky" -> does this mean uniform in
right ascension and uniform in cosine(90-declination)?

-- Your range of injected calibration scale factors is quite narrow,
meaning you only really tested situations where the factor was pretty
close to optimal. Did you test cases where the detectors were very
poorly calibrated? If not, then I highly suggest that you do so, or
adequately comment on expectations from these tests.

SECTION VI

-- "calculated by averaging 64 separate noisy realisations of the PSD
and scaled with the same calibraion scale factor as applied to the
injection and noise" -> firstly the typo in "calibration" should be
corrected. Also, does this mean that the calibration scale factor
appears in "S" as a parameter also during your MCMC? If so, why does
it not explicitly appear in the denominator of the exp[] of equation
6?

-- In the paragraph on page 6 beginning "In figure 3..." there is some
explanation of why the H1 and L1 uncertainties saturate out to ~25%.
The explanation did not seem very clear to me, and does not address
why V1 does not show a similar saturation. What is the significance of
the value of ~25%? Is there some influence of the finiteness of the
detected population at larger distances that is influencing the width
of the uncertainty distribution?

-- In figure 1 the tick labels for the chirp mass are overlaid with
the label. Better labelling or figure formatting is needed here, as
well as units where appropriate. I like the different edge coloring
for the relevant scale factor boxes, but perhaps these red edges could
be thickened for black and white printing.

-- No mention is given to the number of MCMC iterations, the length of
burn-in, or how convergence was ascertained. A brief comment would be
appropriate.

-- Page 9, second paragraph. "this prior is not that same" -> "this
prior is not the same"

### 2016-02-02 docs/response_to_referee.txt
Added repsonse to main comment (and a few minor ones)

@@
-----------------------------------------------------------------------
-Report of the Referee -- DL11679/Pitkin
-----------------------------------------------------------------------
+We would like to thank the referee for their very useful comments.
+We have included response to them inline below and make note any
+changes to the manuscript. These changes are displayed in red in the
+resubmitted version of the manuscript.
 
-This article proposes a novel method of calibrating ground-based
-gravitational-wave detectors by employing compact-binary systems as
-standard sirens. I find it to be scientifically interesting, novel,
-and well explained. Provided that the authors can address my following
-points, I suggest this article for publication subject to minor
-revision.
-
-** MAJOR COMMENTS
-
--- In section 5, I have a problem with the assumption of the sGRB jet
+Major comment: "In section 5, I have a problem with the assumption of the sGRB jet
 opening angle being modelled as a half-normal distribution. Figure 3
@@
 better prior for the inclination angle, or alternatively provide me
-with a better justification of your current choice.
+with a better justification of your current choice."
 
-** MINOR COMMENTS
+Response: Our poor description regarding how we use the opening angle (theta_j)
+distribution to constrain the distribution on the orientation angle (iota) of
+the systems has caused confusion here. We are not really interested
+in the jet opening angle distribution, but are just trying to somehow use
+it as a proxy for getting an idea of the orientation angle distribution,
+which would not be expected to look the same. Under the assumption that to
+observe a sGRB we must have theta_j >= |iota|, and assuming a Rayleigh
+distribution on theta_j, the distribution on |iota| does become a half-normal
+distribution. In fact our choice of a half-normal for |iota| with a sigma value of
+14.8 degs is rather conservative (in that it extends to larger iota values
+[this is something we would plan to correct in future studies]) compared to the
+half-normal distribution that you truly get based a Rayleigh distribution in 
+theta_j with a median of 10 degs. We have added text to the manuscipt to
+hopefully make this clearer. If the referee would like to see more detail on
+these we can send them further information.
 
-INTRODUCTION
+Minor comments
+--------------
 
--- The first sentence should be followed by a reference to Abadie et
+Comment: "The first sentence should be followed by a reference to Abadie et
 al. (2010), or a similar updated article on predicted detection rates
-if available.
+if available."
 
--- "will be detected out to a maximum range of ≈ 450 Mpc." -> clarify
-whether this is the horizon distance or sensemon distance.
+Response:
 
--- "Observations of sGRBs give an inferred rate of sGRB producing CBC
+Comment: ""will be detected out to a maximum range of ≈ 450 Mpc." -> clarify
+whether this is the horizon distance or sensemon distance."
+
+Response:
+
+Comment: ""Observations of sGRBs give an inferred rate of sGRB producing CBC
 mergers" -> mention that the sGRB rate must be corrected for beaming
-effects to map to CBC merger rate.
+effects to map to CBC merger rate."
 
--- "may be possible to identify the host galaxy of the source" -> is a
+Response:
+
+Comment: ""may be possible to identify the host galaxy of the source" -> is a
 host galaxy necessary for redshift identification? Can the redshift
-not be attained from spectroscopic analysis of the sGRB light curve?
+not be attained from spectroscopic analysis of the sGRB light curve?"
 
--- "With knowledge of the redshift, using current best estimates for
+Response:
+
+Comment: ""With knowledge of the redshift, using current best estimates for
 the Hubble constant and other cosmological parameters, the true
 luminosity distance can be estimated to ∼ 1% accuracy." -> reference
-needed for current estimates of cosmological parameters.
+needed for current estimates of cosmological parameters."
 
-SECTION II
+Response:
 
--- Semi-colon in first sentence can be simply replaced by a comma.
+Comment: "Semi-colon in first sentence can be simply replaced by a comma."
 
--- "at the frequency range used in this project" --> explicitly
-mention the frequency range.
+Response:
 
--- "error in the amplitude calibration" -> is this the strain
-amplitude or the error signal output. Clarify either way.
+Comment: ""at the frequency range used in this project" --> explicitly
+mention the frequency range."
 
--- "within a control feedback loop (see [7]), which are subject to
+Response:
+
+Comment: ""error in the amplitude calibration" -> is this the strain
+amplitude or the error signal output. Clarify either way."
+
+Response:
+
+Comment: ""within a control feedback loop (see [7]), which are subject to
 uncertainties." -> This is quite a vague sentence. What kind of
-uncertainties?
+uncertainties?"
 
--- Even though the math is quite simple, I would advocate explicitly
+Response:
+
+Comment: "Even though the math is quite simple, I would advocate explicitly
 stating the relationship between the true length response function and
-the estimated one via the calibration scale factor.
+the estimated one via the calibration scale factor."
 
--- "larger amplitude (e.g. be closer)" -> it would only appear to be
+Response:
+
+Comment: ""larger amplitude (e.g. be closer)" -> it would only appear to be
 closer if we had a handle on the other parameters affecting the
 amplitude e.g. the chirp mass. We obviously get the latter from the
-signal's phase evolution, but this point should be clarified.
+signal's phase evolution, but this point should be clarified."
 
--- How realistic is the assumption of a constant calibration scale
+Response:
+
+Comment: "How realistic is the assumption of a constant calibration scale
 factor over the full LIGO band? What physical mechanism would lead us
@@
 would other calibration tests fail to mitigate the effect dealt with
-in this article?
+in this article?"
 
-SECTION III
+Response:
 
--- "different signal arrival times at each detector in the network" ->
+Comment: ""different signal arrival times at each detector in the network" ->
 the concept of a network has not been mentioned or discussed before. A
 simple sentence explaining the purpose and composition of the kind of
-detector network expected in the advanced era is necessary.
+detector network expected in the advanced era is necessary."
 
--- A reference for the Einstein Telescope is necessary, whether this
+Response:
+
+Comment: "A reference for the Einstein Telescope is necessary, whether this
 be the original design study, or possibly studies of arm orientation
-preferences.
+preferences."
 
--- "The use of GW events with sGRB counterparts for use as a
+Response:
+
+Comment: ""The use of GW events with sGRB counterparts for use as a
 cosmological tool was investigated in [11] and [12] with respect to
@@
 the concept was also explored for advanced era detectors in a later
-section of Taylor, Gair, Mandel (2012).
+section of Taylor, Gair, Mandel (2012)."
 
--- "These instead use either the distribution [...] assumed neutron
+Response:
+
+Comment: ""These instead use either the distribution [...] assumed neutron
 star (NS) mass" -> to clarify, Taylor, Gair, Mandel (2012) assume the
 *form* of the NS mass distribution, but search over the parameters of
-the distribution as hyperparameters in a hierarchical modeling scheme.
+the distribution as hyperparameters in a hierarchical modeling scheme."
 
--- "specifically for the Hubble constant, which is has uncertainties
-of O(1%)" -> reference to Planck results or other appropriate sources.
+Response:
 
-SECTION IV
+Comment: ""specifically for the Hubble constant, which is has uncertainties
+of O(1%)" -> reference to Planck results or other appropriate sources."
 
--- "If the event exhibits an optical afterglow then the host galaxy
+Response:
+
+Comment: ""If the event exhibits an optical afterglow then the host galaxy
 can be identified and from which a redshift can be obtained" -> I
 reiterate my question from earlier: why can the redshift not be
-determined from the sGRB light curve itself?
+determined from the sGRB light curve itself?"
 
--- "The fraction of events with associated redshifts is ∼1/3" -> do
-you have a reference for this?
+Response:
 
--- "equal to the horizon range of the advanced GW detector network for
-BNS systems." -> what is the network composition in this case?
+Comment: ""The fraction of events with associated redshifts is ∼1/3" -> do
+you have a reference for this?"
 
--- "derived from three main sources, population synthesis models..."
--> "derived from three main sources: population synthesis models..."
+Response:
 
--- "In [3] using the rate of observed sGRBs and assuming a beaming
+Comment: ""equal to the horizon range of the advanced GW detector network for
+BNS systems." -> what is the network composition in this case?"
+
+Response:
+
+Comment: ""derived from three main sources, population synthesis models..."
+-> "derived from three main sources: population synthesis models...""
+
+Response:
+
+Comment: ""In [3] using the rate of observed sGRBs and assuming a beaming
 angle of 15◦ a similar event rate of 1–180 is obtained." -> This
@@
 [3], where the rate of observed sGRBs was converted to the rate of BNS
-events by assuming a beaming angle of 15 degrees."
+events by assuming a beaming angle of 15 degrees.""
 
--- Is the beaming angle here the full opening angle, or the half
-opening angle?
+Response:
 
--- "For the GW followup of sGRB scenario the merger time for BNS/NSBH
+Comment: "Is the beaming angle here the full opening angle, or the half
+opening angle?"
+
+Response:
+
+Comment: ""For the GW followup of sGRB scenario the merger time for BNS/NSBH
 systems will be estimated from the sGRB to within a few seconds" ->
@@
 BNS/NSBH systems will be estimated from the sGRB to within a few
-seconds."
+seconds.""
 
--- "However, as discussed earlier, the property of beaming severely
+Response:
+
+Comment: ""However, as discussed earlier, the property of beaming severely
 impacts the probable rate of such joint observations." -> So am I to
@@
 signals will be comparatively higher than the regular
-GW-detection-only SNR distribution?
+GW-detection-only SNR distribution?"
 
-SECTION V
+Response:
 
--- In equation 3, I believe the inclination term in hx should simply
+Comment: "In equation 3, I believe the inclination term in hx should simply
 be cosine(iota) rather than the square. Please proof check all
-equations to ensure they match your intent.
+equations to ensure they match your intent."
 
--- "the sky position of the source α and δ." -> I assume these
-correspond to RA and DEC, but this should be made clear.
+Response: The referee is correct. This has been fixed.
 
--- See my comment on the modeling of the inclination angle
-distribution above.
+Comment: ""the sky position of the source α and δ." -> I assume these
+correspond to RA and DEC, but this should be made clear."
 
--- "for the black hole we use a prior based on the canonical mass
+Response: alpha and delta have now been defined as right ascension and declinarion respectively.
+
+Comment: ""for the black hole we use a prior based on the canonical mass
 distribution used for the rate results in [34] with a mean of 5 M and
@@
 you are targeting is significantly different than your assumptions
-here?
+here?"
 
--- In equation 7, do your choices mean p(C=0.1) = p(C=10) or
-p(lnC=ln(0.1)) = p(lnC=ln(10))?
+Response:
 
--- "uniform distribution on the sky" -> does this mean uniform in
-right ascension and uniform in cosine(90-declination)?
+Comment: "In equation 7, do your choices mean p(C=0.1) = p(C=10) or
+p(lnC=ln(0.1)) = p(lnC=ln(10))?"
 
--- Your range of injected calibration scale factors is quite narrow,
+Response: The choice mean p(C=0.1) = p(C=10). This has been fixed.
+
+Comment: ""uniform distribution on the sky" -> does this mean uniform in
+right ascension and uniform in cosine(90-declination)?"
+
+Response:
+
+Comment: "Your range of injected calibration scale factors is quite narrow,
 meaning you only really tested situations where the factor was pretty
@@
 poorly calibrated? If not, then I highly suggest that you do so, or
-adequately comment on expectations from these tests.
+adequately comment on expectations from these tests."
 
-SECTION VI
+Response:
 
--- "calculated by averaging 64 separate noisy realisations of the PSD
+Comment: ""calculated by averaging 64 separate noisy realisations of the PSD
 and scaled with the same calibraion scale factor as applied to the
@@
 it not explicitly appear in the denominator of the exp[] of equation
-6?
+6?"
 
--- In the paragraph on page 6 beginning "In figure 3..." there is some
+Response: The typo has been corrected. The calibration scale factor does
+not appear in "S" during the MCMC or the denominator of equation 6. The
+description in this section is saying how the simulated data was created,
+whereas during the MCMC it is the only the expect signal waveform
+that has the scale factor applied to it to try and match the data (to which
+the calibration factor has already been applied during its creation).
+
+Comment: "In the paragraph on page 6 beginning "In figure 3..." there is some
 explanation of why the H1 and L1 uncertainties saturate out to ~25%.
@@
 detected population at larger distances that is influencing the width
-of the uncertainty distribution?
+of the uncertainty distribution?"
 
--- In figure 1 the tick labels for the chirp mass are overlaid with
+Response:
+
+Comment: "In figure 1 the tick labels for the chirp mass are overlaid with
 the label. Better labelling or figure formatting is needed here, as
@@
 for the relevant scale factor boxes, but perhaps these red edges could
-be thickened for black and white printing.
+be thickened for black and white printing."
 
--- No mention is given to the number of MCMC iterations, the length of
+Response:
+
+Comment: "No mention is given to the number of MCMC iterations, the length of
 burn-in, or how convergence was ascertained. A brief comment would be
-appropriate.
+appropriate."
 
--- Page 9, second paragraph. "this prior is not that same" -> "this
-prior is not the same"
+Response:
+
+Comment: "Page 9, second paragraph. "this prior is not that same" -> "this
+prior is not the same""
+
+Response: This has been fixed.

### 2016-02-02 docs/cosmic_cal.tex
cosmic_cal.tex: add refresp macro for adding any changes in response to the referees report
cosmic_cal.tex: added some responses to main referee comment and a few minor ones

@@
 \newcommand{\dcc}{LIGO-P1500198}
-\newcommand{\cm}[1]{\textcolor{red}{CM: #1}}
-\newcommand{\MP}[1]{\textcolor{blue}{MP: #1}}
-\newcommand{\lw}[1]{\textcolor{green}{LW: #1}}
+\newcommand{\refresp}[1]{\textcolor{red}{#1}}
 
@@
 \gws_{\times}(f) &\propto
-\left(\cos^{2}\iota\right)D_{L}^{-1}\mathcal{M}^{5/6} f^{-7/6}e^{-i\Psi(f, \mathcal{M},
+\cos{\iota}D_{L}^{-1}\mathcal{M}^{5/6} f^{-7/6}e^{-i\Psi(f, \mathcal{M},
 t_\mathrm{c}, \phi_{\mathrm{c}})} 
@@
 are dependent upon the polarisation angle $\psi$ and the sky position of the
-source $\alpha$ and $\delta$.
+source \refresp{with right ascension} $\alpha$ and \refresp{declination} $\delta$.
 
@@
 in~\cite{2014ApJ...780..118F} provides a median opening angle value of $\sim
-10^{\circ}$. We assume that the beam opening angle is a half-normal
-distribution, for which the median is given by $\sqrt{2}\erf{}^{-1}(1/2)\sigma
-\sim 0.67\sigma$. It follows that a median opening angle of a $\sim 10^{\circ}$
-corresponds to $\sigma \sim 14.8^{\circ}$. We therefore place a Gaussian
-prior on the inclination angle $\iota$ with zero mean and a standard deviation
-$\sigma=14.8^{\circ}$. This, and the choice of polarisation angle prior range given below, 
+10^{\circ}$. \refresp{We use this median opening angle as a rather conservative
+guide to form a Gaussian prior distribution on the system inclination angle,
+$\iota$, by converting it into an equivalent standard devaition for a half-normal
+distribution. We therefore place a Gaussian prior on $\iota$ with zero mean and a standard deviation
+$\sigma=14.8^{\circ}$.} This, and the choice of polarisation angle prior range given below, 
 explicitly fixes the handedness of the system's rotation (or alternatively whether it is face-on or 
@@
 also (na\"{i}vely) assume that the prior \ac{PDF} of the calibration scale factor is either
-larger or smaller than unity by an equivalent factor, e.g.\ $p(\scf|I)=10$ is
-equivalent to $p(\scf|I)=0.1$. For the scale factor in each detector we 
+larger or smaller than unity by an equivalent factor, e.g.\ \refresp{$p(\scf=10|I)$ is
+equivalent to $p(\scf=0.1|I)$}. For the scale factor in each detector we 
 therefore use the following log-normal distribution as our prior
@@
 64 separate noisy realisations of the \ac{PSD} and scaled
-with the same calibraion scale factor as applied to the injection and noise\footnote{We do not account for there 
+with the same \refresp{calibration} scale factor as applied to the injection and noise\footnote{We do not account for there 
 being a potential difference
@@
 larger distances the \acp{PDF} on \scft for V1 are not well constrained and start to become dominated
-by the prior. As we have previously noted this prior is not that same as the distribution used
+by the prior. As we have previously noted this prior is not \refresp{the} same as the distribution used
 to generate the \scft values, but is instead far broader and more conservative. So, this bias is

### 2016-02-03 docs/response_to_referee.txt
Most of my comments have been addressed in the response and manuscript

@@
 We would like to thank the referee for their very useful comments.
-We have included response to them inline below and make note any
-changes to the manuscript. These changes are displayed in red in the
-resubmitted version of the manuscript.
+We have included response to them inline below and make note where this
+has resulted in changes to the manuscript. These changes are displayed in red
+in the resubmitted version of the manuscript.
 
@@
 which would not be expected to look the same. Under the assumption that to
-observe a sGRB we must have theta_j >= |iota|, and assuming a Rayleigh
+observe a sGRB we must have theta_j/2 >= |iota|, and assuming a Rayleigh
 distribution on theta_j, the distribution on |iota| does become a half-normal
 distribution. In fact our choice of a half-normal for |iota| with a sigma value of
-14.8 degs is rather conservative (in that it extends to larger iota values
+14.8 degs is rather conservative* (in that it extends to larger iota values
 [this is something we would plan to correct in future studies]) compared to the
@@
 these we can send them further information.
+
+*this conservative nature is in part due to one of us (M Pitkin) mistakenly
+assuming the opening angle referred to the half width of the beam rather than
+the full beam width.
 
@@
 
-Response:
+Response: We agree and have added the Abadie et al. (2010) reference here.
 
@@
 
-Response:
+Response: Done.
 
@@
 
-Response:
+Response: This has been changed to "...within a control feedback loop (see [7]),
+each of which are subject to measurement uncertainties."
 
@@
 
-Response:
+Response: The relationship has now been explictly added.
 
@@
 
-Response:
+Response: Our analysis uses the frequency range from 20-400Hz, so our assumption
+of a constant calibration scale factor only relates to this band rather than the
+full LIGO bandwidth. However, having just one scale factor should be a
+conservative case and in the future we would like to look at how having
+some frequency dependence to the calibration scale factor would effects results.
+
+We do not consider what physical effects could cause a poor estimation of the
+length response function. We mainly put our proposal forward as an independent
+way of estimating the length response function.
 
@@
 
-Response:
+Response: See response to the first comment.
 
@@
 
-Response:
+Response: We have tested situations were the detectors were very poorly
+calibrated e.g. scale factors ranging from ~0.1 to ~10. Even for these
+very poorly calibrated cases the widths of the scale factor distributions
+stay consistent with cases were the scale factor is close to 1. We have
+added some additional text to footnote 7 about this.
 
@@
 
-Response:
+Response: We have added the additional text stating "the weakest signals that could still
+be detected would always have a single detector \ac{SNR} of $\sim 5.5$ no matter
+their distance, hence the plateau.
+
+The issue of the V1 result not saturating is addressed in the manuscript in the lines
+"The largest SNR contribution will generally come from the two aLIGO detectors and thus
+the detection criteria (SNR threshold) will not apply to the AdV result. Hence, the SNR
+in AdV can be small and thus the ability to constrain its scale factor becomes poor (although it
+still provides information that the calibration is not grossly inaccurate)."
 
@@
 
-Response:
+Response: As a proxy to check for convergence of the chains we check that the
+scale parameter posterior histograms do not contain many disjoint modes. We have
+added this as a footnote into the paper. The actual tests we use are that 90% or
+more of the histogram bins are non-zero and the standard deviation
+of the posterior is less than 1/6th of the total histogram range.
 

### 2016-02-03 docs/cosmic_cal.tex
Most of my comments have been addressed in the response and manuscript

@@
 It is expected that the advanced generation of interferometric \ac{GW}
-detectors will detect waves emitted from $O(10s)$ of \acp{CBC} per year. One
+detectors will detect waves emitted from $O(10s)$ of \acp{CBC} per year \refresp{\cite{2010CQGra..27q3001A}}. One
 such class of these cataclysmic events, the inspiral and merger of \ac{BNS}
@@
 The technique used throughout \ac{GW} research in calibrating the detectors is carried out through 
-a complicated system of physical manipulations of the Fabry-P\'{e}rot Michelson interferometer; 
+a complicated system of physical manipulations of the Fabry-P\'{e}rot Michelson interferometer\refresp{,}
 where an elaborate feedback system is used to sustain a defined measurement in arm length 
@@
 time-scales). Calculation of $R$ requires the measurement of various functions within a control 
-feedback loop (see \cite{2010NIMPA.624..223A}), which are subject to uncertainties. In this 
+feedback loop (see \cite{2010NIMPA.624..223A}), \refresp{each of which are subject to measurement uncertainties}. In this 
 study we will assume an estimate of $R$ is available (although in theory we could take on the role 
@@
 \end{equation}
-where $h(f)$ is the true strain and $h_{\rm m}(f)$ is the measured strain. With this definition it
+where $h(f)$ is the true strain and $h_{\rm m}(f)$ is the measured strain \refresp{(i.e.\
+the measured length response function $R(f)$ is related to the true length response function $R_{\rm true}(f)$
+via $R(f) = \scf R_{\rm true}(f)$)}. With this definition it
 means that if $\scf > 1$ then a signal would appear to have a larger amplitude (e.g.\ be 
@@
 uncertainties that are expected via experimental methods, but is not the same as
-the prior that we assume on $\mathcal{C}$ (see Eq.~\ref{eq:scale_prior}) when recovering 
+the prior that we assume on $\scf$ (see Eq.~\ref{eq:scale_prior}) when recovering 
 signals\footnote{The reason behind this is that we wanted to use a fairly na\"{i}ve and
 conservative prior for signal recovery that had little impact in the results, i.e.\
-to produce results domintaed by the likelihood for these parameters.}.
+to produce results domintaed by the likelihood for these parameters. \refresp{However,
+limited studies using $\scf$ values within the much larger range of
+$\sim 0.1$ to 10 show distributions with consistent relative widths to
+those we find in our main study.}}.
 
@@
 an upper limit on the calibration scale factor.}. This has provided posterior probability
-distributions on the calibration scale factors for each detector.  Examples of
+distributions\refresp{\footnote{\refresp{As a proxy to check for convergence of the MCMC chains
+we check that the calibration scale factor posterior histograms do not contain many disjoint modes.}}}
+on the calibration scale factors for each detector.  Examples of
 the marginalised posterior probabilities for a \ac{BNS} system and a \ac{NSBH}
@@
 those sources with a high enough \ac{SNR} that we would consider them
-detectable. There will therefore be fewer sources with \ac{SNR} higher than
+detectable\refresp{, i.e.\ the weakest signals that could still be detected would
+always have a single detector \ac{SNR} of $\sim 5.5$ no matter their distance, hence the
+plateau}. \refresp{In addition there will also} be fewer sources with \ac{SNR} higher than
 this criteria at large distances, giving us a narrower range, and we

### 2016-02-04 docs/cosmic_cal.tex
Updated response to comment about plots

@@
  \caption{\label{fig:bnspost} The marginalised posterior probability distributions for some of the unknown
- parameters of a \ac{BNS} system, including calibration scaling factors (surrounded by the red 
+ parameters of a \ac{BNS} system, including calibration scaling factors (surrounded by the 
+\refresp{thicker} red 
 borders) for the three detectors (H1, L1 and V1). The simulated signal was at a distance of 
-250\,Mpc, had \ac{SNR} of 7.9, 9.2 and 5.2 and percentage relative uncertainties in the calibration scale 
-factors of 17\%, 15\% and 24\% for each of the detectors respectively.}
+250\,Mpc, had \ac{SNR} of 7.9, 9.1 and 5.2 and percentage relative uncertainties in the calibration 
+scale factors of 15\%, 14\% and 22\% for each of the detectors respectively.}
 \end{figure*}
@@
  parameters of a \ac{NSBH} system, including calibration scaling factors for the three detectors 
-(H1, L1 and V1). The simulated signal was at a distance of 450\,Mpc, had \ac{SNR} of 9.8, 10.3 and 
-3.8 and percentage relative uncertainties in the calibration scale factors of 15\%, 14\% and 36\% for each of 
-the detectors respectively.}
+(H1, L1 and V1). The simulated signal was at a distance of 450\,Mpc, had \ac{SNR} of 9.7, 10.3 and 
+3.8 and percentage relative uncertainties in the calibration scale factors of 14\%, 13\% and 43\% 
+for each of the detectors respectively.}
 \end{figure*}

### 2016-02-04 docs/response_to_referee.txt
Updated response to comment about plots
First complete set of answers to the referee comments.

@@
 
-Response:
+Response: This is the horizon distance and has been clarified in the
+manuscript.
 
@@
 
-Response:
+Response: This has now been clarified in the text.
 
@@
 
-Response:
+Response: No, unfortunately the redshift cannot be otained via the sGRB itself
+since it contains no transmission or absorbtion lines. CHRIS: Need a bit more
+than this.
 
@@
 
-Response:
+Response: We have added a reference to the most recent Planck cosmology
+results.
 
@@
 
-Response:
+Response: It is the strain amplitude and is now clarified in the text.
 
@@
 
-Response:
+Response: The text has been changed to state "(i.e.\ for a given system, would
+seem closer)"
 
@@
 
-Response:
+Response: The network is first mentioned in the previous sentence but we have
+now added extra clarification.
 
@@
 
-Response:
+Response: We have added a reference to Punturo et al 2010 "The third generation
+of gravitational wave observatories and their science reach"
 
@@
 
-Response:
+Response: This omission has now been corrected in the text.
 
@@
 
-Response:
+Response: Corrected.
 
@@
 
-Response:
+Response: We have added a reference to the latest Planck cosmology results.
 
@@
 
-Response:
+Response: See previous answer.
 
@@
 
-Response:
+Response: We don't know of a particular publication that includes this number
+but we have added a reference to the NASA Swift GRB database from which it can
+be seen that only ~1/3 have a measured redshift.
 
@@
 
-Response:
+Response: The present statement indicates only an approximate matching between
+the closest known sGRB and the advanced detector network range. We feel that
+stating a specific network configuration (LIGO Hanford, LIGO Livingston, Virgo,
+Kagra, LIGO India) would be misleading and therefore not necessary.
 
@@
 
-Response:
+Response: Done.
 
 Comment: ""In [3] using the rate of observed sGRBs and assuming a beaming
-angle of 15◦ a similar event rate of 1–180 is obtained." -> This
+angle of 15 degrees a similar event rate of 1–180 is obtained." -> This
 sentence was a bit difficult to parse. I suggest restructuring it to
@@
 
-Response:
+Response: No. The text at present is actually misleading/wrong. The truth is
+that jointly observed systems will be face-on and therefore better oriented. At
+a given distance such an event will likely have higher SNR than an arbitrarily
+oriented GW-only event. A meaningful way to express this is to state that the
+senstive range (and hence the sensitive volume) is increased for these jointly
+observed systems. We have modified the text to read:
+
+"This prerequisite property limits us to systems that are approximately
+``face-on'' and therefore effectively optimally oriented. At given distances
+this biases us to higher SNR signals but it is more meaningful to think of this
+as an increase in the sensitivity range of the detectors."
 
@@
 
-Response:
+Response: Our choices of signal parameter prior primarily effect the ensemble
+statistics, i.e. the distributions of calibration uncertainties at given
+distances. These give the reader an indication of the likely accuracy of an
+astrophysical calibration scheme. In "reality" we would first hope to be able
+to make this measurement for single detections in which case the true prior is out
+of our control and we must simply deal with what the universe gives us. In this
+case our assumed mass priors would be dominated by the information content of
+the likelihood due to the mass dependent phasing information.  Hence for
+individual detections we believe that any choice of conservative priors will
+have negligible effect on the result. If in the future one were to attempt this
+with a population of detections then it would be necessary to assume a
+hierachical scheme with some form for the mass distribution as the referee
+suggests.  However, we feel that this is beyond the scope of this work.
 
@@
 
-Response:
+Response: Yes.
 
@@
 
-Response:
+Response: MATT: I think this is done now right?
 

### 2016-02-04 docs/cosmic_cal.tex
First complete set of answers to the referee comments.

--- 2016-02-03 docs/cosmic_cal.tex
@@
 It is expected that the advanced generation of interferometric \ac{GW}
-detectors will detect waves emitted from $O(10s)$ of \acp{CBC} per year \refresp{\cite{2010CQGra..27q3001A}}. One
-such class of these cataclysmic events, the inspiral and merger of \ac{BNS}
-systems will be detected out to a maximum range of $\approx 450$ Mpc. Assuming
-the current best estimates for the cosmological parameters, this is equivalent
-to a redshift $z\approx 0.1$. As noted by~\citet{1986Natur.323..310S}, \ac{CBC}
+detectors will detect waves emitted from $O(10s)$ of \acp{CBC} per year
+\refresp{\cite{2010CQGra..27q3001A}}. One such class of these cataclysmic
+events, the inspiral and merger of \ac{BNS} systems will be detected out to a
+maximum range \refresp{(horizon distance)} of $\approx 450$ Mpc. Assuming the
+current best estimates for the cosmological parameters, this is equivalent to a
+redshift $z\approx 0.1$. As noted by~\citet{1986Natur.323..310S}, \ac{CBC}
 systems can be used as cosmological distance markers, otherwise known as
@@
 produce tightly beamed \ac{EM} emission parallel to the orbital angular
-momentum vector of the system. Additional evidence for the coincidence
-of \acp{sGRB} with \ac{GW} events are the estimated astrophysical rates of both
+momentum vector of the system. Additional evidence for the coincidence of
+\acp{sGRB} with \ac{GW} events are the estimated astrophysical rates of both
 phenomena. Observations of \acp{sGRB} give an inferred rate of \ac{sGRB}
 producing \ac{CBC} mergers of $8\times 10^{-9}$--$1.1\times
-10^{-6}$\,Mpc$^{-3}$\,yr$^{-1}$~\cite{2012MNRAS.425.2668C} which can be
-compared to independently obtained estimates of \ac{BNS} merger rates of
-$10^{-8}$--$10^{-5}$\,Mpc$^{-3}$\,yr$^{-1}$ \cite{2010CQGra..27q3001A}, or
-\ac{NSBH} merger rates of
+10^{-6}$\,Mpc$^{-3}$\,yr$^{-1}$~\cite{2012MNRAS.425.2668C} \refresp{after accounting for
+beaming effects. This} can be compared to independently obtained estimates of
+\ac{BNS} merger rates of $10^{-8}$--$10^{-5}$\,Mpc$^{-3}$\,yr$^{-1}$
+\cite{2010CQGra..27q3001A}, or \ac{NSBH} merger rates of
 $6\,\times\,10^{-10}$--$1\,\times\,10^{-6}$\,Mpc$^{-3}$\,yr$^{-1}$
@@
 If a single \ac{GW} event is observed in coincidence with a \ac{sGRB} it may be
-possible to identify the host galaxy of the source.  With this
-information a spectroscopic redshift can be very accurately obtained (see
-values quoted in e.g.\ \cite{0004-637X-664-2-1000} and references therein). With
-knowledge of the redshift, using current best estimates for the Hubble constant
-and other cosmological parameters, the true luminosity distance can be
-estimated to $\sim 1\%$ accuracy.  The \ac{GW} measurement acts as a standard
-siren also giving us a direct measurement of the luminosity distance to the
-source. The accuracy of such a measurement depends on a number of factors
-including the accuracy with which the \ac{GW} detector has been calibrated.
-Hence, by comparison with the distance estimate from the \ac{sGRB} we can
-re-calibrate (or validate) the existing experimentally obtained calibration.
+possible to identify the host galaxy of the source.  With this information a
+spectroscopic redshift can be very accurately obtained (see values quoted in
+e.g.\ \cite{0004-637X-664-2-1000} and references therein). With knowledge of
+the redshift, using current best estimates for the Hubble constant and other
+cosmological parameters, the true luminosity distance can be estimated to $\sim
+1\%$ accuracy~\refresp{\cite{2015arXiv150201589P}}.  The \ac{GW} measurement
+acts as a standard siren also giving us a direct measurement of the luminosity
+distance to the source. The accuracy of such a measurement depends on a number
+of factors including the accuracy with which the \ac{GW} detector has been
+calibrated.  Hence, by comparison with the distance estimate from the \ac{sGRB}
+we can re-calibrate (or validate) the existing experimentally obtained
+calibration.
 
@@
 
-The technique used throughout \ac{GW} research in calibrating the detectors is carried out through 
-a complicated system of physical manipulations of the Fabry-P\'{e}rot Michelson interferometer\refresp{,}
-where an elaborate feedback system is used to sustain a defined measurement in arm length 
-difference between the moving mirrors. A comprehensive description of the calibration procedure (in 
-particular for the LIGO detectors during their fifth science run) can be found in 
-\cite{2010NIMPA.624..223A}, and a brief description is given in \cite{Vitale:2012} and references 
-therein. For \ac{aLIGO} and \ac{AdV}, at the frequency range used in this project, the error in the 
-amplitude calibration is expected to be roughly $10\%$ \cite{Vitale:2012}. This is a benchmark set 
-for the error estimation using the proposed method in this project.
+The technique used throughout \ac{GW} research in calibrating the detectors is
+carried out through a complicated system of physical manipulations of the
+Fabry-P\'{e}rot Michelson interferometer\refresp{,} where an elaborate feedback
+system is used to sustain a defined measurement in arm length difference
+between the moving mirrors. A comprehensive description of the calibration
+procedure (in particular for the LIGO detectors during their fifth science run)
+can be found in \cite{2010NIMPA.624..223A}, and a brief description is given in
+\cite{Vitale:2012} and references therein. For \ac{aLIGO} and \ac{AdV}, at the
+frequency range used in this analysis \refresp{(20--400 Hz)}, the error in the
+\refresp{strain} amplitude calibration is expected to be roughly $10\%$
+\cite{Vitale:2012}. This is a benchmark set for the error estimation using the
+proposed method in this project.
 
 % Describe our model for the calibration
-The \ac{GW} strain is measured through the differential arm length, $\Delta L$, changes of the 
-interferometer via $h(f,t) = \Delta L(f,t) / L$, where $L$ is the full arm length. Calibration is 
-required to relate the actual measured interferometer error signal output $e(f, t)$ to $\Delta L$. 
-This relation is known as the length response function, $R(f)$, defined such that
+The \ac{GW} strain is measured through the differential arm length, $\Delta L$,
+changes of the interferometer via $h(f,t) = \Delta L(f,t) / L$, where $L$ is
+the full arm length. Calibration is required to relate the actual measured
+interferometer error signal output $e(f, t)$ to $\Delta L$.  This relation is
+known as the length response function, $R(f)$, defined such that
+%
 \begin{equation}
@@
 \end{equation}
-where we assume $R$ varies only slowly in time (in comparison to transient signal 
-time-scales). Calculation of $R$ requires the measurement of various functions within a control 
-feedback loop (see \cite{2010NIMPA.624..223A}), \refresp{each of which are subject to measurement uncertainties}. In this 
-study we will assume an estimate of $R$ is available (although in theory we could take on the role 
-of estimating $R$ itself), but that it differs from the truth through some unknown scale factor, 
-$\scf$, so that at a particular time we have
+%
+where we assume $R$ varies only slowly in time (in comparison to transient
+signal time-scales). Calculation of $R$ requires the measurement of various
+functions within a control feedback loop (see \cite{2010NIMPA.624..223A}),
+\refresp{each of which are subject to measurement uncertainties}. In this study
+we will assume an estimate of $R$ is available (although in theory we could
+take on the role of estimating $R$ itself), but that it differs from the truth
+through some unknown scale factor, $\scf$, so that at a particular time we have
+%
 \begin{equation}\label{eq:scalefactor}
@@
 \end{equation}
-where $h(f)$ is the true strain and $h_{\rm m}(f)$ is the measured strain \refresp{(i.e.\
-the measured length response function $R(f)$ is related to the true length response function $R_{\rm true}(f)$
-via $R(f) = \scf R_{\rm true}(f)$)}. With this definition it
-means that if $\scf > 1$ then a signal would appear to have a larger amplitude (e.g.\ be 
-closer) than in reality\footnote{However, the \ac{SNR} would be the correct value as the signal and 
-the noise will both contain the same scale factor.}, whereas if $\scf < 1$ it would appear to have a 
-lower amplitude than reality. In this analysis we will simplify the situation by assuming that 
-$\scf$ is a constant with respect to frequency (and is real, so has no phase component), but in future studies that assumption could be 
-dropped and $\scf$ could take some functional form, or piece-wise fit, with respect to $f$ (e.g.\ 
-in a similar way to the method of \cite{2013PhRvD..88h4044L} used for fitting differences in 
-\ac{PSD} estimates).
+%
+where $h(f)$ is the true strain and $h_{\rm m}(f)$ is the measured strain
+\refresp{(i.e.\ the measured length response function $R(f)$ is related to the
+true length response function $R_{\rm true}(f)$ via $R(f) = \scf R_{\rm
+true}(f)$)}. With this definition it means that if $\scf > 1$ then a signal
+would appear to have a larger amplitude \refresp{(i.e.\ for a given system,
+would seem closer)} than in reality\footnote{However, the \ac{SNR} would be
+the correct value as the signal and the noise will both contain the same scale
+factor.}, whereas if $\scf < 1$ it would appear to have a lower amplitude than
+reality.  In this analysis we will simplify the situation by assuming that
+$\scf$ is a constant with respect to frequency (and is real, so has no phase
+component), but in future studies that assumption could be dropped and $\scf$
+could take some functional form, or piece-wise fit, with respect to $f$ (e.g.\
+in a similar way to the method of \cite{2013PhRvD..88h4044L} used for fitting
+differences in \ac{PSD} estimates).
 
@@
 function of the component masses) and the binary orientation with respect to
-the detector network. However, measurement of the phase evolution of such an
-event allows accurate determination of the chirp mass\footnote{In reality it
-is the redshifted chirp mass that is measured.}. Timing information from the
-different signal arrival times at each detector in the network allows
-some level of sky position determination~\cite{2014ApJ...795..105S}.
-Finally, amplitude variation between differently oriented interferometers
-allows some (weaker) level of determination of the binary system orientation.
+the \refresp{global network of \ac{GW} detectors}. However, measurement of the
+phase evolution of such an event allows accurate determination of the chirp
+mass\footnote{In reality it is the redshifted chirp mass that is measured.}.
+Timing information from the different signal arrival times at each detector in
+the network allows some level of sky position
+determination~\cite{2014ApJ...795..105S}.  Finally, amplitude variation between
+differently oriented interferometers allows some (weaker) level of
+determination of the binary system orientation.
 
@@
 in~\cite{2010CQGra..27u5006S} and~\cite{2011PhRvD..83b3005Z} with respect to
-the 3rd generation \ac{GW} interferometer, the \ac{ET}. In this case, the
-redshift information obtained from the \ac{sGRB} host galaxy, provides a
-complementary measurement allowing each event to inform the distance-redshift
-relationship and hence obtain cosmological parameter estimates. Other methods
-for cosmological inference (see~\cite{1996PhRvD..53.2878F,2012PhRvD..85b3535T,
+the 3rd generation \ac{GW} interferometer, the
+\ac{ET}~\refresp{\cite{2010CQGra..27h4007P} and in ~\cite{2012PhRvD..85b3535T}
+for the Advanced detector network}. In this case, the redshift information
+obtained from the \ac{sGRB} host galaxy, provides a complementary measurement
+allowing each event to inform the distance-redshift relationship and hence
+obtain cosmological parameter estimates. Other methods for cosmological
+inference (see~\cite{1996PhRvD..53.2878F,2012PhRvD..85b3535T,
 2012PhRvD..86b3502T,Messenger:2011ux,2013arXiv1312.1862M}) have been proposed
 for \ac{BNS} systems that do not rely on any redshift measurements. These
-instead use either the distribution of measured \acp{SNR} and assumed \ac{NS}
-mass and spatial distributions or exploit the features of the tidal and
-post-merger hyper-massive \ac{NS} stages of the \ac{GW} waveform.
+instead use either the distribution of measured \acp{SNR} and assumed
+\refresp{form} of \ac{NS} mass and spatial distributions or exploit the
+features of the tidal and post-merger hyper-massive \ac{NS} stages of the
+\ac{GW} waveform.
 
@@
 cosmological problem upside-down. We assume that the existing \ac{EM}
-cosmological parameter estimation (specifically for the Hubble constant, which is has 
-uncertainties of $O(1\%)$) is
+cosmological parameter estimation (specifically for the Hubble constant, which
+is has uncertainties of $O(1\%)$~\refresp{\cite{2015arXiv150201589P}}) is
 accurately determined and therefore a known quantity. We then assume that the
@@
 It is believed that \acp{sGRB} (those gamma-ray bursts with duration ${<}2$ sec) are
-emitted during the merger of
-\ac{BNS} or \ac{NSBH} systems ~\cite{1989Natur.340..126E,1992ApJ...395L..83N}. The emission
-from these events is highly beamed along the binary rotation axis and hence
-only potentially observable for a fraction of the mergers. If the event
-exhibits an optical afterglow then the host galaxy can be identified and from which
-a redshift can be obtained (e.g.~\cite{2005Natur.437..851G}).  The fraction of events with 
-associated redshifts is ${\sim}1/3$. The range to which the \emph{Swift} \ac{XRT} has detected
-\acp{sGRB} is $z{\sim}2$ and the nearest event with associated redshift is at
+emitted during the merger of \ac{BNS} or \ac{NSBH} systems
+~\cite{1989Natur.340..126E,1992ApJ...395L..83N}. The emission from these events
+is highly beamed along the binary rotation axis and hence only potentially
+observable for a fraction of the mergers. If the event exhibits an optical
+afterglow then the host galaxy can be identified and from which a redshift can
+be obtained (e.g.~\cite{2005Natur.437..851G}).  The fraction of events with
+associated redshifts is
+${\sim}1/3$~\refresp{\footnote{http://swift.gsfc.nasa.gov/archive/grb\_table/}}.
+The range to which the \emph{Swift} \ac{XRT} has detected \acp{sGRB} is
+$z{\sim}2$ and the nearest event with associated redshift is at
 $z{\sim}0.1$~\cite{2015GCN..17278...1C}, which is approximately equal to the
@@
 The predicted detection rates of \ac{BNS} events (irrespective of an \ac{sGRB}
-counterpart) are derived from three main sources, population synthesis models,
-extrapolations based on known galactic \ac{BNS} systems, and from observations
-of \ac{sGRB} events. Various studies using the first two methods have been
-compiled into an overall detection rate of $0.2$--$200$ \ac{BNS} events per
+counterpart) are derived from three main sources\refresp{:} population
+synthesis models, extrapolations based on known galactic \ac{BNS} systems, and
+from observations of \ac{sGRB} events. Various studies using the first two
+methods have been compiled into an overall detection rate of $0.2$--$200$
+\ac{BNS} events per
 year~\cite{2013arXiv1304.0670L,2010CQGra..27q3001A}\footnote{We note that this
@@
 also note the study mentioned in Sec.~\ref{sec:intro} that potentially lowers
-the expected rate by a factor of five.}.  In~\cite{2012MNRAS.425.2668C} using
-the rate of observed \acp{sGRB} and assuming a beaming angle of $15^{\circ}$ a
-similar event rate of $1$--$180$ is obtained. The beaming angle is the primary
-factor in determining the fraction of detected \ac{GW} \ac{BNS} signals with an
-\ac{EM} \ac{sGRB} counterpart and is unfortunately highly uncertain.  Various
-authors have provided estimates on the rate of joint detections to be in the
-range $0.02$--$7$ per
+the expected rate by a factor of five.}. \refresp{A similar event rate of
+$1$--$180$ was obtained in~\cite{2012MNRAS.425.2668C}, where the rate of
+observed \acp{sGRB} was converted to the rate of \ac{BNS} events by assuming a
+beaming angle of $15^{\circ}$.} The beaming angle is the primary factor in
+determining the fraction of detected \ac{GW} \ac{BNS} signals with an \ac{EM}
+\ac{sGRB} counterpart and is unfortunately highly uncertain.  Various authors
+have provided estimates on the rate of joint detections to be in the range
+$0.02$--$7$ per
 year~\cite{2013PhRvL.111r1101C,Kelley:2012fl,2014arXiv1405.2254S,2015MNRAS.448.3026W}.
@@
 % observing a coincident event
-The most likely scenario in which a coincident \ac{GW}--\ac{sGRB} event would be identified is 
-through the targeted follow-up of an observed \ac{GW} or by post-facto matching of \ac{GW} 
-trigger lists with known (or subsequently found \cite{2015ApJS..217....8B}) \ac{sGRB} events.  
-The likelihood of being able to follow-up \ac{GW} 
-events using gamma-ray telescopes with low enough latency to catch a \ac{sGRB} is low.  Compounding 
-this issue is the relatively large \ac{GW} sky error-box giving a field-of-view for the \ac{EM} 
-observatories to search spanning $\sim 100$'s of square degrees (e.g.\ 
-\cite{2011CQGra..28j5021F,2013arXiv1304.0670L,2014ApJ...795..105S}).  For the \ac{GW} 
-follow-up of \ac{sGRB} scenario the merger time for \ac{BNS}/\ac{NSBH} systems will be estimated 
-from the \ac{sGRB} to within a few seconds (see e.g.\ the discussion in Sec.\ 2.2 of 
-\cite{2012ApJ...760...12A}). This makes the follow-up search less 
-computationally expensive since it is performed over a smaller range of data using potentially 
-fewer numbers of waveform templates.  This computational saving enables the use of a more 
-computationally expensive multi-detector coherent scheme rather than the cheaper coincidence 
-methods used in the untargeted searches.
+The most likely scenario in which a coincident \ac{GW}--\ac{sGRB} event would
+be identified is through the targeted follow-up of an observed \ac{GW} or by
+post-facto matching of \ac{GW} trigger lists with known (or subsequently found
+\cite{2015ApJS..217....8B}) \ac{sGRB} events.  The likelihood of being able to
+follow-up \ac{GW} events using gamma-ray telescopes with low enough latency to
+catch a \ac{sGRB} is low.  Compounding this issue is the relatively large
+\ac{GW} sky error-box giving a field-of-view for the \ac{EM} observatories to
+search spanning $\sim 100$'s of square degrees (e.g.\
+\cite{2011CQGra..28j5021F,2013arXiv1304.0670L,2014ApJ...795..105S}).
+\refresp{For the scenario of a \ac{GW}-followup of an observed \ac{sGRB}, the
+merger time for \ac{BNS}/\ac{NSBH} systems will be estimated from the \ac{sGRB}
+to within a few seconds (see e.g.\ 340 the discussion in Sec.\ 2.2 of
+\cite{2012ApJ...760...12A})}. This makes the follow-up search less
+computationally expensive since it is performed over a
+smaller range of data using potentially fewer numbers of waveform templates.
+This computational saving enables the use of a more computationally expensive
+multi-detector coherent scheme rather than the cheaper coincidence methods used
+in the untargeted searches.
 
 % discuss the inclination angle issue
-A fortunate consequence of a joint \ac{GW}--\ac{sGRB} observation will be the fact that in order 
-for such an event to be observed, the \ac{BNS}/\ac{NSBH} system must have had its orbital angular 
-momentum vector pointing towards (or away) from the detector.  The actual inclination angle of the 
-system, defined as the angle between the orbital angular momentum vector and the line of sight, 
-must be $<$
-half of the \ac{sGRB} beaming angle.  This prerequisite property limits us to systems that are 
-approximately ``face-on'' and therefore biases us to higher \ac{SNR} signals.  However, as 
-discussed earlier, the property of beaming severely impacts the probable rate of such joint 
+A fortunate consequence of a joint \ac{GW}--\ac{sGRB} observation will be the
+fact that in order for such an event to be observed, the \ac{BNS}/\ac{NSBH}
+system must have had its orbital angular momentum vector pointing towards (or
+away) from the detector.  The actual inclination angle of the system, defined
+as the angle between the orbital angular momentum vector and the line of sight,
+must be $<$ half of the \ac{sGRB} beaming angle.  This prerequisite property
+limits us to systems that are approximately ``face-on'' \refresp{and therefore
+effectively optimally oriented. At given distances this biases us to  higher
+\ac{SNR} signals but it is more meaningful to think of this as an increase in
+the sensitivity range of the detectors}. However, as discussed earlier, the
+property of beaming severely impacts the probable rate of such joint
 observations.

### 2016-02-08 docs/response_to_referee.txt
Near final changes to the referee responses.

@@
 theta_j with a median of 10 degs. We have added text to the manuscipt to
-hopefully make this clearer. If the referee would like to see more detail on
-these we can send them further information.
+hopefully make this clearer.
 
-*this conservative nature is in part due to one of us (M Pitkin) mistakenly
-assuming the opening angle referred to the half width of the beam rather than
-the full beam width.
+*this conservative nature is in part due to mistakenly assuming the opening
+angle referred to the half width of the beam rather than the full beam width.
 
@@
 Response: No, unfortunately the redshift cannot be otained via the sGRB itself
-since it contains no transmission or absorbtion lines. CHRIS: Need a bit more
-than this.
+since it contains no specific spectral features to allow redshift
+determination, e.g. transmission or absorbtion lines.
 
@@
 distances. These give the reader an indication of the likely accuracy of an
-astrophysical calibration scheme. In "reality" we would first hope to be able
-to make this measurement for single detections in which case the true prior is out
-of our control and we must simply deal with what the universe gives us. In this
-case our assumed mass priors would be dominated by the information content of
-the likelihood due to the mass dependent phasing information.  Hence for
-individual detections we believe that any choice of conservative priors will
-have negligible effect on the result. If in the future one were to attempt this
-with a population of detections then it would be necessary to assume a
-hierachical scheme with some form for the mass distribution as the referee
-suggests.  However, we feel that this is beyond the scope of this work.
+astrophysical calibration scheme. We specifically discuss the effect of higher
+mass BH systems on our results just before Section VIA.  However, in that
+discussion we do not mention bias but rather the effect of different choices of
+source population where the appropriate prior would be used. With regards to
+bias, in "reality" we would first hope to be able to make a calibration
+measurement for single detections in which case the true source distribution is
+unknown to us. In this case our assumed mass priors would be dominated by the
+information content of the likelihood due to the mass dependent phasing
+information.  Hence for individual detections we believe that any choice of
+conservative priors will have negligible effect on the result. If in the future
+one were to attempt this with a population of detections then it would be
+necessary to assume a hierachical scheme with some form for the mass
+distribution as the referee suggests.  However, we feel that this is beyond the
+scope of this work.
 
@@
 
-Response: The choice mean p(C=0.1) = p(C=10). This has been fixed.
+Response: The choice means p(C=0.1) = p(C=10). This has been fixed.
 
@@
 
-Response: We have added the additional text stating "the weakest signals that could still
-be detected would always have a single detector \ac{SNR} of $\sim 5.5$ no matter
-their distance, hence the plateau.
+Response: We have added the additional text stating "the weakest signals that
+could still be detected would always have a single detector \ac{SNR} of $\sim
+5.5$ no matter their distance, hence the plateau.
 
-The issue of the V1 result not saturating is addressed in the manuscript in the lines
-"The largest SNR contribution will generally come from the two aLIGO detectors and thus
-the detection criteria (SNR threshold) will not apply to the AdV result. Hence, the SNR
-in AdV can be small and thus the ability to constrain its scale factor becomes poor (although it
-still provides information that the calibration is not grossly inaccurate)."
+The issue of the V1 result not saturating is addressed in the manuscript in the
+lines "The largest SNR contribution will generally come from the two aLIGO
+detectors and thus the detection criteria (SNR threshold) will not apply to the
+AdV result. Hence, the SNR in AdV can be small and thus the ability to
+constrain its scale factor becomes poor (although it still provides information
+that the calibration is not grossly inaccurate)."
 
@@
 
-Response: MATT: I think this is done now right?
+Response: We have ammended the plots to produce non-overlapping make the chirp
+mass tick labels, give the chirp mass units, and thicker red borders on the
+scale factor subplots.
 

### 2016-02-08 docs/cosmic_cal.tex
Near final changes to the referee responses.

@@
 guide to form a Gaussian prior distribution on the system inclination angle,
-$\iota$, by converting it into an equivalent standard devaition for a half-normal
-distribution. We therefore place a Gaussian prior on $\iota$ with zero mean and a standard deviation
-$\sigma=14.8^{\circ}$.} This, and the choice of polarisation angle prior range given below, 
-explicitly fixes the handedness of the system's rotation (or alternatively whether it is face-on or 
-back-on). If we were to allow the inclination to have two modes (one for face-on and one 
-for back-on), or alternatively to double the range of the polarisation angle, it would have no 
-effect on our ability to estimate the signal amplitude as the two modes should be identical with 
-regards to the amplitude parameter probability distributions. However, we are consistent between the 
-range of simulated signals discussed in Sec.~\ref{sec:simulations} and the initial ranges stated.
+$\iota$, by converting it into an equivalent standard deviation for a
+half-normal distribution. We therefore place a Gaussian prior on $\iota$ with
+zero mean and a standard deviation $\sigma=14.8^{\circ}$.} This, and the choice
+of polarisation angle prior range given below, explicitly fixes the handedness
+of the system's rotation (or alternatively whether it is face-on or back-on).
+If we were to allow the inclination to have two modes (one for face-on and one
+for back-on), or alternatively to double the range of the polarisation angle,
+it would have no effect on our ability to estimate the signal amplitude as the
+two modes should be identical with regards to the amplitude parameter
+probability distributions. However, we are consistent between the range of
+simulated signals discussed in Sec.~\ref{sec:simulations} and the initial
+ranges stated.
 

### 2016-02-23 docs/cosmic_cal.tex
Removed refresp macro as referee has accepted all changes

@@
 \newcommand{\dcc}{LIGO-P1500198}
-\newcommand{\refresp}[1]{\textcolor{red}{#1}}
 
@@
 detectors will detect waves emitted from $O(10s)$ of \acp{CBC} per year
-\refresp{\cite{2010CQGra..27q3001A}}. One such class of these cataclysmic
+\cite{2010CQGra..27q3001A}. One such class of these cataclysmic
 events, the inspiral and merger of \ac{BNS} systems will be detected out to a
-maximum range \refresp{(horizon distance)} of $\approx 450$ Mpc. Assuming the
+maximum range (horizon distance) of $\approx 450$ Mpc. Assuming the
 current best estimates for the cosmological parameters, this is equivalent to a
@@
 producing \ac{CBC} mergers of $8\times 10^{-9}$--$1.1\times
-10^{-6}$\,Mpc$^{-3}$\,yr$^{-1}$~\cite{2012MNRAS.425.2668C} \refresp{after accounting for
-beaming effects. This} can be compared to independently obtained estimates of
+10^{-6}$\,Mpc$^{-3}$\,yr$^{-1}$~\cite{2012MNRAS.425.2668C} after accounting for
+beaming effects. This can be compared to independently obtained estimates of
 \ac{BNS} merger rates of $10^{-8}$--$10^{-5}$\,Mpc$^{-3}$\,yr$^{-1}$
@@
 cosmological parameters, the true luminosity distance can be estimated to $\sim
-1\%$ accuracy~\refresp{\cite{2015arXiv150201589P}}.  The \ac{GW} measurement
+1\%$ accuracy~\cite{2015arXiv150201589P}.  The \ac{GW} measurement
 acts as a standard siren also giving us a direct measurement of the luminosity
@@
 carried out through a complicated system of physical manipulations of the
-Fabry-P\'{e}rot Michelson interferometer\refresp{,} where an elaborate feedback
+Fabry-P\'{e}rot Michelson interferometer, where an elaborate feedback
 system is used to sustain a defined measurement in arm length difference
@@
 \cite{Vitale:2012} and references therein. For \ac{aLIGO} and \ac{AdV}, at the
-frequency range used in this analysis \refresp{(20--400 Hz)}, the error in the
-\refresp{strain} amplitude calibration is expected to be roughly $10\%$
+frequency range used in this analysis (20--400 Hz), the error in the
+strain amplitude calibration is expected to be roughly $10\%$
 \cite{Vitale:2012}. This is a benchmark set for the error estimation using the
@@
 functions within a control feedback loop (see \cite{2010NIMPA.624..223A}),
-\refresp{each of which are subject to measurement uncertainties}. In this study
+each of which are subject to measurement uncertainties. In this study
 we will assume an estimate of $R$ is available (although in theory we could
@@
 where $h(f)$ is the true strain and $h_{\rm m}(f)$ is the measured strain
-\refresp{(i.e.\ the measured length response function $R(f)$ is related to the
+(i.e.\ the measured length response function $R(f)$ is related to the
 true length response function $R_{\rm true}(f)$ via $R(f) = \scf R_{\rm
-true}(f)$)}. With this definition it means that if $\scf > 1$ then a signal
-would appear to have a larger amplitude \refresp{(i.e.\ for a given system,
-would seem closer)} than in reality\footnote{However, the \ac{SNR} would be
+true}(f)$). With this definition it means that if $\scf > 1$ then a signal
+would appear to have a larger amplitude (i.e.\ for a given system,
+would seem closer) than in reality\footnote{However, the \ac{SNR} would be
 the correct value as the signal and the noise will both contain the same scale
@@
 function of the component masses) and the binary orientation with respect to
-the \refresp{global network of \ac{GW} detectors}. However, measurement of the
+the global network of \ac{GW} detectors. However, measurement of the
 phase evolution of such an event allows accurate determination of the chirp
@@
 the 3rd generation \ac{GW} interferometer, the
-\ac{ET}~\refresp{\cite{2010CQGra..27h4007P} and in ~\cite{2012PhRvD..85b3535T}
-for the Advanced detector network}. In this case, the redshift information
+\ac{ET}~\cite{2010CQGra..27h4007P} and in ~\cite{2012PhRvD..85b3535T}
+for the Advanced detector network. In this case, the redshift information
 obtained from the \ac{sGRB} host galaxy, provides a complementary measurement
@@
 instead use either the distribution of measured \acp{SNR} and assumed
-\refresp{form} of \ac{NS} mass and spatial distributions or exploit the
+form of \ac{NS} mass and spatial distributions or exploit the
 features of the tidal and post-merger hyper-massive \ac{NS} stages of the
@@
 cosmological parameter estimation (specifically for the Hubble constant, which
-is has uncertainties of $O(1\%)$~\refresp{\cite{2015arXiv150201589P}}) is
+is has uncertainties of $O(1\%)$~\cite{2015arXiv150201589P}) is
 accurately determined and therefore a known quantity. We then assume that the
@@
 associated redshifts is
-${\sim}1/3$~\refresp{\footnote{http://swift.gsfc.nasa.gov/archive/grb\_table/}}.
+${\sim}1/3$~\footnote{http://swift.gsfc.nasa.gov/archive/grb\_table/}.
 The range to which the \emph{Swift} \ac{XRT} has detected \acp{sGRB} is
@@
 The predicted detection rates of \ac{BNS} events (irrespective of an \ac{sGRB}
-counterpart) are derived from three main sources\refresp{:} population
+counterpart) are derived from three main sources: population
 synthesis models, extrapolations based on known galactic \ac{BNS} systems, and
@@
 also note the study mentioned in Sec.~\ref{sec:intro} that potentially lowers
-the expected rate by a factor of five.}. \refresp{A similar event rate of
+the expected rate by a factor of five.}. A similar event rate of
 $1$--$180$ was obtained in~\cite{2012MNRAS.425.2668C}, where the rate of
 observed \acp{sGRB} was converted to the rate of \ac{BNS} events by assuming a
-beaming angle of $15^{\circ}$.} The beaming angle is the primary factor in
+beaming angle of $15^{\circ}$. The beaming angle is the primary factor in
 determining the fraction of detected \ac{GW} \ac{BNS} signals with an \ac{EM}
@@
 \cite{2011CQGra..28j5021F,2013arXiv1304.0670L,2014ApJ...795..105S}).
-\refresp{For the scenario of a \ac{GW}-followup of an observed \ac{sGRB}, the
+For the scenario of a \ac{GW}-followup of an observed \ac{sGRB}, the
 merger time for \ac{BNS}/\ac{NSBH} systems will be estimated from the \ac{sGRB}
 to within a few seconds (see e.g.\ 340 the discussion in Sec.\ 2.2 of
-\cite{2012ApJ...760...12A})}. This makes the follow-up search less
+\cite{2012ApJ...760...12A}). This makes the follow-up search less
 computationally expensive since it is performed over a
@@
 must be $<$ half of the \ac{sGRB} beaming angle.  This prerequisite property
-limits us to systems that are approximately ``face-on'' \refresp{and therefore
+limits us to systems that are approximately ``face-on'' and therefore
 effectively optimally oriented. At given distances this biases us to  higher
 \ac{SNR} signals but it is more meaningful to think of this as an increase in
-the sensitivity range of the detectors}. However, as discussed earlier, the
+the sensitivity range of the detectors. However, as discussed earlier, the
 property of beaming severely impacts the probable rate of such joint
@@
 are dependent upon the polarisation angle $\psi$ and the sky position of the
-source \refresp{with right ascension} $\alpha$ and \refresp{declination} $\delta$.
+source with right ascension $\alpha$ and declination $\delta$.
 
@@
 in~\cite{2014ApJ...780..118F} provides a median opening angle value of $\sim
-10^{\circ}$. \refresp{We use this median opening angle as a rather conservative
+10^{\circ}$. We use this median opening angle as a rather conservative
 guide to form a Gaussian prior distribution on the system inclination angle,
@@
 half-normal distribution. We therefore place a Gaussian prior on $\iota$ with
-zero mean and a standard deviation $\sigma=14.8^{\circ}$.} This, and the choice
+zero mean and a standard deviation $\sigma=14.8^{\circ}$. This, and the choice
 of polarisation angle prior range given below, explicitly fixes the handedness
@@
 also (na\"{i}vely) assume that the prior \ac{PDF} of the calibration scale factor is either
-larger or smaller than unity by an equivalent factor, e.g.\ \refresp{$p(\scf=10|I)$ is
-equivalent to $p(\scf=0.1|I)$}. For the scale factor in each detector we 
+larger or smaller than unity by an equivalent factor, e.g.\ $p(\scf=10|I)$ is
+equivalent to $p(\scf=0.1|I)$. For the scale factor in each detector we 
 therefore use the following log-normal distribution as our prior
@@
 conservative prior for signal recovery that had little impact in the results, i.e.\
-to produce results domintaed by the likelihood for these parameters. \refresp{However,
+to produce results domintaed by the likelihood for these parameters. However,
 limited studies using $\scf$ values within the much larger range of
 $\sim 0.1$ to 10 show distributions with consistent relative widths to
-those we find in our main study.}}.
+those we find in our main study.}.
 
@@
 64 separate noisy realisations of the \ac{PSD} and scaled
-with the same \refresp{calibration} scale factor as applied to the injection and noise\footnote{We do not account for there 
+with the same calibration scale factor as applied to the injection and noise\footnote{We do not account for there 
 being a potential difference
@@
 an upper limit on the calibration scale factor.}. This has provided posterior probability
-distributions\refresp{\footnote{\refresp{As a proxy to check for convergence of the MCMC chains
-we check that the calibration scale factor posterior histograms do not contain many disjoint modes.}}}
+distributions\footnote{As a proxy to check for convergence of the MCMC chains
+we check that the calibration scale factor posterior histograms do not contain many disjoint modes.}
 on the calibration scale factors for each detector.  Examples of
@@
  parameters of a \ac{BNS} system, including calibration scaling factors (surrounded by the 
-\refresp{thicker} red 
+thicker red 
 borders) for the three detectors (H1, L1 and V1). The simulated signal was at a distance of 
@@
 those sources with a high enough \ac{SNR} that we would consider them
-detectable\refresp{, i.e.\ the weakest signals that could still be detected would
+detectable, i.e.\ the weakest signals that could still be detected would
 always have a single detector \ac{SNR} of $\sim 5.5$ no matter their distance, hence the
-plateau}. \refresp{In addition there will also} be fewer sources with \ac{SNR} higher than
+plateau. In addition there will also be fewer sources with \ac{SNR} higher than
 this criteria at large distances, giving us a narrower range, and we
@@
 larger distances the \acp{PDF} on \scft for V1 are not well constrained and start to become dominated
-by the prior. As we have previously noted this prior is not \refresp{the} same as the distribution used
+by the prior. As we have previously noted this prior is not the same as the distribution used
 to generate the \scft values, but is instead far broader and more conservative. So, this bias is

### 2016-03-03 docs/cosmic_cal.tex
Some minor changes based on proofs received

@@
 calibration. We use simulations of individual detectable gravitational-wave signals from binary 
-neutron star (BNS) or neutron star-black hole (NSBH) systems, which we assume to be found in 
+neutron star (BNS) or neutron star-black hole systems, which we assume to be found in 
 coincidence with short gamma-ray bursts, to estimate any discrepancy in the overall scaling of the 
@@
 \acrodef{CBC}[CBC]{compact binary coalescence}
-\acrodef{EM}[EM]{electro-magnetic}
+\acrodef{EM}[EM]{electromagnetic}
 \acrodef{aLIGO}[aLIGO]{Advanced LIGO}
@@
 ``standard sirens'' (analogous to \ac{EM} standard candles). The nature of
-these sources will allow us to estimate, amongst other parameters, their
+these sources will allow us to estimate, among other parameters, their
 luminosity distance without the need for calibration against the cosmic
@@
 momentum vector of the system. Additional evidence for the coincidence of
-\acp{sGRB} with \ac{GW} events are the estimated astrophysical rates of both
+\acp{sGRB} with \ac{GW} events is the estimated astrophysical rates of both
 phenomena. Observations of \acp{sGRB} give an inferred rate of \ac{sGRB}
@@
 \cite{2015MNRAS.448..928K}, which studied known galactic \ac{BNS} systems in
-more detail suggest a rate approximately five times lower.}.
+more detail, suggests a rate approximately five times lower.}.
 
@@
 calibrated.  Hence, by comparison with the distance estimate from the \ac{sGRB}
-we can re-calibrate (or validate) the existing experimentally obtained
+we can recalibrate (or validate) the existing experimentally obtained
 calibration.
@@
 sirens and their proposed coincident \ac{EM} signatures, the \acp{sGRB} in
-Secs.~\ref{sec:sirens} and~\ref{sec:GRB}. In Secs.~\ref{sec:analysis} and \ref{sec:results} we 
+Secs.~\ref{sec:sirens}--\ref{sec:GRB}. In Secs.~\ref{sec:analysis}--\ref{sec:results} we 
 discuss our analysis method and the results, and in Sec.~\ref{sec:discussion} we 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Gravitational wave detector calibration\label{sec:calibration}}
+\section{Gravitational-wave detector calibration\label{sec:calibration}}
 
@@
 where we assume $R$ varies only slowly in time (in comparison to transient
-signal time-scales). Calculation of $R$ requires the measurement of various
+signal time scales). Calculation of $R$ requires the measurement of various
 functions within a control feedback loop (see \cite{2010NIMPA.624..223A}),
 each of which are subject to measurement uncertainties. In this study
-we will assume an estimate of $R$ is available (although in theory we could
+we assume an estimate of $R$ is available (although in theory we could
 take on the role of estimating $R$ itself), but that it differs from the truth
@@
 factor.}, whereas if $\scf < 1$ it would appear to have a lower amplitude than
-reality.  In this analysis we will simplify the situation by assuming that
+reality.  In this analysis we simplify the situation by assuming that
 $\scf$ is a constant with respect to frequency (and is real, so has no phase
 component), but in future studies that assumption could be dropped and $\scf$
-could take some functional form, or piece-wise fit, with respect to $f$ (e.g.\
+could take some functional form, or piecewise fit, with respect to $f$ [e.g.\
 in a similar way to the method of \cite{2013PhRvD..88h4044L} used for fitting
-differences in \ac{PSD} estimates).
+differences in \ac{PSD} estimates].
 
@@
 in~\cite{1986Natur.323..310S}. Unlike the primary \ac{EM} standard candle
-event, Type Ia Supernovae, the measured luminosity of a \ac{CBC} in \acp{GW} is
+event, type Ia supernovae, the measured luminosity of a \ac{CBC} in \acp{GW} is
 not only a function of distance. It also depends upon the chirp mass (a
@@
 they give us a direct measure of the absolute luminosity distance to sources.
-This is distinct from Type Ia Supernovae standard candles that only provide
+This is distinct from type Ia supernovae standard candles that only provide
 relative luminosity distance measures and require calibration via other methods
@@
 \ac{sGRB} counterparts for use as a cosmological tool was investigated
-in~\cite{2010CQGra..27u5006S} and~\cite{2011PhRvD..83b3005Z} with respect to
-the 3rd generation \ac{GW} interferometer, the
-\ac{ET}~\cite{2010CQGra..27h4007P} and in ~\cite{2012PhRvD..85b3535T}
+in~\cite{2010CQGra..27u5006S}--\cite{2011PhRvD..83b3005Z} with respect to
+the third generation \ac{GW} interferometer, the
+Einstein Telescope~\cite{2010CQGra..27h4007P} and in ~\cite{2012PhRvD..85b3535T}
 for the Advanced detector network. In this case, the redshift information
-obtained from the \ac{sGRB} host galaxy, provides a complementary measurement
+obtained from the \ac{sGRB} host galaxy provides a complementary measurement
 allowing each event to inform the distance-redshift relationship and hence
@@
 form of \ac{NS} mass and spatial distributions or exploit the
-features of the tidal and post-merger hyper-massive \ac{NS} stages of the
+features of the tidal and postmerger hypermassive \ac{NS} stages of the
 \ac{GW} waveform.
@@
 described in this manuscript we are essentially turning the standard
-cosmological problem upside-down. We assume that the existing \ac{EM}
-cosmological parameter estimation (specifically for the Hubble constant, which
-is has uncertainties of $O(1\%)$~\cite{2015arXiv150201589P}) is
+cosmological problem upside down. We assume that the existing \ac{EM}
+cosmological parameter estimation [specifically for the Hubble constant, which
+has uncertainties of $O(1\%)$~\cite{2015arXiv150201589P}] is
 accurately determined and therefore a known quantity. We then assume that the
@@
 merger time for \ac{BNS}/\ac{NSBH} systems will be estimated from the \ac{sGRB}
-to within a few seconds (see e.g.\ 340 the discussion in Sec.\ 2.2 of
+to within a few seconds (see e.g.\ the discussion in Sec.\ 2.2 of
 \cite{2012ApJ...760...12A}). This makes the follow-up search less
@@
 limits us to systems that are approximately ``face-on'' and therefore
-effectively optimally oriented. At given distances this biases us to  higher
+effectively optimally oriented. At given distances this makes us biased towards higher
 \ac{SNR} signals but it is more meaningful to think of this as an increase in

### 2016-03-04 docs/cosmic_cal.tex
Some futher minor chnages from the proofing stage

@@
 observable for a fraction of the mergers. If the event exhibits an optical
-afterglow then the host galaxy can be identified and from which a redshift can
+afterglow then the host galaxy can be identified, from which a redshift can
 be obtained (e.g.~\cite{2005Natur.437..851G}).  The fraction of events with
@@
 ${\sim}1/3$~\footnote{http://swift.gsfc.nasa.gov/archive/grb\_table/}.
-The range to which the \emph{Swift} \ac{XRT} has detected \acp{sGRB} is
+The range to which the \emph{Swift} x-ray telescope has detected \acp{sGRB} is
 $z{\sim}2$ and the nearest event with associated redshift is at
@@
 synthesis models, extrapolations based on known galactic \ac{BNS} systems, and
-from observations of \ac{sGRB} events. Various studies using the first two
+observations of \ac{sGRB} events. Various studies using the first two
 methods have been compiled into an overall detection rate of $0.2$--$200$
@@
 year~\cite{2013arXiv1304.0670L,2010CQGra..27q3001A}\footnote{We note that this
-rate is that given in Table 1 of \cite{2013arXiv1304.0670L} for an \ac{aLIGO}
+rate is that given in Table I of \cite{2013arXiv1304.0670L} for an \ac{aLIGO}
 and \ac{AdV} network consisting to two \ac{aLIGO} detectors operating at design
@@
 be identified is through the targeted follow-up of an observed \ac{GW} or by
-post-facto matching of \ac{GW} trigger lists with known (or subsequently found
+\emph{post facto} matching of \ac{GW} trigger lists with known (or subsequently found
 \cite{2015ApJS..217....8B}) \ac{sGRB} events.  The likelihood of being able to
@@
 catch a \ac{sGRB} is low.  Compounding this issue is the relatively large
-\ac{GW} sky error-box giving a field-of-view for the \ac{EM} observatories to
-search spanning $\sim 100$'s of square degrees (e.g.\
+\ac{GW} sky error box giving a field of view for the \ac{EM} observatories to
+search spanning $\sim 100$s of square degrees (e.g.\
 \cite{2011CQGra..28j5021F,2013arXiv1304.0670L,2014ApJ...795..105S}).
@@
 and references therein) with a 3.5 post-Newtonian expansion in phase for
-modelling our signal (in both simulations and signal recovery).  For \ac{BNS}
-systems we use non-spinning waveforms under the assumption that the component spins will be small 
+modeling our signal (in both simulations and signal recovery).  For \ac{BNS}
+systems we use nonspinning waveforms under the assumption that the component spins will be small 
 and have a negligible effect on the analysis\footnote{In future studies we could create our 
-simulations including small spins and confirm that it does not have a significant effect on our 
-results if recovery is with non-spinning templates.} (see discussion in \cite{2014ApJ...795..105S} 
+simulations including small spins and confirm that this does not have a significant effect on our 
+results if recovery is with nonspinning templates.} (see discussion in \cite{2014ApJ...795..105S} 
 and studies in \cite{2012PhRvD..86h4017B}). For \ac{NSBH} systems we use a spinning waveform, but in 
@@
 
-For a non-spinning system the general form of the frequency-domain polarisation
+For a nonspinning system the general form of the frequency-domain polarization
 amplitudes is
@@
 %
-where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=Mq^{3/5}$,
+where the chirp mass $\mathcal{M}$ is defined as $\mathcal{M}=Mq^{3/5}$,
 with the symmetric mass ratio $q=m_{1}m_{2}/M^{2}$, total mass
@@
 where $F^{k}_{+}$ and $F^{k}_{\times}$ are the antenna response functions which
-are dependent upon the polarisation angle $\psi$ and the sky position of the
+are dependent upon the polarization angle $\psi$ and the sky position of the
 source with right ascension $\alpha$ and declination $\delta$.
@@
 % Define the likelihood function we will be using
-We consider that the measured data in any detector is defined as
+We consider that the measured data in any detector are defined as
 %
@@
 $i_{\rm low}$ and $i_{\rm high}$ corresponding to the lower and upper range in
-frequencies used for our analysis, which were 20\,Hz to 400\,Hz respectively.
+frequencies used for our analysis, which were 20 to 400\,Hz respectively.
 This frequency range was chosen as the vast majority of the \ac{SNR} for the
@@
 by using a 10 to 1500\,Hz frequency range), so going to lower
-or higher frequencies provides very little additional information whilst increasing
+or higher frequencies provides very little additional information while increasing
 the computational cost\footnote{In a real analysis, rather than this case study, the additional
@@
 Other information relevant for a coincident \ac{sGRB} is that the system is
-likely to be relatively close to face-on (and therefore nearly circularly
-polarised), which allows us to place prior constraints on the system
+likely to be relatively close to face on (and therefore nearly circularly
+polarized), which allows us to place prior constraints on the system
 inclination angle. Beam opening angles for \acp{sGRB} are relatively poorly
 constrained as they are based on only a few jet-break detections. A comparative
-study of estimates and lower-limits of opening angles for 13 \acp{sGRB} given
+study of estimates and lower limits of opening angles for 13 \acp{sGRB} given
 in~\cite{2014ApJ...780..118F} provides a median opening angle value of $\sim
@@
 zero mean and a standard deviation $\sigma=14.8^{\circ}$. This, and the choice
-of polarisation angle prior range given below, explicitly fixes the handedness
-of the system's rotation (or alternatively whether it is face-on or back-on).
-If we were to allow the inclination to have two modes (one for face-on and one
-for back-on), or alternatively to double the range of the polarisation angle,
+of polarization angle prior range given below, explicitly fixes the handedness
+of the system's rotation (or alternatively whether it is face on or back on).
+If we were to allow the inclination to have two modes (one for face on and one
+for back on), or alternatively to double the range of the polarization angle,
 it would have no effect on our ability to estimate the signal amplitude as the
@@
 spins are small enough that they will be negligible). We use a uniform prior on
-the normalised aligned spin magnitude between $-1$ and $1$.
+the normalized aligned spin magnitude between $-1$ and $1$.
 
@@
 from the identification of the host galaxy. This fixes the time delay between detectors to a known 
-value and, along with the polarisation angle, defines the antenna response patterns.
+value and, along with the polarization angle, defines the antenna response patterns.
 
@@
 $\phi_{\mathrm{c}}$ a uniform prior on the range $(0,\pi]$ is used, and for the
-polarisation angle we use a uniform prior on the range $(0,\pi/2]$. As the
-signals we use are generally close to being circularly polarised (face-on) the
-phase and polarisation angle will be largely degenerate and their exact
+polarization angle we use a uniform prior on the range $(0,\pi/2]$. As the
+signals we use are generally close to being circularly polarized (face on) the
+phase and polarization angle will be largely degenerate and their exact
 combination will have little effect on the result.
@@
 We make the assumption that in general the calibration applied to the detector
-data will be close to being correct, so want to use a prior that is peaked at $\scf_k=1$. We
+data will be close to being correct, so we want to use a prior that is peaked at $\scf_k=1$. We
 also (na\"{i}vely) assume that the prior \ac{PDF} of the calibration scale factor is either
@@
 equivalent to $p(\scf=0.1|I)$. For the scale factor in each detector we 
-therefore use the following log-normal distribution as our prior
+therefore use the following log-normal distribution as our prior,
 %
@@
 prior does not give a symmetric amount of probability about unity (so our prior
-favours \scf\xspace values greater than 1), but given our simulation criterion
-described in Section~\ref{sec:simulations} we find that the likelihood
+favors \scf\xspace values greater than 1), but given our simulation criterion
+described in Sec.~\ref{sec:simulations} we find that the likelihood
 generally overwhelms the prior and our na\"{i}ve assumptions have little effect.
@@
 \ac{BNS} and \ac{NSBH} signals. These were simulated
-at a range of distances: 50\,Mpc to 500\,Mpc with 50\,Mpc increments for the
-\ac{BNS} signals and 100\,Mpc to 900\,Mpc with 100\,Mpc increments for the
+at a range of distances: 50 to 500\,Mpc with 50\,Mpc increments for the
+\ac{BNS} signals and 100 to 900\,Mpc with 100\,Mpc increments for the
 \ac{NSBH} signals. The simulations used the two \ac{aLIGO} detectors (H1 and
 L1) and the \ac{AdV} detector (V1) and were all performed in the frequency
-domain and spanned a frequency range from 20\,Hz to 400\,Hz, which is a range that contains the 
-vast majority of the \ac{SNR} and will therefore be all that is required to provide good 
+domain and spanned a frequency range from 20 to 400\,Hz, which is a range that contains the 
+vast majority of the \ac{SNR} and is therefore all that is required to provide good 
 constraints on \scf. We assumed all
-detectors were operating at their design sensitivity (as given by Figure~1 of
-\cite{2013arXiv1304.0670L}). As such we added coloured Gaussian noise to each
+detectors were operating at their design sensitivity (as given by Fig.~1 of
+\cite{2013arXiv1304.0670L}). As such we added colored Gaussian noise to each
 simulation based on the sensitivity curve, but scaled with the associated
@@
 Separately for the \ac{BNS} and \ac{NSBH} systems we have performed a large
-number (several hundred up to $O(1000)$) of simulations at each distance value.
+number [several hundred up to $O(1000)$] of simulations at each distance value.
 When proposing injections for each simulation the source parameters were
 randomly drawn from the prior distributions described in
-Section~\ref{sec:priors}, with the sky position being drawn randomly from a
+Sec.~\ref{sec:priors}, with the sky position being drawn randomly from a
 uniform distribution on the sky, the coalescence times held fixed within the
-centre of the prior window, and with the calibration scaling factors for each
+center of the prior window, and with the calibration scaling factors for each
 detector drawn from a Gaussian distribution with a mean of one and standard
@@
 that this distribution is broadly consistent with the scale of calibration
-uncertainties that are expected via experimental methods, but is not the same as
-the prior that we assume on $\scf$ (see Eq.~\ref{eq:scale_prior}) when recovering 
+uncertainties that is expected via experimental methods, but is not the same as
+the prior that we assume on $\scf$ [see Eq.~\ref{eq:scale_prior}] when recovering 
 signals\footnote{The reason behind this is that we wanted to use a fairly na\"{i}ve and
@@
 % describe the detection cut 
-In accepting a proposed simulation as one to be analysed we introduced a
-criteria that the signal be ``detectable'', based on that used in
-\cite{2012PhRvD..85h2002A}. We therefore only analysed simulations that had an
+In accepting a proposed simulation as one to be analyzed we introduced a
+criterion that the signal be ``detectable'', based on that used in
+\cite{2012PhRvD..85h2002A}. We therefore only analyzed simulations that had an
 \ac{SNR} of $\geq 5.5$ in at least two detectors. Since we assume that these
@@
 Therefore, at larger distances the population of injected
-sources was not uniformly distributed over the sky and instead favoured locations with better
+sources was not uniformly distributed over the sky and instead favored locations with better
 antenna response.  Similarly, despite already being constrained by the prior to
 small inclination angles, the accepted simulations show preference for sources
-very close to circularly polarised.
+very close to circularly polarized.
 
@@
 a roughly Gaussian probability distribution) should be given by $\sim 1/{\rm
-SNR}$ for each detector. As we will show below this assumption is reasonable,
+SNR}$ for each detector. As we show below this assumption is reasonable,
 but small correlations do exist between parameters meaning that it does not
@@
 unknown source parameters using the priors as discussed in
-Section~\ref{sec:priors}. In each case when calculating the likelihood we used
+Sec.~\ref{sec:priors}. In each case when calculating the likelihood we used
 an estimate of the noise \ac{PSD} based on the advanced detector design
 sensitivities (using those given in~\cite{2013arXiv1304.0670L}), but calculated by averaging 
-64 separate noisy realisations of the \ac{PSD} and scaled
+64 separate noisy realizations of the \ac{PSD} and scaled
 with the same calibration scale factor as applied to the injection and noise\footnote{We do not account for there 
@@
 on the calibration scale factors for each detector.  Examples of
-the marginalised posterior probabilities for a \ac{BNS} system and a \ac{NSBH}
+the marginalized posterior probabilities for a \ac{BNS} system and a \ac{NSBH}
 system observed with the three detector network are shown in
-figures~\ref{fig:bnspost} and \ref{fig:nsbhpost} respectively.
+Figs.~\ref{fig:bnspost} and \ref{fig:nsbhpost} respectively.
 
@@
 (i.e.\ $1\sigma$) of these scale factors' confidence intervals compared to the
-true value. These are shown as boxplots in figures~\ref{fig:bnsresults} and
+true value. These are shown as boxplots in Figs.~\ref{fig:bnsresults} and
 \ref{fig:nsbhresults} for the \ac{BNS} and \ac{NSBH} systems respectively. The
-boxes show the extent from the lower to upper quartile of the values, whilst
-the whiskers extend from the 5$^{\rm th}$ to 95$^{\rm th}$ percentile. The
+boxes show the extent from the lower to upper quartile of the values, while
+the whiskers extend from the fifth to 95$^{\rm th}$ percentile. The
 black line within each box gives the median value and the star gives the mean
 value. Also shown on each plot as the dashed magenta line is the percentage of
-signals drawn from the prior distribution that fulfill the detection criterion.
+signals drawn from the prior distribution that fulfils the detection criterion.
 
@@
  \end{center}
- \caption{\label{fig:bnspost} The marginalised posterior probability distributions for some of the unknown
+ \caption{\label{fig:bnspost} The marginalized posterior probability distributions for some of the unknown
  parameters of a \ac{BNS} system, including calibration scaling factors (surrounded by the 
@@
  \end{center}
- \caption{\label{fig:nsbhpost} The marginalised posterior probability distributions for some of the unknown
+ \caption{\label{fig:nsbhpost} The marginalized posterior probability distributions for some of the unknown
  parameters of a \ac{NSBH} system, including calibration scaling factors for the three detectors 
@@
  \caption{\label{fig:bnsresults} Distributions of the percentage accuracy at which the calibration 
-scale factors can be determined for a three detector network if using
-\ac{BNS} systems (provided a coincident \ac{GRB} is observed and can yield a
+scale factors can be determined for a three detector network using
+\ac{BNS} systems (provided a coincident GRB is observed and can yield a
 distance estimate). The boxplots span the lower to upper quartile range of the
@@
 scale factors can be determined for a three detector network if using \ac{NSBH} systems 
-(provided a coincident \ac{GRB} is observed and can yield a distance estimate). The plot contents 
-are the same as in figure~\ref{fig:bnsresults}.}
+(provided a coincident GRB is observed and can yield a distance estimate). The plot contents 
+are the same as in Fig.~\ref{fig:bnsresults}.}
 \end{figure*}
@@
 % discuss the BNS plots - general behaviour
-In figure~\ref{fig:bnsresults} we see that on average the scale factor can be
+In Fig.~\ref{fig:bnsresults} we see that on average the scale factor can be
 recovered to equivalent precision in both \ac{aLIGO} detectors, as would be
@@
 $\gtrsim 250$\,Mpc the upper extent of the uncertainty for H1 and L1 hits a
-maximum at $\sim 25\%$, whilst the width of boxes narrows. This is due to our
-``detectibility'' criteria, where for all distances we will only be seeing
+maximum at $\sim 25\%$, while the width of boxes narrows. This is due to our
+``detectibility'' criteria, where for all distances we only see
 those sources with a high enough \ac{SNR} that we would consider them
@@
 $\iota$ even within the constrained prior range, and slight correlations
-between the scale factors for H1 and L1 as seen in figure~\ref{fig:bnspost}.
+between the scale factors for H1 and L1 as seen in Fig.~\ref{fig:bnspost}.
 
 % discuss the NSBH sources
-In figure~\ref{fig:nsbhresults} we see very similar results for the \ac{NSBH}
+In Fig.~\ref{fig:nsbhresults} we see very similar results for the \ac{NSBH}
 systems although the higher \ac{SNR} of the signals means that we can provide
-$\sim 20\mbox{--}25\%$ uncertainties on the calibration scale factors for H1
+$\sim 20\%\mbox{--}25\%$ uncertainties on the calibration scale factors for H1
 and L1 out to greater distances. The distributions generally appear slightly
@@
 results.  The second factor is that there appears to be a modest effect due to
-the population of sources that are detectable. For distances at which the
+the population of sources that is detectable. For distances at which the
 \ac{BNS} and \ac{NSBH} systems would give the same percentage of observable
@@
 10$\,M$_{\odot}$. We expect these would produce similar results to those we see
-in figure~\ref{fig:nsbhresults}, but again extending to further distances. However, at higher 
+in Fig.~\ref{fig:nsbhresults}, but again extending to further distances. However, at higher 
 masses we would encounter the problem that the chirp mass will be less well constrained due to 
@@
 
-It is also interested to see if our analysis has any bias on the recovered \scft \acp{PDF}, or whether they
+It is also interesting to see if our analysis has any bias on the recovered \scft \acp{PDF}, or whether they
 are recovered in a way that is consistent with the simulated values.
@@
 of this type of plot to check consistency of \ac{GW} analyes)
-for the \ac{BNS} systems at 50 and 500\,Mpc (see figure~\ref{fig:ppplots}).
+for the \ac{BNS} systems at 50 and 500\,Mpc (see Fig.~\ref{fig:ppplots}).
 Self-consistent \acp{PDF} should show that a fraction of injected values falls within the corresponding credible interval,
@@
 that for sources at 50\,Mpc there is no highly significant bias on the recovery of \scft for any of
-the detectors. However, in figure~\ref{subfig:pp500}, with sources at 500\,Mpc, whilst
+the detectors. However, in Fig.~\ref{subfig:pp500}, with sources at 500\,Mpc, while
 H1 and L1 follow the diagonal well, V1 goes significantly above the diagonal.
@@
 by the prior. As we have previously noted this prior is not the same as the distribution used
-to generate the \scft values, but is instead far broader and more conservative. So, this bias is
-showing up due to the simulated values of \scft being generated from a much narrow range.
+to generate the \scft values, but is instead far broader and more conservative. So this bias is
+showing up due to the simulated values of \scft being generated from a much more narrow range.
 
 Despite the inconsistency in V1 it is satisfying to see that for the detectors for which the
-\scft values are reasonably constrained (i.e.\ H1 and L1 in figure~\ref{fig:bnsresults}) the
+\scft values are reasonably constrained (i.e.\ H1 and L1 in Fig.~\ref{fig:bnsresults}) the
 actual value is recovered consistently, with little effect from the prior.
@@
 \end{center}
-\caption{\label{fig:ppplots} Both figures shows the cumulative fraction of true \scft values
+\caption{\label{fig:ppplots} Both figures show the cumulative fraction of true \scft values
 found within a
-given fractional credible interval versus fractional credible interval for each detector for
+given fractional credible interval versus the fractional credible interval for each detector for
 the \ac{BNS} simulations at (\subref{subfig:pp50}) 50\,Mpc and (\subref{subfig:pp500})
@@
 inaccurate and is fairly consistent between detectors.} consistency check that the calibration is 
-reasonably accurate in addition to the existing calbiration methods \cite{2010NIMPA.624..223A}.  We acknowledge that we have made a number of simplifying assumptions, 
+reasonably accurate in addition to the existing calibration methods \cite{2010NIMPA.624..223A}.
+We acknowledge that we have made a number of simplifying assumptions, 
 including that the waveforms we use accurately model the true signal, and that the overall 
@@
 sources and in future work we intend to apply related methods for astrophysical
-calibration.  We also have neglected any phase uncertainty in the calibration.
+calibration.  We have also neglected any phase uncertainty in the calibration.
 
@@
 A further application of astrophysical calibration is the relative calibration
-of \ac{GW} detectors.  In this case we would be interested in obtaining
+of \ac{GW} detectors.  In this case we are interested in obtaining
 posterior distributions on the ratios of calibration scale factors and phases between
@@
 the sky position estimates for \ac{CBC} and burst sources (Vitale, Messenger \& Pitkin, in preparation). In this case,
-amplitude information is a secondary, but important factor, in breaking
+amplitude information is a secondary, but important, factor in breaking
 degeneracies on sky position estimates.
@@
 % discuss PSD uncertainty
-We emphasise that our work has assumed that the \ac{PSD} used in the likelihood
+We emphasize that our work has assumed that the \ac{PSD} used in the likelihood
 function is an accurate representation of the \ac{PSD} at the time of the
@@
 each. However, since the scale factor uncertainty is related to the \ac{SNR}
-then the \ac{SNR} contribution in each frequency section will be of relevance.
+then the \ac{SNR} contribution in each frequency section is of relevance.
 In this case Bayesian model selection may be applicable for selecting the
@@
 analysis. Indeed these methods, and those of \cite{Vitale:2012} are already
-being used for marginalising over calibration uncertainties, but do not yet
+being used to marginalize over calibration uncertainties, but do not yet
 attempt to estimate them.
@@
 the inclination of the system is not known and distances are fairly uncertain, so a precise 
-calibration assessment may not be possible. Although there is at least one currently known 
+calibration assessment may not be possible. However, there is at least one currently known 
 eclipsing white dwarf binary (J0651) \cite{2041-8205-737-1-L23}, which does provide the system 
@@
 
-We would like to acknowledge useful discussions within the LIGO-Virgo
-Collaboration and specifically with Salvatore Vitale. LW was part funded for
+We acknowledge useful discussions within the LIGO-Virgo
+Collaboration and specifically with Salvatore Vitale. L.W.\ was part funded for
 this work through a Royal Astronomical Society Undergraduate Research Bursary.
-MP and CM are funded by the Science and Technology Facilities Council (STFC) under grant
-number ST/L000946/1. CM is also supported by a Glasgow University Lord Kelvin
-Adam Smith Fellowship. We are grateful for computational resources provided by
-Cardiff University, and funded by STFC grant ST/I006285/1 supporting UK Involvement in
-the operation of Advanced LIGO. This document has LIGO DCC number \dcc.
+M.P.\ and C.M.\ are funded by the Science and Technology Facilities Council (STFC) under Grant.
+No.\ ST/L000946/1. C.M.\ is also supported by a Glasgow University Lord Kelvin
+Adam Smith fellowship. We are grateful for computational resources provided by
+Cardiff University, and funded by STFC Grant No.\ ST/I006285/1 supporting United Kingdom involvement in
+the operation of Advanced LIGO. This document has LIGO DCC Report No. \dcc.
 
